### 2018-09-28 index.Rmd
initial commit

---
title: Case study validations of automatic bullet matching

# to produce blinded version set to 1
blinded: 0

authors: 
- name: Author 1
  thanks: The authors gratefully acknowledge XXX NIST funding for CSAFE XXX access to bullets from Hamby set 44, XXX Tylor, Melissa and Kasi. Scanning. Hamby advice. 
  affiliation: Department of YYY, University of XXX

- name: Author 2
  affiliation: Department of ZZZ, University of WWW

keywords:
- 3 to 6 keywords
- that do not appear in the title

abstract: |

bibliography: bibliography
biblio-style: apsr

output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: template.tex

---
\newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
\noindent

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  error = FALSE, 
  collapse = TRUE,
  comment = "",
  fig.height = 6,
  fig.width = 6,
  fig.align = "center",
  out.width= '\\textwidth',
  cache = FALSE,
  fig.path='figures/',
  echo=FALSE
)
options(knitr.table.format = "latex")

library(tidyverse)
library(scales)
library(multidplyr) # install_github("hadley/multidplyr")
library(bulletxtrctr) # install_github("csafe-isu/bulletr")
library(gridExtra)
library(kableExtra)
```

```{r functions}
estimateBeta <- function(x, method = "mm", probs=c(0.25, 0.75)) {
  if (method =="mm") {
    mu = mean(x, na.rm=TRUE)
    sd = sd(x, na.rm=TRUE)
    lambda <- mu/sd*(1-mu)/sd - 1
    alpha <- mu*lambda
    beta <- (1-mu)*lambda
  }
  if (method =="LB") {
    quantile1=list(p=probs[1],x=quantile(probs=probs[1], x))
    quantile2=list(p=probs[2],x=quantile(probs=probs[2], x))
    bs <- LearnBayes::beta.select(quantile1,quantile2)
    alpha = bs[1]
    beta = bs[2]
  }
  return(c(alpha=alpha, beta=beta))
}
```
# Idea of the paper

In \citet{aoas} an algorithm has been laid out to quantify the similarity between land engraved areas of bullets. 
In this paper we present results of the algorithm on several case studies where ground truth is known.

# Introduction and Background
In current practice, firearms and toolmark examiners (FTE) evaluate the similiarity of striae on bullets by placing the evidence together with a bullet from the suspected gun under a comparison microscope and visually classify similarity according to the theory of firearms identification \citep{identification} as one of identification, inconclusive or exclusion. Exact guidelines for this classification vary from lab to lab, e.g.\ some labs will exclude only on the basis of non-matching class characteristics (such as direction of the twist in rifling, land length or number of lands, or type of rifling). In some labs,  CMS (consecutively matching striae) as defined by \citet{biasotti} is used as a way to quantify the strength of a match.

In the algorithm  similarity of bullet land engraved areas is evaluated based on 3d topographic scans of land engraved areas (LEAs). From each land engraved area a signature is derived by
\begin{itemize}
\item[(a)] identifying an area where striae are expressed, 
\item[(b)] discarding extraneous/contaminated data, such as data from groove engraved area and areas affected by break-off or contact with objects after the bullet exited the barrel  (``tank rash"), 
\item[(c)] removing bullet curvature. 
\end{itemize}
A signature for a land engraved area is then defined as the sequence $z_i$, $i = 1, ..., I$, where $I$ is the number of observed locations across the base of the bullet. \autoref{fig:signatures} shows a set of six signatures corresponding to the six land engraved areas of bullets 1 and 2 from barrel 1 of set 44 of the Hamby study \citep{hamby}.

```{r signatures, fig.width = 10, fig.height = 4, fig.cap="Signatures of all six lands of bullets 1 and 2 from barrel 1 of the Hamby set 44."}
# signatures for one bullet
h44cc <- readRDS("data/h44-ccdata.rda")
br1 <- h44cc[1:12]
dframe <- 1:12 %>% purrr::map_df(.f = function(i) {
  d <- br1[[i]]
  d <- d %>% group_by(y) %>% summarize(
    value = mean(value, na.rm=TRUE)
  )
  d$id <- i
  d %>% ungroup(y) %>% rename(
    x = y
  )
})

dframe$bullet <- rep(c("Bullet 1", "Bullet 2"), each = 6)[dframe$id]
dframe$land <- paste("Land", (dframe$id %% 6 + 1))

dframe <- dframe %>% group_by(bullet, land) %>% nest()
names(dframe)[3] <- "ccdata"

library(bulletxtrctr)
dframe <- dframe %>% mutate(
  grooves = ccdata %>% 
    purrr::map(.f = cc_locate_grooves, method = "middle", 
               adjust = 30, return_plot = TRUE)
)
dframe <- dframe %>% mutate(
  sigs = purrr::map2(
    .x = ccdata, .y = grooves, 
    .f = function(x, y) {
      cc_get_signature(
        ccdata = x, grooves = y, span1 = 0.75, span2 = 0.03)
    })
)
dframe$barrelland <- dframe$land
idx <- which(dframe$bullet=="Bullet 2")
dframe$landid <- parse_number(dframe$land)
dframe$landid[idx] <- (dframe$landid[idx] +1) %% 6 + 1
dframe$land <- paste("Land", dframe$landid)

signatures <- dframe %>% select(bullet, land, sigs) %>% tidyr::unnest()
signatures %>% 
  ggplot(aes(x = x/1000)) + 
  geom_line(aes(y = raw_sig), colour = "grey70") +
  geom_line(aes(y = sig), colour = "grey30") +
  facet_grid(bullet~land) +
  ylim(c(-5,5)) +
  theme_bw() +
  ylab("Surface measurement (in microns)") +
  xlab("Relative Location (in millimeters)") 
```

Quantitative features describing the similarity of two signature are extracted. Features include cross-correlation, Euclidean distance, depth and number of matching striae, as well as the number of consecutively matching striae (CMS).

# Validation Sets

The algorithm in \citet{aoas} is trained on scans from Hamby sets 252 and 173 made available through \citet{nist}. (Set 173 was originally published as Hamby set 44. This has been corrected now.)

Here, we are considering three validation sets:
\begin{itemize}
\item {\bf Hamby set 44} 
Hamby set 44 is one set of the Hamby study \citep{hamby}. Each Hamby set  consists of a total of 35 bullets fired through ten consecutively manufactured barrels of Rugers P85. Each set consists of 20 known bullets (two from each of the ten barrels) and 15 questioned bullets of unknown origin. Note that all Hamby sets are closed sets, i.e. all questioned bullets are fired through one of the ten barrels. The ammunition used for this set were 9 mm Luger 115 Grain Full Metal Jacket  from the Winchester Ammunition Company.
\item {\bf Phoenix PD}
Tylor Klep from Phoenix PD provided sets of known and unkown bullets: the set of known bullets consists of three test fires (B1, B2, B3) from each of eight different barrels (A9, C8, F6, L5, M2, P7, R3, U10). Ten unknown bullets were provided (B, E, H, J, K N, Q, T, Y, Z). Land engraved areas for each of the six lands of each bullet were scanned by Bill Henderson (Sensofar).
\item {\bf Houston FSI}
This study was set up by Melissa Nally and Kasi Kirksey from FSI Houston. Three test sets based on ten consecutively rifled Ruger LCP barrels (A, B, C, D, E, F, G, H, I, J) and three other, non-consecutively rifled Ruger LCP barrels (R1, R2, R3). Each test set consists of three test fires each from five consecutively rifled barrels. Additionally, ten unknowns bullets are provided for each kit. The ammunition used in both test fires and the unknowns were Remington UMC 9mm Luger Full Metal Jackets.
\end{itemize}

Scans of all land engraved areas were taken on a Sensofar Confocal Light microscope at 20x magnification resulting a resolution of 0.645 microns per pixel. If not indicated otherwise, scans were taken at the Roy J Carver high resolution microscopy lab at Iowa State University.

Case studies were chosen such that difficulty for the matching algorithm increases from one to the other: Hamby set 44 is part of the Hamby study. The algorithm in \citet{aoas} was trained on sets 173 and 252, so the bullets in Hamby 44 are of the same type of ammunition and are fired through the same barrels as the bullets in the training set.

For the Phoenix PD set, Ruger XXX barrels were used. Obviously, the barrels in this set are different from the barrels in the training, but the barrels are similarly rifled as the Ruger P-85 barrels.

For the Houston Sets, Ruger LCP barrels were used. These barrels are first rifled traditionally, i.e. similar to the Ruger P-85, but are double fired. The second firing happens after the rifling. This second firing has the potential to introduce some subclass characteristics, i.e. anomalies that are shared between similarly barrels. This should make the automatic matching, and in particular, the classification as different-source land engraved areas/bullets, harder.

# From land-to-land scores to bullet summary scores

```{r, echo=FALSE, results = 'hide', warning = FALSE, message = FALSE}
features <- read.csv("data/fsi-features.csv")
features <- features %>% mutate(
  bullet1 = gsub("ullet ","", bullet1),
  bullet2 = gsub("ullet ","", bullet2)
)
b12 <- features %>% filter(group1 == "G1", barrel1=="KA", bullet1== "B1",
                           group2 == "G1", barrel2=="KA", bullet2== "B2")
p1 <- b12 %>% ggplot(aes(x = land1, y = land2, fill = rfscore)) + 
  geom_tile() + ggtitle ("Same barrel") +  
  scale_fill_gradient2("RF score: ", midpoint=0.3, low = "darkgrey", high = "darkorange", limits=c(0,1)) +
  theme_bw() + 
  xlab ("Bullet 2") + ylab("Bullet 1") +
  theme(legend.position = "bottom") + coord_equal()

bab <- features %>% filter(group1 == "G1", barrel1=="KA", bullet1== "B1",
                           group2 == "G1", barrel2=="KB", bullet2== "B1")
p2 <- bab %>% ggplot(aes(x = land1, y = land2, fill = rfscore)) + 
  geom_tile() + ggtitle ("Different barrel") +  
  scale_fill_gradient2("RF score: ", midpoint=0.3, low = "darkgrey", high = "darkorange", limits=c(0,1)) +
  theme_bw() + 
  xlab ("Bullet 2") + ylab("Bullet 1") +
  theme(legend.position = "bottom") + coord_equal()
```

Land-land comparisons lead to a whole set of scores for bullet-bullet comparisons. \autoref{fig:b2b} shows two matrices of scores for a pair of bullet-bullet matches. On the left, a matrix is shown that is typical for scores from two bullets from the same barrel. On the right, values for a pair of known non-matching bullets are shown.

```{r b2b, echo = FALSE, fig.width = 9, fig.height = 5,  out.width='.6\\textwidth', fig.cap="Overview of land-land matching scores for two pairs of bullet-bullet comparisons. On the left the two bullets are known to come from the same source (barrel) on the right, the bullets are from two different sources (barrels)"}
grid.arrange(p1, p2, ncol=2, padding=2)
```

When imaging bullets, operators scan one land at a time in a clockwise (left twisted rifling) or anti-clockwise (right twisted rifling) sequence. The order in which scans are acquired is kept as meta-information. Let us assume that lands on a bullet are labelled $\ell_i$ with $i = 1, .., p$. A match between two bullets therefore results in an expected additional $p-1$ matches between pairs of lands. These lands are also expected to be in a sequence, i.e. if there is a match between lands $\ell_i$ on bullet 1 and $\ell_j$ on bullet 2, we also expect lands $\ell_{i\oplus s}$ and $\ell_{j\oplus s}$ to match for all integers $s$, where $\oplus$ is defined as $a \oplus b \equiv \left((a + b)\mod p\right) + 1$. This relationship gives rise to the sequence average maximum (SAM) to quantify a bullet-to-bullet match.

\begin{definition}{Sequence Average and its Maximum}
Let $A$ be a square real-valued matrix of dimensions $p \times p$. For the purpose of this paper, $A$ consists of scores describing the similarity between two sets of land engraved areas.
The $k$th \emph{sequence average} $SA(A, k)$ for $k = 0, ... p-1$ is defined  as 
$$SA(A, k) = \frac{1}{p} \sum_{i=1}^{p} a_{i,i \oplus k}, \text{ where } i\oplus k := \left((i + k)\mod p\right) + 1.$$
The \emph{Sequence Average Maximum} \citep[SAM, ][]{sam} of square matrix $A$ of scores is defined as
$$SAM (A) = \max_{k = 1}^{p} SA(A, k).$$
\end{definition}

Note that the sequence average maximum of the correlation between lands is used in SensoComp to capture the similarity between bullets. The correlation based SAM score has also been called the 'average correlation calculated at the max phase' in \citet{chu}.

```{r sam-sketch, out.width='\\textwidth', fig.width=12, fig.height = 2.5, fig.cap="Sketch of all six land-to-land sequences between two bullets with six lands. "}
df2 <- data.frame(expand.grid(x = 1:6, y = 1:6))
df2$case1 <- df2$x == ((df2$y+5) %% 6) + 1
df2$case2 <- df2$x == ((df2$y+0) %% 6) + 1
df2$case3 <- df2$x == ((df2$y+1) %% 6) + 1
df2$case4 <- df2$x == ((df2$y+2) %% 6) + 1
df2$case5 <- df2$x == ((df2$y+3) %% 6) + 1
df2$case6 <- df2$x == ((df2$y+4) %% 6) + 1

df2 %>% gather("case", "values", starts_with("case")) %>% 
  mutate(case = gsub("case", "", case)) %>%
  mutate(case = paste0("k = ", as.numeric(case)-1)) %>%
  ggplot(aes(x = factor(x), y = factor(y, levels=1:6), fill = values)) + 
  geom_tile(colour = "grey20") + 
  facet_wrap(~case, ncol = 6) +
  xlab("Bullet 2") + ylab("Bullet 1") +
  scale_x_discrete(labels = paste0("L",1:6)) +
  scale_y_discrete(labels = paste0("L",1:6)) +
  scale_fill_manual("Cell included in SA(A,k)", values = c("white", "grey20")) + 
  coord_equal() +
  theme(legend.position="bottom")
```

Looking back at Figure \ref{fig:b2b}, we see that for the two bullets from the same barrel, the sequence average for 
$k=2$ is higher than the other sequence averages, and also higher than the sequence averages for the other pair of bullets shown on the right of the figure.

SAM scores allow us to define a single quantity for each pair of bullets that describes the similarity between these two bullets. 
This leads us to two minimal requirements the scores should fulfill:
\begin{itemize}
\item[(R1)] {\bf Monotonicity:} a higher score is indicative of higher similarity between a pair of bullets, in particular, similarity scores of same-source pairs of bullets are higher than different-source pairs.
\item[(R2)] {\bf Stability:} the same score leads to the same conclusion. 
\end{itemize}

# Results from the validation sets

## Hamby Set 44

```{r h44, echo=FALSE, fig.width = 10, fig.height=5, fig.cap="Hamby set 44: overview of all bullet-to-bullet matches between all pairs of questioned bullets ($y$-axis) and known test fires from ten barrels ($x$ axis)."}
h44features <- read.csv("data/h44-features.csv")

h44nest <- h44features %>% 
  filter(barrel1=="Unk") %>%
  filter(barrel1 != barrel2 | bullet1 != bullet2) %>%
  mutate(
    barrel1 = "Unknown",
    barrel2 = as.character(barrel2),
    barrel2 = replace(barrel2, barrel2=="Unk", "Unknown")
  ) %>%
  group_by(bullet1, barrel2, bullet2) %>%  nest()

h44nest <- h44nest %>% mutate(
  sam_ccf = data %>% purrr::map_dbl(.f=function(d) {
    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$ccf)
    max(scores)
  }),
  sam_rf = data %>% purrr::map_dbl(.f=function(d) {
    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore)
    max(scores)
  })
  )
h44nest <- h44nest %>% 
  mutate(
    barrel2 = factor(barrel2, levels =c(1:10, "Unknown")),
    bullet1 = factor(bullet1, levels=rev(c("K", "O","L", "P",
                                           "J",
                                         "H", "I", "Y", "G",
                                         "E", "U", "X", "F", 
                                         "T", "S"))),
    bullet2 = factor(bullet2, levels=c(1:2, rev(levels(bullet1))))
  )
h44 <-  h44nest %>%
  ggplot(aes(y = bullet1, x = bullet2, 
             fill = sam_rf)) + 
  geom_tile(size=1) + 
  facet_grid(.~barrel2, space="free", scales = "free") +
  ylab("Questioned bullets") + 
  xlab("Testfires from barrels 1 to 10 (left), questioned bullets (right)") +
  scale_fill_gradient2("RF score ", low="darkgrey", 
                       high="darkorange", midpoint=0.375, limits=c(0,1)) + 
  scale_colour_manual("same source", values="darkorange") +

  #  scale_x_discrete(position = "top") +
  scale_y_discrete() +
  geom_tile(aes(colour = TRUE), size = .5, data = filter(h44nest, sam_rf > 0.3)) +
  guides(colour = guide_legend(override.aes = list(fill = NA))) + ggtitle("Hamby set 44") +
  theme_bw() +
    theme(legend.position="bottom")

h44nest$samesource <- h44nest$sam_rf > 0.3

h44
```

\autoref{fig:h44} shows an overview of all scores from pairs of questioned bullets with all other bullets. On the left of \autoref{fig:h44} there are ten strips labelled 1 through 10. These strips correspond to known barrels 1 through 10. For each of these barrels, two test fires were done. Each of the colored tiles corresponds to a pair of bullets between a questioned bullet (shown along the $y$-axis) and one of the other bullets. The similarity score is encoded in color: grey colors correspond to low similarity scores, orange colors correspond to high similarity scores. Ground truth is encoded in this figure as a thin orange frame for all pairs of same-source bullets. 
What we want to see for the relationship between questioned bullets and the ten barrels, is one barrel with orange-filled, orange-framed tiles for both bullets from the same barrel, and grey tiles across the other barrels. This expectation is met for all questioned bullets, i.e. the automatic matching  identifies the correct barrel for all questioned bullets. On the right of \autoref{fig:h44} the relationship between all pairs of questioned bullets is shown. Note that questioned bullets are not compared to themselves, leaving white squares on the diagonal.
Some of the questioned bullets match the same barrel, e.g. questioned bullets 'P' and 'J' both match barrel 5. Therefore bullets 'P' and 'J' are also matching each other in the square on the right hand side.

## Phoenix PD

```{r pd, echo=FALSE, fig.width = 10, fig.height=5, fig.cap="Phoenix study: overview of all bullet-to-bullet matches between all pairs of questioned bullets ($y$-axis) and known test fires from ten barrels ($x$ axis). The order of the bullets on the $y$ axis is determined by matching barrel."}
features <- read.csv("data/pd-features.csv.gz")
f2 <- features %>% 
  separate(first, into = c("foo1", "foo2", "barrel1", "bullet1", "land1", "foo3"), remove = FALSE) %>% 
  separate(second, into = c("foo4", "foo5", "bullet2", "land2", "foo6"), remove = FALSE) %>%
  select(-foo1, -foo2, -foo3, -foo4, -foo5, -foo6)

pdnest <- f2 %>% group_by(bullet2, barrel1, bullet1) %>%  nest()

pdnest <- pdnest %>% mutate(
  sam_ccf = data %>% purrr::map_dbl(.f=function(d) {
    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$ccf)
    max(scores)
  }),
  sam_rf = data %>% purrr::map_dbl(.f=function(d) {
    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore)
    max(scores)
  }), 
  KM = data %>% purrr::map_lgl(.f=function(d) { any(d$KM) })
)

pdnest <- pdnest %>% 
  mutate(
    bullet2 = factor(bullet2, levels=rev(c("N", "B", "E", "T", "H", "J", "K", "Q", "Y", "Z")))
  )
p3 <- pdnest %>%
  ggplot(aes(y = bullet2, x = bullet1, fill = sam_rf)) + 
  geom_tile(size=1) + 
  facet_grid(.~barrel1) +
  ylab("Questioned bullets") + 
  xlab("Known bullets") +
  scale_fill_gradient2("RF score ", low="darkgrey", 
                       high="darkorange", midpoint=0.45, limits=c(0,1)) + 
  scale_colour_manual("same source", values="darkorange") + 
  scale_y_discrete() +
  theme(legend.position="bottom") +
  coord_equal() +
  geom_tile(aes(colour = TRUE), size = .5,
            data = filter(pdnest, KM)) +
  guides(colour = guide_legend(override.aes = list(fill = NA)))

pdnest$samesource <- pdnest$KM

p3 + ggtitle("Phoenix PD set")
```

\autoref{fig:pd} shows an overview of similarity scores for all pairs of questioned bullets and  test fires. The color encoding is the same as in the previous figure. What we see here, are matches of one of the questioned bullets to all three bullets of one of the barrels each. None of the test fires from barrel U10 shows a match to any of the questioned bullets. SImilarly, questioned bullets 'Q', 'Y', and 'Z' do not match any of the knwon barrels. This is the sign of an open set. 
The results from automatic matching, again, correctly match barrel and questioned bullets.

## Houston FSI

\autoref{fig:hou} shows an overview the three sets of the scores for all pairs of questioned bullets with all other bullets

```{r hou, fig.width=9, fig.height = 3, fig.cap="Overview of matching scores for all pairs of questioned bullets to known bullets from five barrels (on the left) and questioned bullets to themselves (tiles on the right).", fig.subcap=c("Set 1", "Set 2", "Set 3"), fig.align="middle", fig.ncol=1}
features <- read.csv("data/fsi-features.csv")

features <- features %>% mutate(
  bullet1 = gsub("ullet ","", bullet1),
  bullet2 = gsub("ullet ","", bullet2)
)

groups <- features %>% 
  filter(group1 == group2) %>%
  filter(barrel1 == "Unknowns") %>%
  filter(barrel1 != barrel2 | bullet1 != bullet2) %>%
  group_by(group1, barrel1, bullet1, barrel2, bullet2) %>%
  nest()

hounest <- groups %>% mutate(
  sam_ccf = data %>% purrr::map_dbl(.f=function(d) {
    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$ccf)
    max(scores)
  }),
  sam_rf = data %>% purrr::map_dbl(.f=function(d) {
    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore)
    max(scores)
  })
)

hounest$samesource <- hounest$sam_rf > 0.3

hou1nest <- hounest %>% filter(group1 == "G1") %>%
  mutate(
    bullet1 = factor(bullet1, levels=rev(c("U40", "U36", "U42", "U10", "U77", "U37", "U28", "U15"))),
    bullet2 = factor(bullet2, levels=c("B1", "B2", "B3", levels(bullet1)))
  )

hou1 <- hou1nest %>%
  ggplot(aes(y = bullet1, x = bullet2, fill = sam_rf)) + 
  geom_tile(size=1) + 
  facet_grid(.~barrel2, scales = "free", space="free") +
  ylab("Questioned bullets") + 
  xlab("Known bullets and Questioned bullets") +
  scale_fill_gradient2("RF score ", low="darkgrey", 
                       high="darkorange", midpoint=0.375, limits=c(0,1)) + 
  scale_y_discrete() +
  scale_colour_manual("same source", values="darkorange") +
  geom_tile(aes(colour = TRUE), size = .5, data = filter(hou1nest, sam_rf > 0.3)) +
  guides(colour = guide_legend(override.aes = list(fill = NA))) + ggtitle("Houston set 1")

hou2nest <- hounest %>% filter(group1 == "G2") %>%
  mutate(
    bullet1 = factor(bullet1, levels=rev(c("U23", "U41", "U61", "U98", "U73", "U34", "U63", "U66"))),
    bullet2 = factor(bullet2, levels=c("B1", "B2", "B3", levels(bullet1)))
  )
hou2 <-  hou2nest %>%
  ggplot(aes(y = bullet1, x = bullet2, fill = sam_rf)) + 
  geom_tile(size=1) + 
  facet_grid(.~barrel2, scales = "free", space="free") +
  ylab("Questioned bullets") + 
  xlab("Known bullets and Questioned bullets") +
  scale_fill_gradient2("RF score ", low="darkgrey", 
                       high="darkorange", midpoint=0.375, limits=c(0,1)) + 
  scale_y_discrete() +
    scale_colour_manual("same source", values="darkorange") +
  geom_tile(aes(colour = TRUE), size = .5, data = filter(hou2nest, sam_rf > 0.3)) +
  guides(colour = guide_legend(override.aes = list(fill = NA))) + ggtitle("Houston set 2")

hou3nest <- hounest %>% filter(group1 == "G3") %>%
  mutate(
    bullet1 = factor(bullet1, levels=rev(c("U27", "U33", "U36", "U49", "U56", "U65", "U14", "U45"))),
    bullet2 = factor(bullet2, levels=c("B1", "B2", "B3", levels(bullet1)))
  )

hou3 <- hou3nest %>%
  ggplot(aes(y = bullet1, x = bullet2, fill = sam_rf)) + 
  geom_tile(size=1) + 
  facet_grid(.~barrel2, scales = "free", space="free") +
  ylab("Questioned bullets") + 
  xlab("Known bullets and Questioned bullets") +
  scale_fill_gradient2("RF score ", low="darkgrey", 
                       high="darkorange", midpoint=0.375, limits=c(0,1)) + 
  scale_y_discrete() +
    scale_colour_manual("same source", values="darkorange") +
  geom_tile(aes(colour = TRUE), size = .5, data = filter(hou3nest, sam_rf > 0.3)) +
  guides(colour = guide_legend(override.aes = list(fill = NA))) + ggtitle("Houston set 3")

hou1

hou2

hou3
```

# Comparison of SAM scores

\autoref{fig:compare} shows density curves for the scores of each of the studies, comparing the SAM scores based on cross-correlation (top) and the random forest score (bottom). Color indicates ground truth. Besides densities, small vertical lines below the $x$-axis indicate observed scores in a so-called **rug-plot**. The plot shows that for all three case studies and both measures, i.e. SAM scores based on cross-correlation as well as SAM scores based on the random forest scores, all similarity scores are higher for pairs of bullets from the same source, i.e. bullets fired through the same barrel, than pairs of bullets from different sources (fired through different barrels).

```{r compare, fig.width = 8, fig.height = 5, fig.cap="Density curves of similarity scores from cross-correlation (top) and RF scores (bottom). Different colors indicate same source versus different source. Ideally all scores of different source pairs should be much lower than scores for same source pairs."}
nests <- rbind(
  h44nest %>% select(sam_rf, sam_ccf, samesource) %>%
    mutate(study = "Hamby-44", group1=NA),
  pdnest %>% select(sam_rf, sam_ccf, samesource) %>%
    mutate(study = "Phoenix PD", group1=NA),
  hounest %>% select(sam_rf, sam_ccf, samesource, group1) %>%
    mutate(study = "Houston FSI")
)
nests$study <- factor(nests$study, levels=c("Hamby-44", "Phoenix PD", "Houston FSI"))
nestslong <- nests %>% gather(key="measure", value="value", sam_rf, sam_ccf)
nestslong$measure <- factor(nestslong$measure)
levels(nestslong$measure) <- c("Cross-correlation", "RF score")

nestslong %>% ggplot(aes(x = value, fill=samesource, 
                     colour=samesource)) +
  geom_density(alpha=0.6) +
  geom_rug(alpha=0.5) +
  facet_grid(measure~study) +
  theme_bw() +
  scale_fill_manual("Same source", values=c("darkgrey", "darkorange")) +
  scale_colour_manual("Same source", values=c("darkgrey", "darkorange")) +
  scale_x_continuous("SAM scores", limits=c(0,1)) +
  theme(legend.position="bottom") +
  geom_hline(yintercept =0) 
```

What we find in \autoref{fig:compare} is that neither of the measure fulfills both of the minimal requirements laid out above. Both measures fulfill (R1), i.e. all  scores  of  different-source pairs are lower than scores of same-source pairs for SAM scores based on cross-correlation and RF scores.
However, SAM scores based on the multivariate random forest score show better separation between same-source and different-source scores than SAM scores based on cross-correlation. This can be seen from the larger horizontal distance between the modes of the density curves of RF based SAM scores than cross-correlation SAM scores.

Regarding requirement (R2), \autoref{fig:compare} shows that for neither of the two measures  a single cutoff exists across different studies that   separates same-source scores from different-source scores without introducing errors.

Potential improvements are two-fold: we can try to fine tune the matching  algorithm to take make and model of the firearm and ammunition used into account. Downside is that we need to massively expand the database used for training.
Another solution would be to expand on the summarization. SAM scores only take the scores of the maximum sequence into account and ignore scores associated with pairs of land engraved areas from different sources. Those scores might be useful in establishing a baseline that better expresses similarity between pairs of bullets.

# Discussion and Conclusions

What is quite surprising, that what we anticipated to be the easiest case of Hamby 44 turned out to be the hardest to solve.

# References

### 2018-09-28 index.Rmd
link to fsi

@@
 ```{r, echo=FALSE, results = 'hide', warning = FALSE, message = FALSE}
-features <- read.csv("data/fsi-features.csv")
+features <- read.csv("data/fsi-features.csv.gz")
 features <- features %>% mutate(
@@
 ```{r hou, fig.width=9, fig.height = 3, fig.cap="Overview of matching scores for all pairs of questioned bullets to known bullets from five barrels (on the left) and questioned bullets to themselves (tiles on the right).", fig.subcap=c("Set 1", "Set 2", "Set 3"), fig.align="middle", fig.ncol=1}
-features <- read.csv("data/fsi-features.csv")
+features <- read.csv("data/fsi-features.csv.gz")
 

### 2018-09-28 index.Rmd
change in rda

@@
 # signatures for one bullet
-h44cc <- readRDS("data/h44-ccdata.rda")
-br1 <- h44cc[1:12]
-dframe <- 1:12 %>% purrr::map_df(.f = function(i) {
-  d <- br1[[i]]
-  d <- d %>% group_by(y) %>% summarize(
-    value = mean(value, na.rm=TRUE)
-  )
-  d$id <- i
-  d %>% ungroup(y) %>% rename(
-    x = y
-  )
-})
+# h44cc <- readRDS("data/h44-ccdata.rda")
+# br1 <- h44cc[1:12]
+# dframe <- 1:12 %>% purrr::map_df(.f = function(i) {
+#   d <- br1[[i]]
+#   d <- d %>% group_by(y) %>% summarize(
+#     value = mean(value, na.rm=TRUE)
+#   )
+#   d$id <- i
+#   d %>% ungroup(y) %>% rename(
+#     x = y
+#   )
+# })
+# 
+# dframe$bullet <- rep(c("Bullet 1", "Bullet 2"), each = 6)[dframe$id]
+# dframe$land <- paste("Land", (dframe$id %% 6 + 1))
+# 
+# dframe <- dframe %>% group_by(bullet, land) %>% nest()
+# names(dframe)[3] <- "ccdata"
+# 
+# library(bulletxtrctr)
+# dframe <- dframe %>% mutate(
+#   grooves = ccdata %>% 
+#     purrr::map(.f = cc_locate_grooves, method = "middle", 
+#                adjust = 30, return_plot = TRUE)
+# )
+# dframe <- dframe %>% mutate(
+#   sigs = purrr::map2(
+#     .x = ccdata, .y = grooves, 
+#     .f = function(x, y) {
+#       cc_get_signature(
+#         ccdata = x, grooves = y, span1 = 0.75, span2 = 0.03)
+#     })
+# )
+# dframe$barrelland <- dframe$land
+# idx <- which(dframe$bullet=="Bullet 2")
+# dframe$landid <- parse_number(dframe$land)
+# dframe$landid[idx] <- (dframe$landid[idx] +1) %% 6 + 1
+# dframe$land <- paste("Land", dframe$landid)
 
-dframe$bullet <- rep(c("Bullet 1", "Bullet 2"), each = 6)[dframe$id]
-dframe$land <- paste("Land", (dframe$id %% 6 + 1))
-
-dframe <- dframe %>% group_by(bullet, land) %>% nest()
-names(dframe)[3] <- "ccdata"
-
-library(bulletxtrctr)
-dframe <- dframe %>% mutate(
-  grooves = ccdata %>% 
-    purrr::map(.f = cc_locate_grooves, method = "middle", 
-               adjust = 30, return_plot = TRUE)
-)
-dframe <- dframe %>% mutate(
-  sigs = purrr::map2(
-    .x = ccdata, .y = grooves, 
-    .f = function(x, y) {
-      cc_get_signature(
-        ccdata = x, grooves = y, span1 = 0.75, span2 = 0.03)
-    })
-)
-dframe$barrelland <- dframe$land
-idx <- which(dframe$bullet=="Bullet 2")
-dframe$landid <- parse_number(dframe$land)
-dframe$landid[idx] <- (dframe$landid[idx] +1) %% 6 + 1
-dframe$land <- paste("Land", dframe$landid)
-
+signatures <- read.csv("data/signatures.csv")
 signatures <- dframe %>% select(bullet, land, sigs) %>% tidyr::unnest()

### 2018-09-28 index.Rmd
comment out dframe line

@@
 signatures <- read.csv("data/signatures.csv")
-signatures <- dframe %>% select(bullet, land, sigs) %>% tidyr::unnest()
+# signatures <- dframe %>% select(bullet, land, sigs) %>% tidyr::unnest()
 signatures %>% 

### 2018-09-28 index.Rmd
phoenix unknown to unknown comparison included

--- 2018-09-28 index.Rmd
@@
 # signatures for one bullet
-h44cc <- readRDS("data/h44-ccdata.rda")
-br1 <- h44cc[1:12]
-dframe <- 1:12 %>% purrr::map_df(.f = function(i) {
-  d <- br1[[i]]
-  d <- d %>% group_by(y) %>% summarize(
-    value = mean(value, na.rm=TRUE)
-  )
-  d$id <- i
-  d %>% ungroup(y) %>% rename(
-    x = y
-  )
-})
+# h44cc <- readRDS("data/h44-ccdata.rda")
+# br1 <- h44cc[1:12]
+# dframe <- 1:12 %>% purrr::map_df(.f = function(i) {
+#   d <- br1[[i]]
+#   d <- d %>% group_by(y) %>% summarize(
+#     value = mean(value, na.rm=TRUE)
+#   )
+#   d$id <- i
+#   d %>% ungroup(y) %>% rename(
+#     x = y
+#   )
+# })
+# 
+# dframe$bullet <- rep(c("Bullet 1", "Bullet 2"), each = 6)[dframe$id]
+# dframe$land <- paste("Land", (dframe$id %% 6 + 1))
+# 
+# dframe <- dframe %>% group_by(bullet, land) %>% nest()
+# names(dframe)[3] <- "ccdata"
+# 
+# library(bulletxtrctr)
+# dframe <- dframe %>% mutate(
+#   grooves = ccdata %>% 
+#     purrr::map(.f = cc_locate_grooves, method = "middle", 
+#                adjust = 30, return_plot = TRUE)
+# )
+# dframe <- dframe %>% mutate(
+#   sigs = purrr::map2(
+#     .x = ccdata, .y = grooves, 
+#     .f = function(x, y) {
+#       cc_get_signature(
+#         ccdata = x, grooves = y, span1 = 0.75, span2 = 0.03)
+#     })
+# )
+# dframe$barrelland <- dframe$land
+# idx <- which(dframe$bullet=="Bullet 2")
+# dframe$landid <- parse_number(dframe$land)
+# dframe$landid[idx] <- (dframe$landid[idx] +1) %% 6 + 1
+# dframe$land <- paste("Land", dframe$landid)
+# signatures <- dframe %>% select(bullet, land, sigs) %>% tidyr::unnest()
 
-dframe$bullet <- rep(c("Bullet 1", "Bullet 2"), each = 6)[dframe$id]
-dframe$land <- paste("Land", (dframe$id %% 6 + 1))
-
-dframe <- dframe %>% group_by(bullet, land) %>% nest()
-names(dframe)[3] <- "ccdata"
-
-library(bulletxtrctr)
-dframe <- dframe %>% mutate(
-  grooves = ccdata %>% 
-    purrr::map(.f = cc_locate_grooves, method = "middle", 
-               adjust = 30, return_plot = TRUE)
-)
-dframe <- dframe %>% mutate(
-  sigs = purrr::map2(
-    .x = ccdata, .y = grooves, 
-    .f = function(x, y) {
-      cc_get_signature(
-        ccdata = x, grooves = y, span1 = 0.75, span2 = 0.03)
-    })
-)
-dframe$barrelland <- dframe$land
-idx <- which(dframe$bullet=="Bullet 2")
-dframe$landid <- parse_number(dframe$land)
-dframe$landid[idx] <- (dframe$landid[idx] +1) %% 6 + 1
-dframe$land <- paste("Land", dframe$landid)
-
-signatures <- dframe %>% select(bullet, land, sigs) %>% tidyr::unnest()
+signatures <- read.csv("data/signatures.csv")
 signatures %>% 
@@
 
-```{r pd, echo=FALSE, fig.width = 10, fig.height=5, fig.cap="Phoenix study: overview of all bullet-to-bullet matches between all pairs of questioned bullets ($y$-axis) and known test fires from ten barrels ($x$ axis).
+```{r pd, echo=FALSE, fig.width = 9, fig.height=4, fig.cap="Phoenix study: overview of all bullet-to-bullet matches between all pairs of questioned bullets ($y$-axis) and known test fires from ten barrels ($x$ axis).
 The order of the bullets on the $y$ axis is determined by matching barrel."}
 features <- read.csv("data/pd-features.csv.gz")
+
 f2 <- features %>% 
@@
 
-pdnest <- f2 %>% group_by(bullet2, barrel1, bullet1) %>%  nest()
+unknowns <- read.csv("data/pd-features-unknown.csv")
+unknowns <- unknowns %>% filter(bullet1 != bullet2)
+unknowns <- unknowns %>% mutate(
+  bullet2 = gsub("Unknown 1-", "", bullet2),
+  bullet1 = gsub("Unknown 1-", "", bullet1)
+)
+f3 <- rbind(
+  f2 %>% select(b1, b2, barrel1, bullet1, 
+                land1, bullet2, land2, ccf, rfscore, KM) %>%
+    mutate(barrel2 = "Unknown"),
+  unknowns %>% select(b1, b2, bullet1, 
+                land1, bullet2, land2, ccf, rfscore) %>%
+    mutate(barrel1 ="Unknown", barrel2="Unknown", KM=FALSE)
+  )
+pdnest <- f3 %>% group_by(bullet2, barrel1, bullet1) %>%  nest()
 
@@
   mutate(
-    bullet2 = factor(bullet2, levels=rev(c("N", "B", "E", "T", "H", "J", "K", "Q", "Y", "Z")))
+    bullet2 = factor(bullet2, levels=rev(c("N", "B", "E", "T", "H", "J", "K", "Q", "Y", "Z"))),
+    bullet1 = factor(bullet1, levels=c("B1", "B2", "B3", rev(levels(bullet2))))
   )
@@
   geom_tile(size=1) + 
-  facet_grid(.~barrel1) +
+  facet_grid(.~barrel1, scales="free", space="free") +
   ylab("Questioned bullets") + 
-  xlab("Known bullets") +
+  xlab("Known bullets and Questioned bullets") +
   scale_fill_gradient2("RF score ", low="darkgrey", 
@@
   scale_y_discrete() +
+  theme_bw() +
   theme(legend.position="bottom") +
-  coord_equal() +
+#  coord_equal() +
   geom_tile(aes(colour = TRUE), size = .5,

### 2018-10-02 index.Rmd
typos

@@
 # Introduction and Background
-In current practice, firearms and toolmark examiners (FTE) evaluate the similiarity of striae on bullets by placing the evidence together with a bullet from the suspected gun under a comparison microscope and visually classify similarity according to the theory of firearms identification \citep{identification} as one of identification, inconclusive or exclusion.
+In current practice, firearms and toolmark examiners (FTE) evaluate the similarity of striae on bullets by placing the evidence together with a bullet from the suspected gun under a comparison microscope and visually classify similarity according to the theory of firearms identification \citep{identification} as one of identification, inconclusive or exclusion.
 Exact guidelines for this classification vary from lab to lab, e.g.\ some labs will exclude only on the basis of non-matching class characteristics (such as direction of the twist in rifling, land length or number of lands, or type of rifling).
@@
 \item {\bf Phoenix PD}
-Tylor Klep from Phoenix PD provided sets of known and unkown bullets: the set of known bullets consists of three test fires (B1, B2, B3) from each of eight different barrels (A9, C8, F6, L5, M2, P7, R3, U10).
+Tylor Klep from Phoenix PD provided sets of known and unknown bullets: the set of known bullets consists of three test fires (B1, B2, B3) from each of eight different barrels (A9, C8, F6, L5, M2, P7, R3, U10).
 Ten unknown bullets were provided (B, E, H, J, K N, Q, T, Y, Z).
@@
 None of the test fires from barrel U10 shows a match to any of the questioned bullets.
-SImilarly, questioned bullets 'Q', 'Y', and 'Z' do not match any of the knwon barrels.
+Similarly, questioned bullets 'Q', 'Y', and 'Z' do not match any of the known barrels.
 This is the sign of an open set.
@@
 Downside is that we need to massively expand the database used for training.
-Another solution would be to expand on the summarization.
+Another solution would be to expand on the summation.
 SAM scores only take the scores of the maximum sequence into account and ignore scores associated with pairs of land engraved areas from different sources.

### 2018-10-02 index.Rmd
updated citation

@@
 
-The algorithm in \citet{aoas} is trained on scans from Hamby sets 252 and 173 made available through \citet{nist}. (Set 173 was originally published as Hamby set 44. This has been corrected now.)
+The algorithm in \citet{aoas} is trained on scans from Hamby sets 252 and 173 made available through \citet{nist}. Set 173 was originally published as Hamby set 44. This has been rectified now.
 
@@
 \item {\bf Phoenix PD}
-Tylor Klep from Phoenix PD provided sets of known and unknown bullets: the set of known bullets consists of three test fires (B1, B2, B3) from each of eight different barrels (A9, C8, F6, L5, M2, P7, R3, U10).
-Ten unknown bullets were provided (B, E, H, J, K N, Q, T, Y, Z).
+Tylor Klep from Phoenix PD provided sets of known test fires and questioned bullets: the set of known bullets consists of three test fires (B1, B2, B3) from each of eight different barrels (A9, C8, F6, L5, M2, P7, R3, U10).
+Ten questioned bullets were provided (B, E, H, J, K N, Q, T, Y, Z).
+This set is an open set, i.e.\ it is not known in advance whether all (or any) of the questioned bullets are fired from the known barrels.
 Land engraved areas for each of the six lands of each bullet were scanned by Bill Henderson (Sensofar).

### 2018-10-02 index.Rmd
open set/closed set

@@
 This set is an open set, i.e.\ it is not known in advance whether all (or any) of the questioned bullets are fired from the known barrels.
+
 Land engraved areas for each of the six lands of each bullet were scanned by Bill Henderson (Sensofar).
@@
 Each test set consists of three test fires each from five consecutively rifled barrels.
-Additionally, ten unknowns bullets are provided for each kit.
-The ammunition used in both test fires and the unknowns were Remington UMC 9mm Luger Full Metal Jackets.
+Additionally, ten questioned bullets are provided for each kit.
+The ammunition used in both test fires and the questioned bullets were Remington UMC 9mm Luger Full Metal Jackets.
+All three of the test sets are open, i.e.\ not every one of the questioned bullets is fired from the five known barrels in each of the test set.
 \end{itemize}
@@
 Similarly, questioned bullets 'Q', 'Y', and 'Z' do not match any of the known barrels.
-This is the sign of an open set.
-The results from automatic matching, again, correctly match barrel and questioned bullets.
+This is a sign of the previously mentioned open set characteristic of the study.
+The results from automatic matching, again, correctly pair questioned bullets to their corresponding barrels.
 

### 2018-10-02 index.Rmd
define plus

@@
 
-Land-land comparisons lead to a whole set of scores for bullet-bullet comparisons.
+Land-to-land comparisons lead to a whole set of scores for bullet-to-bullet comparisons.
 \autoref{fig:b2b} shows two matrices of scores for a pair of bullet-bullet matches.
@@
 These lands are also expected to be in a sequence, i.e.
-if there is a match between lands $\ell_i$ on bullet 1 and $\ell_j$ on bullet 2, we also expect lands $\ell_{i\oplus s}$ and $\ell_{j\oplus s}$ to match for all integers $s$, where $\oplus$ is defined as $a \oplus b \equiv \left((a + b)\mod p\right) + 1$.
+if there is a match between lands $\ell_i$ on bullet 1 and $\ell_j$ on bullet 2, we also expect lands $\ell_{i\oplus s}$ and $\ell_{j\oplus s}$ to match for all integers $s$, where $\oplus$ is defined as $a \oplus b \equiv \left((a + b - 1)\mod p\right) + 1$.
 This relationship gives rise to the sequence average maximum (SAM) to quantify a bullet-to-bullet match.
@@
 The $k$th \emph{sequence average} $SA(A, k)$ for $k = 0, ... p-1$ is defined  as 
-$$SA(A, k) = \frac{1}{p} \sum_{i=1}^{p} a_{i,i \oplus k}, \text{ where } i\oplus k := \left((i + k)\mod p\right) + 1.$$
+$$SA(A, k) = \frac{1}{p} \sum_{i=1}^{p} a_{i,i \oplus k}, \text{ where } i\oplus k := \left((i + k - 1)\mod p\right) + 1.$$
 The \emph{Sequence Average Maximum} \citep[SAM, ][]{sam} of square matrix $A$ of scores is defined as

### 2018-10-02 index.Rmd
Susan's comments incorporated

@@
 Quantitative features describing the similarity of two signature are extracted.
-Features include cross-correlation, Euclidean distance, depth and number of matching striae, as well as the number of consecutively matching striae (CMS).
+Features include characteristics of the signature that firearms and toolmark examiners base their visual assessment on, such as  depth of striae, number of matching striae, and the number of consecutively matching striae (CMS).
+We also include statistical features, such as cross-correlation and Euclidean distance between the signatures.
 
@@
 
-Potential improvements are two-fold: we can try to fine tune the matching  algorithm to take make and model of the firearm and ammunition used into account.
-Downside is that we need to massively expand the database used for training.
-Another solution would be to expand on the summation.
-SAM scores only take the scores of the maximum sequence into account and ignore scores associated with pairs of land engraved areas from different sources.
+There are several approaches for potential improvements: we can try to fine tune the matching  algorithm to take make and model of the firearm and ammunition used into account or expand the traing of the algorithm to include a wider variety of makes and models.
+The downside of this is that we need to considerably expand the database used for training.
+Another solution would be to expand on the aggregation used to get from land-to-land scores to bullet scores: SAM scores only take the scores of the maximum sequence into account and ignore scores associated with pairs of land engraved areas from different sources.
 Those scores might be useful in establishing a baseline that better expresses similarity between pairs of bullets.

### 2018-10-02 index.Rmd
land-to-land comparisons for six bullets from the Hamby set

@@
 
+```{r, fig.width=7.5, fig.height=2.8, fig.cap="Examples of land-to-land comparisons from the Hamby-44 set.
+Along the top are the three bullets from different source with the highest RF scores, along the bottom are the three bullets from same source with the lowest RF score.", fig.subcap=c("Ground truth: different source", "Ground truth: same-source"), fig.align="middle", fig.ncol=1}
+ex2 <- h44nest %>% filter(samesource == FALSE & sam_rf >= 0.295)
+rfs2 <- ex2$sam_rf
+ex2 <- ex2 %>% mutate(
+  data = data %>% purrr::map(.f = function(d) {
+ #   browser()
+    index <- which.max(bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore))
+    d$same <- ((d$land2 - d$land1) %% 6)  == index-1
+
+  #   d %>% ggplot(aes(x = land1, y=land2, fill=rfscore)) + 
+  #                  geom_tile() +
+  #     geom_tile(colour="darkorange", data = d %>% filter(same), size=0.5) +
+  #     scale_fill_gradient2(low="darkgrey", high = "darkorange",
+  #                      midpoint = 0.5) +
+  #     scale_colour_manual(values=c( "darkorange")) +
+  # scale_x_discrete("") +
+  # scale_y_discrete("")
+    d
+  })
+)
+ex2$example <- with(ex2, paste0(bullet1, " vs ", barrel2, "-", bullet2))
+ex2 <- unnest(ex2, data)
+ex2 %>% 
+  ggplot(aes(x = land1, y = land2, fill=rfscore)) + 
+  geom_tile() +
+  theme_bw() +
+  scale_fill_gradient2("RF score", low="darkgrey", high = "darkorange",
+                       midpoint = 0.5) +
+#  geom_tile(colour="darkgrey", data = ex2 %>% filter(same), size=0.5) +
+  facet_grid(.~example) +
+  scale_x_discrete("") +
+  scale_y_discrete("") +
+  annotate(geom="text", x = 0.5, y = 6.75, label=sprintf("SAM RF: %.3f", rfs2), hjust = 0)
+
+ex1 <- h44nest %>% filter(samesource == TRUE & sam_rf < 0.4)
+rfs <- ex1$sam_rf
+ex1 <- ex1 %>% mutate(
+  data = data %>% purrr::map(.f = function(d) {
+ #   browser()
+    index <- which.max(bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore))
+    d$same <- ((d$land2 - d$land1) %% 6)  == index-1
+
+  #   d %>% ggplot(aes(x = land1, y=land2, fill=rfscore)) + 
+  #                  geom_tile() +
+  #     geom_tile(colour="darkorange", data = d %>% filter(same), size=0.5) +
+  #     scale_fill_gradient2(low="darkgrey", high = "darkorange",
+  #                      midpoint = 0.5) +
+  #     scale_colour_manual(values=c( "darkorange")) +
+  # scale_x_discrete("") +
+  # scale_y_discrete("")
+    d
+  })
+)
+ex1$example <- with(ex1, paste0(bullet1, " vs ", barrel2, "-", bullet2))
+ex1 <- unnest(ex1, data)
+ex1 %>% 
+  ggplot(aes(x = land1, y = land2, fill=rfscore)) + 
+  geom_tile() +
+  theme_bw() +
+  scale_fill_gradient2("RF score", low="darkgrey", high = "darkorange",
+                       midpoint = 0.5) +
+  geom_tile(colour="darkorange", data = ex1 %>% filter(same), size=0.5) +
+  facet_grid(.~example) +
+  scale_x_discrete("") +
+  scale_y_discrete("") +
+  annotate(geom="text", x = 0.5, y = 6.75, label=sprintf("SAM RF: %.3f", rfs), hjust = 0)
+```
+
 # Discussion and Conclusions

### 2018-10-04 index.Rmd
overview of six close cases

@@
 the automatic matching  identifies the correct barrel for all questioned bullets.
+For two of the questioned bullets, 'I' and 'F' the similarity scores to the matching barrels are considerably smaller than for the other questioned bullets.
+Some of these close decisions are further investigated in \autoref{fig:close}.
 On the right of \autoref{fig:h44} the relationship between all pairs of questioned bullets is shown.
@@
 Therefore bullets 'P' and 'J' are also matching each other in the square on the right hand side.
+
+```{r close, fig.width=11.25, fig.height=4.2, fig.cap="Examples of land-to-land comparisons from the Hamby-44 set.
+Along the top are the three bullets from different source with the highest RF scores, along the bottom are the three bullets from same source with the lowest RF score.", fig.subcap=c("Ground truth: different source", "Ground truth: same-source"), fig.align="middle", fig.ncol=1}
+ex2 <- h44nest %>% filter(samesource == FALSE & sam_rf >= 0.295)
+rfs2 <- ex2$sam_rf
+ex2 <- ex2 %>% mutate(
+  data = data %>% purrr::map(.f = function(d) {
+ #   browser()
+    index <- which.max(bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore))
+    d$same <- ((d$land2 - d$land1) %% 6)  == index-1
+
+  #   d %>% ggplot(aes(x = land1, y=land2, fill=rfscore)) + 
+  #                  geom_tile() +
+  #     geom_tile(colour="darkorange", data = d %>% filter(same), size=0.5) +
+  #     scale_fill_gradient2(low="darkgrey", high = "darkorange",
+  #                      midpoint = 0.5) +
+  #     scale_colour_manual(values=c( "darkorange")) +
+  # scale_x_discrete("") +
+  # scale_y_discrete("")
+    d
+  })
+)
+ex2$example <- with(ex2, paste0(bullet1, " vs ", barrel2, "-", bullet2))
+ex2 <- unnest(ex2, data)
+ex2 %>% 
+  ggplot(aes(x = land1, y = land2, fill=rfscore)) + 
+  geom_tile() +
+  theme_bw() +
+  scale_fill_gradient2("RF score", low="darkgrey", high = "darkorange",
+                       midpoint = 0.5) +
+#  geom_tile(colour="darkgrey", data = ex2 %>% filter(same), size=0.5) +
+  facet_grid(.~example) +
+  scale_x_discrete("") +
+  scale_y_discrete("") +
+  annotate(geom="text", x = 0.5, y = 6.75, label=sprintf("SAM RF: %.3f", rfs2), hjust = 0)
+
+ex1 <- h44nest %>% filter(samesource == TRUE & sam_rf < 0.4)
+rfs <- ex1$sam_rf
+ex1 <- ex1 %>% mutate(
+  data = data %>% purrr::map(.f = function(d) {
+ #   browser()
+    index <- which.max(bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore))
+    d$same <- ((d$land2 - d$land1) %% 6)  == index-1
+
+  #   d %>% ggplot(aes(x = land1, y=land2, fill=rfscore)) + 
+  #                  geom_tile() +
+  #     geom_tile(colour="darkorange", data = d %>% filter(same), size=0.5) +
+  #     scale_fill_gradient2(low="darkgrey", high = "darkorange",
+  #                      midpoint = 0.5) +
+  #     scale_colour_manual(values=c( "darkorange")) +
+  # scale_x_discrete("") +
+  # scale_y_discrete("")
+    d
+  })
+)
+ex1$example <- with(ex1, paste0(bullet1, " vs ", barrel2, "-", bullet2))
+ex1 <- unnest(ex1, data)
+ex1 %>% 
+  ggplot(aes(x = land1, y = land2, fill=rfscore)) + 
+  geom_tile() +
+  theme_bw() +
+  scale_fill_gradient2("RF score", low="darkgrey", high = "darkorange",
+                       midpoint = 0.5) +
+  geom_tile(colour="darkorange", data = ex1 %>% filter(same), size=0.5) +
+  facet_grid(.~example) +
+  scale_x_discrete("") +
+  scale_y_discrete("") +
+  annotate(geom="text", x = 0.5, y = 6.75, label=sprintf("SAM RF: %.3f", rfs), hjust = 0)
+```
 
@@
 
-```{r, fig.width=7.5, fig.height=2.8, fig.cap="Examples of land-to-land comparisons from the Hamby-44 set.
-Along the top are the three bullets from different source with the highest RF scores, along the bottom are the three bullets from same source with the lowest RF score.", fig.subcap=c("Ground truth: different source", "Ground truth: same-source"), fig.align="middle", fig.ncol=1}
-ex2 <- h44nest %>% filter(samesource == FALSE & sam_rf >= 0.295)
-rfs2 <- ex2$sam_rf
-ex2 <- ex2 %>% mutate(
-  data = data %>% purrr::map(.f = function(d) {
- #   browser()
-    index <- which.max(bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore))
-    d$same <- ((d$land2 - d$land1) %% 6)  == index-1
-
-  #   d %>% ggplot(aes(x = land1, y=land2, fill=rfscore)) + 
-  #                  geom_tile() +
-  #     geom_tile(colour="darkorange", data = d %>% filter(same), size=0.5) +
-  #     scale_fill_gradient2(low="darkgrey", high = "darkorange",
-  #                      midpoint = 0.5) +
-  #     scale_colour_manual(values=c( "darkorange")) +
-  # scale_x_discrete("") +
-  # scale_y_discrete("")
-    d
-  })
-)
-ex2$example <- with(ex2, paste0(bullet1, " vs ", barrel2, "-", bullet2))
-ex2 <- unnest(ex2, data)
-ex2 %>% 
-  ggplot(aes(x = land1, y = land2, fill=rfscore)) + 
-  geom_tile() +
-  theme_bw() +
-  scale_fill_gradient2("RF score", low="darkgrey", high = "darkorange",
-                       midpoint = 0.5) +
-#  geom_tile(colour="darkgrey", data = ex2 %>% filter(same), size=0.5) +
-  facet_grid(.~example) +
-  scale_x_discrete("") +
-  scale_y_discrete("") +
-  annotate(geom="text", x = 0.5, y = 6.75, label=sprintf("SAM RF: %.3f", rfs2), hjust = 0)
-
-ex1 <- h44nest %>% filter(samesource == TRUE & sam_rf < 0.4)
-rfs <- ex1$sam_rf
-ex1 <- ex1 %>% mutate(
-  data = data %>% purrr::map(.f = function(d) {
- #   browser()
-    index <- which.max(bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore))
-    d$same <- ((d$land2 - d$land1) %% 6)  == index-1
-
-  #   d %>% ggplot(aes(x = land1, y=land2, fill=rfscore)) + 
-  #                  geom_tile() +
-  #     geom_tile(colour="darkorange", data = d %>% filter(same), size=0.5) +
-  #     scale_fill_gradient2(low="darkgrey", high = "darkorange",
-  #                      midpoint = 0.5) +
-  #     scale_colour_manual(values=c( "darkorange")) +
-  # scale_x_discrete("") +
-  # scale_y_discrete("")
-    d
-  })
-)
-ex1$example <- with(ex1, paste0(bullet1, " vs ", barrel2, "-", bullet2))
-ex1 <- unnest(ex1, data)
-ex1 %>% 
-  ggplot(aes(x = land1, y = land2, fill=rfscore)) + 
-  geom_tile() +
-  theme_bw() +
-  scale_fill_gradient2("RF score", low="darkgrey", high = "darkorange",
-                       midpoint = 0.5) +
-  geom_tile(colour="darkorange", data = ex1 %>% filter(same), size=0.5) +
-  facet_grid(.~example) +
-  scale_x_discrete("") +
-  scale_y_discrete("") +
-  annotate(geom="text", x = 0.5, y = 6.75, label=sprintf("SAM RF: %.3f", rfs), hjust = 0)
-```
-
 # Discussion and Conclusions

### 2018-10-05 index.Rmd
labelling added

--- 2018-10-02 index.Rmd
@@
 h44features <- read.csv("data/h44-features.csv")
+# tank rash lands:
+h44features <- h44features %>% mutate(
+  rfscore = replace(rfscore, (bullet1 == "I" & land1 == 6) | (bullet2 == "I" & land2 == 6), NA)
+)
+h44features <- h44features %>% mutate(
+  rfscore = replace(rfscore, (barrel1==8 & bullet1 == "1" & land1 == 6) | (barrel2==8 & bullet2 == "1" & land2 == 6), NA)
+)
 
@@
 
-h44nest <- h44nest %>% mutate(
-  sam_ccf = data %>% purrr::map_dbl(.f=function(d) {
-    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$ccf)
-    max(scores)
-  }),
+h44nest <- h44nest %>% filter(bullet1=="I") %>%  mutate(
+  # sam_ccf = data %>% purrr::map_dbl(.f=function(d) {
+  #   scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$ccf)
+  #   max(scores)
+  # }),
   sam_rf = data %>% purrr::map_dbl(.f=function(d) {
@@
   scale_fill_gradient2("RF score ", low="darkgrey", 
-                       high="darkorange", midpoint=0.375, limits=c(0,1)) + 
+                       high="darkorange", midpoint=0.35, limits=c(0,1)) + 
   scale_colour_manual("same source", values="darkorange") +
@@
 the automatic matching  identifies the correct barrel for all questioned bullets.
+For two of the questioned bullets, 'I' and 'F' the similarity scores to the matching barrels are considerably smaller than for the other questioned bullets.
 On the right of \autoref{fig:h44} the relationship between all pairs of questioned bullets is shown.
@@
 Therefore bullets 'P' and 'J' are also matching each other in the square on the right hand side.
+
+```{r close, eval=FALSE, fig.width=11.25, fig.height=4.2, fig.cap="Examples of land-to-land comparisons from the Hamby-44 set.
+Along the top are the three bullets from different source with the highest RF scores, along the bottom are the three bullets from same source with the lowest RF score.", fig.subcap=c("Ground truth: different source", "Ground truth: same-source"), fig.align="middle", fig.ncol=1}
+# ex2 <- h44nest %>% filter(samesource == FALSE & sam_rf >= 0.287)
+# rfs2 <- ex2$sam_rf
+# ex2 <- ex2 %>% mutate(
+#   data = data %>% purrr::map(.f = function(d) {
+#  #   browser()
+#     index <- which.max(bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore))
+#     d$same <- ((d$land2 - d$land1) %% 6)  == index-1
+#     
+#   #   d %>% ggplot(aes(x = land1, y=land2, fill=rfscore)) + 
+#   #                  geom_tile() +
+#   #     geom_tile(colour="darkorange", data = d %>% filter(same), size=0.5) +
+#   #     scale_fill_gradient2(low="darkgrey", high = "darkorange",
+#   #                      midpoint = 0.5) +
+#   #     scale_colour_manual(values=c( "darkorange")) +
+#   # scale_x_discrete("") +
+#   # scale_y_discrete("")
+#     d
+#   })
+# )
+# ex2$example <- with(ex2, paste0(bullet1, " vs ", barrel2, "-", bullet2))
+# ex2 <- unnest(ex2, data)
+# ex2 %>% 
+#   ggplot(aes(x = land1, y = land2, fill=rfscore)) + 
+#   geom_tile() +
+#   theme_bw() +
+#   scale_fill_gradient2("RF score", low="darkgrey", high = "darkorange",
+#                        midpoint = 0.5) +
+# #  geom_tile(colour="darkgrey", data = ex2 %>% filter(same), size=0.5) +
+#   facet_grid(.~example) +
+#   scale_x_discrete("") +
+#   scale_y_discrete("") +
+#   annotate(geom="text", x = 0.5, y = 6.75, label=sprintf("SAM RF: %.3f", rfs2), hjust = 0)
+# 
+# 
+# 
+# ex1 <- h44nest %>% filter(samesource == TRUE & sam_rf < 0.4)
+# rfs <- ex1$sam_rf
+# ex1 <- ex1 %>% mutate(
+#   data = data %>% purrr::map(.f = function(d) {
+#  #   browser()
+#     index <- which.max(bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore))
+#     d$same <- ((d$land2 - d$land1) %% 6)  == index-1
+#     
+#   #   d %>% ggplot(aes(x = land1, y=land2, fill=rfscore)) + 
+#   #                  geom_tile() +
+#   #     geom_tile(colour="darkorange", data = d %>% filter(same), size=0.5) +
+#   #     scale_fill_gradient2(low="darkgrey", high = "darkorange",
+#   #                      midpoint = 0.5) +
+#   #     scale_colour_manual(values=c( "darkorange")) +
+#   # scale_x_discrete("") +
+#   # scale_y_discrete("")
+#     d
+#   })
+# )
+# ex1$example <- with(ex1, paste0(bullet1, " vs ", barrel2, "-", bullet2))
+# ex1 <- unnest(ex1, data)
+# ex1 %>% 
+#   ggplot(aes(x = land1, y = land2, fill=rfscore)) + 
+#   geom_tile() +
+#   theme_bw() +
+#   scale_fill_gradient2("RF score", low="darkgrey", high = "darkorange",
+#                        midpoint = 0.5, na.value = NA) +
+#   geom_tile(colour="darkorange", data = ex1 %>% filter(same), size=0.5) +
+#   facet_grid(.~example) +
+#   scale_x_discrete("") +
+#   scale_y_discrete("") +
+#   annotate(geom="text", x = 0.5, y = 6.75, label=sprintf("SAM RF: %.3f", rfs), hjust = 0)
+
+# \autoref{fig:close} shows an example of six comparisons of bullet pairs at the land-to-land level.
+The panels along the top of the figure show three bullet pairs that are known to be from different barrels, the three bullet pairs along the bottom are known to be from the same barrel.
+The light orange frame in the bottom three panels indicate which lands correspond to one-another in the matching.
+It is obvious, that not all of the scores are as high as we would usually see in same-source comparisons.
+```
 
@@
 
+labels <- h44nest %>% filter( samesource, sam_rf < 0.5)
+labels$comparison <- with(labels, paste(bullet1,"vs", paste(barrel2,bullet2,sep="-")))
+labels$measure <- "RF score"
+labels$study <- "Hamby-44"
+labels$y <- 5+4:1
 nestslong %>% ggplot(aes(x = value, fill=samesource, 
@@
   theme(legend.position="bottom") +
-  geom_hline(yintercept =0) 
+  geom_hline(yintercept =0) +
+  geom_text(aes(x = sam_rf, y = y, label=comparison), 
+            hjust = 0,
+            data = labels) +
+  geom_segment(aes(x = sam_rf, xend = sam_rf, y = 0, yend = y), 
+            size=0.1,
+            colour="grey20",
+            data = labels)
+
 ```

### 2018-10-06 index.Rmd
tank rash bullets, added some conclusions

@@
 
-h44nest <- h44nest %>% filter(bullet1=="I") %>%  mutate(
-  # sam_ccf = data %>% purrr::map_dbl(.f=function(d) {
-  #   scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$ccf)
-  #   max(scores)
-  # }),
+h44nest <- h44nest %>%  mutate(
+  sam_ccf = data %>% purrr::map_dbl(.f=function(d) {
+    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$ccf)
+    max(scores)
+  }),
   sam_rf = data %>% purrr::map_dbl(.f=function(d) {
@@
 
+# Appendix
+
+\autoref{fig:tankrash} shows an overview of all lands of bullets in the Hamby-44 with major deficiencies, such as 'tank rash' or extreme pitting (holes caused by direct contact with burning gun powder).
+```{r tankrash, fig.cap="Overview of bullet lands with prominent deficiency such as tank rash or extreme pitting in the Hamby-44 study.", fig.width=4, fig.height = 2, out.width='47.5%', fig.subcap=c("Br1-B1-L6", "Br2-B2-L5", "Br3-B1-L5", "Br8-B1-L6", "Br8-B2-L2", "Br8-B2-L6", "Unk-E-L6", "Unk-I-L6"), fig.align="middle", fig.ncol=2}
+knitr::include_graphics(
+  c('./images/B1-B1-L6.png', './images/B2-B2-L5.png',
+    './images/B3-B1-L5.png', './images/B8-B1-L6.png',
+    './images/B8-B2-L2.png', './images/B8-B2-L6.png',
+    './images/NA-BE-L6.png', './images/NA-BI-L6.png')
+)
+
+```
+
+Three lands across the two known bullets of barrel 8 are affected. This is the likely cause for the low scores for barrel 8 in \autoref{fig:h44}.
+
+### Limitations
+All of the studies involve the same firearm.
+We see some variations in scores across different models of Rugers.
+Likely, we are going to see even bigger variations with different brands of firearms, where there might be different number of lands of different lengths.
+We also know \cite{???} that different firearms mark differently well.
+Rugers are among the firearms that mark very well, so we would correspondingly expect to see lowered similarity scores for other firearms.
+The matching algorithm works on matching lands.
+Barrels with polygonal rifling, such as Glocks, do not introduce well defined lands on bullets, which makes an application of the matching algorithm impossible.
+
+XXX Should we include some Sig Sauer matches?
+
 # References

### 2018-10-09 index.Rmd
Add comment color, use styler to clean up code in the file according to tidyverse style guidelines

@@
 \newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
+\newcommand{\svp}[1]{{\textcolor{darkgreen}{#1}}}
 \noindent
@@
   fig.align = "center",
-  out.width= '\\textwidth',
+  out.width = "\\textwidth",
   cache = FALSE,
-  fig.path='figures/',
-  echo=FALSE
+  fig.path = "figures/",
+  echo = FALSE
 )
@@
 ```{r functions}
-estimateBeta <- function(x, method = "mm", probs=c(0.25, 0.75)) {
-  if (method =="mm") {
-    mu = mean(x, na.rm=TRUE)
-    sd = sd(x, na.rm=TRUE)
-    lambda <- mu/sd*(1-mu)/sd - 1
-    alpha <- mu*lambda
-    beta <- (1-mu)*lambda
+estimateBeta <- function(x, method = "mm", probs = c(0.25, 0.75)) {
+  if (method == "mm") {
+    mu <- mean(x, na.rm = TRUE)
+    sd <- sd(x, na.rm = TRUE)
+    lambda <- mu / sd * (1 - mu) / sd - 1
+    alpha <- mu * lambda
+    beta <- (1 - mu) * lambda
   }
-  if (method =="LB") {
-    quantile1=list(p=probs[1],x=quantile(probs=probs[1], x))
-    quantile2=list(p=probs[2],x=quantile(probs=probs[2], x))
-    bs <- LearnBayes::beta.select(quantile1,quantile2)
-    alpha = bs[1]
-    beta = bs[2]
+  if (method == "LB") {
+    quantile1 <- list(p = probs[1], x = quantile(probs = probs[1], x))
+    quantile2 <- list(p = probs[2], x = quantile(probs = probs[2], x))
+    bs <- LearnBayes::beta.select(quantile1, quantile2)
+    alpha <- bs[1]
+    beta <- bs[2]
   }
-  return(c(alpha=alpha, beta=beta))
+  return(c(alpha = alpha, beta = beta))
 }
@@
 signatures %>%
-  ggplot(aes(x = x/1000)) + 
+  ggplot(aes(x = x / 1000)) +
   geom_line(aes(y = raw_sig), colour = "grey70") +
   geom_line(aes(y = sig), colour = "grey30") +
-  facet_grid(bullet~land) +
-  ylim(c(-5,5)) +
+  facet_grid(bullet ~ land) +
+  ylim(c(-5, 5)) +
   theme_bw() +
@@
 features <- features %>% mutate(
-  bullet1 = gsub("ullet ","", bullet1),
-  bullet2 = gsub("ullet ","", bullet2)
+  bullet1 = gsub("ullet ", "", bullet1),
+  bullet2 = gsub("ullet ", "", bullet2)
 )
-b12 <- features %>% filter(group1 == "G1", barrel1=="KA", bullet1== "B1",
-                           group2 == "G1", barrel2=="KA", bullet2== "B2")
-p1 <- b12 %>% ggplot(aes(x = land1, y = land2, fill = rfscore)) + 
-  geom_tile() + ggtitle ("Same barrel") +  
-  scale_fill_gradient2("RF score: ", midpoint=0.3, low = "darkgrey", high = "darkorange", limits=c(0,1)) +
+b12 <- features %>% filter(
+  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
+  group2 == "G1", barrel2 == "KA", bullet2 == "B2"
+)
+p1 <- b12 %>%
+  ggplot(aes(x = land1, y = land2, fill = rfscore)) +
+  geom_tile() + ggtitle("Same barrel") +
+  scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
   theme_bw() +
-  xlab ("Bullet 2") + ylab("Bullet 1") +
+  xlab("Bullet 2") + ylab("Bullet 1") +
   theme(legend.position = "bottom") + coord_equal()
 
-bab <- features %>% filter(group1 == "G1", barrel1=="KA", bullet1== "B1",
-                           group2 == "G1", barrel2=="KB", bullet2== "B1")
-p2 <- bab %>% ggplot(aes(x = land1, y = land2, fill = rfscore)) + 
-  geom_tile() + ggtitle ("Different barrel") +  
-  scale_fill_gradient2("RF score: ", midpoint=0.3, low = "darkgrey", high = "darkorange", limits=c(0,1)) +
+bab <- features %>% filter(
+  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
+  group2 == "G1", barrel2 == "KB", bullet2 == "B1"
+)
+p2 <- bab %>%
+  ggplot(aes(x = land1, y = land2, fill = rfscore)) +
+  geom_tile() + ggtitle("Different barrel") +
+  scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
   theme_bw() +
-  xlab ("Bullet 2") + ylab("Bullet 1") +
+  xlab("Bullet 2") + ylab("Bullet 1") +
   theme(legend.position = "bottom") + coord_equal()
@@
 On the left the two bullets are known to come from the same source (barrel) on the right, the bullets are from two different sources (barrels)"}
-grid.arrange(p1, p2, ncol=2, padding=2)
+grid.arrange(p1, p2, ncol = 2, padding = 2)
 ```
@@
 df2 <- data.frame(expand.grid(x = 1:6, y = 1:6))
-df2$case1 <- df2$x == ((df2$y+5) %% 6) + 1
-df2$case2 <- df2$x == ((df2$y+0) %% 6) + 1
-df2$case3 <- df2$x == ((df2$y+1) %% 6) + 1
-df2$case4 <- df2$x == ((df2$y+2) %% 6) + 1
-df2$case5 <- df2$x == ((df2$y+3) %% 6) + 1
-df2$case6 <- df2$x == ((df2$y+4) %% 6) + 1
+df2$case1 <- df2$x == ((df2$y + 5) %% 6) + 1
+df2$case2 <- df2$x == ((df2$y + 0) %% 6) + 1
+df2$case3 <- df2$x == ((df2$y + 1) %% 6) + 1
+df2$case4 <- df2$x == ((df2$y + 2) %% 6) + 1
+df2$case5 <- df2$x == ((df2$y + 3) %% 6) + 1
+df2$case6 <- df2$x == ((df2$y + 4) %% 6) + 1
 
-df2 %>% gather("case", "values", starts_with("case")) %>% 
+df2 %>%
+  gather("case", "values", starts_with("case")) %>%
   mutate(case = gsub("case", "", case)) %>%
-  mutate(case = paste0("k = ", as.numeric(case)-1)) %>%
-  ggplot(aes(x = factor(x), y = factor(y, levels=1:6), fill = values)) + 
+  mutate(case = paste0("k = ", as.numeric(case) - 1)) %>%
+  ggplot(aes(x = factor(x), y = factor(y, levels = 1:6), fill = values)) +
   geom_tile(colour = "grey20") +
@@
   xlab("Bullet 2") + ylab("Bullet 1") +
-  scale_x_discrete(labels = paste0("L",1:6)) +
-  scale_y_discrete(labels = paste0("L",1:6)) +
+  scale_x_discrete(labels = paste0("L", 1:6)) +
+  scale_y_discrete(labels = paste0("L", 1:6)) +
   scale_fill_manual("Cell included in SA(A,k)", values = c("white", "grey20")) +
   coord_equal() +
-  theme(legend.position="bottom")
+  theme(legend.position = "bottom")
 ```
@@
 h44features <- h44features %>% mutate(
-  rfscore = replace(rfscore, (barrel1==8 & bullet1 == "1" & land1 == 6) | (barrel2==8 & bullet2 == "1" & land2 == 6), NA)
+  rfscore = replace(rfscore, (barrel1 == 8 & bullet1 == "1" & land1 == 6) | (barrel2 == 8 & bullet2 == "1" & land2 == 6), NA)
 )
@@
 h44nest <- h44features %>%
-  filter(barrel1=="Unk") %>%
+  filter(barrel1 == "Unk") %>%
   filter(barrel1 != barrel2 | bullet1 != bullet2) %>%
@@
     barrel2 = as.character(barrel2),
-    barrel2 = replace(barrel2, barrel2=="Unk", "Unknown")
+    barrel2 = replace(barrel2, barrel2 == "Unk", "Unknown")
   ) %>%
-  group_by(bullet1, barrel2, bullet2) %>%  nest()
+  group_by(bullet1, barrel2, bullet2) %>%
+  nest()
 
 h44nest <- h44nest %>% mutate(
-  sam_ccf = data %>% purrr::map_dbl(.f=function(d) {
+  sam_ccf = data %>% purrr::map_dbl(.f = function(d) {
     scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$ccf)
@@
   }),
-  sam_rf = data %>% purrr::map_dbl(.f=function(d) {
+  sam_rf = data %>% purrr::map_dbl(.f = function(d) {
     scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore)
@@
   mutate(
-    barrel2 = factor(barrel2, levels =c(1:10, "Unknown")),
-    bullet1 = factor(bullet1, levels=rev(c("K", "O","L", "P",
+    barrel2 = factor(barrel2, levels = c(1:10, "Unknown")),
+    bullet1 = factor(bullet1, levels = rev(c(
+      "K", "O", "L", "P",
       "J",
@@
       "E", "U", "X", "F",
-                                         "T", "S"))),
-    bullet2 = factor(bullet2, levels=c(1:2, rev(levels(bullet1))))
+      "T", "S"
+    ))),
+    bullet2 = factor(bullet2, levels = c(1:2, rev(levels(bullet1))))
   )
 h44 <- h44nest %>%
-  ggplot(aes(y = bullet1, x = bullet2, 
-             fill = sam_rf)) + 
-  geom_tile(size=1) + 
-  facet_grid(.~barrel2, space="free", scales = "free") +
+  ggplot(aes(
+    y = bullet1, x = bullet2,
+    fill = sam_rf
+  )) +
+  geom_tile(size = 1) +
+  facet_grid(. ~ barrel2, space = "free", scales = "free") +
   ylab("Questioned bullets") +
   xlab("Testfires from barrels 1 to 10 (left), questioned bullets (right)") +
-  scale_fill_gradient2("RF score ", low="darkgrey", 
-                       high="darkorange", midpoint=0.35, limits=c(0,1)) + 
-  scale_colour_manual("same source", values="darkorange") +
+  scale_fill_gradient2("RF score ",
+    low = "darkgrey",
+    high = "darkorange", midpoint = 0.35, limits = c(0, 1)
+  ) +
+  scale_colour_manual("same source", values = "darkorange") +
 
@@
   theme_bw() +
-    theme(legend.position="bottom")
+  theme(legend.position = "bottom")
 
@@
 f3 <- rbind(
-  f2 %>% select(b1, b2, barrel1, bullet1, 
-                land1, bullet2, land2, ccf, rfscore, KM) %>%
+  f2 %>%
+    select(
+      b1, b2, barrel1, bullet1,
+      land1, bullet2, land2, ccf, rfscore, KM
+    ) %>%
     mutate(barrel2 = "Unknown"),
-  unknowns %>% select(b1, b2, bullet1, 
-                land1, bullet2, land2, ccf, rfscore) %>%
-    mutate(barrel1 ="Unknown", barrel2="Unknown", KM=FALSE)
+  unknowns %>%
+    select(
+      b1, b2, bullet1,
+      land1, bullet2, land2, ccf, rfscore
+    ) %>%
+    mutate(barrel1 = "Unknown", barrel2 = "Unknown", KM = FALSE)
 )
@@
 pdnest <- pdnest %>% mutate(
-  sam_ccf = data %>% purrr::map_dbl(.f=function(d) {
+  sam_ccf = data %>% purrr::map_dbl(.f = function(d) {
     scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$ccf)
@@
   }),
-  KM = data %>% purrr::map_lgl(.f=function(d) { any(d$KM) })
+  KM = data %>% purrr::map_lgl(.f = function(d) {
+    any(d$KM)
+  })
 )
@@
   mutate(
-    bullet2 = factor(bullet2, levels=rev(c("N", "B", "E", "T", "H", "J", "K", "Q", "Y", "Z"))),
-    bullet1 = factor(bullet1, levels=c("B1", "B2", "B3", rev(levels(bullet2))))
+    bullet2 = factor(bullet2, levels = rev(c("N", "B", "E", "T", "H", "J", "K", "Q", "Y", "Z"))),
+    bullet1 = factor(bullet1, levels = c("B1", "B2", "B3", rev(levels(bullet2))))
   )
@@
   ggplot(aes(y = bullet2, x = bullet1, fill = sam_rf)) +
-  geom_tile(size=1) + 
-  facet_grid(.~barrel1, scales="free", space="free") +
+  geom_tile(size = 1) +
+  facet_grid(. ~ barrel1, scales = "free", space = "free") +
   ylab("Questioned bullets") +
   xlab("Known bullets and Questioned bullets") +
-  scale_fill_gradient2("RF score ", low="darkgrey", 
-                       high="darkorange", midpoint=0.45, limits=c(0,1)) + 
-  scale_colour_manual("same source", values="darkorange") + 
+  scale_fill_gradient2("RF score ",
+    low = "darkgrey",
+    high = "darkorange", midpoint = 0.45, limits = c(0, 1)
+  ) +
+  scale_colour_manual("same source", values = "darkorange") +
   scale_y_discrete() +
   theme_bw() +
-  theme(legend.position="bottom") +
+  theme(legend.position = "bottom") +
   #  coord_equal() +
-  geom_tile(aes(colour = TRUE), size = .5,
-            data = filter(pdnest, KM)) +
+  geom_tile(aes(colour = TRUE),
+    size = .5,
+    data = filter(pdnest, KM)
+  ) +
   guides(colour = guide_legend(override.aes = list(fill = NA)))
@@
 features <- features %>% mutate(
-  bullet1 = gsub("ullet ","", bullet1),
-  bullet2 = gsub("ullet ","", bullet2)
+  bullet1 = gsub("ullet ", "", bullet1),
+  bullet2 = gsub("ullet ", "", bullet2)
 )
@@
 hounest <- groups %>% mutate(
-  sam_ccf = data %>% purrr::map_dbl(.f=function(d) {
+  sam_ccf = data %>% purrr::map_dbl(.f = function(d) {
     scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$ccf)
@@
 
-hou1nest <- hounest %>% filter(group1 == "G1") %>%
+hou1nest <- hounest %>%
+  filter(group1 == "G1") %>%
   mutate(
-    bullet1 = factor(bullet1, levels=rev(c("U40", "U36", "U42", "U10", "U77", "U37", "U28", "U15"))),
-    bullet2 = factor(bullet2, levels=c("B1", "B2", "B3", levels(bullet1)))
+    bullet1 = factor(bullet1, levels = rev(c("U40", "U36", "U42", "U10", "U77", "U37", "U28", "U15"))),
+    bullet2 = factor(bullet2, levels = c("B1", "B2", "B3", levels(bullet1)))
   )
@@
   ggplot(aes(y = bullet1, x = bullet2, fill = sam_rf)) +
-  geom_tile(size=1) + 
-  facet_grid(.~barrel2, scales = "free", space="free") +
+  geom_tile(size = 1) +
+  facet_grid(. ~ barrel2, scales = "free", space = "free") +
   ylab("Questioned bullets") +
   xlab("Known bullets and Questioned bullets") +
-  scale_fill_gradient2("RF score ", low="darkgrey", 
-                       high="darkorange", midpoint=0.375, limits=c(0,1)) + 
+  scale_fill_gradient2("RF score ",
+    low = "darkgrey",
+    high = "darkorange", midpoint = 0.375, limits = c(0, 1)
+  ) +
   scale_y_discrete() +
-  scale_colour_manual("same source", values="darkorange") +
+  scale_colour_manual("same source", values = "darkorange") +
   geom_tile(aes(colour = TRUE), size = .5, data = filter(hou1nest, sam_rf > 0.3)) +
@@
 
-hou2nest <- hounest %>% filter(group1 == "G2") %>%
+hou2nest <- hounest %>%
+  filter(group1 == "G2") %>%
   mutate(
-    bullet1 = factor(bullet1, levels=rev(c("U23", "U41", "U61", "U98", "U73", "U34", "U63", "U66"))),
-    bullet2 = factor(bullet2, levels=c("B1", "B2", "B3", levels(bullet1)))
+    bullet1 = factor(bullet1, levels = rev(c("U23", "U41", "U61", "U98", "U73", "U34", "U63", "U66"))),
+    bullet2 = factor(bullet2, levels = c("B1", "B2", "B3", levels(bullet1)))
   )
@@
   ggplot(aes(y = bullet1, x = bullet2, fill = sam_rf)) +
-  geom_tile(size=1) + 
-  facet_grid(.~barrel2, scales = "free", space="free") +
+  geom_tile(size = 1) +
+  facet_grid(. ~ barrel2, scales = "free", space = "free") +
   ylab("Questioned bullets") +
   xlab("Known bullets and Questioned bullets") +
-  scale_fill_gradient2("RF score ", low="darkgrey", 
-                       high="darkorange", midpoint=0.375, limits=c(0,1)) + 
+  scale_fill_gradient2("RF score ",
+    low = "darkgrey",
+    high = "darkorange", midpoint = 0.375, limits = c(0, 1)
+  ) +
   scale_y_discrete() +
-    scale_colour_manual("same source", values="darkorange") +
+  scale_colour_manual("same source", values = "darkorange") +
   geom_tile(aes(colour = TRUE), size = .5, data = filter(hou2nest, sam_rf > 0.3)) +
@@
 
-hou3nest <- hounest %>% filter(group1 == "G3") %>%
+hou3nest <- hounest %>%
+  filter(group1 == "G3") %>%
   mutate(
-    bullet1 = factor(bullet1, levels=rev(c("U27", "U33", "U36", "U49", "U56", "U65", "U14", "U45"))),
-    bullet2 = factor(bullet2, levels=c("B1", "B2", "B3", levels(bullet1)))
+    bullet1 = factor(bullet1, levels = rev(c("U27", "U33", "U36", "U49", "U56", "U65", "U14", "U45"))),
+    bullet2 = factor(bullet2, levels = c("B1", "B2", "B3", levels(bullet1)))
   )
@@
   ggplot(aes(y = bullet1, x = bullet2, fill = sam_rf)) +
-  geom_tile(size=1) + 
-  facet_grid(.~barrel2, scales = "free", space="free") +
+  geom_tile(size = 1) +
+  facet_grid(. ~ barrel2, scales = "free", space = "free") +
   ylab("Questioned bullets") +
   xlab("Known bullets and Questioned bullets") +
-  scale_fill_gradient2("RF score ", low="darkgrey", 
-                       high="darkorange", midpoint=0.375, limits=c(0,1)) + 
+  scale_fill_gradient2("RF score ",
+    low = "darkgrey",
+    high = "darkorange", midpoint = 0.375, limits = c(0, 1)
+  ) +
   scale_y_discrete() +
-    scale_colour_manual("same source", values="darkorange") +
+  scale_colour_manual("same source", values = "darkorange") +
   geom_tile(aes(colour = TRUE), size = .5, data = filter(hou3nest, sam_rf > 0.3)) +
@@
 nests <- rbind(
-  h44nest %>% select(sam_rf, sam_ccf, samesource) %>%
-    mutate(study = "Hamby-44", group1=NA),
-  pdnest %>% select(sam_rf, sam_ccf, samesource) %>%
-    mutate(study = "Phoenix PD", group1=NA),
-  hounest %>% select(sam_rf, sam_ccf, samesource, group1) %>%
+  h44nest %>%
+    select(sam_rf, sam_ccf, samesource) %>%
+    mutate(study = "Hamby-44", group1 = NA),
+  pdnest %>%
+    select(sam_rf, sam_ccf, samesource) %>%
+    mutate(study = "Phoenix PD", group1 = NA),
+  hounest %>%
+    select(sam_rf, sam_ccf, samesource, group1) %>%
     mutate(study = "Houston FSI")
 )
-nests$study <- factor(nests$study, levels=c("Hamby-44", "Phoenix PD", "Houston FSI"))
-nestslong <- nests %>% gather(key="measure", value="value", sam_rf, sam_ccf)
+nests$study <- factor(nests$study, levels = c("Hamby-44", "Phoenix PD", "Houston FSI"))
+nestslong <- nests %>% gather(key = "measure", value = "value", sam_rf, sam_ccf)
 nestslong$measure <- factor(nestslong$measure)
@@
 
-labels <- h44nest %>% filter( samesource, sam_rf < 0.5)
-labels$comparison <- with(labels, paste(bullet1,"vs", paste(barrel2,bullet2,sep="-")))
+labels <- h44nest %>% filter(samesource, sam_rf < 0.5)
+labels$comparison <- with(labels, paste(bullet1, "vs", paste(barrel2, bullet2, sep = "-")))
 labels$measure <- "RF score"
 labels$study <- "Hamby-44"
-labels$y <- 5+4:1
-nestslong %>% ggplot(aes(x = value, fill=samesource, 
-                     colour=samesource)) +
-  geom_density(alpha=0.6) +
-  geom_rug(alpha=0.5) +
-  facet_grid(measure~study) +
+labels$y <- 5 + 4:1
+nestslong %>%
+  ggplot(aes(
+    x = value, fill = samesource,
+    colour = samesource
+  )) +
+  geom_density(alpha = 0.6) +
+  geom_rug(alpha = 0.5) +
+  facet_grid(measure ~ study) +
   theme_bw() +
-  scale_fill_manual("Same source", values=c("darkgrey", "darkorange")) +
-  scale_colour_manual("Same source", values=c("darkgrey", "darkorange")) +
-  scale_x_continuous("SAM scores", limits=c(0,1)) +
-  theme(legend.position="bottom") +
-  geom_hline(yintercept =0) +
-  geom_text(aes(x = sam_rf, y = y, label=comparison), 
+  scale_fill_manual("Same source", values = c("darkgrey", "darkorange")) +
+  scale_colour_manual("Same source", values = c("darkgrey", "darkorange")) +
+  scale_x_continuous("SAM scores", limits = c(0, 1)) +
+  theme(legend.position = "bottom") +
+  geom_hline(yintercept = 0) +
+  geom_text(aes(x = sam_rf, y = y, label = comparison),
     hjust = 0,
-            data = labels) +
+    data = labels
+  ) +
   geom_segment(aes(x = sam_rf, xend = sam_rf, y = 0, yend = y),
-            size=0.1,
-            colour="grey20",
-            data = labels)
-
+    size = 0.1,
+    colour = "grey20",
+    data = labels
+  )
 ```
@@
 knitr::include_graphics(
-  c('./images/B1-B1-L6.png', './images/B2-B2-L5.png',
-    './images/B3-B1-L5.png', './images/B8-B1-L6.png',
-    './images/B8-B2-L2.png', './images/B8-B2-L6.png',
-    './images/NA-BE-L6.png', './images/NA-BI-L6.png')
+  c(
+    "./images/B1-B1-L6.png", "./images/B2-B2-L5.png",
+    "./images/B3-B1-L5.png", "./images/B8-B1-L6.png",
+    "./images/B8-B2-L2.png", "./images/B8-B2-L6.png",
+    "./images/NA-BE-L6.png", "./images/NA-BI-L6.png"
   )
-
+)
 ```

### 2018-10-09 index.Rmd
references updated

--- 2018-10-06 index.Rmd
@@
 What is quite surprising, that what we anticipated to be the easiest case of Hamby 44 turned out to be the hardest to solve. 
-
-# Appendix
-
-\autoref{fig:tankrash} shows an overview of all lands of bullets in the Hamby-44 with major deficiencies, such as 'tank rash' or extreme pitting (holes caused by direct contact with burning gun powder).
+This is likely due to several lands in the Hamby-44 set having major deficiencies,  such as 'tank rash' (a collision of the bullet after exiting the barrel with a surface causing markings on top of the striations from the barrel) or extreme pitting (holes caused by direct contact with burning gun powder).
+\autoref{fig:tankrash} gives an overview of all lands of bullets in the Hamby-44 set affected. 
 ```{r tankrash, fig.cap="Overview of bullet lands with prominent deficiency such as tank rash or extreme pitting in the Hamby-44 study.", fig.width=4, fig.height = 2, out.width='47.5%', fig.subcap=c("Br1-B1-L6", "Br2-B2-L5", "Br3-B1-L5", "Br8-B1-L6", "Br8-B2-L2", "Br8-B2-L6", "Unk-E-L6", "Unk-I-L6"), fig.align="middle", fig.ncol=2}
@@
 
-Three lands across the two known bullets of barrel 8 are affected. This is the likely cause for the low scores for barrel 8 in \autoref{fig:h44}.
+Three lands from bullets known to be fired from barrel 8 are affected.
+This is the likely cause for the low scores for barrel 8 in \autoref{fig:h44}.
+We also see that questioned bullet 'E' is affected, explaining some of the low similarity scores for this bullet.
 
@@
 Likely, we are going to see even bigger variations with different brands of firearms, where there might be different number of lands of different lengths.
-We also know \cite{???} that different firearms mark differently well.
+We also know  that different firearms mark differently well \cite{???}.
 Rugers are among the firearms that mark very well, so we would correspondingly expect to see lowered similarity scores for other firearms.
@@
 
-XXX Should we include some Sig Sauer matches?
+As seen in the earlier discussion, similarity scores are affected by deficiencies on lands.
+It has been shown \citet{lpr} that parts of lands can be used for successful identifications.
+The  algorithm proposed by \citet{aoas} is not capable of  automatically detecting parts of lands with well expressed striae.
+XXX citation for neural network algorithm
 

### 2018-10-09 index.Rmd
wording changes

--- 2018-10-09 index.Rmd
@@
 \newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
-\newcommand{\svp}[1]{{\textcolor{darkgreen}{#1}}}
+\newcommand{\svp}[1]{{\textcolor{teal}{#1}}}
 \noindent
@@
 Exact guidelines for this classification vary from lab to lab, e.g.\ some labs will exclude only on the basis of non-matching class characteristics (such as direction of the twist in rifling, land length or number of lands, or type of rifling).
-In some labs,  CMS (consecutively matching striae) as defined by \citet{biasotti} is used as a way to quantify the strength of a match.
+In some labs,  CMS (consecutively matching striae) as defined by \citet{biasotti} \svp{are} used as a way to quantify the strength of a match.
 
-In the algorithm  similarity of bullet land engraved areas is evaluated based on 3d topographic scans of land engraved areas (LEAs). From each land engraved area a signature is derived by
+In the \svp{matching algorithm,} similarity of bullet land engraved areas is evaluated based on 3d topographic scans of land engraved areas (LEAs).
+From each land engraved area a signature is derived by
 \begin{itemize}
@@
 Quantitative features describing the similarity of two signature are extracted.
-Features include characteristics of the signature that firearms and toolmark examiners base their visual assessment on, such as  depth of striae, number of matching striae, and the number of consecutively matching striae (CMS).
+Features include characteristics of the signature that firearms and toolmark examiners base their visual assessment on, such as \svp{striae depth}, number of matching striae, and the number of consecutively matching striae (CMS).
 We also include statistical features, such as cross-correlation and Euclidean distance between the signatures.
+```{r, eval = F}
+# TODO: Add correlation and euclidian distance pictures?
+```
 
@@
 
-The algorithm in \citet{aoas} is trained on scans from Hamby sets 252 and 173 made available through \citet{nist}. Set 173 was originally published as Hamby set 44. This has been rectified now.
+The algorithm in \citet{aoas} is trained on scans from Hamby sets 252 and 173 made available through \citet{nist}.
+Set 173 was originally published as Hamby set 44\svp{, but was mislabeled.} \svp{Probably need to mention that the NIST scans have a different resolution, which may affect how the results generalize?}
 
@@
 Each set consists of 20 known bullets (two from each of the ten barrels) and 15 questioned bullets of unknown origin.
-Note that all Hamby sets are closed sets, i.e.
-all questioned bullets are fired through one of the ten barrels.
+Note that all Hamby sets are closed sets\svp{; that is,} all questioned bullets are fired through one of the ten barrels.
 The ammunition used for this set were 9 mm Luger 115 Grain Full Metal Jacket  from the Winchester Ammunition Company.
@@
 Ten questioned bullets were provided (B, E, H, J, K N, Q, T, Y, Z).
-This set is an open set, i.e.\ it is not known in advance whether all (or any) of the questioned bullets are fired from the known barrels.
+This set is an open set\svp{; that is,}\ it is not known in advance whether all (or any) of the questioned bullets are fired from the known barrels.
 
@@
 The ammunition used in both test fires and the questioned bullets were Remington UMC 9mm Luger Full Metal Jackets.
-All three of the test sets are open, i.e.\ not every one of the questioned bullets is fired from the five known barrels in each of the test set.
+All three of the test sets are open\svp{; that is,}\ not every one of the questioned bullets is fired from the five known barrels in each of the test set.
 \end{itemize}
 
-Scans of all land engraved areas were taken on a Sensofar Confocal Light microscope at 20x magnification resulting a resolution of 0.645 microns per pixel.
+Scans of all land engraved areas were taken on a Sensofar Confocal Light microscope at 20x magnification resulting \svp{in} a resolution of 0.645 microns per pixel.
 If not indicated otherwise, scans were taken at the Roy J Carver high resolution microscopy lab at Iowa State University.
 
-Case studies were chosen such that difficulty for the matching algorithm increases from one to the other: Hamby set 44 is part of the Hamby study.
+Case studies were chosen such that difficulty for the matching algorithm increases \svp{with each case study}: Hamby set 44 is part of the Hamby study.
 The algorithm in \citet{aoas} was trained on sets 173 and 252, so the bullets in Hamby 44 are of the same type of ammunition and are fired through the same barrels as the bullets in the training set.
 
-For the Phoenix PD set, Ruger XXX barrels were used. Obviously, the barrels in this set are different from the barrels in the training, but the barrels are similarly rifled as the Ruger P-85 barrels.
+\svp{The }Phoenix PD set \svp{uses} Ruger XXX barrels, \svp{which }are different from the barrels in the \svp{Hamby sets used for }training\svp{ the algorithm}, but the barrels are \svp{rifled similarly to} the Ruger P-85 barrels.
 
-For the Houston Sets, Ruger LCP barrels were used.
+\svp{The} Houston sets\svp{ use} Ruger LCP barrels.
 These barrels are first rifled traditionally, i.e.
 similar to the Ruger P-85, but are double fired.
-The second firing happens after the rifling.
+\svp{\{I still don't understand what double fired means in this context\}} The second firing happens after the rifling.
 This second firing has the potential to introduce some subclass characteristics, i.e.
-anomalies that are shared between similarly barrels.
+anomalies that are shared between similarly \svp{rifled?
+manufactured?} barrels.
 This should make the automatic matching, and in particular, the classification as different-source land engraved areas/bullets, harder.

### 2018-10-09 index.Rmd
discussion
Phrasing changes

@@
 
+XXX 3d technology
+
 In the \svp{matching algorithm,} similarity of bullet land engraved areas is evaluated based on 3d topographic scans of land engraved areas (LEAs).
 From each land engraved area a signature is derived by
+
 \begin{itemize}
@@
 the automatic matching  identifies the correct barrel for all questioned bullets.
-For two of the questioned bullets, 'I' and 'F' the similarity scores to the matching barrels are considerably smaller than for the other questioned bullets.
-On the right of \autoref{fig:h44} the relationship between all pairs of questioned bullets is shown.
-Note that questioned bullets are not compared to themselves, leaving white squares on the diagonal.
+For two of the questioned bullets, 'I' and 'F', the similarity scores to the matching barrels are considerably smaller than for the other questioned bullets.
+\svp{The right side of \autoref{fig:h44} shows the relationship between all pairs of questioned bullets.} Note that questioned bullets are not compared to themselves, leaving white squares on the diagonal.
 Some of the questioned bullets match the same barrel, e.g.
 questioned bullets 'P' and 'J' both match barrel 5.
-Therefore bullets 'P' and 'J' are also matching each other in the square on the right hand side.
+Therefore bullets 'P' and 'J' \svp{also match} each other in the square on the right hand side.
 
@@
 \autoref{fig:pd} shows an overview of similarity scores for all pairs of questioned bullets and  test fires.
-The color encoding is the same as in the previous figure.
-What we see here, are matches of one of the questioned bullets to all three bullets of one of the barrels each.
-None of the test fires from barrel U10 shows a match to any of the questioned bullets.
+The color encoding is the same as in the previous figures.
+\svp{Here, we see that some of the questioned bullets match all 3 bullets fired out of the same barrel.} None of the test fires from barrel U10 \svp{match} any of the questioned bullets.
 Similarly, questioned bullets 'Q', 'Y', and 'Z' do not match any of the known barrels.
 This is a sign of the previously mentioned open set characteristic of the study.
-The results from automatic matching, again, correctly pair questioned bullets to their corresponding barrels.
+\svp{Once again, the results from automatic matching correctly pair questioned bullets with their corresponding barrels.}
 
@@
 
-\autoref{fig:hou} shows an overview the three sets of the scores for all pairs of questioned bullets with all other bullets
+\autoref{fig:hou} shows an overview the three sets of the scores for all pairs of questioned bullets with all other bullets \svp{in the set.
+In set 1, five of the questioned bullets can be matched to known bullets in the set.
+None of the questioned bullets in set 1 match bullets fired from barrels KC or KD.
+Additionally, two of the questioned bullets which do not match any of the barrels in the set match each other.
+In set 2, four of the questioned bullets can be matched to known bullets in the set; of the additional four questioned bullets which do not match any known bullets, two were fired from the same barrel.
+In set 3, six of the questioned bullets match known bullets, and the remaining two questioned bullets match each other but do not match any known bullets in the set.
+Thus, we can conclude that the algorithm performs well on open set studies, as it can identify matches among questioned bullets which do not match any known bullets.}
 
@@
 However, SAM scores based on the multivariate random forest score show better separation between same-source and different-source scores than SAM scores based on cross-correlation.
-This can be seen from the larger horizontal distance between the modes of the density curves of RF based SAM scores than cross-correlation SAM scores.
+This can be seen from the larger horizontal distance between the modes\svp{(peaks)} of the density curves of RF based SAM scores \svp{compared to the} cross-correlation SAM scores.
 
-Regarding requirement (R2), \autoref{fig:compare} shows that for neither of the two measures  a single cutoff exists across different studies that separates same-source scores from different-source scores without introducing errors.
+Regarding requirement (R2), \autoref{fig:compare} shows that \svp{a single cutoff value does not exist that separates same-source scores from different-source scores across all three studies, without introducing additional errors}.
 
@@
 The downside of this is that we need to considerably expand the database used for training.
+
 Another solution would be to expand on the aggregation used to get from land-to-land scores to bullet scores: SAM scores only take the scores of the maximum sequence into account and ignore scores associated with pairs of land engraved areas from different sources.
@@
 
-What is quite surprising, that what we anticipated to be the easiest case of Hamby 44 turned out to be the hardest to solve. 
-This is likely due to several lands in the Hamby-44 set having major deficiencies,  such as 'tank rash' (a collision of the bullet after exiting the barrel with a surface causing markings on top of the striations from the barrel) or extreme pitting (holes caused by direct contact with burning gun powder).
+What is quite surprising \svp{is} that what we anticipated to be the easiest case of Hamby 44 turned out to be the hardest to solve.
+This is likely due to several lands in the Hamby-44 set \svp{that have} major deficiencies,  such as 'tank rash' (a collision of the bullet after exiting the barrel with a surface causing markings on top of the striations from the barrel) or extreme pitting (holes caused by direct contact with burning gun powder).
 \autoref{fig:tankrash} gives an overview of all lands of bullets in the Hamby-44 set affected.
+\svp{These issues affect the algorithm's ability to identify a stable signature, which impacts the subsequent match scores.}
 ```{r tankrash, fig.cap="Overview of bullet lands with prominent deficiency such as tank rash or extreme pitting in the Hamby-44 study.", fig.width=4, fig.height = 2, out.width='47.5%', fig.subcap=c("Br1-B1-L6", "Br2-B2-L5", "Br3-B1-L5", "Br8-B1-L6", "Br8-B2-L2", "Br8-B2-L6", "Unk-E-L6", "Unk-I-L6"), fig.align="middle", fig.ncol=2}
@@
 
-Three lands from bullets known to be fired from barrel 8 are affected.
-This is the likely cause for the low scores for barrel 8 in \autoref{fig:h44}.
-We also see that questioned bullet 'E' is affected, explaining some of the low similarity scores for this bullet.
+Three lands from bullets known to be fired from barrel 8 are affected\svp{, producing} low scores for barrel 8 in \autoref{fig:h44}.
+We also see that questioned bullet 'E' is affected, explaining some of the low similarity scores for this bullet.\svp{You've marked I vs.
+5 and F vs.
+8 in \autoref{fig:compare}; should E be marked?}
 

### 2018-10-09 index.Rmd
lit review

@@
 
-XXX 3d technology
+## 3D Imaging Technology
+\svp{3D characterization of bullet surfaces has been considered since 1958 \citeauthor{davis1968introduction}, but the technology was not sufficiently advanced for analysis based on 3D measurements to be a viable option.
+Approaches to utilizing 3D measurment data were proposed starting in 1999 \citeauthor{dekinderAutomatedComparisonsBullet1999}, and have been further developed by \citet{bachrachDevelopment3DbasedAutomated2002,xieAutomatedBulletidentificationSystem2009, chuPilotStudyAutomated2010}.
+These approaches utilize 3D surface measurements directly to characterize the topology of bullet surfaces, rather than using visual or image comparison techniques.
+In many cases, they also provide some level of automation of the comparison process, with the goal of reducing human biases by augmenting the visual information with 3D measurements.
+This allows for examination not only of the peaks in the bullet signature, but also for comparison of the valleys, which are more difficult to assess visually except in relation to the peaks.
+}
 
-In the \svp{matching algorithm,} similarity of bullet land engraved areas is evaluated based on 3d topographic scans of land engraved areas (LEAs).
+\svp{Commonly, approaches derived from 3D surface measurements use some features which are developed to mimic the perception of FTEs\citep{luAutomatedBulletIdentification2014}.
+Class characteristics, which are shared by a group of firearms with the same rifling design, manufacturer, and tooling process, are typically evaluated first, as a mismatch on class charcteristics is sufficient to conclude that two samples do not match.
+Automated approaches to estimation of width of the land engraved area and twist angle were described in \citet{chuPilotStudyAutomated2010}.
+Individual characteristics, which are not shared by all members of a class, can also be automatically assessed from 3D surface measurements; in some cases, these features mimic features used by FTEs during image analysis.
+One of the most common features used to assess the strength of a match based on individual characteristics is the cross-correlation function (ccf), which is described in several studies \citep{maNISTBulletSignature2004, vorburgerApplicationsCrosscorrelationFunctions2011, chuPilotStudyAutomated2010}.
+Additional features proposed for assessing match strength also include Signature Distance \citep{maNISTBulletSignature2004} and automatically determined Consecutive Matching Striae \citep{chuAutomaticIdentificationBullet2013}.}
+
+### Need to transition to explanation of the match algorithm...
+In the \svp{matching algorithm proposed in \citet{aoas},} similarity of bullet land engraved areas is evaluated based on 3d topographic scans of land engraved areas (LEAs).
 From each land engraved area a signature is derived by
@@
 \svp{These issues affect the algorithm's ability to identify a stable signature, which impacts the subsequent match scores.}
-```{r tankrash, fig.cap="Overview of bullet lands with prominent deficiency such as tank rash or extreme pitting in the Hamby-44 study.", fig.width=4, fig.height = 2, out.width='47.5%', fig.subcap=c("Br1-B1-L6", "Br2-B2-L5", "Br3-B1-L5", "Br8-B1-L6", "Br8-B2-L2", "Br8-B2-L6", "Unk-E-L6", "Unk-I-L6"), fig.align="middle", fig.ncol=2}
+```{r tankrash, fig.cap="Overview of bullet lands with prominent deficiency such as tank rash or extreme pitting in the Hamby-44 study.", fig.width=4, fig.height = 2, out.width='47.5%', fig.subcap=c("Br1-B1-L6 (breakaway)", "Br2-B2-L5 (tank rash)", "Br3-B1-L5 (tank rash)", "Br8-B1-L6 (tank rash)", "Br8-B2-L2 (pitting)", "Br8-B2-L6 (breakaway)", "Unk-E-L6 (breakaway)", "Unk-I-L6 (breakaway)"), fig.align="middle", fig.ncol=2}
 knitr::include_graphics(

### 2018-10-10 index.Rmd
Transition, additional citations

@@
 One of the most common features used to assess the strength of a match based on individual characteristics is the cross-correlation function (ccf), which is described in several studies \citep{maNISTBulletSignature2004, vorburgerApplicationsCrosscorrelationFunctions2011, chuPilotStudyAutomated2010}.
-Additional features proposed for assessing match strength also include Signature Distance \citep{maNISTBulletSignature2004} and automatically determined Consecutive Matching Striae \citep{chuAutomaticIdentificationBullet2013}.}
+Additional features proposed for automatically assessing match strength also include signature distance \citep{maNISTBulletSignature2004}, striae depth and width \citep{ComputerIdentificationBullets1978}, and consecutive matching striae \citep{chuAutomaticIdentificationBullet2013}.}
 
-### Need to transition to explanation of the match algorithm...
-In the \svp{matching algorithm proposed in \citet{aoas},} similarity of bullet land engraved areas is evaluated based on 3d topographic scans of land engraved areas (LEAs).
+In the \svp{matching algorithm proposed in \citet{aoas}, many of the features described above are combined to produce a similarity score} based on 3d topographic scans of land engraved areas (LEAs).
 From each land engraved area a signature is derived by
@@
 We also include statistical features, such as cross-correlation and Euclidean distance between the signatures.
-```{r, eval = F}
-# TODO: Add correlation and euclidian distance pictures?
-```
 
@@
 The algorithm in \citet{aoas} is trained on scans from Hamby sets 252 and 173 made available through \citet{nist}.
-Set 173 was originally published as Hamby set 44\svp{, but was mislabeled.} \svp{Probably need to mention that the NIST scans have a different resolution, which may affect how the results generalize?}
+Set 173 was originally published as Hamby set 44\svp{, but was mislabeled.} \svp{The scans used to fit the model were taken at a resolution of \svp{Probably need to mention that the NIST scans have a different resolution, which may affect how the results generalize?}
 

### 2018-10-10 index.Rmd
added references

@@
 
-There are several approaches for potential improvements: we can try to fine tune the matching algorithm to take make and model of the firearm and ammunition used into account or expand the traing of the algorithm to include a wider variety of makes and models.
+There are several approaches for potential improvements: we can try to fine tune the matching algorithm to take make and model of the firearm and ammunition used into account or expand the trainings base of the algorithm to include a wider variety of makes and models.
 The downside of this is that we need to considerably expand the database used for training.
@@
 Those scores might be useful in establishing a baseline that better expresses similarity between pairs of bullets.
+One approach that uses scores for both same source pairs and different source pairs are score-based likelihood ratios \citep{Bunch:2013if, Morrison:2018fh}.
+\citet{Morrison:2018km}  stress the need to additionally include the typicality of a signature.
 

### 2018-10-11 index.Rmd
addressing first couple of edits

@@
 Exact guidelines for this classification vary from lab to lab, e.g.\ some labs will exclude only on the basis of non-matching class characteristics (such as direction of the twist in rifling, land length or number of lands, or type of rifling).
-In some labs,  CMS (consecutively matching striae) as defined by \citet{biasotti} \svp{are} used as a way to quantify the strength of a match.
+In some labs,  CMS (consecutively matching striae) as defined by \citet{biasotti} is used as a measure to quantify the strength of a match.
 
 ## 3D Imaging Technology
-\svp{3D characterization of bullet surfaces has been considered since 1958 \citeauthor{davis1968introduction}, but the technology was not sufficiently advanced for analysis based on 3D measurements to be a viable option.
-Approaches to utilizing 3D measurment data were proposed starting in 1999 \citeauthor{dekinderAutomatedComparisonsBullet1999}, and have been further developed by \citet{bachrachDevelopment3DbasedAutomated2002,xieAutomatedBulletidentificationSystem2009, chuPilotStudyAutomated2010}.
+3D characterization of bullet surfaces has been considered since 1958 \citeauthor{davis1968introduction}, but the technology was not sufficiently advanced for analysis based on 3D measurements to be a viable option.
+Approaches to utilizing 3D measurement data were proposed starting in 1999 \citeauthor{dekinderAutomatedComparisonsBullet1999}, and have been further developed by \citet{bachrachDevelopment3DbasedAutomated2002,xieAutomatedBulletidentificationSystem2009, chuPilotStudyAutomated2010}.
 These approaches utilize 3D surface measurements directly to characterize the topology of bullet surfaces, rather than using visual or image comparison techniques.
@@
 This allows for examination not only of the peaks in the bullet signature, but also for comparison of the valleys, which are more difficult to assess visually except in relation to the peaks.
-}
 
-\svp{Commonly, approaches derived from 3D surface measurements use some features which are developed to mimic the perception of FTEs\citep{luAutomatedBulletIdentification2014}.
-Class characteristics, which are shared by a group of firearms with the same rifling design, manufacturer, and tooling process, are typically evaluated first, as a mismatch on class charcteristics is sufficient to conclude that two samples do not match.
+Commonly, approaches derived from 3D surface measurements use some features which are developed to mimic the perception of FTEs\citep{luAutomatedBulletIdentification2014}.
+Class characteristics, which are shared by a group of firearms with the same rifling design, manufacturer, and tooling process, are typically evaluated first, as a mismatch on class charcteristics is sufficient for an exclusion.
 Automated approaches to estimation of width of the land engraved area and twist angle were described in \citet{chuPilotStudyAutomated2010}.
@@
 One of the most common features used to assess the strength of a match based on individual characteristics is the cross-correlation function (ccf), which is described in several studies \citep{maNISTBulletSignature2004, vorburgerApplicationsCrosscorrelationFunctions2011, chuPilotStudyAutomated2010}.
-Additional features proposed for automatically assessing match strength also include signature distance \citep{maNISTBulletSignature2004}, striae depth and width \citep{ComputerIdentificationBullets1978}, and consecutive matching striae \citep{chuAutomaticIdentificationBullet2013}.}
+Additional features proposed for automatically assessing match strength also include signature distance \citep{maNISTBulletSignature2004}, striae depth and width \citep{ComputerIdentificationBullets1978}, and consecutive matching striae \citep{chuAutomaticIdentificationBullet2013}.
 
-In the \svp{matching algorithm proposed in \citet{aoas}, many of the features described above are combined to produce a similarity score} based on 3d topographic scans of land engraved areas (LEAs).
+In the matching algorithm proposed in \citet{aoas}, several of the features described above are combined to produce a similarity score based on 3d topographic scans of land engraved areas (LEAs).
 From each land engraved area a signature is derived by
@@
 Quantitative features describing the similarity of two signature are extracted.
-Features include characteristics of the signature that firearms and toolmark examiners base their visual assessment on, such as \svp{striae depth}, number of matching striae, and the number of consecutively matching striae (CMS).
+Features include characteristics of the signature that firearms and toolmark examiners base their visual assessment on, such as striae depth, number of matching striae, and the number of consecutively matching striae (CMS).
 We also include statistical features, such as cross-correlation and Euclidean distance between the signatures.
@@
 The algorithm in \citet{aoas} is trained on scans from Hamby sets 252 and 173 made available through \citet{nist}.
-Set 173 was originally published as Hamby set 44\svp{, but was mislabeled.} \svp{The scans used to fit the model were taken at a resolution of \svp{Probably need to mention that the NIST scans have a different resolution, which may affect how the results generalize?}
+Set 173 was originally published as Hamby set 44.
+This mislabelling has been correct now.
+The scans used to train the model were taken at 20 fold magnification for a resolution of 1.5625 microns per pixel.
+\hh{While magnification is generally of interest in microscopy, {\it resolution} -- usually measured in microns per pixel -- is of more interest for 3D topographic measurements, as it determines the operative level of available data.}
+
+\hh{XXX We need a gentler intro to the validation sets - maybe some discussion of training and test sets?}
 
@@
 Each set consists of 20 known bullets (two from each of the ten barrels) and 15 questioned bullets of unknown origin.
-Note that all Hamby sets are closed sets\svp{; that is,} all questioned bullets are fired through one of the ten barrels.
+Note that all Hamby sets are closed sets; that is, all questioned bullets are fired through one of the ten barrels.
 The ammunition used for this set were 9 mm Luger 115 Grain Full Metal Jacket  from the Winchester Ammunition Company.
@@
 Ten questioned bullets were provided (B, E, H, J, K N, Q, T, Y, Z).
-This set is an open set\svp{; that is,}\ it is not known in advance whether all (or any) of the questioned bullets are fired from the known barrels.
+This set is an open set; that is,\ it is not known in advance whether all (or any) of the questioned bullets are fired from the known barrels.
 
@@
 The ammunition used in both test fires and the questioned bullets were Remington UMC 9mm Luger Full Metal Jackets.
-All three of the test sets are open\svp{; that is,}\ not every one of the questioned bullets is fired from the five known barrels in each of the test set.
+All three of the test sets are open; that is,\ not every one of the questioned bullets is fired from the five known barrels in each of the test set.
 \end{itemize}
 
-Scans of all land engraved areas were taken on a Sensofar Confocal Light microscope at 20x magnification resulting \svp{in} a resolution of 0.645 microns per pixel.
+Scans of all land engraved areas were taken on a Sensofar Confocal Light microscope at 20x magnification resulting in a resolution of 0.645 microns per pixel.
 If not indicated otherwise, scans were taken at the Roy J Carver high resolution microscopy lab at Iowa State University.
 
-Case studies were chosen such that difficulty for the matching algorithm increases \svp{with each case study}: Hamby set 44 is part of the Hamby study.
+Case studies were chosen such that difficulty for the matching algorithm increases with each case study: Hamby set 44 is part of the Hamby study.
 The algorithm in \citet{aoas} was trained on sets 173 and 252, so the bullets in Hamby 44 are of the same type of ammunition and are fired through the same barrels as the bullets in the training set.
 
-\svp{The }Phoenix PD set \svp{uses} Ruger XXX barrels, \svp{which }are different from the barrels in the \svp{Hamby sets used for }training\svp{ the algorithm}, but the barrels are \svp{rifled similarly to} the Ruger P-85 barrels.
+The Phoenix PD set uses Ruger XXX barrels, which are different from the barrels in the Hamby sets used for training the algorithm, but the barrels are rifled similarly to the Ruger P-85 barrels.
 
@@
 These barrels are first rifled traditionally, i.e.
-similar to the Ruger P-85, but are double fired.
-\svp{\{I still don't understand what double fired means in this context\}} The second firing happens after the rifling.
-This second firing has the potential to introduce some subclass characteristics, i.e.
-anomalies that are shared between similarly \svp{rifled?
-manufactured?} barrels.
-This should make the automatic matching, and in particular, the classification as different-source land engraved areas/bullets, harder.
+similar to the Ruger P-85, but are treated to a secondary round of heating after barreling.
+This second round of heat exposure has the potential to introduce some subclass characteristics, i.e.
+anomalies that are shared between different barrels manufactured in this manner.
+This should make the automatic matching harder, and in particular, the classification as different-source land engraved areas/bullets.
 

### 2018-10-11 index.Rmd
journaltitle is not recognized by bib style

@@
 \svp{These issues affect the algorithm's ability to identify a stable signature, which impacts the subsequent match scores.}
-```{r tankrash, fig.cap="Overview of bullet lands with prominent deficiency such as tank rash or extreme pitting in the Hamby-44 study.", fig.width=4, fig.height = 2, out.width='47.5%', fig.subcap=c("Br1-B1-L6 (breakaway)", "Br2-B2-L5 (tank rash)", "Br3-B1-L5 (tank rash)", "Br8-B1-L6 (tank rash)", "Br8-B2-L2 (pitting)", "Br8-B2-L6 (breakaway)", "Unk-E-L6 (breakaway)", "Unk-I-L6 (breakaway)"), fig.align="middle", fig.ncol=2}
+```{r tankrash, fig.cap="Overview of bullet lands with prominent deficiency such as tank rash or extreme pitting in the Hamby-44 study.", fig.width=4, fig.height = 2, out.width='47.5%', fig.subcap=c("Br1-B1-L6 (tank rask)", "Br2-B2-L5 (tank rash)", "Br3-B1-L5 (tank rash)", "Br8-B1-L6 (tank rash)", "Br8-B2-L2 (pitting)", "Br8-B2-L6 (tank rash)", "Unk-E-L6 (tank rash)", "Unk-I-L6 (tank rash and break-off)"), fig.align="middle", fig.ncol=2}
 knitr::include_graphics(

### 2018-10-11 index.Rmd
Slight tweaks and bib fixes

@@
 ## 3D Imaging Technology
-3D characterization of bullet surfaces has been considered since 1958 \citeauthor{davis1968introduction}, but the technology was not sufficiently advanced for analysis based on 3D measurements to be a viable option.
-Approaches to utilizing 3D measurement data were proposed starting in 1999 \citeauthor{dekinderAutomatedComparisonsBullet1999}, and have been further developed by \citet{bachrachDevelopment3DbasedAutomated2002,xieAutomatedBulletidentificationSystem2009, chuPilotStudyAutomated2010}.
+3D characterization of bullet surfaces has been considered since 1958 (\citeauthor{davis1968introduction}), but the technology was not sufficiently advanced for analysis based on 3D measurements to be a viable option.
+Approaches to utilizing 3D measurement data were proposed starting in 1999 (\citeauthor{dekinderAutomatedComparisonsBullet1999}), and have been further developed by \citet{bachrachDevelopment3DbasedAutomated2002,xieAutomatedBulletidentificationSystem2009, chuPilotStudyAutomated2010}.
 These approaches utilize 3D surface measurements directly to characterize the topology of bullet surfaces, rather than using visual or image comparison techniques.
@@
 
-```{r signatures, fig.width = 10, fig.height = 4, fig.cap="Signatures of all six lands of bullets 1 and 2 from barrel 1 of the Hamby set 44."}
+```{r signatures, fig.width = 10, fig.height = 4, fig.cap="Signatures of all six lands of bullets 1 and 2 from barrel 1 of the Hamby set 44.
+\\svp{The raw signature is shown in light gray and the refined signature in dark gray.}"}
 # signatures for one bullet

### 2018-10-12 index.Rmd
fixing references and removing duplicates
incorporated other edits

@@
 
-```{r signatures, fig.width = 10, fig.height = 4, fig.cap="Signatures of all six lands of bullets 1 and 2 from barrel 1 of the Hamby set 44.
-\\svp{The raw signature is shown in light gray and the refined signature in dark gray.}"}
+```{r signatures, fig.cap="Signatures of all six lands of bullets 1 and 2 from barrel 1 of the Hamby set 44.", fig.height=4, fig.width=10}
 # signatures for one bullet
@@
   ggplot(aes(x = x / 1000)) +
-  geom_line(aes(y = raw_sig), colour = "grey70") +
+#  geom_line(aes(y = raw_sig), colour = "grey70") +
   geom_line(aes(y = sig), colour = "grey30") +
@@
 
-\svp{The} Houston sets\svp{ use} Ruger LCP barrels.
+The Houston sets  use Ruger LCP barrels.
 These barrels are first rifled traditionally, i.e.
@@
 Note that the sequence average maximum of the correlation between lands is used in SensoComp to capture the similarity between bullets.
-The correlation based SAM score has also been called the 'average correlation calculated at the max phase' in \citet{chu}.
+The correlation based SAM score has also been called the 'average correlation calculated at the max phase' in \citet{chuPilotStudyAutomated2010}.
 
@@
 For two of the questioned bullets, 'I' and 'F', the similarity scores to the matching barrels are considerably smaller than for the other questioned bullets.
-\svp{The right side of \autoref{fig:h44} shows the relationship between all pairs of questioned bullets.} Note that questioned bullets are not compared to themselves, leaving white squares on the diagonal.
-Some of the questioned bullets match the same barrel, e.g.
-questioned bullets 'P' and 'J' both match barrel 5.
-Therefore bullets 'P' and 'J' \svp{also match} each other in the square on the right hand side.
+The right side of \autoref{fig:h44} shows the relationship between all pairs of questioned bullets.
+Note that questioned bullets are not compared to themselves, leaving white squares on the diagonal.
+Some of the questioned bullets match the same barrel, e.g.\ questioned bullets 'P' and 'J' both match barrel 5.
+Therefore bullets 'P' and 'J' also match each other in the square on the right hand side.
 
@@
 The color encoding is the same as in the previous figures.
-\svp{Here, we see that some of the questioned bullets match all 3 bullets fired out of the same barrel.} None of the test fires from barrel U10 \svp{match} any of the questioned bullets.
-Similarly, questioned bullets 'Q', 'Y', and 'Z' do not match any of the known barrels.
-This is a sign of the previously mentioned open set characteristic of the study.
-\svp{Once again, the results from automatic matching correctly pair questioned bullets with their corresponding barrels.}
+Here, we see that questioned bullets either match all 3 bullets fired out of the same barrel for exactly one of the barrels or none of the known barrels.
+Questioned bullets 'Q', 'Y', and 'Z', do not match any of the known barrels.
+None of the test fires from barrel U10 match any of the questioned bullets.  This is a sign of the previously mentioned open set characteristic of the study. 
+Once again, the results from automatic matching correctly pair questioned bullets with their corresponding barrels.
 
@@
 
-\autoref{fig:hou} shows an overview the three sets of the scores for all pairs of questioned bullets with all other bullets \svp{in the set.
+\autoref{fig:hou} shows an overview the three sets of the scores for all pairs of questioned bullets with all other bullets in the set.
 In set 1, five of the questioned bullets can be matched to known bullets in the set.
@@
 In set 3, six of the questioned bullets match known bullets, and the remaining two questioned bullets match each other but do not match any known bullets in the set.
-Thus, we can conclude that the algorithm performs well on open set studies, as it can identify matches among questioned bullets which do not match any known bullets.}
+Thus, we can conclude that the algorithm performs well on open set studies, as it can identify matches among questioned bullets which do not match any known bullets.
 
@@
 However, SAM scores based on the multivariate random forest score show better separation between same-source and different-source scores than SAM scores based on cross-correlation.
-This can be seen from the larger horizontal distance between the modes\svp{(peaks)} of the density curves of RF based SAM scores \svp{compared to the} cross-correlation SAM scores.
+This can be seen from the larger horizontal distance between the modes (peaks) of the density curves of RF based SAM scores compared to the cross-correlation SAM scores.
 
-Regarding requirement (R2), \autoref{fig:compare} shows that \svp{a single cutoff value does not exist that separates same-source scores from different-source scores across all three studies, without introducing additional errors}.
+Regarding requirement (R2), \autoref{fig:compare} shows that a single cutoff value does not exist that separates same-source scores from different-source scores across all three studies, without introducing additional errors.
 
@@
 
-What is quite surprising \svp{is} that what we anticipated to be the easiest case of Hamby 44 turned out to be the hardest to solve.
-This is likely due to several lands in the Hamby-44 set \svp{that have} major deficiencies,  such as 'tank rash' (a collision of the bullet after exiting the barrel with a surface causing markings on top of the striations from the barrel) or extreme pitting (holes caused by direct contact with burning gun powder).
+What is quite surprising is that what we anticipated to be the easiest case of Hamby 44 turned out to be the hardest to solve.
+This is likely due to several lands in the Hamby-44 set that have major deficiencies,  such as 'tank rash' (a collision of the bullet after exiting the barrel with a surface causing markings on top of the striations from the barrel) or extreme pitting (holes caused by direct contact with burning gun powder).
 \autoref{fig:tankrash} gives an overview of all lands of bullets in the Hamby-44 set affected.
-\svp{These issues affect the algorithm's ability to identify a stable signature, which impacts the subsequent match scores.}
+These issues affect the algorithm's ability to identify a stable signature, which impacts the subsequent match scores.
 ```{r tankrash, fig.cap="Overview of bullet lands with prominent deficiency such as tank rash or extreme pitting in the Hamby-44 study.", fig.width=4, fig.height = 2, out.width='47.5%', fig.subcap=c("Br1-B1-L6 (tank rask)", "Br2-B2-L5 (tank rash)", "Br3-B1-L5 (tank rash)", "Br8-B1-L6 (tank rash)", "Br8-B2-L2 (pitting)", "Br8-B2-L6 (tank rash)", "Unk-E-L6 (tank rash)", "Unk-I-L6 (tank rash and break-off)"), fig.align="middle", fig.ncol=2}
@@
 
-Three lands from bullets known to be fired from barrel 8 are affected\svp{, producing} low scores for barrel 8 in \autoref{fig:h44}.
-We also see that questioned bullet 'E' is affected, explaining some of the low similarity scores for this bullet.\svp{You've marked I vs.
-5 and F vs.
-8 in \autoref{fig:compare}; should E be marked?}
+Three lands from bullets known to be fired from barrel 8 are affected, producing low scores for barrel 8 in \autoref{fig:h44}.
+We also see that questioned bullet 'I' is affected, explaining some of the low similarity scores for this bullet.
 

### 2018-10-15 index.Rmd
clarification

@@
 
-Regarding requirement (R2), \autoref{fig:compare} shows that a single cutoff value does not exist that separates same-source scores from different-source scores across all three studies, without introducing additional errors.
+Regarding requirement (R2), \autoref{fig:compare} shows that a single cutoff value does not exist that separates same-source scores from different-source scores across all three studies, without introducing errors.
 
@@
 ### Limitations
-All of the studies involve the same firearm.
+All of the studies involve the same type of firearm.
 We see some variations in scores across different models of Rugers.

### 2018-10-16 index.Rmd
messing about with validation cases

@@
 This mislabelling has been correct now.
-The scans used to train the model were taken at 20 fold magnification for a resolution of 1.5625 microns per pixel.
-\hh{While magnification is generally of interest in microscopy, {\it resolution} -- usually measured in microns per pixel -- is of more interest for 3D topographic measurements, as it determines the operative level of available data.}
 
@@
 
-Scans of all land engraved areas were taken on a Sensofar Confocal Light microscope at 20x magnification resulting in a resolution of 0.645 microns per pixel.
+The scans used to train the model were taken at 20 fold magnification for a resolution of 1.5625 microns per pixel.
+While magnification is generally of interest in microscopy, {\it resolution} -- usually measured in microns per pixel -- is of more interest for 3D topographic measurements, as it determines the operative level of available data.
+
+Scans of all land engraved areas for the validation data were taken on a Sensofar Confocal Light microscope at 20x magnification resulting in a resolution of 0.645 microns per pixel.
 If not indicated otherwise, scans were taken at the Roy J Carver high resolution microscopy lab at Iowa State University.

### 2018-10-23 index.Rmd
adding land-level results

@@
 
+\hh{The automatic matching algorithm is based on a random forest [@breiman], which has a built-in internal testing mechanism to prevent potential overfitting.
+Error rates reported by random forest algorithms are based on these internal test sets.
+However, there is some discussion on whether or in which situations these error rates are biased [@rfOver2; @rfOver].
+Neither of these papers adresses another issue with internal test sets: internal test sets are constructed to have the same distribution as the training data (apart from sampling variability).
+For any machine learning method the true test comes with testing the performance of an algorithm on an external test data.
+Good performance (in terms of low error rates) on external test data validates a ML algorithm.
+External data allows an assessment of the algorithm's sensitivity to distributional changes.
+Using performance on external test data also allows us to test limitations of the algorithm by going outside the parameters of the training data.
+}
+
 The algorithm in \citet{aoas} is trained on scans from Hamby sets 252 and 173 made available through \citet{nist}.
 Set 173 was originally published as Hamby set 44.
-This mislabelling has been correct now.
+This mislabelling has been corrected now.
 
-\hh{XXX We need a gentler intro to the validation sets - maybe some discussion of training and test sets?}
-
-Here, we are considering three validation sets: 
+\hh{For the validation of the automatic matching algorithm we are considering three validation sets of -- what should be -- increasing difficulty level:} 
 \begin{itemize}
@@
 All three of the test sets are open; that is,\ not every one of the questioned bullets is fired from the five known barrels in each of the test set.
+The structure of these three sets is similar to a forthcoming study by Nally and Kirksey, but the results here come from preliminary test sets made available to us.
 \end{itemize}
@@
 The scans used to train the model were taken at 20 fold magnification for a resolution of 1.5625 microns per pixel.
-While magnification is generally of interest in microscopy, {\it resolution} -- usually measured in microns per pixel -- is of more interest for 3D topographic measurements, as it determines the operative level of available data.
+While magnification is generally of interest in microscopy, **resolution** -- usually measured in microns per pixel -- is of more interest for 3D topographic measurements, as it determines the operative level of available data.
 
@@
 
-The Phoenix PD set uses Ruger XXX barrels, which are different from the barrels in the Hamby sets used for training the algorithm, but the barrels are rifled similarly to the Ruger P-85 barrels.
+The Phoenix PD set uses Ruger P-95 barrels, which are different from the barrels in the Hamby sets used for training the algorithm, but the barrels are rifled similarly to the Ruger P-85 barrels.
 
@@
 
-\autoref{fig:hou} shows an overview the three sets of the scores for all pairs of questioned bullets with all other bullets in the set.
+\autoref{fig:hou} shows an overview of the three sets of scores for all pairs of questioned bullets with all other bullets in the set.
 In set 1, five of the questioned bullets can be matched to known bullets in the set.
@@
 In set 3, six of the questioned bullets match known bullets, and the remaining two questioned bullets match each other but do not match any known bullets in the set.
-Thus, we can conclude that the algorithm performs well on open set studies, as it can identify matches among questioned bullets which do not match any known bullets.
-
+%Thus, we can conclude that the algorithm performs well on open set studies, as it can identify matches among questioned bullets which do not match any known bullets. 
+While all bullet-to-bullet matches are correctly identified, and no known matches are missed, it is obvious from the generally lighter shades of the tiles corresponding to matching bullets in \autoref{fig:hou} for set 2 that  the algorithm is not performing as well as for the other sets (see also \autoref{tab:aucs}).
 ```{r hou, fig.width=9, fig.height = 3, fig.cap="Overview of matching scores for all pairs of questioned bullets to known bullets from five barrels (on the left) and questioned bullets to themselves (tiles on the right).", fig.subcap=c("Set 1", "Set 2", "Set 3"), fig.align="middle", fig.ncol=1}
@@
 
-groups <- features %>%
+features <- features %>%
   filter(group1 == group2) %>%
   filter(barrel1 == "Unknowns") %>%
-  filter(barrel1 != barrel2 | bullet1 != bullet2) %>%
+  filter(barrel1 != barrel2 | bullet1 != bullet2)
+
+groups <- features %>%
   group_by(group1, barrel1, bullet1, barrel2, bullet2) %>%
@@
     select(sam_rf, sam_ccf, samesource) %>%
-    mutate(study = "Hamby-44", group1 = NA),
+    mutate(study = "Hamby 44", group1 = NA),
   pdnest %>%
@@
 )
-nests$study <- factor(nests$study, levels = c("Hamby-44", "Phoenix PD", "Houston FSI"))
+nests$study <- factor(nests$study, levels = c("Hamby 44", "Phoenix PD", "Houston FSI"))
 nestslong <- nests %>% gather(key = "measure", value = "value", sam_rf, sam_ccf)
@@
 labels$measure <- "RF score"
-labels$study <- "Hamby-44"
+labels$study <- "Hamby 44"
 labels$y <- 5 + 4:1
@@
 
+### Comparison of land-to-land scores
+
+\hh{XXX error rates on land to land matches? XXX}
+
+```{r, cache=TRUE}
+hounest <- hounest %>% mutate(
+  data = data %>% purrr::map(
+    .f = function(d) {
+      d$sameland = bullet_to_land_predict(
+        land1 = d$land1, land2 = d$land2, 
+        d$rfscore, difference=0.1)
+      d
+    })
+)
+houlands <- hounest %>% unnest(data)
+
+h44nest <- h44nest %>% mutate(
+  data = data %>% purrr::map(
+    .f = function(d) {
+      d$sameland = bullet_to_land_predict(
+        land1 = d$land1, land2 = d$land2, 
+        d$rfscore, difference=0.1)
+      d
+    })
+)
+h44lands <- h44nest %>% unnest(data)
+
+pdnest <- pdnest %>% mutate(
+  data = data %>% purrr::map(
+    .f = function(d) {
+      d$sameland = bullet_to_land_predict(
+        land1 = d$land1, land2 = d$land2, 
+        d$rfscore, difference=0.1)
+      d
+    })
+)
+pdlands <- pdnest %>% unnest(data)
+
+all_lands <- rbind(
+  h44lands %>%
+    select(rfscore, ccf, sameland, samesource) %>%
+    mutate(study = "Hamby 44", group1 = ""),
+  pdlands %>%
+    select(rfscore, ccf, sameland, samesource) %>%
+    mutate(study = "Phoenix PD", group1 = ""),
+  houlands %>%
+    select(rfscore, ccf, sameland, samesource, group1) %>%
+    mutate(study = "Houston FSI")
+)
+all_lands$study <- factor(all_lands$study, levels = c("Hamby 44", "Phoenix PD", "Houston FSI"))
+all_lands <- all_lands %>%
+  mutate(
+    sameland = replace(sameland, !samesource, FALSE)
+  )
+```
+
+```{r rocs, fig.caption="ROC curves for all studies based on land-to-land comparisons."}
+library(plotROC)
+library(pROC)
+all_lands %>% 
+  ggplot(aes(colour = interaction(study, group1))) +
+  geom_roc(aes(m = rfscore, d = as.numeric(sameland)), 
+           labels =FALSE, pointsize = 0) 
+```
+
+```{r aucs, results="asis"}
+aucs <- all_lands %>% group_by(study, group1) %>% nest() %>%
+  mutate(
+    auc_rf = data %>% purrr::map_dbl(function(d) {
+      auc(d$sameland, d$rfscore)
+    }),
+    auc_ccf = data %>% purrr::map_dbl(function(d) {
+      auc(d$sameland, d$ccf)
+    })
+
+  )
+#
+library(xtable)
+names(aucs)[2] <- "set"
+names(aucs)[4] <- "RF score"
+names(aucs)[5] <- "Cross-correlation"
+print(xtable(aucs %>% select(-data), digits = 3, caption = "Area under the curve (AUC) for each (set) of the studies.
+At the land level there is no difference in prediction power between RF score and cross-correlation."), include.rownames=FALSE, label="tab:aucs")
+```
+
 # Discussion and Conclusions
@@
 What is quite surprising is that what we anticipated to be the easiest case of Hamby 44 turned out to be the hardest to solve.
-This is likely due to several lands in the Hamby-44 set that have major deficiencies,  such as 'tank rash' (a collision of the bullet after exiting the barrel with a surface causing markings on top of the striations from the barrel) or extreme pitting (holes caused by direct contact with burning gun powder).
+This is likely due to several lands in Hamby set 44 that have major deficiencies,  such as 'tank rash' (a collision of the bullet after exiting the barrel with a surface causing markings on top of the striations from the barrel) or extreme pitting (holes caused by direct contact with burning gun powder).
 \autoref{fig:tankrash} gives an overview of all lands of bullets in the Hamby-44 set affected.

### 2018-10-23 index.Rmd
land to land scores

@@
 
-### Comparison of land-to-land scores
+# Digging deeper: comparison of land-to-land scores
 
-\hh{XXX error rates on land to land matches? XXX}
+\hh{XXX error rates on land to land matches? ROC curves in \autoref{fig:rocs} and corresponding AUC results for all studies in \autoref{tab:rocs} XXX}
 
@@
         land1 = d$land1, land2 = d$land2, 
-        d$rfscore, difference=0.1)
+        d$rfscore, difference=0.05)
       d
@@
 
-```{r rocs, fig.caption="ROC curves for all studies based on land-to-land comparisons."}
+```{r rocs, fig.caption="ROC curves for all studies based on land-to-land comparisons.", fig.width = 4.5, fig.height = 3, out.width = ".5\\textwidth"}
 library(plotROC)
@@
 all_lands %>% 
-  ggplot(aes(colour = interaction(study, group1))) +
+  ggplot(aes(colour = interaction(study, group1, sep = " "))) +
   geom_roc(aes(m = rfscore, d = as.numeric(sameland)), 
-           labels =FALSE, pointsize = 0) 
+           labels =FALSE, pointsize = 0, alpha = 0.8) +
+  theme_bw() +
+  xlab("Fraction False Positives") +
+  ylab("Fraction True Positives") +
+  scale_colour_manual("Study", values = c("#fe9929", "#8856a7",
+                                         "#74a9cf", "#2b8cbe", "#045a8d"))
 ```
 
-```{r aucs, results="asis"}
+```{r aucs, eval = FALSE}
 aucs <- all_lands %>% group_by(study, group1) %>% nest() %>%
@@
 ```
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{llrr}
+  \hline
+\bf study & \bf set & \bf AUC(RF score) & \bf AUC(Cross-correlation) \\ 
+  \hline
+\bf Hamby 44 &  & 0.956 & 0.926 \\ 
+\bf   Phoenix PD &  & 0.976 & 0.986 \\ 
+\bf   Houston FSI & G1 & 0.991 & 0.988 \\ 
+\bf   Houston FSI & G2 & 0.898 & 0.902 \\ 
+\bf   Houston FSI & G3 & 0.986 & 0.987 \\ 
+   \hline
+\end{tabular}
+\caption{Area under the curve (AUC) for each (set) of the studies.
+As seen in the ROC curves, the algorithm performs the least well on set 2 of the Houston FSI study.
+Overall the perform on all sets is very good to excellent.
+At the land level there is no difference in prediction power between RF score and cross-correlation.} 
+\end{table}
 

### 2018-10-23 index.Rmd
references

@@
 
-```{r rocs, fig.caption="ROC curves for all studies based on land-to-land comparisons.", fig.width = 4.5, fig.height = 3, out.width = ".5\\textwidth"}
+```{r rocs, fig.cap="ROC curves for all studies based on land-to-land comparisons.", fig.width = 4.5, fig.height = 3, out.width = ".5\\textwidth"}
 library(plotROC)
@@
 Overall the perform on all sets is very good to excellent.
-At the land level there is no difference in prediction power between RF score and cross-correlation.} 
+At the land level there is no difference in prediction power between RF score and cross-correlation.\label{tab:aucs}} 
 \end{table}

### 2018-10-23 index.Rmd
bibliography and a few ref changes

@@
     template: template.tex
+---
 
----
 \newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
@@
   out.width = "\\textwidth",
-  cache = FALSE,
+  cache = TRUE,
   fig.path = "figures/",
@@
 This is likely due to several lands in Hamby set 44 that have major deficiencies,  such as 'tank rash' (a collision of the bullet after exiting the barrel with a surface causing markings on top of the striations from the barrel) or extreme pitting (holes caused by direct contact with burning gun powder).
-\autoref{fig:tankrash} gives an overview of all lands of bullets in the Hamby-44 set affected.
+\autoref{fig:tankrash} gives an overview of all lands of affected bullets in the Hamby-44 set.
 These issues affect the algorithm's ability to identify a stable signature, which impacts the subsequent match scores.
@@
 ### Limitations
-All of the studies involve the same type of firearm.
-We see some variations in scores across different models of Rugers.
-Likely, we are going to see even bigger variations with different brands of firearms, where there might be different number of lands of different lengths.
-We also know  that different firearms mark differently well \cite{???}.
+All of the studies \svp{presented here} involve the same type of firearm.
+We see some variations in scores across different models of Rugers, and it is probable that we will see even bigger variations with different brands of firearms that vary in number of lands and land length.
+We also know that different firearms and ammunition combinations mark differently well \citep{bolton-king_preventing_2016}.
 Rugers are among the firearms that mark very well, so we would correspondingly expect to see lowered similarity scores for other firearms.
@@
 The algorithm proposed by \citet{aoas} is not capable of  automatically detecting parts of lands with well expressed striae.
-XXX citation for neural network algorithm
+\svp{It may be useful to couple the \citet{aoas} algorithm with an algorithm which assesses the quality of the input data and determines which portions of the data to use for comparison.
+This would emulate the process used by examiners, who first assess the quality of the evidence and whether there is enough information present to attempt a match.} XXX citation for neural network algorithm \svp{Do you want to release this idea into the wild at this point?
+}
 

### 2018-10-23 index.Rmd
color change?

@@
   ylab("Fraction True Positives") +
-  scale_colour_manual("Study", values = c("#fe9929", "#8856a7",
+  scale_colour_manual("Study", values = c("#fe9929", "#aa0000",
                                          "#74a9cf", "#2b8cbe", "#045a8d"))

### 2018-10-23 index.Rmd
spelling check, grammar

@@
 However, there is some discussion on whether or in which situations these error rates are biased [@rfOver2; @rfOver].
-Neither of these papers adresses another issue with internal test sets: internal test sets are constructed to have the same distribution as the training data (apart from sampling variability).
+Neither of these papers addresses another issue with internal test sets: internal test sets are constructed to have the same distribution as the training data (apart from sampling variability).
 For any machine learning method the true test comes with testing the performance of an algorithm on an external test data.
@@
 Set 173 was originally published as Hamby set 44.
-This mislabelling has been corrected now.
+This mis-labeling has been corrected now.
 
@@
 
-What we find in \autoref{fig:compare} is that neither of the measure fulfills both of the minimal requirements laid out above.
+What we find in \autoref{fig:compare} is that neither of the measures fulfill both of the minimal requirements laid out above.
 Both measures fulfill (R1), i.e.
@@
 
-There are several approaches for potential improvements: we can try to fine tune the matching algorithm to take make and model of the firearm and ammunition used into account or expand the trainings base of the algorithm to include a wider variety of makes and models.
+There are several approaches for potential improvements: we can try to fine tune the matching algorithm to take make and model of the firearm and ammunition used into account or expand the training base of the algorithm to include a wider variety of makes and models.
 The downside of this is that we need to considerably expand the database used for training.

### 2018-10-23 index.Rmd
wording

@@
 ## 3D Imaging Technology
-3D characterization of bullet surfaces has been considered since 1958 (\citeauthor{davis1968introduction}), but the technology was not sufficiently advanced for analysis based on 3D measurements to be a viable option.
-Approaches to utilizing 3D measurement data were proposed starting in 1999 (\citeauthor{dekinderAutomatedComparisonsBullet1999}), and have been further developed by \citet{bachrachDevelopment3DbasedAutomated2002,xieAutomatedBulletidentificationSystem2009, chuPilotStudyAutomated2010}.
+3D characterization of bullet surfaces has been considered since at least 1958 (\citeauthor{davis1968introduction}), but the technology was not sufficiently advanced for analysis based on 3D measurements to be a viable option.
+More recently, approaches using 3D measurement data were proposed starting in 1999 (\citeauthor{dekinderAutomatedComparisonsBullet1999}), and have been further developed by \citet{bachrachDevelopment3DbasedAutomated2002,xieAutomatedBulletidentificationSystem2009, chuPilotStudyAutomated2010}.
 These approaches utilize 3D surface measurements directly to characterize the topology of bullet surfaces, rather than using visual or image comparison techniques.

### 2018-10-24 index.Rmd
Add reference and comment

@@
 As seen in the ROC curves, the algorithm performs the least well on set 2 of the Houston FSI study.
-Overall the perform on all sets is very good to excellent.
-At the land level there is no difference in prediction power between RF score and cross-correlation.\label{tab:aucs}} 
+Overall the performance on all sets is very good to excellent.
+At the land level there is no difference in prediction power between RF score and cross-correlation.
+\svp{It has been observed \citep{marzban_roc_2004} that AUC differentiates good models from bad, but does not effectively distinguish between good models to determine which model is better.} \label{tab:aucs}}
 \end{table}

### 2018-10-25 index.Rmd
added table with AUC

--- 2018-10-23 index.Rmd
@@
 
-```{r functions}
+```{r functions, eval=FALSE}
 estimateBeta <- function(x, method = "mm", probs = c(0.25, 0.75)) {
@@
 # Introduction and Background
-In current practice, firearms and toolmark examiners (FTE) evaluate the similarity of striae on bullets by placing the evidence together with a bullet from the suspected gun under a comparison microscope and visually classify similarity according to the theory of firearms identification \citep{identification} as one of identification, inconclusive or exclusion.
+In current practice, firearms and toolmark examiners (FTE) evaluate the similarity of striae on bullets by placing the evidence together with a bullet from the suspected gun under a comparison microscope.
+They then visually classify similarity according to the theory of firearms identification \citep{identification} as one of identification, inconclusive or exclusion.
 Exact guidelines for this classification vary from lab to lab, e.g.\ some labs will exclude only on the basis of non-matching class characteristics (such as direction of the twist in rifling, land length or number of lands, or type of rifling).
@@
 ## 3D Imaging Technology
-3D characterization of bullet surfaces has been considered since 1958 (\citeauthor{davis1968introduction}), but the technology was not sufficiently advanced for analysis based on 3D measurements to be a viable option.
-Approaches to utilizing 3D measurement data were proposed starting in 1999 (\citeauthor{dekinderAutomatedComparisonsBullet1999}), and have been further developed by \citet{bachrachDevelopment3DbasedAutomated2002,xieAutomatedBulletidentificationSystem2009, chuPilotStudyAutomated2010}.
-These approaches utilize 3D surface measurements directly to characterize the topology of bullet surfaces, rather than using visual or image comparison techniques.
+More explicit characterization of bullet surfaces has been discussed since at least 1958 (\citeauthor{davis1968introduction}), but the technology was not sufficiently advanced for analysis based on 3D measurements to be a viable option.
+\svp{Use of a scanning electron microscope to quantitatively examine and compare bullet striae was demonstrated in \citet{ComputerIdentificationBullets1978}}.
+More recently, approaches using 3D measurement data were explicitly described in 1999 (\citeauthor{dekinderAutomatedComparisonsBullet1999}), and have been further developed by \citet{bachrachDevelopment3DbasedAutomated2002,xieAutomatedBulletidentificationSystem2009, chuPilotStudyAutomated2010}.
+These approaches utilize 3D surface measurements directly to characterize the topology of land engraved areas (LEAs), rather than using visual or image comparison techniques.
 In many cases, they also provide some level of automation of the comparison process, with the goal of reducing human biases by augmenting the visual information with 3D measurements.
-This allows for examination not only of the peaks in the bullet signature, but also for comparison of the valleys, which are more difficult to assess visually except in relation to the peaks.
+Utilizing the 3D measurements allows for examination not only of the peaks in the bullet signature, but also for comparison of the valleys, which are more difficult to assess visually except in relation to the peaks.
+\svp{In addition, there is significantly more information to be gained from quantitative assessment of peak height and valley depth than from the qualitative assessment of the same information.}
 
-Commonly, approaches derived from 3D surface measurements use some features which are developed to mimic the perception of FTEs\citep{luAutomatedBulletIdentification2014}.
+Commonly, approaches derived from 3D surface measurements use some features which are \svp{similar to those qualitatively analyzed by FTEs}\citep{luAutomatedBulletIdentification2014}.
 Class characteristics, which are shared by a group of firearms with the same rifling design, manufacturer, and tooling process, are typically evaluated first, as a mismatch on class charcteristics is sufficient for an exclusion.
 Automated approaches to estimation of width of the land engraved area and twist angle were described in \citet{chuPilotStudyAutomated2010}.
-Individual characteristics, which are not shared by all members of a class, can also be automatically assessed from 3D surface measurements; in some cases, these features mimic features used by FTEs during image analysis.
+Individual characteristics, which are not shared by all members of a class, can also be automatically assessed from 3D surface measurements.
 One of the most common features used to assess the strength of a match based on individual characteristics is the cross-correlation function (ccf), which is described in several studies \citep{maNISTBulletSignature2004, vorburgerApplicationsCrosscorrelationFunctions2011, chuPilotStudyAutomated2010}.
@@
 
-In the matching algorithm proposed in \citet{aoas}, several of the features described above are combined to produce a similarity score based on 3d topographic scans of land engraved areas (LEAs).
+In the matching algorithm proposed in \citet{aoas}, several of the features described above are combined using a random forest\citep{breiman} to produce a similarity score based on 3d topographic scans of land engraved areas (LEAs).
 From each land engraved area a signature is derived by
@@
 
-\hh{The automatic matching algorithm is based on a random forest [@breiman], which has a built-in internal testing mechanism to prevent potential overfitting.
+\hh{The automatic matching algorithm is based on a random forest \citep{breiman}, which has a built-in internal testing mechanism to prevent potential overfitting.
 Error rates reported by random forest algorithms are based on these internal test sets.
-However, there is some discussion on whether or in which situations these error rates are biased [@rfOver2; @rfOver].
+However, there is some discussion on whether or in which situations these error rates are biased \citep{rfOver2, rfOver}.
 Neither of these papers addresses another issue with internal test sets: internal test sets are constructed to have the same distribution as the training data (apart from sampling variability).
@@
 anomalies that are shared between different barrels manufactured in this manner.
-This should make the automatic matching harder, and in particular, the classification as different-source land engraved areas/bullets.
+This should make the automatic matching harder, and in particular, \svp{is expected to complicate} the classification as different-source land engraved areas/bullets.
 
@@
 
-Looking back at Figure \ref{fig:b2b}, we see that for the two bullets from the same barrel, the sequence average for 
+Looking back at \autoref{fig:b2b}, we see that for the two bullets from the same barrel, the sequence average for 
 $k=2$ is higher than the other sequence averages, and also higher than the sequence averages for the other pair of bullets shown on the right of the figure.
@@
 # tank rash lands:
-h44features <- h44features %>% mutate(
-  rfscore = replace(rfscore, (bullet1 == "I" & land1 == 6) | (bullet2 == "I" & land2 == 6), NA)
-)
+#h44features <- h44features %>% mutate(
+#  rfscore = replace(rfscore, (bullet1 == "I" & land1 == 6) | (bullet2 == #"I" & land2 == 6), NA)
+#)
 h44features <- h44features %>% mutate(
@@
     scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore)
+    max(scores)
+  }),
+  sam_cms = data %>% purrr::map_dbl(.f = function(d) {
+    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$cms)
     max(scores)
@@
 These strips correspond to known barrels 1 through 10.
-For each of these barrels, two test fires were done.
-Each of the colored tiles corresponds to a pair of bullets between a questioned bullet (shown along the $y$-axis) and one of the other bullets.
+\svp{Each barrel was test-fired twice, so }each of the colored tiles corresponds to a \svp{comparison} between a questioned bullet (shown along the $y$-axis) and one of the \svp{known} bullets.
 The similarity score is encoded in color: grey colors correspond to low similarity scores, orange colors correspond to high similarity scores.
 Ground truth is encoded in this figure as a thin orange frame for all pairs of same-source bullets.
-What we want to see for the relationship between questioned bullets and the ten barrels, is one barrel with orange-filled, orange-framed tiles for both bullets from the same barrel, and grey tiles across the other barrels.
-This expectation is met for all questioned bullets, i.e.
+\svp{We want to see one barrel with two orange-filled, orange-framed tiles for each questioned bullet, and grey tiles for all other barrels, indicating a match between a questioned bullet and a single barrel.} This expectation is met for all questioned bullets, i.e.
 the automatic matching  identifies the correct barrel for all questioned bullets.
 For two of the questioned bullets, 'I' and 'F', the similarity scores to the matching barrels are considerably smaller than for the other questioned bullets.
+
 The right side of \autoref{fig:h44} shows the relationship between all pairs of questioned bullets.
@@
       b1, b2, barrel1, bullet1,
-      land1, bullet2, land2, ccf, rfscore, KM
+      land1, bullet2, land2, ccf, rfscore, cms, KM
     ) %>%
@@
       b1, b2, bullet1,
-      land1, bullet2, land2, ccf, rfscore
+      land1, bullet2, land2, ccf, rfscore, cms, 
     ) %>%
@@
     any(d$KM)
+  }),
+  sam_cms = data %>% purrr::map_dbl(.f = function(d) {
+    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$cms)
+    max(scores)
   })
@@
 In set 3, six of the questioned bullets match known bullets, and the remaining two questioned bullets match each other but do not match any known bullets in the set.
-%Thus, we can conclude that the algorithm performs well on open set studies, as it can identify matches among questioned bullets which do not match any known bullets. 
+<!--Thus, we can conclude that the algorithm performs well on open set studies, as it can identify matches among questioned bullets which do not match any known bullets. -->
 While all bullet-to-bullet matches are correctly identified, and no known matches are missed, it is obvious from the generally lighter shades of the tiles corresponding to matching bullets in \autoref{fig:hou} for set 2 that  the algorithm is not performing as well as for the other sets (see also \autoref{tab:aucs}).
@@
 
-\autoref{fig:compare} shows density curves for the scores of each of the studies, comparing the SAM scores based on cross-correlation (top) and the random forest score (bottom).
+\autoref{fig:compare} shows density curves for the scores bullet-to-bullet matches of each of the studies, comparing the SAM scores based on cross-correlation (top) and the random forest score (bottom).
 Color indicates ground truth.
-Besides densities, small vertical lines below the $x$-axis indicate observed scores in a so-called **rug-plot**.
+Besides densities, small vertical lines below the $x$-axis indicate observed scores in a **rug-plot**.
 The plot shows that for all three case studies and both measures, i.e.
@@
   geom_density(alpha = 0.6) +
-  geom_rug(alpha = 0.5) +
+  geom_rug(alpha = 0.5, show.legend = F) +
   facet_grid(measure ~ study) +
@@
   scale_colour_manual("Same source", values = c("darkgrey", "darkorange")) +
-  scale_x_continuous("SAM scores", limits = c(0, 1)) +
+  scale_x_continuous("SAM scores", limits=c(0,1)) +
   theme(legend.position = "bottom") +
-  geom_hline(yintercept = 0) +
+  geom_hline(yintercept = 0, show.legend = F) +
   geom_text(aes(x = sam_rf, y = y, label = comparison),
+            show.legend = F,
             hjust = 0,
@@
   geom_segment(aes(x = sam_rf, xend = sam_rf, y = 0, yend = y),
+               show.legend = F,
                size = 0.1,
@@
 
-\hh{XXX error rates on land to land matches? ROC curves in \autoref{fig:rocs} and corresponding AUC results for all studies in \autoref{tab:rocs} XXX}
+\svp{Examining land-to-land scores provides more details about the matching algorithm's performance.
+\autoref{fig:compare-land-to-land} shows density curves and rug plots for each study.
+There is some overlap between known matching and known non-matching land-to-land scores; in many cases the random forest scores for these overlapping values are more extreme than the corresponding ccf scores.
+This suggests that the matching algorithm is sensitive to the presence of LEAs which have low land-to-land match scores, such as the matches of bullets `I' and `F' in the Hamby 44 study.
+Comparing this to the bullet-to-bullet score density plot shown in \autoref{fig:compare}, we see that there is much greater separation between the same-source and different-source densities for each study in the bullet-to-bullet comparisons than in the land-to-land comparisons, that is, one or two poorly matching lands does not prevent the bullet from showing a matching score, though if there are only weak matches on a few lands, such as the marked Hamby 44 comparisons in \autoref{fig:compare}, the overall score may be affected.
+}
 
-```{r, cache=TRUE}
+```{r, cache=FALSE}
 hounest <- hounest %>% mutate(
@@
 
+set.seed(20140501)
+h44nest$samesource <- h44nest$sam_rf > 0.3
+h44nest <- h44nest %>% 
+  mutate(
+    data = purrr::map2(.x = data, .y = samesource, .f = function(d, y) {
+      d$samesource <- y
+      d
+  }
+  ))
+
 h44nest <- h44nest %>% mutate(
@@
     .f = function(d) {
+      if (any(d$samesource)) {
       d$sameland = bullet_to_land_predict(
         land1 = d$land1, land2 = d$land2, 
-        d$rfscore, difference=0.1)
+        d$rfscore, difference=0.01, alpha =1)
+      } else d$sameland <- FALSE
+      d <- d %>% select(-samesource)
       d
@@
         land1 = d$land1, land2 = d$land2, 
-        d$rfscore, difference=0.1)
+        d$rfscore, difference=0, alpha =1)
       d
@@
   h44lands %>%
-    select(rfscore, ccf, sameland, samesource) %>%
+    select(rfscore, ccf, cms, sameland, samesource) %>%
     mutate(study = "Hamby 44", group1 = ""),
   pdlands %>%
-    select(rfscore, ccf, sameland, samesource) %>%
+    select(rfscore, ccf, cms, sameland, samesource) %>%
     mutate(study = "Phoenix PD", group1 = ""),
   houlands %>%
-    select(rfscore, ccf, sameland, samesource, group1) %>%
+    select(rfscore, ccf, cms, sameland, samesource, group1) %>%
     mutate(study = "Houston FSI")
@@
 
-```{r rocs, fig.cap="ROC curves for all studies based on land-to-land comparisons.", fig.width = 4.5, fig.height = 3, out.width = ".5\\textwidth"}
+```{r check44, eval=FALSE}
+h44lands <- h44lands %>%
+  mutate(
+    sameland = replace(sameland, !samesource, FALSE)
+  )
+h44lands %>% 
+  ggplot(aes(x = land1, y = land2, fill = ccf)) +
+  geom_tile() +
+  geom_tile(aes(colour = sameland), size = 1, 
+            data = h44lands %>% filter(sameland)) +
+  scale_fill_gradient2("RF score: ", midpoint=0.5, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
+  scale_colour_manual(values="darkorange") +
+  facet_grid(bullet1~barrel2+bullet2)
+
+```
+
+```{r checkpd, eval=FALSE}
+pdlands <- pdlands %>%
+  mutate(
+    sameland = replace(sameland, !samesource, FALSE)
+  )
+pdlands %>% 
+  ggplot(aes(x = land1, y = land2, fill = ccf)) +
+  geom_tile() +
+  geom_tile(aes(colour = sameland), size = 1, 
+            data = pdlands %>% filter(sameland)) +
+  scale_fill_gradient2("RF score: ", midpoint=0.5, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
+  scale_colour_manual(values="darkorange") +
+  facet_grid(bullet1~barrel2+bullet2)
+
+```
+
+```{r checkhou, eval=FALSE}
+houlands <- houlands %>%
+  mutate(
+    sameland = replace(sameland, !samesource, FALSE)
+  )
+hou1 <- houlands %>% filter(group1==group2, group1=="G1")
+
+hou1 %>%
+  ggplot(aes(x = land1, y = land2, fill = ccf)) +
+  geom_tile() +
+  geom_tile(aes(colour = sameland), size = 1, 
+            data = hou1 %>% filter(sameland)) +
+  scale_fill_gradient2("RF score: ", midpoint=0.5, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
+  scale_colour_manual(values="darkorange") +
+  facet_grid(bullet1~barrel2+bullet2)
+
+hou2 <- houlands %>% filter(group1==group2, group1=="G2")
+
+hou2 %>%
+  ggplot(aes(x = land1, y = land2, fill = ccf)) +
+  geom_tile() +
+  geom_tile(aes(colour = sameland), size = 1, 
+            data = hou2 %>% filter(sameland)) +
+  scale_fill_gradient2("RF score: ", midpoint=0.5, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
+  scale_colour_manual(values="darkorange") +
+  facet_grid(bullet1~barrel2+bullet2)
+
+hou3 <- houlands %>% filter(group1==group2, group1=="G3")
+
+hou3 %>%
+  ggplot(aes(x = land1, y = land2, fill = ccf)) +
+  geom_tile() +
+  geom_tile(aes(colour = sameland), size = 1, 
+            data = hou3 %>% filter(sameland)) +
+  scale_fill_gradient2("RF score: ", midpoint=0.5, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
+  scale_colour_manual(values="darkorange") +
+  facet_grid(bullet1~barrel2+bullet2)  
+```
+
+```{r compare-land-to-land, fig.width = 8, fig.height = 4, fig.cap="Density curves of land-to-land similarity scores from cross-correlation (top) and RF scores (bottom).
+Different colors indicate same source and phase versus different source or out-of-phase matches.
+The random forest scores tend to be more extreme than the cross correlation scores alone."}
+all_lands %>% 
+  gather(key = "measure", value = "value", ccf, rfscore) %>%
+  mutate(measure = factor(measure, levels = c("ccf", "rfscore"),
+                          labels = c("Cross-correlation", "RF score")),
+         study = paste(study, group1) %>% str_trim() %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"))) %>%
+  ggplot(aes(
+    x = value, fill = sameland,
+    colour = sameland
+  )) +
+  geom_density(alpha = 0.6) +
+  geom_rug(alpha = 0.5/12, show.legend = F) +
+  facet_grid(measure ~ study) +
+  theme_bw() +
+  scale_fill_manual("Same source & best sequence", values = c("darkgrey", "darkorange")) +
+  scale_colour_manual("Same source & best sequence", values = c("darkgrey", "darkorange")) +
+  scale_x_continuous("Land-to-Land scores", limits = c(0, 1)) +
+  theme(legend.position = "bottom") +
+  geom_hline(yintercept = 0, show.legend = F)
+```
+
+\svp{When evaluating classifier performance, it is common to use a receiver operating curve, which plots the false positives against the true positives for each possible value of the cutoff between the two classes (in this case, known match and known non-match).
+The ROC curves for the land-to-land scores are shown in \autoref{fig:rocs}.
+A perfect classifier would have 100\% true positives and 0\% false positives, e.g.
+the ROC curve would be a right angle at (0, 1).
+Classifiers with better performance will be closer to this corner of the plot.
+A random classifier would have an ROC curve that was a straight diagonal line through (0,0) and (1,1).}
+
+```{r rocs, fig.cap="ROC curves for all studies based on land-to-land comparisons \\svp{of the random forest score}.", fig.width = 4.5, fig.height = 3, out.width = ".5\\textwidth"}
 library(plotROC)
@@
 
-```{r aucs, eval = FALSE}
+```{r aucs, eval = TRUE, results = 'asis'}
+
 aucs <- all_lands %>% group_by(study, group1) %>% nest() %>%
@@
       auc(d$sameland, d$ccf)
+    }),
+    auc_cms = data %>% purrr::map_dbl(function(d) {
+      auc(d$sameland, d$cms)
+    }),
+    auc_rf_ci_low = data %>% purrr::map_dbl(function(d) {
+      ci.auc(d$sameland, d$rfscore)[1]
+    }),
+    auc_rf_ci_upper = data %>% purrr::map_dbl(function(d) {
+      ci.auc(d$sameland, d$rfscore)[3]
+    }),
+    auc_ccf_ci_lower = data %>% purrr::map_dbl(function(d) {
+      ci.auc(d$sameland, d$ccf)[1]
+    }),
+    auc_ccf_ci_upper = data %>% purrr::map_dbl(function(d) {
+      ci.auc(d$sameland, d$ccf)[3]
+    }),
+    auc_cms_ci_lower = data %>% purrr::map_dbl(function(d) {
+      ci.auc(d$sameland, d$cms)[1]
+    }),
+    auc_cms_ci_upper = data %>% purrr::map_dbl(function(d) {
+      ci.auc(d$sameland, d$cms)[3]
     })
@@
 names(aucs)[5] <- "Cross-correlation"
-print(xtable(aucs %>% select(-data), digits = 3, caption = "Area under the curve (AUC) for each (set) of the studies.
+names(aucs)[6] <- "CMS"
+
+aucs <- as.data.frame(aucs)
+aucs[,4] <- sprintf("%.3f (%.3f, %.3f)", aucs[,4], aucs[,7], aucs[,8])
+aucs[,5] <- sprintf("%.3f (%.3f, %.3f)", aucs[,5], aucs[,9], aucs[,10])
+aucs[,6] <- sprintf("%.3f (%.3f, %.3f)", aucs[,6], aucs[,11], aucs[,12])
+print(xtable(aucs[,1:6] %>% select(-data), digits = 3, caption = "Area under the curve (AUC) for each (set) of the studies.
 At the land level there is no difference in prediction power between RF score and cross-correlation."), include.rownames=FALSE, label="tab:aucs")
@@
 \centering
-\begin{tabular}{llrr}
+\begin{tabular}{llrrr}
   \hline
-\bf study & \bf set & \bf AUC(RF score) & \bf AUC(Cross-correlation) \\ 
+ && \multicolumn{3}{l}{\bf AUC (95\% Conf. Int)}\\
+\bf Study & \bf Set & \bf RF score & \bf Cross-correlation & \bf CMS \\
   \hline
-\bf Hamby 44 &  & 0.956 & 0.926 \\ 
-\bf   Phoenix PD &  & 0.976 & 0.986 \\ 
-\bf   Houston FSI & G1 & 0.991 & 0.988 \\ 
-\bf   Houston FSI & G2 & 0.898 & 0.902 \\ 
-\bf   Houston FSI & G3 & 0.986 & 0.987 \\ 
+\bf Hamby 44 &  & 0.937 (0.917, 0.957) & 0.916 (0.891, 0.941) & 0.820 (0.787, 0.854) \\
+\bf  Phoenix PD &  & 0.976 (0.957, 0.994) & 0.986 (0.975, 0.996) & 0.879 (0.838, 0.919) \\
+\bf  Houston FSI & Set 1 & 0.991 (0.983, 0.999) & 0.988 (0.979, 0.997) & 0.785 (0.727, 0.842) \\
+\bf  Houston FSI & Set 2 & 0.898 (0.853, 0.943) & 0.902 (0.861, 0.943) & 0.749 (0.684, 0.814) \\
+\bf  Houston FSI & Set 3 & 0.986 (0.973, 0.999) & 0.987 (0.978, 0.997) & 0.911 (0.880, 0.941) \\
    \hline
 \end{tabular}
-\caption{Area under the curve (AUC) for each (set) of the studies.
+\caption{Area under the curve (AUC) for each (set) of the studies for the random forest score, cross-correlation and CMS.
+Overall the perform on all sets is very good to excellent for cross-correlation and the random forest score.
 As seen in the ROC curves, the algorithm performs the least well on set 2 of the Houston FSI study.
-Overall the perform on all sets is very good to excellent.
-At the land level there is no difference in prediction power between RF score and cross-correlation.\label{tab:aucs}} 
+At the land level there is no significant difference in prediction power between RF score and cross-correlation.} \label{tab:aucs}}
 \end{table}
+
+\svp{In \autoref{fig:rocs}, the Houston FSI G1 and G3 curves and the Phoenix PD curve show excellent performance; Houston FSI G2 and Hamby 44 show relatively good performance.
+Area under the curve (AUC) values, which summarize ROC curves, are shown in \autoref{tab:aucs}.
+AUC values are useful for differentiating between poor, good, and excellent model performance, but are not particularly useful when determining which of several models with approximately the same level of performance should be used \citep{marzban_roc_2004}.
+Hamby 44 and Phoenix PD are the only studies with a somewhat noticable difference between the AUC score for RF compared to ccf, and even those differences are relatively small: .03 and -.01, respectively.}
+
+\svp{XXX show ROC for bullet-to-bullet is perfect, just for emphasis?}
+
+```{r why, fig.cap="Density plots for all measures "}
+all_lands %>% gather(measure, score, ccf, cms, rfscore) %>%
+ggplot(aes(x = score, fill = interaction(measure, sameland))) + 
+  geom_density(alpha = 0.5) +
+  scale_fill_brewer(palette = "Paired") +
+  facet_grid(measure~study+group1, scales="free_x") 
+```
 
@@
 
-What is quite surprising is that what we anticipated to be the easiest case of Hamby 44 turned out to be the hardest to solve.
+What is quite surprising is that \svp{while} we anticipated \svp{Hamby 44 would be the easiest set to work with, but it} turned out to be the hardest to solve.
 This is likely due to several lands in Hamby set 44 that have major deficiencies,  such as 'tank rash' (a collision of the bullet after exiting the barrel with a surface causing markings on top of the striations from the barrel) or extreme pitting (holes caused by direct contact with burning gun powder).
@@
 ### Limitations
+
 All of the studies \svp{presented here} involve the same type of firearm.
 We see some variations in scores across different models of Rugers, and it is probable that we will see even bigger variations with different brands of firearms that vary in number of lands and land length.
-We also know that different firearms and ammunition combinations mark differently well \citep{bolton-king_preventing_2016}.
+We also know that different firearms and ammunition combinations mark differently well \citep{bolton-king_preventing_2016, bonfanti_influence_1999}.
 Rugers are among the firearms that mark very well, so we would correspondingly expect to see lowered similarity scores for other firearms.

### 2018-10-25 index.Rmd
table expanded
phrasing

--- 2018-10-24 index.Rmd
@@
 ## 3D Imaging Technology
-3D characterization of bullet surfaces has been considered since at least 1958 (\citeauthor{davis1968introduction}), but the technology was not sufficiently advanced for analysis based on 3D measurements to be a viable option.
-More recently, approaches using 3D measurement data were proposed starting in 1999 (\citeauthor{dekinderAutomatedComparisonsBullet1999}), and have been further developed by \citet{bachrachDevelopment3DbasedAutomated2002,xieAutomatedBulletidentificationSystem2009, chuPilotStudyAutomated2010}.
-These approaches utilize 3D surface measurements directly to characterize the topology of bullet surfaces, rather than using visual or image comparison techniques.
+More explicit characterization of bullet surfaces has been discussed since at least 1958 (\citeauthor{davis1968introduction}), but the technology was not sufficiently advanced for analysis based on 3D measurements to be a viable option.
+\svp{Use of a scanning electron microscope to quantitatively examine and compare bullet striae was demonstrated in \citet{ComputerIdentificationBullets1978}}.
+More recently, approaches using 3D measurement data were explicitly described in 1999 (\citeauthor{dekinderAutomatedComparisonsBullet1999}), and have been further developed by \citet{bachrachDevelopment3DbasedAutomated2002,xieAutomatedBulletidentificationSystem2009, chuPilotStudyAutomated2010}.
+These approaches utilize 3D surface measurements directly to characterize the topology of land engraved areas (LEAs), rather than using visual or image comparison techniques.
 In many cases, they also provide some level of automation of the comparison process, with the goal of reducing human biases by augmenting the visual information with 3D measurements.
-This allows for examination not only of the peaks in the bullet signature, but also for comparison of the valleys, which are more difficult to assess visually except in relation to the peaks.
+Utilizing the 3D measurements allows for examination not only of the peaks in the bullet signature, but also for comparison of the valleys, which are more difficult to assess visually except in relation to the peaks.
+\svp{In addition, there is significantly more information to be gained from quantitative assessment of peak height and valley depth than from the qualitative assessment of the same information.}
 
-Commonly, approaches derived from 3D surface measurements use some features which are developed to mimic the perception of FTEs\citep{luAutomatedBulletIdentification2014}.
+Commonly, approaches derived from 3D surface measurements use some features which are \svp{similar to those qualitatively analyzed by FTEs}\citep{luAutomatedBulletIdentification2014}.
 Class characteristics, which are shared by a group of firearms with the same rifling design, manufacturer, and tooling process, are typically evaluated first, as a mismatch on class charcteristics is sufficient for an exclusion.
 Automated approaches to estimation of width of the land engraved area and twist angle were described in \citet{chuPilotStudyAutomated2010}.
-Individual characteristics, which are not shared by all members of a class, can also be automatically assessed from 3D surface measurements; in some cases, these features mimic features used by FTEs during image analysis.
+Individual characteristics, which are not shared by all members of a class, can also be automatically assessed from 3D surface measurements.
 One of the most common features used to assess the strength of a match based on individual characteristics is the cross-correlation function (ccf), which is described in several studies \citep{maNISTBulletSignature2004, vorburgerApplicationsCrosscorrelationFunctions2011, chuPilotStudyAutomated2010}.

### 2018-10-25 index.Rmd
land-to-land scores

@@
 
-\hh{XXX error rates on land to land matches? ROC curves in \autoref{fig:rocs} and corresponding AUC results for all studies in \autoref{tab:rocs} XXX}
+\svp{Examining land-to-land scores provides more details about the matching algorithm's performance.
+Figure \ref{fig:compare-land-to-land} shows density curves and rug plots for each study.
+There is some overlap between known matching and known non-matching land-to-land scores; in many cases the random forest scores for these overlapping values are more extreme than the corresponding ccf scores.
+This suggests that the matching algorithm is sensitive to the presence of LEAs which have low land-to-land match scores, such as the matches of bullets `I' and `F' in the Hamby 44 study.
+}
+
+\svp{When evaluating classifier performance, it is common to use a receiver operating curve, which plots the false positives against the true positives for each possible value of the cutoff between the two classes (in this case, known match and known non-match).
+The ROC curves for the land-to-land scores are shown in \autoref{fig:rocs}.
+A perfect classifier would have 100\% true positives and 0\% false positives, e.g.
+the ROC curve would be a right angle at (0, 1).
+Classifiers with better performance will be closer to this corner of the plot.
+A random classifier would have an ROC curve that was a straight diagonal line through (0,0) and (1,1).}
+
+\svp{In \autoref{fig:rocs}, the Houston FSI G1 and G3 curves and the Phoenix PD curve show excellent performance; Houston FSI G2 and Hamby 44 show relatively good performance.
+Area under the curve (AUC) values, which summarize ROC curves, are shown in \autoref{tab:aucs}.
+AUC values are useful for differentiating between poor, good, and excellent model performance, but are not particularly useful when determining which of several models with approximately the same level of performance should be used \citep{marzban_roc_2004}.
+Hamby 44 is the only study with a somewhat noticable difference between the AUC score for RF compared to ccf, and even that difference is relatively small.}
+
+\hh{XXX error rates on land to land matches? ROC curves in \autoref{fig:rocs} and corresponding AUC results for all studies in \autoref{tab:aucs} XXX}
 
@@
 
-```{r rocs, fig.cap="ROC curves for all studies based on land-to-land comparisons.", fig.width = 4.5, fig.height = 3, out.width = ".5\\textwidth"}
+```{r compare-land-to-land, fig.width = 8, fig.height = 4, fig.cap="Density curves of land-to-land similarity scores from cross-correlation (top) and RF scores (bottom).
+Different colors indicate same source and phase versus different source or out-of-phase matches.
+The random forest scores tend to be more extreme than the cross correlation scores alone."}
+all_lands %>% 
+  gather(key = "measure", value = "value", ccf, rfscore) %>%
+  mutate(measure = factor(measure, levels = c("ccf", "rfscore"),
+                          labels = c("Cross-correlation", "RF score")),
+         study = paste(study, group1) %>% str_trim() %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"))) %>%
+  ggplot(aes(
+    x = value, fill = sameland,
+    colour = sameland
+  )) +
+  geom_density(alpha = 0.6) +
+  geom_rug(alpha = 0.5/12, show.legend = F) +
+  facet_grid(measure ~ study) +
+  theme_bw() +
+  scale_fill_manual("Same source & best sequence", values = c("darkgrey", "darkorange")) +
+  scale_colour_manual("Same source & best sequence", values = c("darkgrey", "darkorange")) +
+  scale_x_continuous("Land-to-Land scores", limits = c(0, 1)) +
+  theme(legend.position = "bottom") +
+  geom_hline(yintercept = 0, show.legend = F)
+```
+```{r rocs, fig.cap="ROC curves for all studies based on land-to-land comparisons \\svp{of the random forest score}.", fig.width = 4.5, fig.height = 3, out.width = ".5\\textwidth"}
 library(plotROC)
@@
   \hline
-\bf study & \bf set & \bf AUC(RF score) & \bf AUC(Cross-correlation) \\ 
+\bf Study & \bf Set & \bf AUC(RF score) & \bf AUC(Cross-correlation) \\ 
   \hline
@@
 Overall the performance on all sets is very good to excellent.
-At the land level there is no difference in prediction power between RF score and cross-correlation.
-\svp{It has been observed \citep{marzban_roc_2004} that AUC differentiates good models from bad, but does not effectively distinguish between good models to determine which model is better.} \label{tab:aucs}}
+At the land level there is no difference in the predictive power between RF score and cross-correlation. \label{tab:aucs}} 
 \end{table}

### 2018-10-25 index.Rmd
add reference

@@
   theme(legend.position = "bottom") +
-  geom_hline(yintercept = 0) +
+  geom_hline(yintercept = 0, show.legend = F) +
   geom_text(aes(x = sam_rf, y = y, label = comparison),
+            show.legend = F,
             hjust = 0,

### 2018-10-25 index.Rmd
Move text around a bit, phrasing

@@
 \svp{Examining land-to-land scores provides more details about the matching algorithm's performance.
-Figure \ref{fig:compare-land-to-land} shows density curves and rug plots for each study.
+\autoref{fig:compare-land-to-land} shows density curves and rug plots for each study.
 There is some overlap between known matching and known non-matching land-to-land scores; in many cases the random forest scores for these overlapping values are more extreme than the corresponding ccf scores.
 This suggests that the matching algorithm is sensitive to the presence of LEAs which have low land-to-land match scores, such as the matches of bullets `I' and `F' in the Hamby 44 study.
+Comparing this to the bullet-to-bullet score density plot shown in \autoref{fig:compare}, we see that there is much greater separation between the same-source and different-source densities for each study in the bullet-to-bullet comparisons than in the land-to-land comparisons, that is, one or two poorly matching lands does not prevent the bullet from showing a matching score, though if there are only weak matches on a few lands, such as the marked Hamby 44 comparisons in \autoref{fig:compare}, the overall score may be affected.
 }
-
-\svp{When evaluating classifier performance, it is common to use a receiver operating curve, which plots the false positives against the true positives for each possible value of the cutoff between the two classes (in this case, known match and known non-match).
-The ROC curves for the land-to-land scores are shown in \autoref{fig:rocs}.
-A perfect classifier would have 100\% true positives and 0\% false positives, e.g.
-the ROC curve would be a right angle at (0, 1).
-Classifiers with better performance will be closer to this corner of the plot.
-A random classifier would have an ROC curve that was a straight diagonal line through (0,0) and (1,1).}
-
-\svp{In \autoref{fig:rocs}, the Houston FSI G1 and G3 curves and the Phoenix PD curve show excellent performance; Houston FSI G2 and Hamby 44 show relatively good performance.
-Area under the curve (AUC) values, which summarize ROC curves, are shown in \autoref{tab:aucs}.
-AUC values are useful for differentiating between poor, good, and excellent model performance, but are not particularly useful when determining which of several models with approximately the same level of performance should be used \citep{marzban_roc_2004}.
-Hamby 44 is the only study with a somewhat noticable difference between the AUC score for RF compared to ccf, and even that difference is relatively small.}
-
-\hh{XXX error rates on land to land matches? ROC curves in \autoref{fig:rocs} and corresponding AUC results for all studies in \autoref{tab:aucs} XXX}
 
@@
 ```
+
+\svp{When evaluating classifier performance, it is common to use a receiver operating curve, which plots the false positives against the true positives for each possible value of the cutoff between the two classes (in this case, known match and known non-match).
+The ROC curves for the land-to-land scores are shown in \autoref{fig:rocs}.
+A perfect classifier would have 100\% true positives and 0\% false positives, e.g.
+the ROC curve would be a right angle at (0, 1).
+Classifiers with better performance will be closer to this corner of the plot.
+A random classifier would have an ROC curve that was a straight diagonal line through (0,0) and (1,1).}
+
 ```{r rocs, fig.cap="ROC curves for all studies based on land-to-land comparisons \\svp{of the random forest score}.", fig.width = 4.5, fig.height = 3, out.width = ".5\\textwidth"}
@@
 
+\svp{In \autoref{fig:rocs}, the Houston FSI G1 and G3 curves and the Phoenix PD curve show excellent performance; Houston FSI G2 and Hamby 44 show relatively good performance.
+Area under the curve (AUC) values, which summarize ROC curves, are shown in \autoref{tab:aucs}.
+AUC values are useful for differentiating between poor, good, and excellent model performance, but are not particularly useful when determining which of several models with approximately the same level of performance should be used \citep{marzban_roc_2004}.
+Hamby 44 and Phoenix PD are the only studies with a somewhat noticable difference between the AUC score for RF compared to ccf, and even those differences are relatively small: .03 and -.01, respectively.}
+
+\hh{XXX error rates on land to land matches? ROC curves in \autoref{fig:rocs} and corresponding AUC results for all studies in \autoref{tab:aucs} XXX}
+
 # Discussion and Conclusions

### 2018-10-25 index.Rmd
comment out a comment, add one of my own

@@
 
-\hh{XXX error rates on land to land matches? ROC curves in \autoref{fig:rocs} and corresponding AUC results for all studies in \autoref{tab:aucs} XXX}
+<!-- \hh{XXX error rates on land to land matches? ROC curves in \autoref{fig:rocs} and corresponding AUC results for all studies in \autoref{tab:aucs} XXX} -->
+
+\svp{XXX show ROC for bullet-to-bullet is perfect, just for emphasis?}
 

### 2018-10-26 index.Rmd
phrasing
fix compile issues

@@
     max(scores)
+  }),
+  sam_cms = data %>% purrr::map_dbl(.f = function(d) {
+    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$cms)
+    max(scores)
   })
@@
   scale_colour_manual("Same source", values = c("darkgrey", "darkorange")) +
-  scale_x_continuous("SAM scores", limits = c(0, 1)) +
+  scale_x_continuous("SAM scores", limits=c(0,1)) +
   theme(legend.position = "bottom") +
@@
 
-```{r, cache=TRUE}
+```{r, cache=FALSE}
 hounest <- hounest %>% mutate(
@@
     .f = function(d) {
+      if (any(d$samesource)) {
       d$sameland = bullet_to_land_predict(
         land1 = d$land1, land2 = d$land2, 
-        d$rfscore, difference=0.1)
+        d$rfscore, difference=0.01)
+      } else d$sameland <- FALSE
+      d <- d %>% select(-samesource)
       d
@@
         land1 = d$land1, land2 = d$land2, 
-        d$rfscore, difference=0.1)
+        d$rfscore, difference=0)
       d
@@
   )
+```
+
+```{r check44, eval=FALSE}
+h44lands <- h44lands %>%
+  mutate(
+    sameland = replace(sameland, !samesource, FALSE)
+  )
+h44lands %>% 
+  ggplot(aes(x = land1, y = land2, fill = ccf)) +
+  geom_tile() +
+  geom_tile(aes(colour = sameland), size = 1, 
+            data = h44lands %>% filter(sameland)) +
+  scale_fill_gradient2("RF score: ", midpoint=0.5, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
+  scale_colour_manual(values="darkorange") +
+  facet_grid(bullet1~barrel2+bullet2)
+
+```
+
+```{r checkpd, eval=FALSE}
+pdlands <- pdlands %>%
+  mutate(
+    sameland = replace(sameland, !samesource, FALSE)
+  )
+pdlands %>% 
+  ggplot(aes(x = land1, y = land2, fill = ccf)) +
+  geom_tile() +
+  geom_tile(aes(colour = sameland), size = 1, 
+            data = pdlands %>% filter(sameland)) +
+  scale_fill_gradient2("RF score: ", midpoint=0.5, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
+  scale_colour_manual(values="darkorange") +
+  facet_grid(bullet1~barrel2+bullet2)
+
+```
+
+```{r checkhou, eval=FALSE}
+houlands <- houlands %>%
+  mutate(
+    sameland = replace(sameland, !samesource, FALSE)
+  )
+hou1 <- houlands %>% filter(group1==group2, group1=="G1")
+
+hou1 %>%
+  ggplot(aes(x = land1, y = land2, fill = ccf)) +
+  geom_tile() +
+  geom_tile(aes(colour = sameland), size = 1, 
+            data = hou1 %>% filter(sameland)) +
+  scale_fill_gradient2("RF score: ", midpoint=0.5, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
+  scale_colour_manual(values="darkorange") +
+  facet_grid(bullet1~barrel2+bullet2)
+
+hou2 <- houlands %>% filter(group1==group2, group1=="G2")
+
+hou2 %>%
+  ggplot(aes(x = land1, y = land2, fill = ccf)) +
+  geom_tile() +
+  geom_tile(aes(colour = sameland), size = 1, 
+            data = hou2 %>% filter(sameland)) +
+  scale_fill_gradient2("RF score: ", midpoint=0.5, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
+  scale_colour_manual(values="darkorange") +
+  facet_grid(bullet1~barrel2+bullet2)
+
+hou3 <- houlands %>% filter(group1==group2, group1=="G3")
+
+hou3 %>%
+  ggplot(aes(x = land1, y = land2, fill = ccf)) +
+  geom_tile() +
+  geom_tile(aes(colour = sameland), size = 1, 
+            data = hou3 %>% filter(sameland)) +
+  scale_fill_gradient2("RF score: ", midpoint=0.5, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
+  scale_colour_manual(values="darkorange") +
+  facet_grid(bullet1~barrel2+bullet2)  
 ```
@@
 
-```{r aucs, eval = FALSE}
+```{r aucs, eval = FALSE, results = 'asis'}
+
 aucs <- all_lands %>% group_by(study, group1) %>% nest() %>%
@@
 \centering
-\begin{tabular}{llrr}
+\begin{tabular}{llrrr}
   \hline
-\bf Study & \bf Set & \bf AUC(RF score) & \bf AUC(Cross-correlation) \\ 
+ && \multicolumn{3}{l}{\bf AUC (95\% Conf. Int)}\\
+\bf Study & \bf Set & \bf RF score & \bf Cross-correlation & \bf CMS \\
   \hline
-\bf Hamby 44 &  & 0.956 & 0.926 \\ 
-\bf   Phoenix PD &  & 0.976 & 0.986 \\ 
-\bf   Houston FSI & G1 & 0.991 & 0.988 \\ 
-\bf   Houston FSI & G2 & 0.898 & 0.902 \\ 
-\bf   Houston FSI & G3 & 0.986 & 0.987 \\ 
+\bf Hamby 44 &  & 0.937 (0.917, 0.957) & 0.916 (0.891, 0.941) & 0.820 (0.787, 0.854) \\
+\bf  Phoenix PD &  & 0.976 (0.957, 0.994) & 0.986 (0.975, 0.996) & 0.879 (0.838, 0.919) \\
+\bf  Houston FSI & Set 1 & 0.991 (0.983, 0.999) & 0.988 (0.979, 0.997) & 0.785 (0.727, 0.842) \\
+\bf  Houston FSI & Set 2 & 0.898 (0.853, 0.943) & 0.902 (0.861, 0.943) & 0.749 (0.684, 0.814) \\
+\bf  Houston FSI & Set 3 & 0.986 (0.973, 0.999) & 0.987 (0.978, 0.997) & 0.911 (0.880, 0.941) \\
    \hline
 \end{tabular}
-\caption{Area under the curve (AUC) for each (set) of the studies.
+\caption{Area under the curve (AUC) for each (set) of the studies for the random forest score, cross-correlation and CMS.
+Overall the perform on all sets is very good to excellent for cross-correlation and the random forest score.
 As seen in the ROC curves, the algorithm performs the least well on set 2 of the Houston FSI study.
-Overall the performance on all sets is very good to excellent.
-At the land level there is no difference in the predictive power between RF score and cross-correlation. \label{tab:aucs}} 
+At the land level there is no significant difference in prediction power between RF score and cross-correlation.} \label{tab:aucs}
 \end{table}
@@
 
-<!-- \hh{XXX error rates on land to land matches? ROC curves in \autoref{fig:rocs} and corresponding AUC results for all studies in \autoref{tab:aucs} XXX} -->
-
 \svp{XXX show ROC for bullet-to-bullet is perfect, just for emphasis?}
 
+```{r why, fig.cap="Density plots for all measures "}
+all_lands %>% gather(measure, score, ccf, cms, rfscore) %>%
+ggplot(aes(x = score, fill = interaction(measure, sameland))) + 
+  geom_density(alpha = 0.5) +
+  scale_fill_brewer(palette = "Paired") +
+  facet_grid(measure~study+group1, scales="free_x") 
+```
+
 # Discussion and Conclusions
 
-What is quite surprising is that what we anticipated to be the easiest case of Hamby 44 turned out to be the hardest to solve.
+What is quite surprising is that \svp{while} we anticipated \svp{Hamby 44 would be the easiest set to work with, but it} turned out to be the hardest to solve.
 This is likely due to several lands in Hamby set 44 that have major deficiencies,  such as 'tank rash' (a collision of the bullet after exiting the barrel with a surface causing markings on top of the striations from the barrel) or extreme pitting (holes caused by direct contact with burning gun powder).
@@
 \svp{It may be useful to couple the \citet{aoas} algorithm with an algorithm which assesses the quality of the input data and determines which portions of the data to use for comparison.
-This would emulate the process used by examiners, who first assess the quality of the evidence and whether there is enough information present to attempt a match.} XXX citation for neural network algorithm \svp{Do you want to release this idea into the wild at this point?
-}
+This would emulate the process used by examiners, who first assess the quality of the evidence and whether there is enough information present to attempt a match.}
 

### 2018-10-26 index.Rmd
comment

@@
 \end{tabular}
-\caption{Area under the curve (AUC) for each (set) of the studies for the random forest score, cross-correlation and CMS.
+\caption{\svp{Can we turn this into a picture?
+It's really dense and hard to interpret...}Area under the curve (AUC) for each (set) of the studies for the random forest score, cross-correlation and CMS.
 Overall the perform on all sets is very good to excellent for cross-correlation and the random forest score.

### 2018-10-26 index.Rmd
include ccf in ROC, exclude cms from densities

@@
 library(pROC)
-all_lands %>% 
-  ggplot(aes(colour = interaction(study, group1, sep = " "))) +
-  geom_roc(aes(m = rfscore, d = as.numeric(sameland)), 
+all_lands %>% gather(measure, score, rfscore, ccf) %>%
+  ggplot(aes(linetype=measure, colour = interaction(study, group1, sep = " "))) +
+  geom_roc(aes(m = score, d = as.numeric(sameland)), 
            labels =FALSE, pointsize = 0, alpha = 0.8) +
   theme_bw() +
+  facet_wrap(~study) +
   xlab("Fraction False Positives") +
@@
 ```{r why, fig.cap="Density plots for all measures "}
-all_lands %>% gather(measure, score, ccf, cms, rfscore) %>%
+all_lands %>% gather(measure, score, ccf, rfscore) %>%
 ggplot(aes(x = score, fill = interaction(measure, sameland))) + 

### 2018-10-26 index.Rmd
density curves for alnd to land comparisons

@@
 
-```{r why, fig.cap="Density plots for all measures "}
+```{r l2l, fig.cap="Density curves of similarity scores based on land-to-land comparisons for cross-correlation (top) and RF scores (bottom).
+Different colors indicate same source versus different source.
+RF scores for different source comparisons are generally well below 0.5.
+Some same-source comparisons are also below 0.5 for RF scores, indicating potential problems with at least one of the lands involved in the comparison."}
 all_lands %>% gather(measure, score, ccf, rfscore) %>%
-ggplot(aes(x = score, fill = interaction(measure, sameland))) + 
-  geom_density(alpha = 0.5) +
-  scale_fill_brewer(palette = "Paired") +
-  facet_grid(measure~study+group1, scales="free_x") 
+  mutate(
+    measure = factor(measure, labels=c("Cross-correlation", "RF score"))
+  ) %>%
+ggplot(aes(x = score, fill = sameland, colour =sameland)) + 
+  geom_density(alpha = 0.6) +
+  facet_grid(measure ~ study + group1) +
+  theme_bw() +
+  scale_fill_manual("Same source", values = c("darkgrey", "darkorange")) +
+  scale_colour_manual("Same source", values = c("darkgrey", "darkorange")) +
+  scale_x_continuous("Land-to-land scores", limits=c(0,1)) +
+  theme(legend.position = "bottom") +
+  geom_hline(yintercept = 0, show.legend = F) 
 ```

### 2018-10-31 index.Rmd
Getting rid of colored text that's old

@@
 Neither of these papers addresses another issue with internal test sets: internal test sets are constructed to have the same distribution as the training data (apart from sampling variability).
-For any machine learning method the true test comes with testing the performance of an algorithm on an external test data.
+For any machine learning method the true test comes with testing the performance of an algorithm on external test data.
 Good performance (in terms of low error rates) on external test data validates a ML algorithm.
@@
 
-\svp{In \autoref{fig:rocs}, the Houston FSI G1 and G3 curves and the Phoenix PD curve show excellent performance; Houston FSI G2 and Hamby 44 show relatively good performance.
+In \autoref{fig:rocs}, the Houston FSI G1 and G3 curves and the Phoenix PD curve show excellent performance; Houston FSI G2 and Hamby 44 show relatively good performance.
 Area under the curve (AUC) values, which summarize ROC curves, are shown in \autoref{tab:aucs}.
 AUC values are useful for differentiating between poor, good, and excellent model performance, but are not particularly useful when determining which of several models with approximately the same level of performance should be used \citep{marzban_roc_2004}.
-Hamby 44 and Phoenix PD are the only studies with a somewhat noticable difference between the AUC score for RF compared to ccf, and even those differences are relatively small: .03 and -.01, respectively.}
-
-\svp{XXX show ROC for bullet-to-bullet is perfect, just for emphasis?}
+Hamby 44 and Phoenix PD are the only studies with a somewhat noticable difference between the AUC score for RF compared to ccf, and even those differences are relatively small: .03 and -.01, respectively.
 
@@
 
-What is quite surprising is that \svp{while} we anticipated \svp{Hamby 44 would be the easiest set to work with, but it} turned out to be the hardest to solve.
+What is quite surprising is that while we anticipated Hamby 44 would be the easiest set to work with, it turned out to be the hardest to solve.
 This is likely due to several lands in Hamby set 44 that have major deficiencies,  such as 'tank rash' (a collision of the bullet after exiting the barrel with a surface causing markings on top of the striations from the barrel) or extreme pitting (holes caused by direct contact with burning gun powder).
@@
 
-All of the studies \svp{presented here} involve the same type of firearm.
+All of the studies presented here involve the same type of firearm.
 We see some variations in scores across different models of Rugers, and it is probable that we will see even bigger variations with different brands of firearms that vary in number of lands and land length.
@@
 The algorithm proposed by \citet{aoas} is not capable of  automatically detecting parts of lands with well expressed striae.
-\svp{It may be useful to couple the \citet{aoas} algorithm with an algorithm which assesses the quality of the input data and determines which portions of the data to use for comparison.
-This would emulate the process used by examiners, who first assess the quality of the evidence and whether there is enough information present to attempt a match.}
+It may be useful to couple the \citet{aoas} algorithm with an algorithm which assesses the quality of the input data and determines which portions of the data to use for comparison.
+This would emulate the process used by examiners, who first assess the quality of the evidence and whether there is enough information present to attempt a match.
 

### 2018-10-31 index.Rmd
preliminary

@@
 
-abstract: |
+abstract: Recent advances in microscopy have made it possible to collect 3d topographic data, enabling virtual comparisons based on the collected 3D data next to traditional comparison microscopy.
+Automatic matching algorithms have been introduced for various scenarios, such as matching cartridge cases (Tai and Eddy 2018) or matching bullet striae (Hare et al.
+2017b,  Chu et al 2013,  De Kinder and Bonfanti 1999).
+One key aspect of validating automatic matching algorithms is to evaluate the performance of the algorithm on external tests.
+Here, we are presenting a discussion of the performance of the matching algorithm (Hare et al.
+2017b) in three studies.
+We are considering matching performance based on the Random forest score, cross-correlation, and consecutive matching striae (CMS) at the land-to-land level and, using Sequential Average Maxima scores, also at the bullet-to bullet level.
+Cross-correlation and Random Forest scores both result in perfect discrimination of same-source and different-source bullets.
+At the land-to-land level, discrimination (based on area under the curve, AUC) is excellent (???
+0.90).
 
@@
 ```{r compare-land-to-land, fig.width = 8, fig.height = 4, fig.cap="Density curves of land-to-land similarity scores from cross-correlation (top) and RF scores (bottom).
-Different colors indicate same source and phase versus different source or out-of-phase matches.
-The random forest scores tend to be more extreme than the cross correlation scores alone."}
+Different colors indicate same source versus different source for each land.
+RF scores for different source comparisons are generally well below 0.5.
+Some same-source comparisons are also below 0.5 for RF scores, indicating potential problems with at least one of the lands involved in the comparison."}
 all_lands %>% 
@@
   theme_bw() +
-  scale_fill_manual("Same source & best sequence", values = c("darkgrey", "darkorange")) +
-  scale_colour_manual("Same source & best sequence", values = c("darkgrey", "darkorange")) +
+  scale_fill_manual("Same source", values = c("darkgrey", "darkorange")) +
+  scale_colour_manual("Same source", values = c("darkgrey", "darkorange")) +
   scale_x_continuous("Land-to-Land scores", limits = c(0, 1)) +
@@
 
-```{r rocs, fig.cap="ROC curves for all studies based on land-to-land comparisons \\svp{of the random forest score}.", fig.width = 4.5, fig.height = 3, out.width = ".5\\textwidth"}
+```{r rocs, fig.cap="ROC curves for all studies based on land-to-land comparisons of the random forest score.", fig.width = 6, fig.height = 10, out.width = ".5\\textwidth"}
 library(plotROC)
 library(pROC)
-all_lands %>% gather(measure, score, rfscore, ccf) %>%
+rocpic <- all_lands %>% gather(measure, score, rfscore, ccf, cms) %>%
+  mutate(
+    measure = factor(measure, labels = c("CC", "CMS", "RF"))
+  ) %>%
   ggplot(aes(linetype=measure, colour = interaction(study, group1, sep = " "))) +
@@
   theme_bw() +
-  facet_wrap(~study) +
+  facet_wrap(~study+group1, nrow=3) +
   xlab("Fraction False Positives") +
@@
   scale_colour_manual("Study", values = c("#fe9929", "#aa0000",
-                                         "#74a9cf", "#2b8cbe", "#045a8d"))
-```
-
-```{r aucs, eval = FALSE, results = 'asis'}
-
-aucs <- all_lands %>% group_by(study, group1) %>% nest() %>%
-  mutate(
-    auc_rf = data %>% purrr::map_dbl(function(d) {
-      auc(d$sameland, d$rfscore)
+                                         "#74a9cf", "#2b8cbe", "#045a8d"),
+                      guide=FALSE) +
+  theme(legend.position="bottom",
+        legend.key.width = unit(1.5, "cm")) +
+  scale_linetype_manual("Measure", values=c(1,3,2)) 
+```
+
+```{r aucs-pic, fig.width=6, fig.height = 6, out.width='.5\\textwidth', fig.cap=""}
+aucs <- all_lands %>% gather(measure, score, rfscore, ccf, cms) %>% 
+  group_by(study, group1, measure) %>% nest() %>%
+  mutate(
+    auc = data %>% purrr::map_dbl(function(d) {
+      auc(d$sameland, d$score)
     }),
-    auc_ccf = data %>% purrr::map_dbl(function(d) {
-      auc(d$sameland, d$ccf)
+    auc_ci_low = data %>% purrr::map_dbl(function(d) {
+      ci.auc(d$sameland, d$score)[1]
     }),
-    auc_cms = data %>% purrr::map_dbl(function(d) {
-      auc(d$sameland, d$cms)
-    }),
-    auc_rf_ci_low = data %>% purrr::map_dbl(function(d) {
-      ci.auc(d$sameland, d$rfscore)[1]
-    }),
-    auc_rf_ci_upper = data %>% purrr::map_dbl(function(d) {
-      ci.auc(d$sameland, d$rfscore)[3]
-    }),
-    auc_ccf_ci_lower = data %>% purrr::map_dbl(function(d) {
-      ci.auc(d$sameland, d$ccf)[1]
-    }),
-    auc_ccf_ci_upper = data %>% purrr::map_dbl(function(d) {
-      ci.auc(d$sameland, d$ccf)[3]
-    }),
-    auc_cms_ci_lower = data %>% purrr::map_dbl(function(d) {
-      ci.auc(d$sameland, d$cms)[1]
-    }),
-    auc_cms_ci_upper = data %>% purrr::map_dbl(function(d) {
-      ci.auc(d$sameland, d$cms)[3]
+    auc_ci_high = data %>% purrr::map_dbl(function(d) {
+      ci.auc(d$sameland, d$score)[3]
     })
-
-  )
-#
-library(xtable)
-names(aucs)[2] <- "set"
-names(aucs)[4] <- "RF score"
-names(aucs)[5] <- "Cross-correlation"
-names(aucs)[6] <- "CMS"
-
-aucs <- as.data.frame(aucs)
-aucs[,4] <- sprintf("%.3f (%.3f, %.3f)", aucs[,4], aucs[,7], aucs[,8])
-aucs[,5] <- sprintf("%.3f (%.3f, %.3f)", aucs[,5], aucs[,9], aucs[,10])
-aucs[,6] <- sprintf("%.3f (%.3f, %.3f)", aucs[,6], aucs[,11], aucs[,12])
-print(xtable(aucs[,1:6] %>% select(-data), digits = 3, caption = "Area under the curve (AUC) for each (set) of the studies.
-At the land level there is no difference in prediction power between RF score and cross-correlation."), include.rownames=FALSE, label="tab:aucs")
-```
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{llrrr}
-  \hline
- && \multicolumn{3}{l}{\bf AUC (95\% Conf. Int)}\\
-\bf Study & \bf Set & \bf RF score & \bf Cross-correlation & \bf CMS \\
-  \hline
-\bf Hamby 44 &  & 0.937 (0.917, 0.957) & 0.916 (0.891, 0.941) & 0.820 (0.787, 0.854) \\
-\bf  Phoenix PD &  & 0.976 (0.957, 0.994) & 0.986 (0.975, 0.996) & 0.879 (0.838, 0.919) \\
-\bf  Houston FSI & Set 1 & 0.991 (0.983, 0.999) & 0.988 (0.979, 0.997) & 0.785 (0.727, 0.842) \\
-\bf  Houston FSI & Set 2 & 0.898 (0.853, 0.943) & 0.902 (0.861, 0.943) & 0.749 (0.684, 0.814) \\
-\bf  Houston FSI & Set 3 & 0.986 (0.973, 0.999) & 0.987 (0.978, 0.997) & 0.911 (0.880, 0.941) \\
-   \hline
-\end{tabular}
-\caption{\svp{Can we turn this into a picture?
+  )
+
+aucpic <- aucs %>% 
+  mutate(
+    measure = factor(measure, levels = c("cms", "ccf", "rfscore")),
+    measure = factor(measure, labels = c("CMS", "CC", "RF"))
+  ) %>%
+  ggplot(aes(x = auc, y = measure, colour = interaction(study, group1), linetype=measure)) +
+  facet_grid(study+group1~.) +
+  geom_errorbarh(aes(xmin=auc_ci_low, xmax=auc_ci_high), size=1, height=0.3) +
+  geom_point() +
+  theme_bw() +
+  xlab("AUC and 95% confidence intervals") +
+  xlim(c(0.5, 1)) +
+  ylab("") +
+    scale_colour_manual("Study", values = c("#fe9929", "#aa0000",
+                                         "#74a9cf", "#2b8cbe", "#045a8d"),
+                        guide = FALSE) +
+  theme(legend.position = "bottom",
+        legend.key.width = unit(1.5, "cm")) +
+  scale_linetype_manual("Measure", values = c(3,1,2))
+
+```
+
+<!-- \begin{table}[ht] -->
+<!-- \centering -->
+<!-- \begin{tabular}{llrrr} -->
+<!--   \hline -->
+<!--  && \multicolumn{3}{l}{\bf AUC (95\% Conf. Int)}\\ -->
+<!-- \bf Study & \bf Set & \bf RF score & \bf Cross-correlation & \bf CMS \\ -->
+<!--   \hline -->
+<!-- \bf Hamby 44 &  & 0.937 (0.917, 0.957) & 0.916 (0.891, 0.941) & 0.820 (0.787, 0.854) \\ -->
+<!-- \bf  Phoenix PD &  & 0.976 (0.957, 0.994) & 0.986 (0.975, 0.996) & 0.879 (0.838, 0.919) \\ -->
+<!-- \bf  Houston FSI & Set 1 & 0.991 (0.983, 0.999) & 0.988 (0.979, 0.997) & 0.785 (0.727, 0.842) \\ -->
+<!-- \bf  Houston FSI & Set 2 & 0.898 (0.853, 0.943) & 0.902 (0.861, 0.943) & 0.749 (0.684, 0.814) \\ -->
+<!-- \bf  Houston FSI & Set 3 & 0.986 (0.973, 0.999) & 0.987 (0.978, 0.997) & 0.911 (0.880, 0.941) \\ -->
+<!--    \hline -->
+<!-- \end{tabular} -->
+<!-- \caption{\svp{Can we turn this into a picture?
 It's really dense and hard to interpret...}Area under the curve (AUC) for each (set) of the studies for the random forest score, cross-correlation and CMS.
@@
 As seen in the ROC curves, the algorithm performs the least well on set 2 of the Houston FSI study.
-At the land level there is no significant difference in prediction power between RF score and cross-correlation.} \label{tab:aucs}
-\end{table}
+At the land level there is no significant difference in prediction power between RF score and cross-correlation.} \label{tab:aucs} -->
+<!-- \end{table} -->
+
+```{r echo = FALSE, out.width = "0.49\\textwidth", fig.subcap=c("ROC curves", "Area under the curve (AUC)"), fig.cap="ROC curves and AUC for each (set) of the studies for the random forest score (RF), cross-correlation (CC) and consecutively matching striae (CMS).
+As seen in the ROC curves, the algorithm performs the least well on set 2 of the Houston FSI study.
+Based on AUC, the overall performance on all sets is very good to excellent for cross-correlation and the random forest score.
+At the land level there is no significant difference in prediction power between RF score and cross-correlation."}
+  rocpic
+  aucpic
+```
 
@@
 
-```{r l2l, fig.cap="Density curves of similarity scores based on land-to-land comparisons for cross-correlation (top) and RF scores (bottom).
+```{r l2l, eval=FALSE, fig.cap="Density curves of similarity scores based on land-to-land comparisons for cross-correlation (top) and RF scores (bottom).
 Different colors indicate same source versus different source.
@@
 We also see that questioned bullet 'I' is affected, explaining some of the low similarity scores for this bullet.
+\hh{None of the lands of any of the bullets in the other studies are affected by tank rash in a similar manner.}
 

### 2018-10-31 index.Rmd
Finally got the scale where I want it.

@@
 
-```{r rocs, fig.cap="ROC curves for all studies based on land-to-land comparisons of the random forest score.", fig.width = 6, fig.height = 10, out.width = ".5\\textwidth"}
+```{r rocs, fig.cap="ROC curves for all studies based on land-to-land comparisons of the random forest score."}
 library(plotROC)
@@
   mutate(
-    measure = factor(measure, labels = c("CC", "CMS", "RF"))
+    measure = factor(measure, labels = c("CC", "CMS", "RF")),
+    studyg = paste(study, group1) %>% str_trim %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"))
   ) %>%
-  ggplot(aes(linetype=measure, colour = interaction(study, group1, sep = " "))) +
+  ggplot(aes(linetype = measure, colour = studyg)) +
   geom_roc(aes(m = score, d = as.numeric(sameland)), 
-           labels =FALSE, pointsize = 0, alpha = 0.8) +
+           labels = FALSE, pointsize = 0, alpha = 0.8, size = .75) +
   theme_bw() +
-  facet_wrap(~study+group1, nrow=3) +
+  # coord_fixed() + 
+  facet_wrap(~studyg, nrow=3) +
   xlab("Fraction False Positives") +
@@
                       guide=FALSE) +
-  theme(legend.position="bottom",
-        legend.key.width = unit(1.5, "cm")) +
+  theme(legend.position= c(1, 0), legend.justification = c(1, 0), legend.direction = "vertical",
+        legend.key.width = unit(1, "cm")) +
+  guides(linetype = guide_legend(override.aes = list(shape = NA, size = .5))) + 
   scale_linetype_manual("Measure", values=c(1,3,2)) 
@@
 
-```{r aucs-pic, fig.width=6, fig.height = 6, out.width='.5\\textwidth', fig.cap=""}
+```{r aucs-pic, fig.cap=""}
 aucs <- all_lands %>% gather(measure, score, rfscore, ccf, cms) %>% 
@@
     })
+  ) %>% 
+  mutate(
+    studyg = paste(study, group1) %>% str_trim %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3")),
+    measure = factor(measure, levels = c("cms", "ccf", "rfscore"), labels = c("CMS", "CC", "RF")),
+    measurenum = as.numeric(measure),
+    measurechar = as.character(measure)
   )
@@
 aucpic <- aucs %>%
-  mutate(
-    measure = factor(measure, levels = c("cms", "ccf", "rfscore")),
-    measure = factor(measure, labels = c("CMS", "CC", "RF"))
-  ) %>%
-  ggplot(aes(x = auc, y = measure, colour = interaction(study, group1), linetype=measure)) +
-  facet_grid(study+group1~.) +
-  geom_errorbarh(aes(xmin=auc_ci_low, xmax=auc_ci_high), size=1, height=0.3) +
+  ggplot(aes(x = auc, y = measurenum, colour = studyg)) +
+  facet_grid(studyg~.) +
+  geom_segment(aes(x = auc_ci_low, xend = auc_ci_low, y = measurenum - .125, yend = measurenum + .125), size = 1) + 
+  geom_segment(aes(x = auc_ci_high, xend = auc_ci_high, y = measurenum - .125, yend = measurenum + .125), size = 1) + 
+  geom_segment(aes(x=auc_ci_low, xend=auc_ci_high, y = measurenum, yend = measurenum, linetype=measure), size=1) +
   geom_point() +
@@
   xlim(c(0.5, 1)) +
-  ylab("") +
+  scale_y_continuous("", breaks = unique(aucs$measurenum), labels = unique(aucs$measurechar)) + 
   scale_colour_manual("Study", values = c("#fe9929", "#aa0000",
@@
                         guide = FALSE) +
-  theme(legend.position = "bottom",
-        legend.key.width = unit(1.5, "cm")) +
-  scale_linetype_manual("Measure", values = c(3,1,2))
+  # theme(legend.position = "bottom",
+  #       legend.key.width = unit(1.5, "cm")) +
+  scale_linetype_manual("Measure", values = c(3,1,2), guide = F)
 
@@
 
-```{r echo = FALSE, out.width = "0.49\\textwidth", fig.subcap=c("ROC curves", "Area under the curve (AUC)"), fig.cap="ROC curves and AUC for each (set) of the studies for the random forest score (RF), cross-correlation (CC) and consecutively matching striae (CMS).
+```{r echo = FALSE, out.width = c("0.49\\textwidth", ".45\\textwidth"), out.height = c("0.6\\textwidth", "0.6\\textwidth"), fig.width = c(5, 6), fig.height = c(7.5, 7.5), fig.subcap=c("ROC curves", "Area under the curve (AUC)"), fig.cap="ROC curves and AUC for each (set) of the studies for the random forest score (RF), cross-correlation (CC) and consecutively matching striae (CMS).
 As seen in the ROC curves, the algorithm performs the least well on set 2 of the Houston FSI study.

### 2018-10-31 index.Rmd
fix refs

@@
 
-```{r echo = FALSE, out.width = c("0.49\\textwidth", ".45\\textwidth"), out.height = c("0.6\\textwidth", "0.6\\textwidth"), fig.width = c(5, 6), fig.height = c(7.5, 7.5), fig.subcap=c("ROC curves", "Area under the curve (AUC)"), fig.cap="ROC curves and AUC for each (set) of the studies for the random forest score (RF), cross-correlation (CC) and consecutively matching striae (CMS).
+```{r roc-auc, echo = FALSE, out.width = c("0.49\\textwidth", ".45\\textwidth"), out.height = c("0.6\\textwidth", "0.6\\textwidth"), fig.width = c(5, 6), fig.height = c(7.5, 7.5), fig.subcap=c("ROC curves", "Area under the curve (AUC)"), fig.cap="ROC curves and AUC for each (set) of the studies for the random forest score (RF), cross-correlation (CC) and consecutively matching striae (CMS).
 As seen in the ROC curves, the algorithm performs the least well on set 2 of the Houston FSI study.
@@
 
-In \autoref{fig:rocs}, the Houston FSI G1 and G3 curves and the Phoenix PD curve show excellent performance; Houston FSI G2 and Hamby 44 show relatively good performance.
-Area under the curve (AUC) values, which summarize ROC curves, are shown in \autoref{tab:aucs}.
+In \autoref{fig:roc-auc} (a), the Houston FSI G1 and G3 curves and the Phoenix PD curve show excellent performance; Houston FSI G2 and Hamby 44 show relatively good performance.
+Area under the curve (AUC) values, which summarize ROC curves, are shown in \autoref{fig:roc-auc} (b).
 AUC values are useful for differentiating between poor, good, and excellent model performance, but are not particularly useful when determining which of several models with approximately the same level of performance should be used \citep{marzban_roc_2004}.
-Hamby 44 and Phoenix PD are the only studies with a somewhat noticable difference between the AUC score for RF compared to ccf, and even those differences are relatively small: .03 and -.01, respectively.
+Hamby 44 and Phoenix PD are the only studies with a somewhat noticable difference between the AUC score for RF compared to ccf, and even those differences are not statistically significant.
 

### 2018-11-01 index.Rmd
restructuring ideas

@@
 \svp{In addition, there is significantly more information to be gained from quantitative assessment of peak height and valley depth than from the qualitative assessment of the same information.}
+
+## Identification Process
 
@@
 We also include statistical features, such as cross-correlation and Euclidean distance between the signatures.
+
+## Explain RF score, CC and CMS
+
+\hh{We already have a start on this above, but we should focus on these three measures in more detail.}
+
+\hh{In the result section we should probably also include a discussion of EER and the threshold for EER for the three measures.
+Another threshold of interest is CMS=6, because that's what FTEs use often.
+}
 
@@
   geom_density(alpha = 0.6) +
-  geom_rug(alpha = 0.5/12, show.legend = F) +
+#  geom_rug(alpha = 0.5/12, show.legend = F) +
   facet_grid(measure ~ study) +
@@
   scale_colour_manual("Same source", values = c("darkgrey", "darkorange")) +
-  scale_x_continuous("Land-to-Land scores", limits = c(0, 1)) +
+  scale_x_continuous("Land-to-Land scores", limits = c(0, 1), labels = c(0, 0.25, 0.5, 0.75, 1)) +
   theme(legend.position = "bottom") +
@@
 
-<!-- \begin{table}[ht] -->
-<!-- \centering -->
-<!-- \begin{tabular}{llrrr} -->
-<!--   \hline -->
-<!--  && \multicolumn{3}{l}{\bf AUC (95\% Conf. Int)}\\ -->
-<!-- \bf Study & \bf Set & \bf RF score & \bf Cross-correlation & \bf CMS \\ -->
-<!--   \hline -->
-<!-- \bf Hamby 44 &  & 0.937 (0.917, 0.957) & 0.916 (0.891, 0.941) & 0.820 (0.787, 0.854) \\ -->
-<!-- \bf  Phoenix PD &  & 0.976 (0.957, 0.994) & 0.986 (0.975, 0.996) & 0.879 (0.838, 0.919) \\ -->
-<!-- \bf  Houston FSI & Set 1 & 0.991 (0.983, 0.999) & 0.988 (0.979, 0.997) & 0.785 (0.727, 0.842) \\ -->
-<!-- \bf  Houston FSI & Set 2 & 0.898 (0.853, 0.943) & 0.902 (0.861, 0.943) & 0.749 (0.684, 0.814) \\ -->
-<!-- \bf  Houston FSI & Set 3 & 0.986 (0.973, 0.999) & 0.987 (0.978, 0.997) & 0.911 (0.880, 0.941) \\ -->
-<!--    \hline -->
-<!-- \end{tabular} -->
-<!-- \caption{\svp{Can we turn this into a picture?
-It's really dense and hard to interpret...}Area under the curve (AUC) for each (set) of the studies for the random forest score, cross-correlation and CMS.
-Overall the perform on all sets is very good to excellent for cross-correlation and the random forest score.
-As seen in the ROC curves, the algorithm performs the least well on set 2 of the Houston FSI study.
-At the land level there is no significant difference in prediction power between RF score and cross-correlation.} \label{tab:aucs} -->
-<!-- \end{table} -->
-
 ```{r roc-auc, echo = FALSE, out.width = c("0.49\\textwidth", ".45\\textwidth"), out.height = c("0.6\\textwidth", "0.6\\textwidth"), fig.width = c(5, 6), fig.height = c(7.5, 7.5), fig.subcap=c("ROC curves", "Area under the curve (AUC)"), fig.cap="ROC curves and AUC for each (set) of the studies for the random forest score (RF), cross-correlation (CC) and consecutively matching striae (CMS).

### 2018-11-02 index.Rmd
Comment, add barrel info to the phoenix description

@@
 \item {\bf Phoenix PD}
-Tylor Klep from Phoenix PD provided sets of known test fires and questioned bullets: the set of known bullets consists of three test fires (B1, B2, B3) from each of eight different barrels (A9, C8, F6, L5, M2, P7, R3, U10).
+Tylor Klep from Phoenix PD provided sets of known test fires and questioned bullets: the set of known bullets consists of three test fires (B1, B2, B3) from each of eight different Ruger P-95 barrels (A9, C8, F6, L5, M2, P7, R3, U10).
 Ten questioned bullets were provided (B, E, H, J, K N, Q, T, Y, Z).
 This set is an open set; that is,\ it is not known in advance whether all (or any) of the questioned bullets are fired from the known barrels.
-
+\svp{Need to add ammo info here...}
 Land engraved areas for each of the six lands of each bullet were scanned by Bill Henderson (Sensofar).
@@
 
-\svp{When evaluating classifier performance, it is common to use a receiver operating curve, which plots the false positives against the true positives for each possible value of the cutoff between the two classes (in this case, known match and known non-match).
-The ROC curves for the land-to-land scores are shown in \autoref{fig:rocs}.
+When evaluating classifier performance, it is common to use a receiver operating curve, which plots the false positives against the true positives for each possible value of the cutoff between the two classes (in this case, known match and known non-match).
+The ROC curves for the land-to-land scores are shown in \autoref{fig:roc-auc} (a).
 A perfect classifier would have 100\% true positives and 0\% false positives, e.g.
@@
 Classifiers with better performance will be closer to this corner of the plot.
-A random classifier would have an ROC curve that was a straight diagonal line through (0,0) and (1,1).}
+A random classifier would have an ROC curve that was a straight diagonal line through (0,0) and (1,1).
 

### 2018-11-04 index.Rmd
Some explanation of features. Need better CCF stuff...

@@
 
-## Explain RF score, CC and CMS
+## Statistics for Matching Aligned Signatures
+
+\svp{Firearms examiners often use consecutive matching striae (CMS) to describe the strength of a match.
+CMS as used by examiners counts only peaks as striae, as it is difficult to assess vallleys and peaks separately when using a comparison scope.
+If there are at least 6 consecutive matching striae, the bullets are considered to be a good match \citep{biasotti_firearms_1997,nichols_scientific_2006}, according to criteria used by FTEs.
+There are some differences between CMS as defined by FTEs and the method used in \citet{asos} includes both peaks and valleys, as it is easier to assess the features separately using cross-sections.
+XXX Should we go into this, or just leave it as is?
+}
+
+\svp{Cross-correlation between two signatures $S_1(x)$ and $S_2(x)$, $$CC(S_1, S_2, z) = \lim_{L\rightarrow\infty}\left(\frac{1}{L} \int_{-L/2}^{L/2} S_1(x)S_2(x + z) dx \right)/[R_1R_2]$$ where $z$ is the shift difference between the two sequences and $R_1$ and $R_2$ are the root mean square roughness for the two signature \citep{vorburgerApplicationsCrosscorrelationFunctions2011}.
+We maximize $CC$ over values of $z$ to determine the best alignment; this maximal value is used to assess the match strength between the two sequences.
+XXX algorithm for calculating with missing values?}
+
+\svp{XXX No idea what to say about the RF other than what's above?}
 

### 2018-11-05 index.Rmd
eer

@@
 ```{r aucs-pic, fig.cap=""}
+eer <- function(roc){
+  dframe <- data.frame(sens = roc$sensitivities, 
+                       spec = roc$specificities,
+                       threshold = roc$thresholds)
+  dframe$err1 <- 1 - dframe$sens
+  dframe$err2 <- 1 - dframe$spec
+  dframe$diff <- abs(dframe$err1-dframe$err2)
+  dframe$eer <- (dframe$err1+dframe$err2)/2
+  idx <- which.min(dframe$diff)
+  dframe[idx,c("eer", "threshold")]
+}
+
 aucs <- all_lands %>% gather(measure, score, rfscore, ccf, cms) %>% 
@@
       ci.auc(d$sameland, d$score)[3]
-    })
+    }),
+    roc = data %>% purrr::map(function(d) {
+      roc(d$sameland, d$score)
+    }),
+    errors = roc %>% purrr::map(.f = eer)
   ) %>% 
@@
   )
+aucs <- aucs %>% unnest(errors)
 
@@
 
+\hh{\autoref{fig:eer} shows an overview of Equal Error Rates (EER) based on the ROC curves of figure \ref{fig:roc-auc}.
+EERs are error rates when sensitivity and specificity of a test are equal, i.e.
+we see the same error rates of missing an identification for a land-to-land comparison  and wrong identifications, when it is determined that two land engraved areas come from the same source when they are, in fact, known to be from a different source.
+Equal error rates based on CMS of two lands are significantly higher than equal error rates based on cross-correlation or the RF score.
+At best, RF score and cross-correlation have an equal error rate of around 5\% for land-to-land comparisons.
+Interestingly, for consecutively matching striae, an equal error rate is reached for just two matching striae, which from a practical point of view is not sustainable.
+A  CMS of  6 or above is an established cutoff used in identifications.
+This leads to 0 wrong identifications but only identifies 21 out of  715 known same-source lands.
+XXX check after file names are corrected.
+}
+```{r eval=FALSE}
+xtabs(~sameland+I(cms>=12), data = all_lands)
+```
+
+```{r eer, fig.cap="Equal Error Rates for all measures (RF score, Cross-correlation and CMS) under all five (sub-)studies.
+CMS is signifcantly worse than both Cross-correlation and the RF score.
+At best, Cross-correlation and RF score have errors for land-to-land comparison of around 5%.", out.width = '.5\\textwidth'}
+aucs %>% ggplot(aes(y = studyg, x = eer, colour=measure, shape = measure)) + 
+  geom_point(size = 3) +theme_bw() +
+  xlim(c(0, .35)) +
+  xlab("Equal Error Rates (EER)") +
+  ylab("") +
+  scale_color_brewer(palette = "Set2")
+```
+
 # Discussion and Conclusions

### 2018-11-06 index.Rmd
restored cms to integer count of number of peaks

@@
 
-\svp{Firearms examiners often use consecutive matching striae (CMS) to describe the strength of a match.
-CMS as used by examiners counts only peaks as striae, as it is difficult to assess vallleys and peaks separately when using a comparison scope.
+\hh{XXX define signatures mathematically here - as random spatial process; extracted from crosscut by removing big structure first and then applying (cross-validated, \citet{aoas}) 'roughness' filter.
+'roughness' defined by NIST paper.
+}
+
+One of the measures used by firearms examiners to describe the strength of a match quantitatively is the number of consecutive matching striae (CMS) as defined by \citet{biasotti}.
 If there are at least 6 consecutive matching striae, the bullets are considered to be a good match \citep{biasotti_firearms_1997,nichols_scientific_2006}, according to criteria used by FTEs. 
-There are some differences between CMS as defined by FTEs and the method used in \citet{asos} includes both peaks and valleys, as it is easier to assess the features separately using cross-sections.
-XXX Should we go into this, or just leave it as is?
-}
+CMS in this setting are counted as the number of consecutive peaks in signatures of two lands that match.
+
+\hh{HH: let's not go into what we do not do, and instead focus on what we do}
+~~  There are some differences between CMS as defined by FTEs and the method used in \citet{aoas} includes both peaks and valleys, as it is easier to assess the features separately using cross-sections.
+XXX Should we go into this, or just leave it as is?~~
 
@@
 
-\svp{XXX No idea what to say about the RF other than what's above?}
+\svp{XXX No idea what to say about the RF other than what's above? } 
+\hh{XXX that's what I thought, too.}
 
 \hh{We already have a start on this above, but we should focus on these three measures in more detail.}
-
-\hh{In the result section we should probably also include a discussion of EER and the threshold for EER for the three measures.
-Another threshold of interest is CMS=6, because that's what FTEs use often.
-}
 
@@
 h44features <- read.csv("data/h44-features.csv")
+h44features <- h44features %>% mutate(
+  lys = overlap*signature_length*1000/.645,
+  cms = cms2 *lys*.645/1000
+)
 # tank rash lands:
@@
 features <- read.csv("data/pd-features.csv.gz")
-
+features <- features %>% mutate(
+  lys = overlap*signature_length*1000/.645,
+  cms = cms2 *lys*.645/1000
+)
 f2 <- features %>%
@@
 features <- read.csv("data/fsi-features.csv.gz")
+features <- features %>% mutate(
+  lys = overlap*signature_length*1000/.645,
+  cms = cms2 *lys*.645/1000
+)
 

### 2018-11-06 index.Rmd
percentage sign messed up formatting

@@
 
+\svp{Cross-correlation between two signatures $S_1(x)$ and $S_2(x)$, $$CC(S_1, S_2, z) = \lim_{L\rightarrow\infty}\left(\frac{1}{L} \int_{-L/2}^{L/2} S_1(x)S_2(x + z) dx \right)/[R_1R_2]$$ where $z$ is the shift difference between the two sequences and $R_1$ and $R_2$ are the root mean square roughness for the two signature \citep{vorburgerApplicationsCrosscorrelationFunctions2011}.
+We maximize $CC$ over values of $z$ to determine the best alignment; this maximal value is used to assess the match strength between the two sequences.
+XXX algorithm for calculating with missing values?}
+
+\hh{maximum cross-correlation is determined for aligning two signatures [@aoas, @vorburgerApplicationsCrosscorrelationFunctions2011]}
+
 One of the measures used by firearms examiners to describe the strength of a match quantitatively is the number of consecutive matching striae (CMS) as defined by \citet{biasotti}.
@@
 CMS in this setting are counted as the number of consecutive peaks in signatures of two lands that match. 
-
-\hh{HH: let's not go into what we do not do, and instead focus on what we do}
-~~  There are some differences between CMS as defined by FTEs and the method used in \citet{aoas} includes both peaks and valleys, as it is easier to assess the features separately using cross-sections.
-XXX Should we go into this, or just leave it as is?~~
-
-\svp{Cross-correlation between two signatures $S_1(x)$ and $S_2(x)$, $$CC(S_1, S_2, z) = \lim_{L\rightarrow\infty}\left(\frac{1}{L} \int_{-L/2}^{L/2} S_1(x)S_2(x + z) dx \right)/[R_1R_2]$$ where $z$ is the shift difference between the two sequences and $R_1$ and $R_2$ are the root mean square roughness for the two signature \citep{vorburgerApplicationsCrosscorrelationFunctions2011}.
-We maximize $CC$ over values of $z$ to determine the best alignment; this maximal value is used to assess the match strength between the two sequences.
-XXX algorithm for calculating with missing values?}
+\hh{Note that because signatures are moved into phase based on the maximum cross-correlation rather than matching striation marks. This can result in a consecutive matching striae of 0. }
 
@@
 CMS is signifcantly worse than both Cross-correlation and the RF score.
-At best, Cross-correlation and RF score have errors for land-to-land comparison of around 5%.", out.width = '.5\\textwidth'}
+At best, Cross-correlation and RF score have errors for land-to-land comparison of around 5\\%.", out.width = '.5\\textwidth'}
 aucs %>% ggplot(aes(y = studyg, x = eer, colour=measure, shape = measure)) + 

### 2018-11-06 index.Rmd
fixed file error

@@
 )
+# change labels for land 3 and land 4 of bullet B3 in P7
+features <- features %>% mutate(
+  first = as.character(first),
+  first = replace(first, first=="Gun 1-P7/B3/L4.dat", "XXX"),
+  first = replace(first, first=="Gun 1-P7/B3/L3.dat", "Gun 1-P7/B3/L4.dat"),
+  first = replace(first, first=="XXX", "Gun 1-P7/B3/L3.dat")
+)
+
+features <- features %>% mutate(
+  KM = replace(KM, first %in% c("Gun 1-P7/B3/L3.dat", "Gun 1-P7/B3/L4.dat"), FALSE)
+)
+features <- features %>% mutate(
+  KM = ifelse(first == "Gun 1-P7/B3/L3.dat", (second == "Unknown 1-J/L3.dat"), KM),
+  KM = ifelse(first == "Gun 1-P7/B3/L4.dat", (second == "Unknown 1-J/L4.dat"), KM)
+)
+
 f2 <- features %>%
@@
 )
+
 pdnest <- f3 %>% group_by(bullet2, barrel1, bullet1) %>% nest()
@@
 
-In \autoref{fig:roc-auc} (a), the Houston FSI G1 and G3 curves and the Phoenix PD curve show excellent performance; Houston FSI G2 and Hamby 44 show relatively good performance.
+In \autoref{fig:roc-auc} (a), the Houston FSI G1 and G3 curves and the Phoenix PD curve show excellent performance; Houston FSI G2 and Hamby 44 show still a very  good performance.
 Area under the curve (AUC) values, which summarize ROC curves, are shown in \autoref{fig:roc-auc} (b).
 AUC values are useful for differentiating between poor, good, and excellent model performance, but are not particularly useful when determining which of several models with approximately the same level of performance should be used \citep{marzban_roc_2004}.
-Hamby 44 and Phoenix PD are the only studies with a somewhat noticable difference between the AUC score for RF compared to ccf, and even those differences are not statistically significant.
+Hamby 44 and Phoenix PD are the only studies with a somewhat noticable difference between the AUC score for RF compared to cross-correlation, but even those differences are not statistically significant.
 
@@
 A  CMS of  6 or above is an established cutoff used in identifications.
-This leads to 0 wrong identifications but only identifies 21 out of  715 known same-source lands.
+This leads to 16 wrong identifications (out of 49286; for an error rate of about 3 in 10,000) but only identifies 115 out of  715 known same-source lands, for an error rate of 16 in 100.
 XXX check after file names are corrected.
@@
 ```{r eval=FALSE}
-xtabs(~sameland+I(cms>=12), data = all_lands)
+xtabs(~sameland+I(cms>=6), data = all_lands)
 ```

### 2018-11-06 index.Rmd
Slight fixes, affiliations, etc.

--- 2018-11-06 index.Rmd
@@
 authors: 
-- name: Author 1
-  thanks: The authors gratefully acknowledge XXX NIST funding for CSAFE XXX access to bullets from Hamby set 44, XXX Tylor, Melissa and Kasi. Scanning. Hamby advice. 
-  affiliation: Department of YYY, University of XXX
-
-- name: Author 2
-  affiliation: Department of ZZZ, University of WWW
+- name: Heike Hofmann
+  thanks: This work was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.
+The authors wish to thank Jim Hamby, Tylor Klep (Phoenix PD), and Melissa Nally and Kasi Kirksey (FSI Houston) for access to test set bullets.
+XXX Scanning.
+Hamby advice.
+  affiliation: Department of Statistics, Iowa State University
+
+- name: Susan VanderPlas
+  affiliation: Department of Statistics, Iowa State University
 
@@
 
-\hh{XXX define signatures mathematically here - as random spatial process; extracted from crosscut by removing big structure first and then applying (cross-validated, \citet{aoas}) 'roughness' filter.
-'roughness' defined by NIST paper.
+\svp{Firearms examiners often use consecutive matching striae (CMS) to describe the strength of a match.
+CMS as used by examiners counts only peaks as striae, as it is difficult to assess vallleys and peaks separately when using a comparison scope.
+If there are at least 6 consecutive matching striae, the bullets are considered to be a good match \citep{biasotti_firearms_1997,nichols_scientific_2006}, according to criteria used by FTEs.
+There are some differences between CMS as defined by FTEs and the method used in \citet{aoas}.
+The computational method includes counts both peaks and valleys as striae, as it is easier to assess the features separately using cross-sections than it is to do so usingn visual examination.
+XXX Should we go into this, or just leave it as is?
 }
@@
 
-\hh{maximum cross-correlation is determined for aligning two signatures [@aoas, @vorburgerApplicationsCrosscorrelationFunctions2011]}
-
-One of the measures used by firearms examiners to describe the strength of a match quantitatively is the number of consecutive matching striae (CMS) as defined by \citet{biasotti}.
-If there are at least 6 consecutive matching striae, the bullets are considered to be a good match \citep{biasotti_firearms_1997,nichols_scientific_2006}, according to criteria used by FTEs. 
-CMS in this setting are counted as the number of consecutive peaks in signatures of two lands that match. 
-\hh{Note that because signatures are moved into phase based on the maximum cross-correlation rather than matching striation marks. This can result in a consecutive matching striae of 0. }
-
-\svp{XXX No idea what to say about the RF other than what's above? } 
-\hh{XXX that's what I thought, too.}
+\svp{XXX No idea what to say about the RF other than what's above?}
 
 \hh{We already have a start on this above, but we should focus on these three measures in more detail.}
+
+\hh{In the result section we should probably also include a discussion of EER and the threshold for EER for the three measures.
+Another threshold of interest is CMS=6, because that's what FTEs use often.
+}
 
@@
 h44features <- read.csv("data/h44-features.csv")
-h44features <- h44features %>% mutate(
-  lys = overlap*signature_length*1000/.645,
-  cms = cms2 *lys*.645/1000
-)
 # tank rash lands:
@@
 features <- read.csv("data/pd-features.csv.gz")
-features <- features %>% mutate(
-  lys = overlap*signature_length*1000/.645,
-  cms = cms2 *lys*.645/1000
-)
+
 f2 <- features %>%
@@
 features <- read.csv("data/fsi-features.csv.gz")
-features <- features %>% mutate(
-  lys = overlap*signature_length*1000/.645,
-  cms = cms2 *lys*.645/1000
-)
 
@@
 ```{r aucs-pic, fig.cap=""}
-eer <- function(roc){
-  dframe <- data.frame(sens = roc$sensitivities, 
-                       spec = roc$specificities,
-                       threshold = roc$thresholds)
-  dframe$err1 <- 1 - dframe$sens
-  dframe$err2 <- 1 - dframe$spec
-  dframe$diff <- abs(dframe$err1-dframe$err2)
-  dframe$eer <- (dframe$err1+dframe$err2)/2
-  idx <- which.min(dframe$diff)
-  dframe[idx,c("eer", "threshold")]
-}
-
 aucs <- all_lands %>% gather(measure, score, rfscore, ccf, cms) %>% 
@@
       ci.auc(d$sameland, d$score)[3]
-    }),
-    roc = data %>% purrr::map(function(d) {
-      roc(d$sameland, d$score)
-    }),
-    errors = roc %>% purrr::map(.f = eer)
+    })
   ) %>% 
@@
   )
-aucs <- aucs %>% unnest(errors)
 
@@
 
-\hh{\autoref{fig:eer} shows an overview of Equal Error Rates (EER) based on the ROC curves of figure \ref{fig:roc-auc}.
-EERs are error rates when sensitivity and specificity of a test are equal, i.e.
-we see the same error rates of missing an identification for a land-to-land comparison  and wrong identifications, when it is determined that two land engraved areas come from the same source when they are, in fact, known to be from a different source.
-Equal error rates based on CMS of two lands are significantly higher than equal error rates based on cross-correlation or the RF score.
-At best, RF score and cross-correlation have an equal error rate of around 5\% for land-to-land comparisons.
-Interestingly, for consecutively matching striae, an equal error rate is reached for just two matching striae, which from a practical point of view is not sustainable.
-A  CMS of  6 or above is an established cutoff used in identifications.
-This leads to 0 wrong identifications but only identifies 21 out of  715 known same-source lands.
-XXX check after file names are corrected.
-}
-```{r eval=FALSE}
-xtabs(~sameland+I(cms>=12), data = all_lands)
-```
-
-```{r eer, fig.cap="Equal Error Rates for all measures (RF score, Cross-correlation and CMS) under all five (sub-)studies.
-CMS is signifcantly worse than both Cross-correlation and the RF score.
-At best, Cross-correlation and RF score have errors for land-to-land comparison of around 5\\%.", out.width = '.5\\textwidth'}
-aucs %>% ggplot(aes(y = studyg, x = eer, colour=measure, shape = measure)) + 
-  geom_point(size = 3) +theme_bw() +
-  xlim(c(0, .35)) +
-  xlab("Equal Error Rates (EER)") +
-  ylab("") +
-  scale_color_brewer(palette = "Set2")
-```
-
 # Discussion and Conclusions

### 2018-11-06 index.Rmd
starting clean up

--- 2018-11-06 index.Rmd
@@
 \noindent
+
+\hh{
+Cleanup TO-DO
+\begin{itemize}
+\item define false positives and false negatives once and stick to those definitions.
+We should avoid false positive and false negatives, because that only makes sense in a test setting, which we have not defined formally.
+Let's use instead wrong identification and missed identification. 
+\item color: let's not mention 'orange' anywhere in the manuscript - I don't know, whether the printed version of the paper will be in color. 
+\end{itemize}
+}
 
@@
 
-In \autoref{fig:roc-auc} (a), the Houston FSI G1 and G3 curves and the Phoenix PD curve show excellent performance; Houston FSI G2 and Hamby 44 show still a very  good performance.
-Area under the curve (AUC) values, which summarize ROC curves, are shown in \autoref{fig:roc-auc} (b).
+In \autoref{fig:roc-auc}a, the Houston FSI G1 and G3 curves and the Phoenix PD curve show excellent performance; Houston FSI G2 and Hamby 44 show still a very  good performance.
+Area under the curve (AUC) values, which summarize ROC curves, are shown in \autoref{fig:roc-auc}b.
 AUC values are useful for differentiating between poor, good, and excellent model performance, but are not particularly useful when determining which of several models with approximately the same level of performance should be used \citep{marzban_roc_2004}.
@@
 
-\hh{\autoref{fig:eer} shows an overview of Equal Error Rates (EER) based on the ROC curves of figure \ref{fig:roc-auc}.
+\hh{\autoref{fig:eer} shows an overview of Equal Error Rates (EER) and their thresholds based on the ROC curves of \autoref{fig:roc-auc}.
 EERs are error rates when sensitivity and specificity of a test are equal, i.e.
@@
 A  CMS of  6 or above is an established cutoff used in identifications.
-This leads to 16 wrong identifications (out of 49286; for an error rate of about 3 in 10,000) but only identifies 115 out of  715 known same-source lands, for an error rate of 16 in 100.
-XXX check after file names are corrected.
+This leads to 21 wrong identifications (out of 49286; for an error rate of 4 in 10,000) but only identifies 139 out of  715 known same-source lands, for an error rate of 80.6\%.
 }
 ```{r eval=FALSE}
-xtabs(~sameland+I(cms>=6), data = all_lands)
-```
-
-```{r eer, fig.cap="Equal Error Rates for all measures (RF score, Cross-correlation and CMS) under all five (sub-)studies.
+xtabs(~sameland+I(round(cms)>=6), data = all_lands)
+```
+
+```{r eer, fig.width=7, fig.height = 4, fig.cap="Equal Error Rates and thresholds for all measures (RF score, Cross-correlation and CMS) under all five (sub-)studies.
+CMS has an EER threshold of 2 for all studies.
 CMS is signifcantly worse than both Cross-correlation and the RF score.
-At best, Cross-correlation and RF score have errors for land-to-land comparison of around 5\\%.", out.width = '.5\\textwidth'}
-aucs %>% ggplot(aes(y = studyg, x = eer, colour=measure, shape = measure)) + 
+At best, Cross-correlation and RF score have errors for land-to-land comparison of around 5\\%.", out.width = '.8\\textwidth'}
+aucs %>% gather(eer_thresh, value, eer, threshold) %>%
+  filter(!(measure=="CMS" & eer_thresh=="threshold")) %>%
+  mutate(
+    eer_thresh = factor(eer_thresh, labels=c("EER", "Threshold for EER"))
+  ) %>%
+  ggplot(aes(y = studyg, x = value, colour=measure, shape = measure)) + 
   geom_point(size = 3) +theme_bw() +
-  xlim(c(0, .35)) +
-  xlab("Equal Error Rates (EER)") +
+#  xlim(c(0, .35)) +
+  xlab("Equal Error Rates (left) and EER thresholds (right)") +
   ylab("") +
-  scale_color_brewer(palette = "Set2")
+  scale_color_brewer(palette = "Set2") +
+  theme(legend.position = "bottom") +
+  facet_grid(.~eer_thresh, scales="free") +
+  xlim(c(0, NA))
 ```

### 2018-11-06 index.Rmd
light editing

@@
 ```
-# Idea of the paper
-
-In \citet{aoas} an algorithm has been laid out to quantify the similarity between land engraved areas of bullets. 
-In this paper we present results of the algorithm on several case studies where ground truth is known.
 
@@
 In current practice, firearms and toolmark examiners (FTE) evaluate the similarity of striae on bullets by placing the evidence together with a bullet from the suspected gun under a comparison microscope.
-They then visually classify similarity according to the theory of firearms identification \citep{identification} as one of identification, inconclusive or exclusion.
-Exact guidelines for this classification vary from lab to lab, e.g.\ some labs will exclude only on the basis of non-matching class characteristics (such as direction of the twist in rifling, land length or number of lands, or type of rifling).
+XXX this could be a second bullet from the crime scene XXX They then visually classify similarity according to the theory of firearms identification \citep{identification} as one of identification, inconclusive or exclusion.
+Exact guidelines for this classification vary from lab to lab, e.g.\ some labs will exclude only on the basis of non-matching class characteristics, such as direction of the twist in rifling, land length or number of lands, or type of rifling.
 In some labs,  CMS (consecutively matching striae) as defined by \citet{biasotti} is used as a measure to quantify the strength of a match.
@@
 ## 3D Imaging Technology
-More explicit characterization of bullet surfaces has been discussed since at least 1958 (\citeauthor{davis1968introduction}), but the technology was not sufficiently advanced for analysis based on 3D measurements to be a viable option.
+More explicit characterization of bullet surfaces has been discussed since at least 1958 (\citeauthor{davis1968introduction}), but at the time technology was not sufficiently advanced to make an analysis based on 3D measurements a viable option.
 \svp{Use of a scanning electron microscope to quantitatively examine and compare bullet striae was demonstrated in \citet{ComputerIdentificationBullets1978}}.
@@
 These approaches utilize 3D surface measurements directly to characterize the topology of land engraved areas (LEAs), rather than using visual or image comparison techniques.
-In many cases, they also provide some level of automation of the comparison process, with the goal of reducing human biases by augmenting the visual information with 3D measurements.
+In many cases, these papers also provide some level of automation of the comparison process, with the goal of reducing human biases by augmenting the visual information with 3D measurements.
 Utilizing the 3D measurements allows for examination not only of the peaks in the bullet signature, but also for comparison of the valleys, which are more difficult to assess visually except in relation to the peaks.
-\svp{In addition, there is significantly more information to be gained from quantitative assessment of peak height and valley depth than from the qualitative assessment of the same information.}
+\svp{In addition, there is  more information to be gained from quantitative assessment of peak height and valley depth than from the qualitative assessment of the same information.}
+
+\hh{XXX strictly speaking, a comparison microscope is not a qualitative assessment}
 

### 2018-11-09 index.Rmd

@@
 ## Statistics for Matching Aligned Signatures
-
 \hh{XXX define signatures mathematically here - as random spatial process; extracted from crosscut by removing big structure first and then applying (cross-validated, \citet{aoas}) 'roughness' filter.

### 2018-11-09 index.Rmd
fix qualitatively and crosscorrelation definition

@@
 Utilizing the 3D measurements allows for examination not only of the peaks in the bullet signature, but also for comparison of the valleys, which are more difficult to assess visually except in relation to the peaks.
-\svp{In addition, there is  more information to be gained from quantitative assessment of peak height and valley depth than from the qualitative assessment of the same information.}
-
-\hh{XXX strictly speaking, a comparison microscope is not a qualitative assessment}
 
@@
 
-Commonly, approaches derived from 3D surface measurements use some features which are \svp{similar to those qualitatively analyzed by FTEs}\citep{luAutomatedBulletIdentification2014}.
+Commonly, approaches derived from 3D surface measurements use some features which are \svp{similar to those visually assessed by FTEs}\citep{luAutomatedBulletIdentification2014}.
 Class characteristics, which are shared by a group of firearms with the same rifling design, manufacturer, and tooling process, are typically evaluated first, as a mismatch on class charcteristics is sufficient for an exclusion.
@@
 
-\svp{Cross-correlation between two signatures $S_1(x)$ and $S_2(x)$, $$CC(S_1, S_2, z) = \lim_{L\rightarrow\infty}\left(\frac{1}{L} \int_{-L/2}^{L/2} S_1(x)S_2(x + z) dx \right)/[R_1R_2]$$ where $z$ is the shift difference between the two sequences and $R_1$ and $R_2$ are the root mean square roughness for the two signature \citep{vorburgerApplicationsCrosscorrelationFunctions2011}.
-We maximize $CC$ over values of $z$ to determine the best alignment; this maximal value is used to assess the match strength between the two sequences.
-XXX algorithm for calculating with missing values?}
-
-\hh{maximum cross-correlation is determined for aligning two signatures [@aoas, @vorburgerApplicationsCrosscorrelationFunctions2011]}
+\svp{Cross-correlation between two signatures $S_1(x)$ and $S_2(x)$ is defined as $$CC(S_1, S_2, z) = \text{cor}\left(S_1(x), S_2(x + z)\right)$$ where $z$ is the shift difference between the two sequences.
+We maximize $CC$ over values of $z$ to determine the best alignment  \citep{aoas,vorburgerApplicationsCrosscorrelationFunctions2011}; this maximal value is then used to assess the match strength between the two sequences.
+When signatures contain missing values, pairwise complete observations are used to calculate the cross-correlation.}
 

### 2018-11-09 index.Rmd
mathiness

@@
 Cross-correlation and Random Forest scores both result in perfect discrimination of same-source and different-source bullets.
-At the land-to-land level, discrimination (based on area under the curve, AUC) is excellent (???
-0.90).
+At the land-to-land level, discrimination (based on area under the curve, AUC) is excellent ($ge$ 0.90).
 
@@
 
-In the matching algorithm proposed in \citet{aoas}, several of the features described above are combined using a random forest\citep{breiman} to produce a similarity score based on 3d topographic scans of land engraved areas (LEAs).
-From each land engraved area a signature is derived by
+In the matching algorithm proposed in \citet{aoas}, several of the features described above are combined using a random forest \citep{breiman} to produce a similarity score based on 3d topographic scans of land engraved areas (LEAs).
+From each land engraved area a signature is derived \hh{using the process described in detail in \citet{aoas}:}
 
 \begin{itemize}
-\item[(a)] identifying an area where striae are expressed, 
+\item[(a)] Identify an area on the LEA, striae are expressed. 
+XXXX get from area to crosscut XXXX
+\hh{
+Let $z(t)$ describe a digitized toolmark for the striation marks on a land engraved area.
+$z(t)$ is then a spatial process indexed by $t$, where $t, t = 1,...,T$ denotes equally spaced pixel locations for a total length of $T$. }
+
 \item[(b)] discarding extraneous/contaminated data, such as data from groove engraved area and areas affected by break-off or contact with objects after the bullet exited the barrel  (``tank rash"), 

### 2018-11-09 index.Rmd
Add phoenix ammo and remove orange from the description

--- 2018-11-09 index.Rmd
@@
 Let's use instead wrong identification and missed identification. 
-\item color: let's not mention 'orange' anywhere in the manuscript - I don't know, whether the printed version of the paper will be in color. 
+<!-- \item color: let's not mention 'orange' anywhere in the manuscript - I don't know, whether the printed version of the paper will be in color.  -->
 \end{itemize}
@@
 This set is an open set; that is,\ it is not known in advance whether all (or any) of the questioned bullets are fired from the known barrels.
-\svp{Need to add ammo info here...}
+\svp{The ammunition used for this set were 9mm Luger copper jacket from American Eagle.}
 Land engraved areas for each of the six lands of each bullet were scanned by Bill Henderson (Sensofar).
@@
 \svp{Each barrel was test-fired twice, so }each of the colored tiles corresponds to a \svp{comparison} between a questioned bullet (shown along the $y$-axis) and one of the \svp{known} bullets.
-The similarity score is encoded in color: grey colors correspond to low similarity scores, orange colors correspond to high similarity scores.
-Ground truth is encoded in this figure as a thin orange frame for all pairs of same-source bullets.
-\svp{We want to see one barrel with two orange-filled, orange-framed tiles for each questioned bullet, and grey tiles for all other barrels, indicating a match between a questioned bullet and a single barrel.} This expectation is met for all questioned bullets, i.e.
+The similarity score is encoded in color: light grey colors correspond to low similarity scores, dark colors correspond to high similarity scores.
+Ground truth is encoded in this figure as a thin, dark, colored frame for all pairs of same-source bullets.
+\svp{We want to see one barrel with two dark-filled, dark-framed tiles for each questioned bullet, and light grey tiles for all other barrels, indicating a match between a questioned bullet and a single barrel.} This expectation is met for all questioned bullets, i.e.
 the automatic matching  identifies the correct barrel for all questioned bullets.

### 2018-11-09 index.Rmd
rephrase a bit and cite the paper.

@@
 \begin{itemize}
-\item[(a)] Identify an area on the LEA, striae are expressed. 
-XXXX get from area to crosscut XXXX
-\hh{
-Let $z(t)$ describe a digitized toolmark for the striation marks on a land engraved area.
-$z(t)$ is then a spatial process indexed by $t$, where $t, t = 1,...,T$ denotes equally spaced pixel locations for a total length of $T$. }
+\item[(a)] Identify an area on the LEA with expressed striae. Locate a stable crosscut (see \citet{aoas}) in the identified area. 
+<!-- XXXX get from area to crosscut XXXX -->
+<!-- \hh{ -->
+<!-- Let $z(t)$ describe a digitized toolmark for the striation marks on a land engraved area. -->
+<!-- $z(t)$ is then a spatial process indexed by $t$, where $t, t = 1,...,T$ denotes equally spaced pixel locations for a total length of $T$. } -->
 

### 2018-11-12 index.Rmd
phrasing change

@@
 Let's use instead wrong identification and missed identification. 
-<!-- \item color: let's not mention 'orange' anywhere in the manuscript - I don't know, whether the printed version of the paper will be in color.  -->
 \end{itemize}
@@
 XXX this could be a second bullet from the crime scene XXX They then visually classify similarity according to the theory of firearms identification \citep{identification} as one of identification, inconclusive or exclusion.
-Exact guidelines for this classification vary from lab to lab, e.g.\ some labs will exclude only on the basis of non-matching class characteristics, such as direction of the twist in rifling, land length or number of lands, or type of rifling.
-In some labs,  CMS (consecutively matching striae) as defined by \citet{biasotti} is used as a measure to quantify the strength of a match.
+Exact guidelines for this classification vary from lab to lab; some labs will exclude only on the basis of non-matching class characteristics, such as direction of the twist in rifling, land length or number of lands, or type of rifling.
+In other labs, CMS (consecutively matching striae) as defined by \citet{biasotti} is used as a measure to quantify the strength of a match.
 
 ## 3D Imaging Technology
-More explicit characterization of bullet surfaces has been discussed since at least 1958 (\citeauthor{davis1968introduction}), but at the time technology was not sufficiently advanced to make an analysis based on 3D measurements a viable option.
+More explicit characterization of bullet surfaces has been discussed since at least 1958 (\citeauthor{davis1968introduction}), but at the time technology was not sufficiently advanced to make an analysis based on 3D measurements or surface traces a viable option.
 \svp{Use of a scanning electron microscope to quantitatively examine and compare bullet striae was demonstrated in \citet{ComputerIdentificationBullets1978}}.
@@
 
-Commonly, approaches derived from 3D surface measurements use some features which are \svp{similar to those visually assessed by FTEs}\citep{luAutomatedBulletIdentification2014}.
+Commonly, approaches derived from 3D surface measurements use some features which are \svp{similar to those visually assessed by FTEs} \citep{luAutomatedBulletIdentification2014}.
 Class characteristics, which are shared by a group of firearms with the same rifling design, manufacturer, and tooling process, are typically evaluated first, as a mismatch on class charcteristics is sufficient for an exclusion.
@@
 
-\begin{itemize}
-\item[(a)] Identify an area on the LEA with expressed striae. Locate a stable crosscut (see \citet{aoas}) in the identified area. 
 <!-- XXXX get from area to crosscut XXXX -->
@@
 
+\begin{itemize}
+\item[(a)] Identify an area on the LEA with expressed striae. Locate a stable crosscut (see \citet{aoas}) in the identified area.
 \item[(b)] discarding extraneous/contaminated data, such as data from groove engraved area and areas affected by break-off or contact with objects after the bullet exited the barrel  (``tank rash"), 
@@
 CMS in this setting are counted as the number of consecutive peaks in signatures of two lands that match. 
-\hh{Note that because signatures are moved into phase based on the maximum cross-correlation rather than matching striation marks. This can result in a consecutive matching striae of 0. }
-
-\svp{XXX No idea what to say about the RF other than what's above? } 
-\hh{XXX that's what I thought, too.}
-
-\hh{We already have a start on this above, but we should focus on these three measures in more detail.}
+\hh{Note that because signatures are moved into phase based on the maximum cross-correlation rather than matching striation marks.
+This can result in a consecutive matching striae of 0.
+} \svp{CMS are generally considered to be a very conservative criteria, in that bullets fired through the same barrel may not have a high number of consecutively matching striae; in this sense, CMS may be more likely to produce an inconclusive result.}
+
+\svp{The random forest presented in \citet{aoas} takes as input the following characteristics derived from a pair of aligned sequences:
+\begin{itemize}
+\item Cross-correlation 
+\item Number of matching striae 
+\item Total depth of joint striations 
+\item Number of non-matching striae 
+\item Euclidian distance between the sequences 
+\item Consecutive Matching Striae 
+\item Consecutive Non-matching Striae
+\end{itemize}
+The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where 1 indicates that the algorithm believes it to be highly likely that the aligned signatures match, and 0 indicates that the aligned signatures are highly likely to originate from different lands.}
 
@@
 
-\hh{The automatic matching algorithm is based on a random forest \citep{breiman}, which has a built-in internal testing mechanism to prevent potential overfitting.
+\svp{Random forests \citep{breiman} have} \hh{a built-in internal testing mechanism to prevent potential overfitting.
 Error rates reported by random forest algorithms are based on these internal test sets.
 However, there is some discussion on whether or in which situations these error rates are biased \citep{rfOver2, rfOver}.
-Neither of these papers addresses another issue with internal test sets: internal test sets are constructed to have the same distribution as the training data (apart from sampling variability).
-For any machine learning method the true test comes with testing the performance of an algorithm on external test data.
-Good performance (in terms of low error rates) on external test data validates a ML algorithm.
-External data allows an assessment of the algorithm's sensitivity to distributional changes.
-Using performance on external test data also allows us to test limitations of the algorithm by going outside the parameters of the training data.
+Neither of these papers addresses another issue with internal test sets: internal test sets are constructed to have the same distribution as the training data (apart from sampling variability).} \svp{For any machine learning method, a true benchmark of the performance of an algorithm requires testing its' performance on external test data.} \hh{Good performance (in terms of low error rates) on external test data validates a ML algorithm.
+External data also allows an assessment of the algorithm's sensitivity to distributional changes} \svp{as well as testing the algorithm's robustness by going outside the parameters of the training data.
 }
@@
 \begin{itemize}
-\item {\bf Hamby set 44} 
-Hamby set 44 is one set of the Hamby study \citep{hamby}.
+\item {\bf Hamby set 44} is one set of the Hamby study \citep{hamby}.
 Each Hamby set  consists of a total of 35 bullets fired through ten consecutively manufactured barrels of Rugers P85.
@@
 These strips correspond to known barrels 1 through 10.
-\svp{Each barrel was test-fired twice, so }each of the colored tiles corresponds to a \svp{comparison} between a questioned bullet (shown along the $y$-axis) and one of the \svp{known} bullets.
-The similarity score is encoded in color: light grey colors correspond to low similarity scores, dark colors correspond to high similarity scores.
-Ground truth is encoded in this figure as a thin, dark, colored frame for all pairs of same-source bullets.
+\svp{Each barrel was test-fired twice, so each of the questioned bullets (shown along the $y$ axis) matches two bullets fired from a known barrel.
+Colored tiles are used to show the strength of the comparison; light grey colors correspond to low similarity scores, dark colors correspond to high similarity scores.
+squares with a colored frame indicate ground truth.
+} Ground truth is encoded in this figure as a thin, dark, colored frame for all pairs of same-source bullets.
 \svp{We want to see one barrel with two dark-filled, dark-framed tiles for each questioned bullet, and light grey tiles for all other barrels, indicating a match between a questioned bullet and a single barrel.} This expectation is met for all questioned bullets, i.e.

### 2018-11-13 index.Rmd
Phrasing and flow, move the aggregation section up to just after the feature/random forest discussion; keep all of the case study stuff together.

@@
 # Introduction and Background
-In current practice, firearms and toolmark examiners (FTE) evaluate the similarity of striae on bullets by placing the evidence together with a bullet from the suspected gun under a comparison microscope.
-XXX this could be a second bullet from the crime scene XXX They then visually classify similarity according to the theory of firearms identification \citep{identification} as one of identification, inconclusive or exclusion.
+In current practice, firearms and toolmark examiners (FTE) evaluate the similarity of striae on bullets by placing the evidence together with another bullet under a comparison microscope.
+\svp{The second bullet could be a test fire from a weapon recovered during the investigation, or it could be a second bullet from the crime scene.} Examiners visually classify similarity according to the theory of firearms identification \citep{identification} as one of identification, inconclusive or exclusion.
 Exact guidelines for this classification vary from lab to lab; some labs will exclude only on the basis of non-matching class characteristics, such as direction of the twist in rifling, land length or number of lands, or type of rifling.
 In other labs, CMS (consecutively matching striae) as defined by \citet{biasotti} is used as a measure to quantify the strength of a match.
-
-## 3D Imaging Technology
+\svp{In virtually all labs, individual characteristics used to identify matching bullets are derived from visual assessment; some class characteristics may be directly measured, but these are not sufficient for individualization.
+} <!-- Want to make it clear that individual characteristics are visually derived, rather than numerically derived...
+hopefully managed that without being insulting.
+-->
+
+## Identification Using 3D Imaging Technology
 More explicit characterization of bullet surfaces has been discussed since at least 1958 (\citeauthor{davis1968introduction}), but at the time technology was not sufficiently advanced to make an analysis based on 3D measurements or surface traces a viable option.
-\svp{Use of a scanning electron microscope to quantitatively examine and compare bullet striae was demonstrated in \citet{ComputerIdentificationBullets1978}}.
+\svp{\citet{ComputerIdentificationBullets1978} demonstrated use of a scanning electron microscope to quantitatively examine and compare bullet striae.
 More recently, approaches using 3D measurement data were explicitly described in 1999 (\citeauthor{dekinderAutomatedComparisonsBullet1999}), and have been further developed by \citet{bachrachDevelopment3DbasedAutomated2002,xieAutomatedBulletidentificationSystem2009, chuPilotStudyAutomated2010}.
@@
 Utilizing the 3D measurements allows for examination not only of the peaks in the bullet signature, but also for comparison of the valleys, which are more difficult to assess visually except in relation to the peaks.
-
-## Identification Process
 
@@
 
-Quantitative features describing the similarity of two signature are extracted.
-Features include characteristics of the signature that firearms and toolmark examiners base their visual assessment on, such as striae depth, number of matching striae, and the number of consecutively matching striae (CMS).
-We also include statistical features, such as cross-correlation and Euclidean distance between the signatures.
-
-## Statistics for Matching Aligned Signatures
-\hh{XXX define signatures mathematically here - as random spatial process; extracted from crosscut by removing big structure first and then applying (cross-validated, \citet{aoas}) 'roughness' filter.
-'roughness' defined by NIST paper.
+\svp{Once signatures have been extracted from stable regions with expressed striae, they must be aligned in order to assess their match strength.
+\citet{aoas} align signatures by maximizing the cross-correlation between the signatures.
 }
 
-\svp{Cross-correlation between two signatures $S_1(x)$ and $S_2(x)$ is defined as $$CC(S_1, S_2, z) = \text{cor}\left(S_1(x), S_2(x + z)\right)$$ where $z$ is the shift difference between the two sequences.
+\svp{The cross-correlation between two signatures $S_1(x)$ and $S_2(x)$ is defined as $$CC(S_1, S_2, z) = \text{cor}\left(S_1(x), S_2(x + z)\right)$$ where $z$ is the shift difference between the two sequences.
 We maximize $CC$ over values of $z$ to determine the best alignment  \citep{aoas,vorburgerApplicationsCrosscorrelationFunctions2011}; this maximal value is then used to assess the match strength between the two sequences.
@@
 
-One of the measures used by firearms examiners to describe the strength of a match quantitatively is the number of consecutive matching striae (CMS) as defined by \citet{biasotti}.
-If there are at least 6 consecutive matching striae, the bullets are considered to be a good match \citep{biasotti_firearms_1997,nichols_scientific_2006}, according to criteria used by FTEs. 
-CMS in this setting are counted as the number of consecutive peaks in signatures of two lands that match. 
-\hh{Note that because signatures are moved into phase based on the maximum cross-correlation rather than matching striation marks.
-This can result in a consecutive matching striae of 0.
+## Statistics for Matching Aligned Signatures
+<!-- \hh{XXX define signatures mathematically here - as random spatial process; extracted from crosscut by removing big structure first and then applying (cross-validated, \citet{aoas}) 'roughness' filter.
+'roughness' defined by NIST paper.
+} -->
+
+\svp{After two signatures are aligned, q}uantitative features describing the similarity of two signature are extracted.
+Features include characteristics of the signature that firearms and toolmark examiners base their visual assessment on, such as striae depth, number of matching striae, and the number of consecutively matching striae (CMS).
+Statistical features such as cross-correlation and Euclidean distance between the signatures \svp{are also extracted and used to assess the strength of the correspondance between the aligned signatures.
+}
+
+One of the measures firearms examiners use to describe the strength of a match quantitatively is the number of consecutive matching striae (CMS) as defined by \citet{biasotti}.
+If there are at least 6 consecutive matching striae, the bullets are considered to be a good match \citep{biasotti_firearms_1997,nichols_scientific_2006}. 
+CMS in this setting are counted as the number of consecutive peaks in signatures of two lands that match\svp{, in part because it is difficult to visually assess valleys except as relative to peaks}. 
+\hh{Note that because signatures are moved into phase based on the maximum cross-correlation rather than matching striation marks,}\svp{it is possible for two lands to }\hh{have a consecutive matching striae of 0.
 } \svp{CMS are generally considered to be a very conservative criteria, in that bullets fired through the same barrel may not have a high number of consecutively matching striae; in this sense, CMS may be more likely to produce an inconclusive result.}
@@
 \end{itemize}
+
 The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where 1 indicates that the algorithm believes it to be highly likely that the aligned signatures match, and 0 indicates that the aligned signatures are highly likely to originate from different lands.}
+
+# Assessing Match Strength
+\svp{Each of the features described in the previous section, as well as the random forest score, describe comparisons on a land-to-land basis.
+While these comparisons are useful, it is typically more useful to assess the entire physical object (e.g.
+all lands on a single bullet) to describe the correspondance between two pieces of physical evidence.
+In order to aggregate land-to-land match scores, we must consider the physical structure of the evidence; that is, that the sequence of lands stems from spatial information, but the numbering of each land represents the scan order and start point rather than encoding useful information.
+}
+
+When imaging bullets, operators scan one land at a time in a clockwise (left twisted rifling) or anti-clockwise (right twisted rifling) sequence.
+The order in which scans are acquired is kept as meta-information.
+Let us assume that lands on a bullet are labelled $\ell_i$ with $i = 1, .., p$.
+A match between two bullets therefore results in an expected additional $p-1$ matches between pairs of lands.
+These lands are also expected to be in a sequence, i.e.
+if there is a match between lands $\ell_i$ on bullet 1 and $\ell_j$ on bullet 2, we also expect lands $\ell_{i\oplus s}$ and $\ell_{j\oplus s}$ to match for all integers $s$, where $\oplus$ is defined as $a \oplus b \equiv \left((a + b - 1)\mod p\right) + 1$.
+This relationship gives rise to the sequence average maximum (SAM) to quantify a bullet-to-bullet match.
+
+## Sequence Average Maximum
+
+```{r, echo=FALSE, results = 'hide', warning = FALSE, message = FALSE}
+features <- read.csv("data/fsi-features.csv.gz")
+features <- features %>% mutate(
+  bullet1 = gsub("ullet ", "", bullet1),
+  bullet2 = gsub("ullet ", "", bullet2)
+)
+b12 <- features %>% filter(
+  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
+  group2 == "G1", barrel2 == "KA", bullet2 == "B2"
+)
+p1 <- b12 %>%
+  ggplot(aes(x = land1, y = land2, fill = rfscore)) +
+  geom_tile() + ggtitle("Same barrel") +
+  scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
+  theme_bw() +
+  xlab("Bullet 2") + ylab("Bullet 1") +
+  theme(legend.position = "bottom") + coord_equal()
+
+bab <- features %>% filter(
+  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
+  group2 == "G1", barrel2 == "KB", bullet2 == "B1"
+)
+p2 <- bab %>%
+  ggplot(aes(x = land1, y = land2, fill = rfscore)) +
+  geom_tile() + ggtitle("Different barrel") +
+  scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
+  theme_bw() +
+  xlab("Bullet 2") + ylab("Bullet 1") +
+  theme(legend.position = "bottom") + coord_equal()
+```
+
+Land-to-land comparisons lead to a whole set of scores for bullet-to-bullet comparisons.
+\autoref{fig:b2b} shows two matrices of scores for a pair of bullet-bullet matches.
+On the left, a matrix is shown that is typical for scores from two bullets from the same barrel.
+On the right, values for a pair of known non-matching bullets are shown.
+
+```{r b2b, echo = FALSE, fig.width = 9, fig.height = 5,  out.width='.6\\textwidth', fig.cap="Overview of land-land matching scores for two pairs of bullet-bullet comparisons.
+On the left the two bullets are known to come from the same source (barrel) on the right, the bullets are from two different sources (barrels)"}
+grid.arrange(p1, p2, ncol = 2, padding = 2)
+```
+
+\begin{definition}{Sequence Average and its Maximum}
+Let $A$ be a square real-valued matrix of dimensions $p \times p$. For the purpose of this paper, $A$ consists of scores describing the similarity between two sets of land engraved areas.
+The $k$th \emph{sequence average} $SA(A, k)$ for $k = 0, ... p-1$ is defined  as 
+$$SA(A, k) = \frac{1}{p} \sum_{i=1}^{p} a_{i,i \oplus k}, \text{ where } i\oplus k := \left((i + k - 1)\mod p\right) + 1.$$
+The \emph{Sequence Average Maximum} \citep[SAM, ][]{sam} of square matrix $A$ of scores is defined as
+$$SAM (A) = \max_{k = 1}^{p} SA(A, k).$$
+\end{definition}
+
+Note that the sequence average maximum of the correlation between lands is used in SensoComp to capture the similarity between bullets.
+The correlation based SAM score has also been called the 'average correlation calculated at the max phase' in \citet{chuPilotStudyAutomated2010}.
+
+```{r sam-sketch, out.width='\\textwidth', fig.width=12, fig.height = 2.5, fig.cap="Sketch of all six land-to-land sequences between two bullets with six lands. "}
+df2 <- data.frame(expand.grid(x = 1:6, y = 1:6))
+df2$case1 <- df2$x == ((df2$y + 5) %% 6) + 1
+df2$case2 <- df2$x == ((df2$y + 0) %% 6) + 1
+df2$case3 <- df2$x == ((df2$y + 1) %% 6) + 1
+df2$case4 <- df2$x == ((df2$y + 2) %% 6) + 1
+df2$case5 <- df2$x == ((df2$y + 3) %% 6) + 1
+df2$case6 <- df2$x == ((df2$y + 4) %% 6) + 1
+
+df2 %>%
+  gather("case", "values", starts_with("case")) %>%
+  mutate(case = gsub("case", "", case)) %>%
+  mutate(case = paste0("k = ", as.numeric(case) - 1)) %>%
+  ggplot(aes(x = factor(x), y = factor(y, levels = 1:6), fill = values)) +
+  geom_tile(colour = "grey20") +
+  facet_wrap(~case, ncol = 6) +
+  xlab("Bullet 2") + ylab("Bullet 1") +
+  scale_x_discrete(labels = paste0("L", 1:6)) +
+  scale_y_discrete(labels = paste0("L", 1:6)) +
+  scale_fill_manual("Cell included in SA(A,k)", values = c("white", "grey20")) +
+  coord_equal() +
+  theme(legend.position = "bottom")
+```
+
+Looking back at \autoref{fig:b2b}, we see that for the two bullets from the same barrel, the sequence average for 
+$k=2$ is higher than the other sequence averages, and also higher than the sequence averages for the other pair of bullets shown on the right of the figure.
+
+SAM scores allow us to define a single quantity for each pair of bullets that describes the similarity between these two bullets. 
+This leads us to two minimal requirements the scores should fulfill:
+\begin{itemize}
+\item[(R1)] {\bf Monotonicity:} a higher score is indicative of higher similarity between a pair of bullets, in particular, similarity scores of same-source pairs of bullets are higher than different-source pairs.
+\item[(R2)] {\bf Stability:} the same score leads to the same conclusion. 
+\end{itemize}
 
@@
 This mis-labeling has been corrected now.
+The scans used to train the model were taken at 20 fold magnification for a resolution of 1.5625 microns per pixel.
+While magnification is generally of interest in microscopy, **resolution** -- usually measured in microns per pixel -- is of more interest for 3D topographic measurements, as it determines the operative level of available data.
 
@@
 
-The scans used to train the model were taken at 20 fold magnification for a resolution of 1.5625 microns per pixel.
-While magnification is generally of interest in microscopy, **resolution** -- usually measured in microns per pixel -- is of more interest for 3D topographic measurements, as it determines the operative level of available data.
-
 Scans of all land engraved areas for the validation data were taken on a Sensofar Confocal Light microscope at 20x magnification resulting in a resolution of 0.645 microns per pixel.
@@
 This should make the automatic matching harder, and in particular, \svp{is expected to complicate} the classification as different-source land engraved areas/bullets.
-
-# From land-to-land scores to bullet summary scores
-
-```{r, echo=FALSE, results = 'hide', warning = FALSE, message = FALSE}
-features <- read.csv("data/fsi-features.csv.gz")
-features <- features %>% mutate(
-  bullet1 = gsub("ullet ", "", bullet1),
-  bullet2 = gsub("ullet ", "", bullet2)
-)
-b12 <- features %>% filter(
-  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
-  group2 == "G1", barrel2 == "KA", bullet2 == "B2"
-)
-p1 <- b12 %>%
-  ggplot(aes(x = land1, y = land2, fill = rfscore)) +
-  geom_tile() + ggtitle("Same barrel") +
-  scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
-  theme_bw() +
-  xlab("Bullet 2") + ylab("Bullet 1") +
-  theme(legend.position = "bottom") + coord_equal()
-
-bab <- features %>% filter(
-  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
-  group2 == "G1", barrel2 == "KB", bullet2 == "B1"
-)
-p2 <- bab %>%
-  ggplot(aes(x = land1, y = land2, fill = rfscore)) +
-  geom_tile() + ggtitle("Different barrel") +
-  scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
-  theme_bw() +
-  xlab("Bullet 2") + ylab("Bullet 1") +
-  theme(legend.position = "bottom") + coord_equal()
-```
-
-Land-to-land comparisons lead to a whole set of scores for bullet-to-bullet comparisons.
-\autoref{fig:b2b} shows two matrices of scores for a pair of bullet-bullet matches.
-On the left, a matrix is shown that is typical for scores from two bullets from the same barrel.
-On the right, values for a pair of known non-matching bullets are shown.
-
-```{r b2b, echo = FALSE, fig.width = 9, fig.height = 5,  out.width='.6\\textwidth', fig.cap="Overview of land-land matching scores for two pairs of bullet-bullet comparisons.
-On the left the two bullets are known to come from the same source (barrel) on the right, the bullets are from two different sources (barrels)"}
-grid.arrange(p1, p2, ncol = 2, padding = 2)
-```
-
-When imaging bullets, operators scan one land at a time in a clockwise (left twisted rifling) or anti-clockwise (right twisted rifling) sequence.
-The order in which scans are acquired is kept as meta-information.
-Let us assume that lands on a bullet are labelled $\ell_i$ with $i = 1, .., p$.
-A match between two bullets therefore results in an expected additional $p-1$ matches between pairs of lands.
-These lands are also expected to be in a sequence, i.e.
-if there is a match between lands $\ell_i$ on bullet 1 and $\ell_j$ on bullet 2, we also expect lands $\ell_{i\oplus s}$ and $\ell_{j\oplus s}$ to match for all integers $s$, where $\oplus$ is defined as $a \oplus b \equiv \left((a + b - 1)\mod p\right) + 1$.
-This relationship gives rise to the sequence average maximum (SAM) to quantify a bullet-to-bullet match.
-
-\begin{definition}{Sequence Average and its Maximum}
-Let $A$ be a square real-valued matrix of dimensions $p \times p$. For the purpose of this paper, $A$ consists of scores describing the similarity between two sets of land engraved areas.
-The $k$th \emph{sequence average} $SA(A, k)$ for $k = 0, ... p-1$ is defined  as 
-$$SA(A, k) = \frac{1}{p} \sum_{i=1}^{p} a_{i,i \oplus k}, \text{ where } i\oplus k := \left((i + k - 1)\mod p\right) + 1.$$
-The \emph{Sequence Average Maximum} \citep[SAM, ][]{sam} of square matrix $A$ of scores is defined as
-$$SAM (A) = \max_{k = 1}^{p} SA(A, k).$$
-\end{definition}
-
-Note that the sequence average maximum of the correlation between lands is used in SensoComp to capture the similarity between bullets.
-The correlation based SAM score has also been called the 'average correlation calculated at the max phase' in \citet{chuPilotStudyAutomated2010}.
-
-```{r sam-sketch, out.width='\\textwidth', fig.width=12, fig.height = 2.5, fig.cap="Sketch of all six land-to-land sequences between two bullets with six lands. "}
-df2 <- data.frame(expand.grid(x = 1:6, y = 1:6))
-df2$case1 <- df2$x == ((df2$y + 5) %% 6) + 1
-df2$case2 <- df2$x == ((df2$y + 0) %% 6) + 1
-df2$case3 <- df2$x == ((df2$y + 1) %% 6) + 1
-df2$case4 <- df2$x == ((df2$y + 2) %% 6) + 1
-df2$case5 <- df2$x == ((df2$y + 3) %% 6) + 1
-df2$case6 <- df2$x == ((df2$y + 4) %% 6) + 1
-
-df2 %>%
-  gather("case", "values", starts_with("case")) %>%
-  mutate(case = gsub("case", "", case)) %>%
-  mutate(case = paste0("k = ", as.numeric(case) - 1)) %>%
-  ggplot(aes(x = factor(x), y = factor(y, levels = 1:6), fill = values)) +
-  geom_tile(colour = "grey20") +
-  facet_wrap(~case, ncol = 6) +
-  xlab("Bullet 2") + ylab("Bullet 1") +
-  scale_x_discrete(labels = paste0("L", 1:6)) +
-  scale_y_discrete(labels = paste0("L", 1:6)) +
-  scale_fill_manual("Cell included in SA(A,k)", values = c("white", "grey20")) +
-  coord_equal() +
-  theme(legend.position = "bottom")
-```
-
-Looking back at \autoref{fig:b2b}, we see that for the two bullets from the same barrel, the sequence average for 
-$k=2$ is higher than the other sequence averages, and also higher than the sequence averages for the other pair of bullets shown on the right of the figure.
-
-SAM scores allow us to define a single quantity for each pair of bullets that describes the similarity between these two bullets. 
-This leads us to two minimal requirements the scores should fulfill:
-\begin{itemize}
-\item[(R1)] {\bf Monotonicity:} a higher score is indicative of higher similarity between a pair of bullets, in particular, similarity scores of same-source pairs of bullets are higher than different-source pairs.
-\item[(R2)] {\bf Stability:} the same score leads to the same conclusion. 
-\end{itemize}
 

### 2018-11-13 index.Rmd
Transitional paragraph, changes in headings for consistency, and other tweaks

@@
 
-\svp{The random forest presented in \citet{aoas} takes as input the following characteristics derived from a pair of aligned sequences:
-\begin{itemize}
+\svp{The random forest presented in \citet{aoas} takes as input the following characteristics derived from a pair of aligned sequences:}
+\svp{\begin{itemize}
 \item Cross-correlation 
@@
 \item Consecutive Non-matching Striae
-\end{itemize}
-
-The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where 1 indicates that the algorithm believes it to be highly likely that the aligned signatures match, and 0 indicates that the aligned signatures are highly likely to originate from different lands.}
+\end{itemize}}
+\svp{The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where 1 indicates that the algorithm believes it to be highly likely that the aligned signatures match, and 0 indicates that the aligned signatures are highly likely to originate from different lands.}
 
@@
 While these comparisons are useful, it is typically more useful to assess the entire physical object (e.g.
-all lands on a single bullet) to describe the correspondance between two pieces of physical evidence.
-In order to aggregate land-to-land match scores, we must consider the physical structure of the evidence; that is, that the sequence of lands stems from spatial information, but the numbering of each land represents the scan order and start point rather than encoding useful information.
-}
+all lands on a single bullet) to describe the correspondance between two pieces of physical evidence.}
+<!-- \svp{In order to aggregate land-to-land match scores, we must consider the physical structure of the evidence; that is, that the sequence of lands stems from spatial information, but the numbering of each land represents the scan order and start point rather than encoding useful information.
+} --><!--This last bit might be redundant-->
+
+```{r, echo=FALSE, results = 'hide', warning = FALSE, message = FALSE}
+features <- read.csv("data/fsi-features.csv.gz")
+features <- features %>% mutate(
+  bullet1 = gsub("ullet ", "", bullet1),
+  bullet2 = gsub("ullet ", "", bullet2)
+)
+b12 <- features %>% filter(
+  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
+  group2 == "G1", barrel2 == "KA", bullet2 == "B2"
+)
+p1 <- b12 %>%
+  ggplot(aes(x = land1, y = land2, fill = rfscore)) +
+  geom_tile() + ggtitle("Same barrel") +
+  scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
+  theme_bw() +
+  xlab("Bullet 2") + ylab("Bullet 1") +
+  theme(legend.position = "bottom") + coord_equal()
+
+bab <- features %>% filter(
+  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
+  group2 == "G1", barrel2 == "KB", bullet2 == "B1"
+)
+p2 <- bab %>%
+  ggplot(aes(x = land1, y = land2, fill = rfscore)) +
+  geom_tile() + ggtitle("Different barrel") +
+  scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
+  theme_bw() +
+  xlab("Bullet 2") + ylab("Bullet 1") +
+  theme(legend.position = "bottom") + coord_equal()
+```
+
+## Land-to-Land Comparisons
+
+Land-to-land comparisons lead to a whole set of scores for bullet-to-bullet comparisons.
+\autoref{fig:b2b} shows two matrices of scores for a pair of bullet-bullet matches.
+On the left, a matrix is shown that is typical for scores from two bullets from the same barrel.
+On the right, values for a pair of known non-matching bullets are shown.
+
+```{r b2b, echo = FALSE, fig.width = 9, fig.height = 5,  out.width='.6\\textwidth', fig.cap="Overview of land-land matching scores for two pairs of bullet-bullet comparisons.
+On the left the two bullets are known to come from the same source (barrel) on the right, the bullets are from two different sources (barrels)"}
+grid.arrange(p1, p2, ncol = 2, padding = 2)
+```
 
@@
 
-## Sequence Average Maximum
-
-```{r, echo=FALSE, results = 'hide', warning = FALSE, message = FALSE}
-features <- read.csv("data/fsi-features.csv.gz")
-features <- features %>% mutate(
-  bullet1 = gsub("ullet ", "", bullet1),
-  bullet2 = gsub("ullet ", "", bullet2)
-)
-b12 <- features %>% filter(
-  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
-  group2 == "G1", barrel2 == "KA", bullet2 == "B2"
-)
-p1 <- b12 %>%
-  ggplot(aes(x = land1, y = land2, fill = rfscore)) +
-  geom_tile() + ggtitle("Same barrel") +
-  scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
-  theme_bw() +
-  xlab("Bullet 2") + ylab("Bullet 1") +
-  theme(legend.position = "bottom") + coord_equal()
-
-bab <- features %>% filter(
-  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
-  group2 == "G1", barrel2 == "KB", bullet2 == "B1"
-)
-p2 <- bab %>%
-  ggplot(aes(x = land1, y = land2, fill = rfscore)) +
-  geom_tile() + ggtitle("Different barrel") +
-  scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
-  theme_bw() +
-  xlab("Bullet 2") + ylab("Bullet 1") +
-  theme(legend.position = "bottom") + coord_equal()
-```
-
-Land-to-land comparisons lead to a whole set of scores for bullet-to-bullet comparisons.
-\autoref{fig:b2b} shows two matrices of scores for a pair of bullet-bullet matches.
-On the left, a matrix is shown that is typical for scores from two bullets from the same barrel.
-On the right, values for a pair of known non-matching bullets are shown.
-
-```{r b2b, echo = FALSE, fig.width = 9, fig.height = 5,  out.width='.6\\textwidth', fig.cap="Overview of land-land matching scores for two pairs of bullet-bullet comparisons.
-On the left the two bullets are known to come from the same source (barrel) on the right, the bullets are from two different sources (barrels)"}
-grid.arrange(p1, p2, ncol = 2, padding = 2)
-```
+## Bullet-to-Bullet Comparisons
 
@@
 \end{itemize}
+
+\svp{In this paper, we validate the algorithm described in \citet{aoas} on three external test sets.
+We assess the results on each set and evaluate whether the minimal requirements described above are fulfilled.
+We compare the random forest algorithm's performance to the performance of other suggested measures for quantitative assessment of match strength, such as cross-correlation and consecutive matching striae.}
 

### 2018-11-13 index.Rmd
phrasing tweaks

@@
 \svp{In this paper, we validate the algorithm described in \citet{aoas} on three external test sets.
-We assess the results on each set and evaluate whether the minimal requirements described above are fulfilled.
-We compare the random forest algorithm's performance to the performance of other suggested measures for quantitative assessment of match strength, such as cross-correlation and consecutive matching striae.}
+We assess the results on each set and evaluate whether the minimal requirements described above are fulfilled for bullet-to-bullet SAM scores.
+We compare the random forest algorithm's performance to the performance of other suggested measures for quantitative assessment of match strength, such as cross-correlation and consecutive matching striae.
+}
 
@@
 The algorithm in \citet{aoas} was trained on sets 173 and 252, so the bullets in Hamby 44 are of the same type of ammunition and are fired through the same barrels as the bullets in the training set.
-
 The Phoenix PD set uses Ruger P-95 barrels, which are different from the barrels in the Hamby sets used for training the algorithm, but the barrels are rifled similarly to the Ruger P-85 barrels.
-
 The Houston sets use Ruger LCP barrels.
 These barrels are first rifled traditionally, i.e.
-similar to the Ruger P-85, but are treated to a secondary round of heating after barreling.
+similar to the Ruger P-85, but are treated to a secondary round of heating after \svp{rifling}.
 This second round of heat exposure has the potential to introduce some subclass characteristics, i.e.
@@
 
-# Results from the validation sets
+# Results
 
@@
 <!--Thus, we can conclude that the algorithm performs well on open set studies, as it can identify matches among questioned bullets which do not match any known bullets. -->
-While all bullet-to-bullet matches are correctly identified, and no known matches are missed, it is obvious from the generally lighter shades of the tiles corresponding to matching bullets in \autoref{fig:hou} for set 2 that  the algorithm is not performing as well as for the other sets (see also \autoref{tab:aucs}).
+While all bullet-to-bullet matches are correctly identified, and no known matches are missed, it is obvious from the generally lighter shades of the tiles corresponding to matching bullets in \autoref{fig:hou}(b) \svp{that the algorithm is not performing as well on set 2 as it performs on sets 1 and 3.}
 ```{r hou, fig.width=9, fig.height = 3, fig.cap="Overview of matching scores for all pairs of questioned bullets to known bullets from five barrels (on the left) and questioned bullets to themselves (tiles on the right).", fig.subcap=c("Set 1", "Set 2", "Set 3"), fig.align="middle", fig.ncol=1}
@@
 Color indicates ground truth.
-Besides densities, small vertical lines below the $x$-axis indicate observed scores in a **rug-plot**.
-The plot shows that for all three case studies and both measures, i.e.
+Small vertical lines below the $x$-axis indicate observed scores in a **rug-plot**.
+The plot shows that for all three case studies and both continuous measures, i.e.
 SAM scores based on cross-correlation as well as SAM scores based on the random forest scores, all similarity scores are higher for pairs of bullets from the same source, i.e.
@@
 
-What we find in \autoref{fig:compare} is that neither of the measures fulfill both of the minimal requirements laid out above.
+\autoref{fig:compare} shows that neither of the measures fulfill both of the minimal requirements laid out\svp{ in \autoref{land-to-land-comparisons}}.
 Both measures fulfill (R1), i.e.

### 2018-11-13 index.Rmd
Add section for identification errors, add paragraph evaluating RF algorithm, tweak wording for flow, slightly restructure discussion/conclusion section

@@
 }
+
+## Identification Errors
+\svp{XXX define wrong identification and missed identification here XXX}
 
@@
 <!--Thus, we can conclude that the algorithm performs well on open set studies, as it can identify matches among questioned bullets which do not match any known bullets. -->
-While all bullet-to-bullet matches are correctly identified, and no known matches are missed, it is obvious from the generally lighter shades of the tiles corresponding to matching bullets in \autoref{fig:hou}(b) \svp{that the algorithm is not performing as well on set 2 as it performs on sets 1 and 3.}
+While all bullet-to-bullet matches are correctly identified, and no known matches are missed, it is obvious from the generally lighter shades of the tiles corresponding to matching bullets in \autoref{fig:hou}\svp{(b) that the algorithm is not performing as well on set 2 as it performs on sets 1 and 3.}
 ```{r hou, fig.width=9, fig.height = 3, fig.cap="Overview of matching scores for all pairs of questioned bullets to known bullets from five barrels (on the left) and questioned bullets to themselves (tiles on the right).", fig.subcap=c("Set 1", "Set 2", "Set 3"), fig.align="middle", fig.ncol=1}
@@
 
-# Comparison of SAM scores
-
+# Evaluating the Random Forest Algorithm
+\svp{The random forest SAM scores correctly separate known matches from known non-matches, using SAM scores to aggregate land-to-land scores into a bullet-to-bullet comparison.
+With no errors at the bullet-to-bullet level, we now need to evaluate the scores in light of the criteria described in \autoref{land-to-land-comparisons}.
+In this section, we will compare the random forest scores to cross-correlation scores; cross-correlation is one of the components of the random forest, but has also been used to quantify the strength of a match in its own right \citep{chuPilotStudyAutomated2010}.
+In addition, both scoring methods use the same scale (0 to 1) and are continuous along that interval; other evaluation methods, such as consecutive matching striae are discrete and more difficult to directly compare to continuous measurements.}
+
+\svp{XXX We may want to add in SAM scores for CMS as well, but that's going to make the the scales on the density plots (and the density plots themselves) a bit wonky. Thoughts?}
+
+## Comparison of SAM scores
 \autoref{fig:compare} shows density curves for the scores bullet-to-bullet matches of each of the studies, comparing the SAM scores based on cross-correlation (top) and the random forest score (bottom).
@@
 
-\autoref{fig:compare} shows that neither of the measures fulfill both of the minimal requirements laid out\svp{ in \autoref{land-to-land-comparisons}}.
+\svp{According to \autoref{fig:compare}, neither cross-correlation nor random forest score} fulfill both of the minimal requirements laid out \svp{in \autoref{land-to-land-comparisons}}.
 Both measures fulfill (R1), i.e.
@@
 There are several approaches for potential improvements: we can try to fine tune the matching algorithm to take make and model of the firearm and ammunition used into account or expand the training base of the algorithm to include a wider variety of makes and models.
-The downside of this is that we need to considerably expand the database used for training.
-
-Another solution would be to expand on the aggregation used to get from land-to-land scores to bullet scores: SAM scores only take the scores of the maximum sequence into account and ignore scores associated with pairs of land engraved areas from different sources.
+The downside of \svp{either of these options} is that we \svp{would} need to considerably expand the database used for training.
+Another solution would be to \svp{augment} the aggregation used to get from land-to-land scores to bullet scores: SAM scores only take the scores of the maximum sequence into account and ignore scores associated with pairs of land engraved areas from different sources.
 Those scores might be useful in establishing a baseline that better expresses similarity between pairs of bullets.
 One approach that uses scores for both same source pairs and different source pairs are score-based likelihood ratios \citep{Bunch:2013if, Morrison:2018fh}.
+
+\svp{XXX This is a bit of an incomplete thought. I think I know where you were going, but can you expand on it a bit so that I'm not putting words in your mouth? XXX}
 \citet{Morrison:2018km}  stress the need to additionally include the typicality of a signature.
 
-# Digging deeper: comparison of land-to-land scores
+## Digging deeper: comparison of land-to-land scores
 
@@
 This suggests that the matching algorithm is sensitive to the presence of LEAs which have low land-to-land match scores, such as the matches of bullets `I' and `F' in the Hamby 44 study.
-Comparing this to the bullet-to-bullet score density plot shown in \autoref{fig:compare}, we see that there is much greater separation between the same-source and different-source densities for each study in the bullet-to-bullet comparisons than in the land-to-land comparisons, that is, one or two poorly matching lands does not prevent the bullet from showing a matching score, though if there are only weak matches on a few lands, such as the marked Hamby 44 comparisons in \autoref{fig:compare}, the overall score may be affected.
+Comparing this to the bullet-to-bullet score density plot shown in \autoref{fig:compare}, we see that there is much greater separation between the same-source and different-source densities for each study in the bullet-to-bullet comparisons than in the land-to-land comparisons, that is, one or two poorly matching lands does not prevent the bullet from showing a matching score, though if there are weak matches on several lands, such as the marked Hamby 44 comparisons in \autoref{fig:compare}, the overall score may be affected.
 }
@@
 
-When evaluating classifier performance, it is common to use a receiver operating curve, which plots the false positives against the true positives for each possible value of the cutoff between the two classes (in this case, known match and known non-match).
-The ROC curves for the land-to-land scores are shown in \autoref{fig:roc-auc} (a).
+\svp{When evaluating classifier performance, it is common to use a receiver operating curve, which plots the false positives against the true positives for each possible value of the cutoff between the two classes (in this case, known match and known non-match).} \svp{As the land-to-land scores are not perfectly separated, we can use ROC curves to evaluate their performance.} \svp{The ROC curves for the land-to-land scores are shown in \autoref{fig:roc-auc} (a).
 A perfect classifier would have 100\% true positives and 0\% false positives, e.g.
@@
 Classifiers with better performance will be closer to this corner of the plot.
-A random classifier would have an ROC curve that was a straight diagonal line through (0,0) and (1,1).
+A random classifier would have an ROC curve that was a straight diagonal line through (0,0) and (1,1).}
 
@@
 
-```{r l2l, eval=FALSE, fig.cap="Density curves of similarity scores based on land-to-land comparisons for cross-correlation (top) and RF scores (bottom).
-Different colors indicate same source versus different source.
-RF scores for different source comparisons are generally well below 0.5.
-Some same-source comparisons are also below 0.5 for RF scores, indicating potential problems with at least one of the lands involved in the comparison."}
-all_lands %>% gather(measure, score, ccf, rfscore) %>%
-  mutate(
-    measure = factor(measure, labels=c("Cross-correlation", "RF score"))
-  ) %>%
-ggplot(aes(x = score, fill = sameland, colour =sameland)) + 
-  geom_density(alpha = 0.6) +
-  facet_grid(measure ~ study + group1) +
-  theme_bw() +
-  scale_fill_manual("Same source", values = c("darkgrey", "darkorange")) +
-  scale_colour_manual("Same source", values = c("darkgrey", "darkorange")) +
-  scale_x_continuous("Land-to-land scores", limits=c(0,1)) +
-  theme(legend.position = "bottom") +
-  geom_hline(yintercept = 0, show.legend = F) 
-```
-
 \hh{\autoref{fig:eer} shows an overview of Equal Error Rates (EER) and their thresholds based on the ROC curves of \autoref{fig:roc-auc}.
@@
 CMS is signifcantly worse than both Cross-correlation and the RF score.
-At best, Cross-correlation and RF score have errors for land-to-land comparison of around 5\\%.", out.width = '.8\\textwidth'}
+At best, Cross-correlation and RF score have errors for land-to-land comparison of around 5\\%.\\svp{I'm uncomfortable using the same scale for EER threshold given that CMS has a different scale; leaving CMS out also seems odd though}", out.width = '.8\\textwidth'}
 aucs %>% gather(eer_thresh, value, eer, threshold) %>%
@@
   ggplot(aes(y = studyg, x = value, colour=measure, shape = measure)) + 
-  geom_point(size = 3) +theme_bw() +
+  geom_point(size = 4) +theme_bw() +
 #  xlim(c(0, .35)) +
@@
   scale_color_brewer(palette = "Set2") +
+  scale_shape_discrete(solid = F) + 
   theme(legend.position = "bottom") +
@@
 
-### Limitations
+## Limitations
 
@@
 This would emulate the process used by examiners, who first assess the quality of the evidence and whether there is enough information present to attempt a match.
+\svp{One major disadvantage of an automated algorithm is that all of the decisions humans make without further thought (recognizing degraded land areas, excluding those areas from consideration, matching only the remaining areas) must be explicitly characterized; however, this explicit characterization means that the process can be scientifically validated to a much higher degree than the human perceptual process
+. }
+
+## Conclusion
+
+\svp{The random forest matching algorithm presented in \citet{aoas} performs well on three different external test sets.
+While a single cutoff value cannot be used to distinguish matches and non-matches across the different test sets, the algorithm makes no errors when a set-specific cutoff value is used on bullet-to-bullet scores aggregated using sequence average maximum.
+XXX Need more here XXX}
 

### 2018-11-14 index.Rmd
more info in phoenix study

@@
 \item {\bf Phoenix PD}
-Tylor Klep from Phoenix PD provided sets of known test fires and questioned bullets: the set of known bullets consists of three test fires (B1, B2, B3) from each of eight different Ruger P-95 barrels (A9, C8, F6, L5, M2, P7, R3, U10).
+Tylor Klep from Phoenix PD provided sets of known test fires and questioned bullets: the set of known bullets consists of three test fires (B1, B2, B3) from each of eight different, consecutively manufactured Ruger P-95 barrels (A9, C8, F6, L5, M2, P7, R3, U10).
 Ten questioned bullets were provided (B, E, H, J, K N, Q, T, Y, Z).
-This set is an open set; that is,\ it is not known in advance whether all (or any) of the questioned bullets are fired from the known barrels.
-\svp{The ammunition used for this set were 9mm Luger copper jacket from American Eagle.}
+This set is an open set; that is,\ it is not known in advance whether all (or any) of the questioned bullets are fired from the known barrels.\hh{ In fact, the results will show, that three of the questioned bullets were fired from three different barrels not included in the knowns.
+These three barrels correspond to an eleventh Ruger P-95, a Ruger P-95C and a Ruger P-85.} \svp{All bullets fired for this study are American Eagle 9mm  copper jacketed Luger.}
 Land engraved areas for each of the six lands of each bullet were scanned by Bill Henderson (Sensofar).

### 2018-11-14 index.Rmd
going over changes to intro

@@
 In current practice, firearms and toolmark examiners (FTE) evaluate the similarity of striae on bullets by placing the evidence together with another bullet under a comparison microscope.
-\svp{The second bullet could be a test fire from a weapon recovered during the investigation, or it could be a second bullet from the crime scene.} Examiners visually classify similarity according to the theory of firearms identification \citep{identification} as one of identification, inconclusive or exclusion.
+The second bullet could be a test fire from a weapon recovered during the investigation, or it could be a second bullet from the crime scene.
+Examiners visually classify similarity according to the theory of firearms identification \citep{identification} as one of identification, inconclusive or exclusion.
 Exact guidelines for this classification vary from lab to lab; some labs will exclude only on the basis of non-matching class characteristics, such as direction of the twist in rifling, land length or number of lands, or type of rifling.
 In other labs, CMS (consecutively matching striae) as defined by \citet{biasotti} is used as a measure to quantify the strength of a match.
-\svp{In virtually all labs, individual characteristics used to identify matching bullets are derived from visual assessment; some class characteristics may be directly measured, but these are not sufficient for individualization.
-} <!-- Want to make it clear that individual characteristics are visually derived, rather than numerically derived...
+In virtually all labs, individual characteristics used to identify matching bullets are derived from visual assessment; some class characteristics may be directly measured, but these are not sufficient for individualization.
+<!-- Want to make it clear that individual characteristics are visually derived, rather than numerically derived...
 hopefully managed that without being insulting.
--->
+HH: reads well -->
 
@@
 More explicit characterization of bullet surfaces has been discussed since at least 1958 (\citeauthor{davis1968introduction}), but at the time technology was not sufficiently advanced to make an analysis based on 3D measurements or surface traces a viable option.
-\svp{\citet{ComputerIdentificationBullets1978} demonstrated use of a scanning electron microscope to quantitatively examine and compare bullet striae.
+\citet{ComputerIdentificationBullets1978} demonstrated use of a scanning electron microscope to quantitatively examine and compare bullet striae.
 More recently, approaches using 3D measurement data were explicitly described in 1999 (\citeauthor{dekinderAutomatedComparisonsBullet1999}), and have been further developed by \citet{bachrachDevelopment3DbasedAutomated2002,xieAutomatedBulletidentificationSystem2009, chuPilotStudyAutomated2010}.
@@
 
-Commonly, approaches derived from 3D surface measurements use some features which are \svp{similar to those visually assessed by FTEs} \citep{luAutomatedBulletIdentification2014}.
+Commonly, approaches derived from 3D surface measurements use some features which are similar to those visually assessed by FTEs \citep{luAutomatedBulletIdentification2014}.
 Class characteristics, which are shared by a group of firearms with the same rifling design, manufacturer, and tooling process, are typically evaluated first, as a mismatch on class charcteristics is sufficient for an exclusion.
@@
 In the matching algorithm proposed in \citet{aoas}, several of the features described above are combined using a random forest \citep{breiman} to produce a similarity score based on 3d topographic scans of land engraved areas (LEAs).
-From each land engraved area a signature is derived \hh{using the process described in detail in \citet{aoas}:}
+From each land engraved area a signature is derived using the process described in detail in \citet{aoas}:
 
@@
 \item[(a)] Identify an area on the LEA with expressed striae. Locate a stable crosscut (see \citet{aoas}) in the identified area.
-\item[(b)] discarding extraneous/contaminated data, such as data from groove engraved area and areas affected by break-off or contact with objects after the bullet exited the barrel  (``tank rash"), 
-\item[(c)] removing bullet curvature. 
+\item[(b)] Discard extraneous/contaminated data, such as data from groove engraved area and areas affected by break-off or contact with objects after the bullet exited the barrel  (``tank rash").
+\item[(c)] Remove bullet curvature \hh{using a Gaussian filter\cite{nist} or non-parametric smooth \citep{aoas}}. 
 \end{itemize}
-A signature for a land engraved area is then defined as the sequence $z_i$, $i = 1, ..., I$, where $I$ is the number of observed locations across the base of the bullet.
+A signature for a land engraved area is then defined as the sequence $S(x)$, $x = 1, ..., I$, where $I$ is the number of observed locations across the base of the bullet.
 \autoref{fig:signatures} shows a set of six signatures corresponding to the six land engraved areas of bullets 1 and 2 from barrel 1 of set 44 of the Hamby study \citep{hamby}.
@@
 
-\svp{The cross-correlation between two signatures $S_1(x)$ and $S_2(x)$ is defined as $$CC(S_1, S_2, z) = \text{cor}\left(S_1(x), S_2(x + z)\right)$$ where $z$ is the shift difference between the two sequences.
-We maximize $CC$ over values of $z$ to determine the best alignment  \citep{aoas,vorburgerApplicationsCrosscorrelationFunctions2011}; this maximal value is then used to assess the match strength between the two sequences.
-When signatures contain missing values, pairwise complete observations are used to calculate the cross-correlation.}
+The cross-correlation function (ccf) between two signatures $S_1(x)$ and $S_2(x)$ is defined as $$CC(S_1, S_2, z) = \text{cor}\left(S_1(x), S_2(x + z)\right)$$ where \hh{$x$ and $x+z$ are integer values appropriately defined within the domains of $S_1$ and $S_2$ and }  $z$ is the lag between the two sequences.
+We maximize $CC$ over values of $z$ to determine the best alignment  \citep{aoas,vorburgerApplicationsCrosscorrelationFunctions2011}; the maximal value of $CC$ is  used to assess the match strength between the two sequences.
+When signatures contain missing values, pairwise complete observations are used to calculate the cross-correlation.
 
@@
 
-\svp{After two signatures are aligned, q}uantitative features describing the similarity of two signature are extracted.
+After two signatures are aligned, quantitative features describing the similarity of two signature are extracted.
 Features include characteristics of the signature that firearms and toolmark examiners base their visual assessment on, such as striae depth, number of matching striae, and the number of consecutively matching striae (CMS).
-Statistical features such as cross-correlation and Euclidean distance between the signatures \svp{are also extracted and used to assess the strength of the correspondance between the aligned signatures.
-}
+Statistical features such as cross-correlation and Euclidean distance between the signatures are also extracted and used to assess the strength of the correspondence between the aligned signatures.
 

### 2018-11-14 index.Rmd
error rates and percentages
Define wrong/missed identifications.

--- 2018-11-13 index.Rmd
@@
 
+## Identification Errors
+\svp{XXX define wrong identification and missed identification here XXX}
+\svp{When evaluating algorithm performance, it is useful to distinguish between two types of errors, which we will refer to as \textbf{wrong identification} and \textbf{missed identification} in this paper.
+Wrong identifications are those in which two bullets from different sources are determined to be from the same source.
+Missed identifications are those in which two bullets from the same source are determined to be from different sources.}
+\svp{XXX Relate to AFTE Theory of Examination criteria? How to handle ``inconclusive"?}
+
 \svp{In this paper, we validate the algorithm described in \citet{aoas} on three external test sets.
@@
 }
-
-## Identification Errors
-\svp{XXX define wrong identification and missed identification here XXX}
 

### 2018-11-14 index.Rmd
add comment

--- 2018-11-14 index.Rmd
@@
 Let's use instead wrong identification and missed identification. 
+\item error rates: we are using rates exchangeably with percentages. We need to fix that. I think percentages might be easier understood. 
 \end{itemize}
@@
 
+## Identification Errors
+\svp{XXX define wrong identification and missed identification here XXX}
+\svp{When evaluating algorithm performance, it is useful to distinguish between two types of errors, which we will refer to as \textbf{wrong identification} and \textbf{missed identification} in this paper.
+Wrong identifications are those in which two bullets from different sources are determined to be from the same source.
+Missed identifications are those in which two bullets from the same source are determined to be from different sources.}
+\svp{XXX Relate to AFTE Theory of Examination criteria? How to handle ``inconclusive"?}
+\svp{XXX May want to have the error type table/chart as a visual representation? XXX}
+
 \svp{In this paper, we validate the algorithm described in \citet{aoas} on three external test sets.

### 2018-11-16 index.Rmd
second ten minute interval of the day

@@
 \item[(b)] Discard extraneous/contaminated data, such as data from groove engraved area and areas affected by break-off or contact with objects after the bullet exited the barrel  (``tank rash").
-\item[(c)] Remove bullet curvature \hh{using a Gaussian filter\cite{nist} or non-parametric smooth \citep{aoas}}. 
+\item[(c)] Remove bullet curvature \hh{using a Gaussian filter \cite{chu?} or non-parametric smooth \citep{aoas}}. 
 \end{itemize}
@@
 
-\svp{Once signatures have been extracted from stable regions with expressed striae, they must be aligned in order to assess their match strength.
-\citet{aoas} align signatures by maximizing the cross-correlation between the signatures.
+\svp{Once signatures have been extracted from stable regions with expressed striae, they must be aligned in order to assess their match strength.} \hh{Maximized cross-correlation is used to pair-wise align signatures \citet{aoas,vorburgerApplicationsCrosscorrelationFunctions2011}.
 }
@@
 The cross-correlation function (ccf) between two signatures $S_1(x)$ and $S_2(x)$ is defined as $$CC(S_1, S_2, z) = \text{cor}\left(S_1(x), S_2(x + z)\right)$$ where \hh{$x$ and $x+z$ are integer values appropriately defined within the domains of $S_1$ and $S_2$ and }  $z$ is the lag between the two sequences.
-We maximize $CC$ over values of $z$ to determine the best alignment  \citep{aoas,vorburgerApplicationsCrosscorrelationFunctions2011}; the maximal value of $CC$ is  used to assess the match strength between the two sequences.
+The lag $z$ used to achieve maximum $CC$ is used to determine the best alignment; the maximal value of $CC$ is  used to assess the match strength between the two sequences.
 When signatures contain missing values, pairwise complete observations are used to calculate the cross-correlation.
@@
 
-After two signatures are aligned, quantitative features describing the similarity of two signature are extracted.
+Once two signatures are aligned, quantitative features describing the similarity of two signature are extracted.
 Features include characteristics of the signature that firearms and toolmark examiners base their visual assessment on, such as striae depth, number of matching striae, and the number of consecutively matching striae (CMS).
@@
 
+\hh{XXX for CMS we need to define peaks, and matching peaks, and for that we need to define the additional filter/smooth.}
+
 \svp{The random forest presented in \citet{aoas} takes as input the following characteristics derived from a pair of aligned sequences:}
+\hh{We need to think what to do about the list - we could either just say that the random forest is based on multiple features, such as, ...
+and list three (and refer to the paper).
+Or we expand quite a bit and define the features in mathy detail using our notation above.}
 \svp{\begin{itemize}
@@
 # Assessing Match Strength
-\svp{Each of the features described in the previous section, as well as the random forest score, describe comparisons on a land-to-land basis.
-While these comparisons are useful, it is typically more useful to assess the entire physical object (e.g.
-all lands on a single bullet) to describe the correspondance between two pieces of physical evidence.}
+Each of the features described in the previous section, as well as the random forest score, describe comparisons on a land-to-land basis.
+While these comparisons are useful, \hh{the question of interest typically involves }  the entire physical object (e.g.
+all lands on a single bullet).
+
+<!--to describe the correspondence between two pieces of physical evidence.-->
 <!-- \svp{In order to aggregate land-to-land match scores, we must consider the physical structure of the evidence; that is, that the sequence of lands stems from spatial information, but the numbering of each land represents the scan order and start point rather than encoding useful information.

### 2018-11-16 index.Rmd
features

@@
 \svp{The random forest presented in \citet{aoas} takes as input the following characteristics derived from a pair of aligned sequences:}
+
 \hh{We need to think what to do about the list - we could either just say that the random forest is based on multiple features, such as, ...
@@
 Or we expand quite a bit and define the features in mathy detail using our notation above.}
+\hh{How about:}
+\hh{The random forest presented in \citet{aoas} is based on a combination of multiple characteristics, such as maximized cross-correlation, number of matching striae, and number of consecutively matching striae among others.
+}
+
 \svp{\begin{itemize}
@@
 \end{itemize}}
-\svp{The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where 1 indicates that the algorithm believes it to be highly likely that the aligned signatures match, and 0 indicates that the aligned signatures are highly likely to originate from different lands.}
+The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where scores close to 1 indicate that the algorithm believes it to be highly likely that the aligned signatures come from the same source, while low scores indicate that the aligned signatures are highly likely to originate from different lands.
 
@@
 While these comparisons are useful, \hh{the question of interest typically involves }  the entire physical object (e.g.
-all lands on a single bullet).
+all lands on a single bullet), \hh{and a conclusion on same or different source should be reached based on the evidence of all lands of the bullets}.
 

### 2018-11-19 index.Rmd
tweaks

@@
 These approaches utilize 3D surface measurements directly to characterize the topology of land engraved areas (LEAs), rather than using visual or image comparison techniques.
-In many cases, these papers also provide some level of automation of the comparison process, with the goal of reducing human biases by augmenting the visual information with 3D measurements.
-Utilizing the 3D measurements allows for examination not only of the peaks in the bullet signature, but also for comparison of the valleys, which are more difficult to assess visually except in relation to the peaks.
+In many cases, these \svp{approaches} also provide some level of automation of the comparison process, with the goal of reducing human biases by augmenting the visual information with 3D measurements.
+Utilizing the 3D measurements allows for examination \svp{of both the peaks and valleys in LEAs; visual examination of valleys is more difficult} except in relation to the peaks.
 
@@
 Individual characteristics, which are not shared by all members of a class, can also be automatically assessed from 3D surface measurements.
-One of the most common features used to assess the strength of a match based on individual characteristics is the cross-correlation function (ccf), which is described in several studies \citep{maNISTBulletSignature2004, vorburgerApplicationsCrosscorrelationFunctions2011, chuPilotStudyAutomated2010}.
-Additional features proposed for automatically assessing match strength also include signature distance \citep{maNISTBulletSignature2004}, striae depth and width \citep{ComputerIdentificationBullets1978}, and consecutive matching striae \citep{chuAutomaticIdentificationBullet2013}.
+One of the most common features used to assess the strength of a match based on individual characteristics is the cross-correlation function, which is \svp{utilized} in several studies \citep{maNISTBulletSignature2004, vorburgerApplicationsCrosscorrelationFunctions2011, chuPilotStudyAutomated2010}.
+Additional features proposed for automatically assessing match strength also include signature distance \citep{maNISTBulletSignature2004}, striae depth and width \citep{ComputerIdentificationBullets1978}, and consecutive matching striae (CMS) \citep{chuAutomaticIdentificationBullet2013}.
 
@@
 \item[(b)] Discard extraneous/contaminated data, such as data from groove engraved area and areas affected by break-off or contact with objects after the bullet exited the barrel  (``tank rash").
-\item[(c)] Remove bullet curvature \hh{using a Gaussian filter \cite{chu?} or non-parametric smooth \citep{aoas}}. 
+\item[(c)] Remove bullet curvature \hh{using a Gaussian filter \cite{chuAutomaticIdentificationBullet2013}\svp{This mentions use of a gaussian filter, but not for the specific purpose of removing curvature.
+Also, since we're discussing the aoas paper, might be better to stick to the non-parametric smooth?} or non-parametric smooth \citep{aoas}}.
 \end{itemize}
@@
 CMS in this setting are counted as the number of consecutive peaks in signatures of two lands that match\svp{, in part because it is difficult to visually assess valleys except as relative to peaks}. 
-\hh{Note that because signatures are moved into phase based on the maximum cross-correlation rather than matching striation marks,}\svp{it is possible for two lands to }\hh{have a consecutive matching striae of 0.
+\hh{Note that because signatures are moved into phase based on the maximum cross-correlation rather than matching striation marks,} \svp{it is possible for two lands to }\hh{have a consecutive matching striae of 0.
 } \svp{CMS are generally considered to be a very conservative criteria, in that bullets fired through the same barrel may not have a high number of consecutively matching striae; in this sense, CMS may be more likely to produce an inconclusive result.}
 
-\hh{XXX for CMS we need to define peaks, and matching peaks, and for that we need to define the additional filter/smooth.}
-
-\svp{The random forest presented in \citet{aoas} takes as input the following characteristics derived from a pair of aligned sequences:}
-
-\hh{We need to think what to do about the list - we could either just say that the random forest is based on multiple features, such as, ...
-and list three (and refer to the paper).
-Or we expand quite a bit and define the features in mathy detail using our notation above.}
-\hh{How about:}
+\hh{XXX for CMS we need to define peaks, and matching peaks, and for that we need to define the additional filter/smooth.}\svp{XXX Can't we just refer to the aoas paper?
+It seems to be defined in there relatively well.}
+
 \hh{The random forest presented in \citet{aoas} is based on a combination of multiple characteristics, such as maximized cross-correlation, number of matching striae, and number of consecutively matching striae among others.
@@
 
-\svp{\begin{itemize}
-\item Cross-correlation 
-\item Number of matching striae 
-\item Total depth of joint striations 
-\item Number of non-matching striae 
-\item Euclidian distance between the sequences 
-\item Consecutive Matching Striae 
-\item Consecutive Non-matching Striae
-\end{itemize}}
 The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where scores close to 1 indicate that the algorithm believes it to be highly likely that the aligned signatures come from the same source, while low scores indicate that the aligned signatures are highly likely to originate from different lands.
@@
 This should make the automatic matching harder, and in particular, \svp{is expected to complicate} the classification as different-source land engraved areas/bullets.
+\svp{In all three studies, the scan resolution is much higher than the scans used to train the algorithm in \citet{aoas}; this difference also provides a test of the algorithm's ability to generalize to scans taken at different resolutions.}
 

### 2018-11-19 index.Rmd
mini changes

@@
 Land-to-land comparisons lead to a whole set of scores for bullet-to-bullet comparisons.
-\autoref{fig:b2b} shows two matrices of scores for a pair of bullet-bullet matches.
+\autoref{fig:b2b} shows two matrices of scores for \hh{two pairs} of bullet-to-bullet matches.
 On the left, a matrix is shown that is typical for scores from two bullets from the same barrel.
@@
 The order in which scans are acquired is kept as meta-information.
-Let us assume that lands on a bullet are labelled $\ell_i$ with $i = 1, .., p$.
-A match between two bullets therefore results in an expected additional $p-1$ matches between pairs of lands.
+Let us assume that lands on a bullet are labelled $\ell_i$ with $i = 1, .., p$, \hh{where $p$ indicates the number of lands a bullet has, as determined by the rifling of a barrel.
+For all of the bullets considered here the number of lands $p$ is 6}.
+A match between  \hh{a pair of lands from two} bullets therefore results in an expected additional $p-1$ matches between pairs of lands.
 These lands are also expected to be in a sequence, i.e.
@@
 SAM scores allow us to define a single quantity for each pair of bullets that describes the similarity between these two bullets. 
-This leads us to two minimal requirements the scores should fulfill:
+This leads us to two minimal requirements the \hh{similarity} scores should fulfill:
 \begin{itemize}
@@
 ## Identification Errors
-\svp{XXX define wrong identification and missed identification here XXX}
-\svp{When evaluating algorithm performance, it is useful to distinguish between two types of errors, which we will refer to as \textbf{wrong identification} and \textbf{missed identification} in this paper.
+When evaluating algorithm performance, it is useful to distinguish between two types of errors, which we will refer to as \textbf{wrong identification} and \textbf{missed identification} in this paper.
 Wrong identifications are those in which two bullets from different sources are determined to be from the same source.
-Missed identifications are those in which two bullets from the same source are determined to be from different sources.}
+Missed identifications are those in which two bullets from the same source are determined to be from different sources.
+\hh{Note that this definition of error is more strict than the AFTE Theory of Identification, in that it does not allow for inconclusive results, but forces a decision based on a cutoff value. }
 \svp{XXX Relate to AFTE Theory of Examination criteria? How to handle ``inconclusive"?}
@@
 However, there is some discussion on whether or in which situations these error rates are biased \citep{rfOver2, rfOver}.
-Neither of these papers addresses another issue with internal test sets: internal test sets are constructed to have the same distribution as the training data (apart from sampling variability).} \svp{For any machine learning method, a true benchmark of the performance of an algorithm requires testing its' performance on external test data.} \hh{Good performance (in terms of low error rates) on external test data validates a ML algorithm.
+Neither of these papers addresses another issue with internal test sets: internal test sets are constructed to have the same distribution as the training data (apart from sampling variability).} \svp{For any machine learning method, a true benchmark of the performance of an algorithm requires testing its performance on external test data.} \hh{Good performance (in terms of low error rates) on external test data validates a ML algorithm.
 External data also allows an assessment of the algorithm's sensitivity to distributional changes} \svp{as well as testing the algorithm's robustness by going outside the parameters of the training data.

### 2018-11-19 index.Rmd
Fix spacing and extra comma

@@
 Ten questioned bullets were provided (B, E, H, J, K N, Q, T, Y, Z).
-This set is an open set; that is,\ it is not known in advance whether all (or any) of the questioned bullets are fired from the known barrels.\hh{ In fact, the results will show, that three of the questioned bullets were fired from three different barrels not included in the knowns.
+This set is an open set; that is,\ it is not known in advance whether all (or any) of the questioned bullets are fired from the known barrels.
+\hh{ In fact, the results will show that three of the questioned bullets were fired from three different barrels not included in the knowns.
 These three barrels correspond to an eleventh Ruger P-95, a Ruger P-95C and a Ruger P-85.} \svp{All bullets fired for this study are American Eagle 9mm  copper jacketed Luger.}

### 2018-11-19 index.Rmd
commas

@@
 This set is an open set; that is,\ it is not known in advance whether all (or any) of the questioned bullets are fired from the known barrels.
-\hh{ In fact, the results will show that three of the questioned bullets were fired from three different barrels not included in the knowns.
+\hh{In fact, the results will show that three of the questioned bullets were fired from three different barrels not included in the knowns.
 These three barrels correspond to an eleventh Ruger P-95, a Ruger P-95C and a Ruger P-85.} \svp{All bullets fired for this study are American Eagle 9mm  copper jacketed Luger.}
@@
 The light orange frame in the bottom three panels indicate which lands correspond to one-another in the matching.
-It is obvious, that not all of the scores are as high as we would usually see in same-source comparisons.
+It is obvious that not all of the scores are as high as we would usually see in same-source comparisons.
 ```

### 2018-11-19 index.Rmd
small edits

@@
 These approaches utilize 3D surface measurements directly to characterize the topology of land engraved areas (LEAs), rather than using visual or image comparison techniques.
-In many cases, these \svp{approaches} also provide some level of automation of the comparison process, with the goal of reducing human biases by augmenting the visual information with 3D measurements.
-Utilizing the 3D measurements allows for examination \svp{of both the peaks and valleys in LEAs; visual examination of valleys is more difficult} except in relation to the peaks.
+In many cases, these approaches also provide some level of automation of the comparison process, with the goal of reducing human biases by augmenting the visual information with 3D measurements.
+Utilizing the 3D measurements allows for examination of both peaks and valleys in LEAs, \hh{it also allows us to take the depth of striae into account, something that is difficult, if not impossible, for visual examination.}
 
@@
 Individual characteristics, which are not shared by all members of a class, can also be automatically assessed from 3D surface measurements.
-One of the most common features used to assess the strength of a match based on individual characteristics is the cross-correlation function, which is \svp{utilized} in several studies \citep{maNISTBulletSignature2004, vorburgerApplicationsCrosscorrelationFunctions2011, chuPilotStudyAutomated2010}.
+One of the most common features used to assess the strength of a match based on individual characteristics is the cross-correlation function, which is utilized in several studies \citep{maNISTBulletSignature2004, vorburgerApplicationsCrosscorrelationFunctions2011, chuPilotStudyAutomated2010}.
 Additional features proposed for automatically assessing match strength also include signature distance \citep{maNISTBulletSignature2004}, striae depth and width \citep{ComputerIdentificationBullets1978}, and consecutive matching striae (CMS) \citep{chuAutomaticIdentificationBullet2013}.
@@
 \item[(b)] Discard extraneous/contaminated data, such as data from groove engraved area and areas affected by break-off or contact with objects after the bullet exited the barrel  (``tank rash").
-\item[(c)] Remove bullet curvature \hh{using a Gaussian filter \cite{chuAutomaticIdentificationBullet2013}\svp{This mentions use of a gaussian filter, but not for the specific purpose of removing curvature.
-Also, since we're discussing the aoas paper, might be better to stick to the non-parametric smooth?} or non-parametric smooth \citep{aoas}}.
+\item[(c)] Remove bullet curvature using a non-parametric smooth. 
 \end{itemize}
@@
 
-\svp{Once signatures have been extracted from stable regions with expressed striae, they must be aligned in order to assess their match strength.} \hh{Maximized cross-correlation is used to pair-wise align signatures \citet{aoas,vorburgerApplicationsCrosscorrelationFunctions2011}.
-}
-
-The cross-correlation function (ccf) between two signatures $S_1(x)$ and $S_2(x)$ is defined as $$CC(S_1, S_2, z) = \text{cor}\left(S_1(x), S_2(x + z)\right)$$ where \hh{$x$ and $x+z$ are integer values appropriately defined within the domains of $S_1$ and $S_2$ and }  $z$ is the lag between the two sequences.
+Once signatures have been extracted from stable regions with expressed striae, they must be aligned in order to assess their match strength.
+Maximized cross-correlation is used to pair-wise align signatures \citet{aoas,vorburgerApplicationsCrosscorrelationFunctions2011}.
+
+The cross-correlation function (ccf) between two signatures $S_1(x)$ and $S_2(x)$ is defined as $$CC(S_1, S_2, z) = \text{cor}\left(S_1(x), S_2(x + z)\right)$$ where \hh{$x$ and $x+z$ are integer values appropriately defined within the domains of $S_1$ and $S_2$ and  $z$ is the lag between the two sequences.
 The lag $z$ used to achieve maximum $CC$ is used to determine the best alignment; the maximal value of $CC$ is  used to assess the match strength between the two sequences.
@@
 If there are at least 6 consecutive matching striae, the bullets are considered to be a good match \citep{biasotti_firearms_1997,nichols_scientific_2006}. 
-CMS in this setting are counted as the number of consecutive peaks in signatures of two lands that match\svp{, in part because it is difficult to visually assess valleys except as relative to peaks}. 
-\hh{Note that because signatures are moved into phase based on the maximum cross-correlation rather than matching striation marks,} \svp{it is possible for two lands to }\hh{have a consecutive matching striae of 0.
-} \svp{CMS are generally considered to be a very conservative criteria, in that bullets fired through the same barrel may not have a high number of consecutively matching striae; in this sense, CMS may be more likely to produce an inconclusive result.}
-
-\hh{XXX for CMS we need to define peaks, and matching peaks, and for that we need to define the additional filter/smooth.}\svp{XXX Can't we just refer to the aoas paper?
-It seems to be defined in there relatively well.}
-
-\hh{The random forest presented in \citet{aoas} is based on a combination of multiple characteristics, such as maximized cross-correlation, number of matching striae, and number of consecutively matching striae among others.
-}
+CMS in this setting are counted as the number of consecutive peaks in signatures of two lands that match, in part because it is difficult to visually assess valleys except as relative to peaks. 
+Note that because signatures are moved into phase based on the maximum cross-correlation rather than matching striation marks, it is possible for two lands to have a consecutive matching striae of 0.
+CMS are generally considered to be a very conservative criteria, in that bullets fired through the same barrel may not have a high number of consecutively matching striae; in this sense, CMS may be more likely to produce an inconclusive result.
+
+<!--\hh{XXX for CMS we need to define peaks, and matching peaks, and for that we need to define the additional filter/smooth.}\svp{XXX Can't we just refer to the aoas paper?
+It seems to be defined in there relatively well.}-->
+
+The random forest presented in \citet{aoas} is based on a combination of multiple characteristics, such as maximized cross-correlation, number of matching striae, and number of consecutively matching striae among others.
 
@@
 Missed identifications are those in which two bullets from the same source are determined to be from different sources.
-\hh{Note that this definition of error is more strict than the AFTE Theory of Identification, in that it does not allow for inconclusive results, but forces a decision based on a cutoff value. }
+\hh{Note that this definition of error is more stringent than the AFTE Theory of Identification, in that it does not allow for inconclusive results, but forces a decision based on a cutoff value. }
 \svp{XXX Relate to AFTE Theory of Examination criteria? How to handle ``inconclusive"?}
@@
 
-\svp{In this paper, we validate the algorithm described in \citet{aoas} on three external test sets.
+In this paper, we validate the algorithm described in \citet{aoas} on three external test sets.
 We assess the results on each set and evaluate whether the minimal requirements described above are fulfilled for bullet-to-bullet SAM scores.
 We compare the random forest algorithm's performance to the performance of other suggested measures for quantitative assessment of match strength, such as cross-correlation and consecutive matching striae.
-}
 
@@
 
-\svp{Random forests \citep{breiman} have} \hh{a built-in internal testing mechanism to prevent potential overfitting.
+Random forests \citep{breiman} have a built-in internal testing mechanism to prevent potential overfitting.
 Error rates reported by random forest algorithms are based on these internal test sets.
 However, there is some discussion on whether or in which situations these error rates are biased \citep{rfOver2, rfOver}.
-Neither of these papers addresses another issue with internal test sets: internal test sets are constructed to have the same distribution as the training data (apart from sampling variability).} \svp{For any machine learning method, a true benchmark of the performance of an algorithm requires testing its performance on external test data.} \hh{Good performance (in terms of low error rates) on external test data validates a ML algorithm.
-External data also allows an assessment of the algorithm's sensitivity to distributional changes} \svp{as well as testing the algorithm's robustness by going outside the parameters of the training data.
-}
+Neither of these papers addresses another issue with internal test sets: internal test sets are constructed to have the same distribution as the training data (apart from sampling variability).
+For any machine learning method, a true benchmark of the performance of an algorithm requires testing its performance on external test data.Good performance (in terms of low error rates) on external test data validates a ML algorithm.
+External data also allows an assessment of the algorithm's sensitivity to distributional changes as well as testing the algorithm's robustness by going outside the parameters of the training data.
 
@@
 
-\hh{For the validation of the automatic matching algorithm we are considering three validation sets of -- what should be -- increasing difficulty level:} 
+For the validation of the automatic matching algorithm we are considering three validation sets of -- what should be -- increasing difficulty level: 
 \begin{itemize}

### 2018-11-19 index.Rmd
rates to percentages

@@
 Once signatures have been extracted from stable regions with expressed striae, they must be aligned in order to assess their match strength.
-Maximized cross-correlation is used to pair-wise align signatures \citet{aoas,vorburgerApplicationsCrosscorrelationFunctions2011}.
+Maximized cross-correlation is used to pair-wise align signatures \citep{aoas,vorburgerApplicationsCrosscorrelationFunctions2011}.
 
@@
 Random forests \citep{breiman} have a built-in internal testing mechanism to prevent potential overfitting.
-Error rates reported by random forest algorithms are based on these internal test sets.
-However, there is some discussion on whether or in which situations these error rates are biased \citep{rfOver2, rfOver}.
+Errors reported by random forest algorithms are based on these internal test sets.
+However, there is some discussion on whether or in which situations these errors are biased \citep{rfOver2, rfOver}.
 Neither of these papers addresses another issue with internal test sets: internal test sets are constructed to have the same distribution as the training data (apart from sampling variability).
-For any machine learning method, a true benchmark of the performance of an algorithm requires testing its performance on external test data.Good performance (in terms of low error rates) on external test data validates a ML algorithm.
+For any machine learning method, a true benchmark of the performance of an algorithm requires testing its performance on external test data.
+Good performance (in terms of wrong and missed identifications) on external test data validates a ML algorithm.
 External data also allows an assessment of the algorithm's sensitivity to distributional changes as well as testing the algorithm's robustness by going outside the parameters of the training data.
@@
 These barrels are first rifled traditionally, i.e.
-similar to the Ruger P-85, but are treated to a secondary round of heating after \svp{rifling}.
+similar to the Ruger P-85, but are treated to a secondary round of heating after rifling.
 This second round of heat exposure has the potential to introduce some subclass characteristics, i.e.
 anomalies that are shared between different barrels manufactured in this manner.
-This should make the automatic matching harder, and in particular, \svp{is expected to complicate} the classification as different-source land engraved areas/bullets.
-\svp{In all three studies, the scan resolution is much higher than the scans used to train the algorithm in \citet{aoas}; this difference also provides a test of the algorithm's ability to generalize to scans taken at different resolutions.}
+This should make the automatic matching harder, and in particular, is expected to complicate the classification as different-source land engraved areas/bullets.
+In all three studies, the scan resolution is much higher than the scans used to train the algorithm in \citet{aoas}; this difference also provides a test of the algorithm's ability to generalize to scans taken at different resolutions.
 
@@
   ylab("Questioned bullets") +
-  xlab("Testfires from barrels 1 to 10 (left), questioned bullets (right)") +
+  xlab("Test fires from barrels 1 to 10 (left), questioned bullets (right)") +
   scale_fill_gradient2("RF score ",
@@
 
-\svp{When evaluating classifier performance, it is common to use a receiver operating curve, which plots the false positives against the true positives for each possible value of the cutoff between the two classes (in this case, known match and known non-match).} \svp{As the land-to-land scores are not perfectly separated, we can use ROC curves to evaluate their performance.} \svp{The ROC curves for the land-to-land scores are shown in \autoref{fig:roc-auc} (a).
-A perfect classifier would have 100\% true positives and 0\% false positives, e.g.
-the ROC curve would be a right angle at (0, 1).
+When evaluating classifier performance, it is common to use a receiver operating characteristic (ROC) curve, which plots the percent of wrong identifications against the percent correct identifications for each possible value of the cutoff between the two classes (in this case, known match and known non-match).
+Land-to-land scores are not perfectly separated, we can therefore use ROC curves to distinguish between the performance of the different methods and studies.
+The ROC curves for the land-to-land scores are shown in \autoref{fig:roc-auc} (a).
+A perfect classifier would have 100\% correct identifications and 0\% wrong identifications, e.g.
+the ROC curve would be a right angle at (0, 100).
 Classifiers with better performance will be closer to this corner of the plot.
-A random classifier would have an ROC curve that was a straight diagonal line through (0,0) and (1,1).}
+A random classifier would have an ROC curve that was a straight diagonal line through (0,0) and (100,100).
 
@@
   facet_wrap(~studyg, nrow=3) +
-  xlab("Fraction False Positives") +
-  ylab("Fraction True Positives") +
+  xlab("Percent Wrong Identifications") +
+  ylab("Percent Correct Identifications") +
   scale_colour_manual("Study", values = c("#fe9929", "#aa0000",
@@
   guides(linetype = guide_legend(override.aes = list(shape = NA, size = .5))) + 
-  scale_linetype_manual("Measure", values=c(1,3,2)) 
+  scale_linetype_manual("Measure", values=c(1,3,2)) +
+  scale_x_continuous(breaks=seq(0,1, by=0.25), labels=c("0", "25\\%", "50\\%", "75\\%", "100\\%")) +
+  scale_y_continuous(breaks=seq(0,1, by=0.25), labels=c("0", "25\\%", "50\\%", "75\\%", "100\\%"))
+
 ```
@@
 Based on AUC, the overall performance on all sets is very good to excellent for cross-correlation and the random forest score.
-At the land level there is no significant difference in prediction power between RF score and cross-correlation."}
+At the land level there is no significant difference in prediction power between RF score and cross-correlation.", cache=FALSE}
   rocpic
@@
 
-\hh{\autoref{fig:eer} shows an overview of Equal Error Rates (EER) and their thresholds based on the ROC curves of \autoref{fig:roc-auc}.
-EERs are error rates when sensitivity and specificity of a test are equal, i.e.
-we see the same error rates of missing an identification for a land-to-land comparison  and wrong identifications, when it is determined that two land engraved areas come from the same source when they are, in fact, known to be from a different source.
-Equal error rates based on CMS of two lands are significantly higher than equal error rates based on cross-correlation or the RF score.
-At best, RF score and cross-correlation have an equal error rate of around 5\% for land-to-land comparisons.
-Interestingly, for consecutively matching striae, an equal error rate is reached for just two matching striae, which from a practical point of view is not sustainable.
+\autoref{fig:eer} shows an overview of Equal Error and their thresholds based on the ROC curves of \autoref{fig:roc-auc}.
+Equal Errors are errors  when sensitivity and specificity of a test are equal, i.e.
+we see the same percentage of missed and wrong identifications for a land-to-land comparison.
+Equal errors based on CMS of two lands are significantly higher than equal errors based on cross-correlation or the RF score.
+At best, RF score and cross-correlation have an equal error of around 5\% for land-to-land comparisons.
+Interestingly, for consecutively matching striae, equal errors are reached for just two matching striae, which from a practical point of view is not sustainable.
 A  CMS of  6 or above is an established cutoff used in identifications.
-This leads to 21 wrong identifications (out of 49286; for an error rate of 4 in 10,000) but only identifies 139 out of  715 known same-source lands, for an error rate of 80.6\%.
-}
+This leads to 21 wrong identifications (out of 49286; for an error of 4 in 10,000 -- or 0.04\%) but only identifies 139 out of  715 known same-source lands, for an error of 80.6\%.
 ```{r eval=FALSE}
@@
 
-```{r eer, fig.width=7, fig.height = 4, fig.cap="Equal Error Rates and thresholds for all measures (RF score, Cross-correlation and CMS) under all five (sub-)studies.
-CMS has an EER threshold of 2 for all studies.
+```{r eer, fig.width=7, fig.height = 4, fig.cap="Equal Error (in Percent)  and corresponding thresholds for all measures (RF score, Cross-correlation and CMS) under all five (sub-)studies.
+CMS has an equal error threshold of 2 for all studies.
 CMS is signifcantly worse than both Cross-correlation and the RF score.
-At best, Cross-correlation and RF score have errors for land-to-land comparison of around 5\\%.\\svp{I'm uncomfortable using the same scale for EER threshold given that CMS has a different scale; leaving CMS out also seems odd though}", out.width = '.8\\textwidth'}
-aucs %>% gather(eer_thresh, value, eer, threshold) %>%
+At best, Cross-correlation and RF score have equal errors for land-to-land comparison of around 5\\%.", out.width = '.8\\textwidth'}
+aucs %>% 
+  mutate(
+    eer = 100*eer
+  ) %>%
+  gather(eer_thresh, value, eer, threshold) %>%
   filter(!(measure=="CMS" & eer_thresh=="threshold")) %>%
   mutate(
-    eer_thresh = factor(eer_thresh, labels=c("EER", "Threshold for EER"))
+    eer_thresh = factor(eer_thresh, labels=c("Equal Error (in Percent)", "Threshold for Equal Error"))
   ) %>%
@@
   geom_point(size = 4) +theme_bw() +
-#  xlim(c(0, .35)) +
-  xlab("Equal Error Rates (left) and EER thresholds (right)") +
+  xlab("Equal Error (in Percent, left) and corresponding thresholds (right)") +
   ylab("") +

### 2018-11-19 index.Rmd
table

--- 2018-11-19 index.Rmd
@@
 \hh{Note that this definition of error is more strict than the AFTE Theory of Identification, in that it does not allow for inconclusive results, but forces a decision based on a cutoff value. }
-\svp{XXX Relate to AFTE Theory of Examination criteria? How to handle ``inconclusive"?}
+
+\begin{table}
+\centering
+\begin{tabular}{c|ccc}
+\multicolumn{2}{c}{} & \multicolumn{2}{c}{Ground Truth}\\\cline{3-4}
+\multicolumn{2}{c}{} & \multicolumn{1}{|c}{Same Source} & \multicolumn{1}{|c|}{Different Source} \\\cline{2-4}
+\multirow{2}{*}{\rotatebox[origin=c]{90}{\hfillResults\hfill}} & 
+  \rotatebox[origin=c]{90}{\hspace{1mm}Exclusion\hspace{1mm}} & 
+  \multicolumn{1}{|c}{Missed ID} & 
+  \multicolumn{1}{|c|}{Correct Exclusion}\\\cline{2-4}
+& \rotatebox[origin=c]{90}{\hspace{1mm}Identification\hspace{1mm}} &
+\multicolumn{1}{|c}{Correct Identification} & \multicolumn{1}{|c|}{Wrong ID}\\\cline{2-4}
+\end{tabular}
+\caption{Something}
+\end{table}
 \svp{XXX May want to have the error type table/chart as a visual representation? XXX}
@@
 Equal error rates based on CMS of two lands are significantly higher than equal error rates based on cross-correlation or the RF score.
-At best, RF score and cross-correlation have an equal error rate of around 5\% for land-to-land comparisons.
+At best, RF score and cross-correlation have an equal error rate of around 5\%}\svp{(EER = 0.05)} \hh{ for land-to-land comparisons.
 Interestingly, for consecutively matching striae, an equal error rate is reached for just two matching striae, which from a practical point of view is not sustainable.
@@
 }
+
 ```{r eval=FALSE}

### 2018-11-19 index.Rmd
resolving merge

--- 2018-11-19 index.Rmd
@@
 Missed identifications are those in which two bullets from the same source are determined to be from different sources.
+\begin{table}
+\centering
+\begin{tabular}{c|ccc}
+\multicolumn{2}{c}{} & \multicolumn{2}{c}{Ground Truth}\\\cline{3-4}
+\multicolumn{2}{c}{} & \multicolumn{1}{|c}{Same Source} & \multicolumn{1}{|c|}{Different Source} \\\cline{2-4}
+\multirow{2}{*}{\rotatebox[origin=c]{90}{\hfill Results\hspace{7mm}\hfill}} & 
+  \rotatebox[origin=c]{90}{\hspace{1mm}Exclusion\hspace{1mm}} & 
+  \multicolumn{1}{|c}{Missed ID} & 
+  \multicolumn{1}{|c|}{Correct Exclusion}\\\cline{2-4}
+& \rotatebox[origin=c]{90}{\hspace{1mm}Identification\hspace{1mm}} &
+\multicolumn{1}{|c}{Correct Identification} & \multicolumn{1}{|c|}{Wrong ID}\\\cline{2-4}
+\end{tabular}
+\caption{Something}
+\end{table}
 \hh{Note that this definition of error is more stringent than the AFTE Theory of Identification, in that it does not allow for inconclusive results, but forces a decision based on a cutoff value. }
 \svp{XXX Relate to AFTE Theory of Examination criteria? How to handle ``inconclusive"?}
-\svp{XXX May want to have the error type table/chart as a visual representation? XXX}
 
@@
   geom_segment(aes(x=auc_ci_low, xend=auc_ci_high, y = measurenum, yend = measurenum, linetype=measure), size=1) +
-  geom_point() +
+  geom_point(size=3) +
   theme_bw() +
@@
 This leads to 21 wrong identifications (out of 49286; for an error of 4 in 10,000 -- or 0.04\%) but only identifies 139 out of  715 known same-source lands, for an error of 80.6\%.
+
 ```{r eval=FALSE}

### 2018-11-19 index.Rmd
shapes added

@@
   geom_segment(aes(x=auc_ci_low, xend=auc_ci_high, y = measurenum, yend = measurenum, linetype=measure), size=1) +
-  geom_point(size=3) +
+  geom_point(size=3.5, aes(shape = measure)) +
   theme_bw() +
@@
   #       legend.key.width = unit(1.5, "cm")) +
-  scale_linetype_manual("Measure", values = c(3,1,2), guide = F)
+  scale_linetype_manual("Measure", values = c(3,1,2), guide = F) +
+  scale_shape_discrete(guide=F)
 
@@
   ggplot(aes(y = studyg, x = value, colour=measure, shape = measure)) + 
-  geom_point(size = 4) +theme_bw() +
+  geom_point(size = 3.5) +
+  theme_bw() +
   xlab("Equal Error (in Percent, left) and corresponding thresholds (right)") +

### 2018-11-20 index.Rmd
Additional text, fixing formatting

@@
 ## Identification Errors
-When evaluating algorithm performance, it is useful to distinguish between two types of errors, which we will refer to as \textbf{wrong identification} and \textbf{missed identification} in this paper.
+When evaluating algorithm performance, it is useful to \svp{enumerate the set of possible outcomes}.
+If ground truth and the algorithm's prediction match, we have either a correct identification or a correct exclusion.
+If ground truth and the algorithm's prediction do not match, we} distinguish between two types of errors, which we will refer to as \textbf{wrong identification} and \textbf{missed identification}.
 Wrong identifications are those in which two bullets from different sources are determined to be from the same source.
 Missed identifications are those in which two bullets from the same source are determined to be from different sources.
+\svp{The full range of possible outcomes is shown in \autoref{tab:possible-outcomes}.}
 \begin{table}
@@
 \begin{tabular}{c|ccc}
-\multicolumn{2}{c}{} & \multicolumn{2}{c}{Ground Truth}\\\cline{3-4}
+\multicolumn{2}{c}{} & \multicolumn{2}{c}{\textbf{Ground Truth}}\\\cline{3-4}
 \multicolumn{2}{c}{} & \multicolumn{1}{|c}{Same Source} & \multicolumn{1}{|c|}{Different Source} \\\cline{2-4}
-\multirow{2}{*}{\rotatebox[origin=c]{90}{\hfill Results\hspace{7mm}\hfill}} & 
-  \rotatebox[origin=c]{90}{\hspace{1mm}Exclusion\hspace{1mm}} & 
-  \multicolumn{1}{|c}{Missed ID} & 
-  \multicolumn{1}{|c|}{Correct Exclusion}\\\cline{2-4}
-& \rotatebox[origin=c]{90}{\hspace{1mm}Identification\hspace{1mm}} &
-\multicolumn{1}{|c}{Correct Identification} & \multicolumn{1}{|c|}{Wrong ID}\\\cline{2-4}
+\multirow{8}{*}{\rotatebox[origin=c]{90}{\hfill \textbf{Results}\hspace{6mm}\hfill}} & 
+  \multirow{4}{*}{\rotatebox[origin=c]{90}{\hspace{1mm}Exclusion\hspace{1mm}}} & \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\
+  & &  \multicolumn{1}{|c}{Missed} & \multicolumn{1}{|c|}{Correct}\\
+  & &  \multicolumn{1}{|c}{Identification} & \multicolumn{1}{|c|}{Exclusion}\\
+  & &  \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{} \\\cline{2-4}
+& \multirow{4}{*}{\rotatebox[origin=c]{90}{\hspace{1mm}Identification\hspace{1mm}}}  & \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\
+  & &  \multicolumn{1}{|c}{Correct} & \multicolumn{1}{|c|}{Wrong}\\
+  & &  \multicolumn{1}{|c}{Identification} & \multicolumn{1}{|c|}{Identification}\\
+  & &  \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{} \\\cline{2-4}
 \end{tabular}
-\caption{Something}
+\caption{Possible outcomes of an examination of two pieces of evidence.\label{tab:possible-outcomes}}
 \end{table}
-\hh{Note that this definition of error is more stringent than the AFTE Theory of Identification, in that it does not allow for inconclusive results, but forces a decision based on a cutoff value. }
-\svp{XXX Relate to AFTE Theory of Examination criteria? How to handle ``inconclusive"?}
+\hh{Note that this definition of error is more stringent than the AFTE Theory of Identification, in that it does not allow for inconclusive results, but forces a decision based on a cutoff value.
+} \svp{This increased rigor may increase the error rate of the model when compared to examiner error rates, but we expect that the increased information available from the 3D measurement data will compensate for this loss.
+A binary decision model is also easier to interpret and provides more clear-cut, definitive results.}
+
+\svp{It is likely that matches which were deemed  inconclusive would instead be marked as exclusions by the algorithm.
+XXX I don't know if we can make this statement, but I think it would help examiners understand how the line gets drawn if we could.
+I have exactly no evidence to back that up (except maybe the Hamby 44 tank rash results?)}
 

### 2018-11-20 index.Rmd
slightly longer caption

@@
 \end{tabular}
-\caption{Possible outcomes of an examination of two pieces of evidence.\label{tab:possible-outcomes}}
+\caption{Possible outcomes of an examination of two pieces of evidence.
+Correct decisions are shown in the top-right and bottom-left corners; incorrect decisions in the top-left and bottom-right corners.\label{tab:possible-outcomes}}
 \end{table}

### 2018-11-20 index.Rmd
restructuring, add sensitivity and specificity

@@
 # Introduction and Background
+
 In current practice, firearms and toolmark examiners (FTE) evaluate the similarity of striae on bullets by placing the evidence together with another bullet under a comparison microscope.
@@
 ## Identification Using 3D Imaging Technology
+
 More explicit characterization of bullet surfaces has been discussed since at least 1958 (\citeauthor{davis1968introduction}), but at the time technology was not sufficiently advanced to make an analysis based on 3D measurements or surface traces a viable option.
@@
 
+## Automated Processing of 3D Scans
+
 In the matching algorithm proposed in \citet{aoas}, several of the features described above are combined using a random forest \citep{breiman} to produce a similarity score based on 3d topographic scans of land engraved areas (LEAs).
@@
 <!-- $z(t)$ is then a spatial process indexed by $t$, where $t, t = 1,...,T$ denotes equally spaced pixel locations for a total length of $T$. } -->
-
 \begin{itemize}
@@
 \autoref{fig:signatures} shows a set of six signatures corresponding to the six land engraved areas of bullets 1 and 2 from barrel 1 of set 44 of the Hamby study \citep{hamby}.
-
 ```{r signatures, fig.cap="Signatures of all six lands of bullets 1 and 2 from barrel 1 of the Hamby set 44.", fig.height=4, fig.width=10}
@@
 
-The cross-correlation function (ccf) between two signatures $S_1(x)$ and $S_2(x)$ is defined as $$CC(S_1, S_2, z) = \text{cor}\left(S_1(x), S_2(x + z)\right)$$ where \hh{$x$ and $x+z$ are integer values appropriately defined within the domains of $S_1$ and $S_2$ and  $z$ is the lag between the two sequences.
-The lag $z$ used to achieve maximum $CC$ is used to determine the best alignment; the maximal value of $CC$ is  used to assess the match strength between the two sequences.
-When signatures contain missing values, pairwise complete observations are used to calculate the cross-correlation.
+The cross-correlation function (ccf) between two signatures $S_1(x)$ and $S_2(x)$ is defined as \[CC(S_1, S_2, z) = \text{cor}\left(S_1(x), S_2(x + z)\right)\] where \hh{$x$ and $x+z$ are integer values appropriately defined within the domains of $S_1$ and $S_2$ and  $z$ is the lag between the two sequences.
+The lag $z$ used to achieve maximum $CC$ is used to determine the best alignment; the maximal value of $CC$ is  used to assess the match strength between the two sequences.} When signatures contain missing values, pairwise complete observations are used to calculate the cross-correlation.
 
 ## Statistics for Matching Aligned Signatures
+
 <!-- \hh{XXX define signatures mathematically here - as random spatial process; extracted from crosscut by removing big structure first and then applying (cross-validated, \citet{aoas}) 'roughness' filter.
@@
 CMS are generally considered to be a very conservative criteria, in that bullets fired through the same barrel may not have a high number of consecutively matching striae; in this sense, CMS may be more likely to produce an inconclusive result.
-
 <!--\hh{XXX for CMS we need to define peaks, and matching peaks, and for that we need to define the additional filter/smooth.}\svp{XXX Can't we just refer to the aoas paper?
 It seems to be defined in there relatively well.}-->
+\svp{XXX Feels like we need to mention something about ccf here just because these are the 3 main features we're using to evaluate the algorithm...?}
 
 The random forest presented in \citet{aoas} is based on a combination of multiple characteristics, such as maximized cross-correlation, number of matching striae, and number of consecutively matching striae among others.
-
 The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where scores close to 1 indicate that the algorithm believes it to be highly likely that the aligned signatures come from the same source, while low scores indicate that the aligned signatures are highly likely to originate from different lands.
 
-# Assessing Match Strength
+## Assessing Match Strength
+
 Each of the features described in the previous section, as well as the random forest score, describe comparisons on a land-to-land basis.
@@
 all lands on a single bullet), \hh{and a conclusion on same or different source should be reached based on the evidence of all lands of the bullets}.
-
 <!--to describe the correspondence between two pieces of physical evidence.-->
@@
 
-## Land-to-Land Comparisons
+### Land-to-Land Comparisons
 
@@
 
-## Bullet-to-Bullet Comparisons
+### Bullet-to-Bullet Comparisons
 
@@
 
-## Identification Errors
+## Identification Errors and Algorithm Evaluation
 When evaluating algorithm performance, it is useful to \svp{enumerate the set of possible outcomes}.
 If ground truth and the algorithm's prediction match, we have either a correct identification or a correct exclusion.
-If ground truth and the algorithm's prediction do not match, we} distinguish between two types of errors, which we will refer to as \textbf{wrong identification} and \textbf{missed identification}.
+If ground truth and the algorithm's prediction do not match, we distinguish between two types of errors, which we will refer to as \textbf{wrong identification} and \textbf{missed identification}.
 Wrong identifications are those in which two bullets from different sources are determined to be from the same source.
@@
 I have exactly no evidence to back that up (except maybe the Hamby 44 tank rash results?)}
+
+\svp{In machine learning and statistics, it is common to evaluate algorithms using sensitivity and specificity; these concepts are related to the error rate.
+\textbf{Sensitivity} is defined as the proportion of actual positives that are correctly identified; that is, when ground truth is same source, the sensitivity is the proportion of correct identifications.
+\textbf{Specificity} is defined as the proportion of actual negatives that are correctly identified; that is, when ground truth is different source, the specificity is the proportion of correct identifications.
+The combination of sensitivity and specificity is sufficient to describe the reliability of an algorithm.
+}
 

### 2018-11-20 index.Rmd
comment
fixing up small things - cross correlation as CC instead of ccf, adding explanation of ROC/AUC on a 0-1 scale, directly referencing R1 and R2 instead of referencing the section, slight phrasing changes

@@
 
-The cross-correlation function (ccf) between two signatures $S_1(x)$ and $S_2(x)$ is defined as \[CC(S_1, S_2, z) = \text{cor}\left(S_1(x), S_2(x + z)\right)\] where \hh{$x$ and $x+z$ are integer values appropriately defined within the domains of $S_1$ and $S_2$ and  $z$ is the lag between the two sequences.
+The cross-correlation (CC) function between two signatures $S_1(x)$ and $S_2(x)$ is defined as \[CC(S_1, S_2, z) = \text{cor}\left(S_1(x), S_2(x + z)\right)\] where \hh{$x$ and $x+z$ are integer values appropriately defined within the domains of $S_1$ and $S_2$ and  $z$ is the lag between the two sequences.
 The lag $z$ used to achieve maximum $CC$ is used to determine the best alignment; the maximal value of $CC$ is  used to assess the match strength between the two sequences.} When signatures contain missing values, pairwise complete observations are used to calculate the cross-correlation.
@@
 This leads us to two minimal requirements the \hh{similarity} scores should fulfill:
-\begin{itemize}
+\begin{description}
 \item[(R1)] {\bf Monotonicity:} a higher score is indicative of higher similarity between a pair of bullets, in particular, similarity scores of same-source pairs of bullets are higher than different-source pairs.
 \item[(R2)] {\bf Stability:} the same score leads to the same conclusion. 
-\end{itemize}
+\end{description}
 
@@
 }
+
+\svp{When evaluating the \citet{aoas} algorithm, we will describe errors as a percentage of correct evaluations.
+Some machine learning tools, such as ROC curves and AUC, will be presented in the decimal (rate) form because they do not easily translate into a percent-form interpretation.
+}
+
+\svp{XXX Should we have some sort of header here to differentiate it from the error section?}
 
@@
 \svp{The random forest SAM scores correctly separate known matches from known non-matches, using SAM scores to aggregate land-to-land scores into a bullet-to-bullet comparison.
-With no errors at the bullet-to-bullet level, we now need to evaluate the scores in light of the criteria described in \autoref{land-to-land-comparisons}.
+With no errors at the bullet-to-bullet level, we now need to evaluate the scores in light of (R1) and (R2).
 In this section, we will compare the random forest scores to cross-correlation scores; cross-correlation is one of the components of the random forest, but has also been used to quantify the strength of a match in its own right \citep{chuPilotStudyAutomated2010}.
@@
 
-\svp{XXX We may want to add in SAM scores for CMS as well, but that's going to make the the scales on the density plots (and the density plots themselves) a bit wonky. Thoughts?}
+<!-- \svp{XXX We may want to add in SAM scores for CMS as well, but that's going to make the the scales on the density plots (and the density plots themselves) a bit wonky. Thoughts?} -->
 
@@
 
-\svp{According to \autoref{fig:compare}, neither cross-correlation nor random forest score} fulfill both of the minimal requirements laid out \svp{in \autoref{land-to-land-comparisons}}.
+\svp{According to \autoref{fig:compare}, neither cross-correlation nor random forest score} fulfill both of the minimal requirements laid out \svp{in the introduction}.
 Both measures fulfill (R1), i.e.
@@
 
-Regarding requirement (R2), \autoref{fig:compare} shows that a single cutoff value does not exist that separates same-source scores from different-source scores across all three studies, without introducing errors.
+Regarding requirement (R2), \autoref{fig:compare} shows that a single cutoff value does not exist that separates same-source scores from different-source scores across all three studies without introducing errors.
 
@@
 \svp{Examining land-to-land scores provides more details about the matching algorithm's performance.
-\autoref{fig:compare-land-to-land} shows density curves and rug plots for each study.
-There is some overlap between known matching and known non-matching land-to-land scores; in many cases the random forest scores for these overlapping values are more extreme than the corresponding ccf scores.
+\autoref{fig:compare-land-to-land} shows density curves for each study.
+There is some overlap between known matching and known non-matching land-to-land scores; in many cases the random forest scores for these overlapping values are more extreme than the corresponding cross-correlation scores.
 This suggests that the matching algorithm is sensitive to the presence of LEAs which have low land-to-land match scores, such as the matches of bullets `I' and `F' in the Hamby 44 study.

### 2018-11-20 index.Rmd
fiddling with figure 9

@@
 Let's use instead wrong identification and missed identification. 
-\item error rates: we are using rates exchangeably with percentages. We need to fix that. I think percentages might be easier understood. 
+\item error rates: we are using rates exchangeably with percentages. We need to fix that. I think percentages might be easier understood. XXX Should be fixed now
 \end{itemize}
@@
 
-The cross-correlation (CC) function between two signatures $S_1(x)$ and $S_2(x)$ is defined as \[CC(S_1, S_2, z) = \text{cor}\left(S_1(x), S_2(x + z)\right)\] where \hh{$x$ and $x+z$ are integer values appropriately defined within the domains of $S_1$ and $S_2$ and  $z$ is the lag between the two sequences.
-The lag $z$ used to achieve maximum $CC$ is used to determine the best alignment; the maximal value of $CC$ is  used to assess the match strength between the two sequences.} When signatures contain missing values, pairwise complete observations are used to calculate the cross-correlation.
+The cross-correlation (CC) function between two signatures $S_1(x)$ and $S_2(x)$ is defined as \[CC(S_1, S_2, z) = \text{cor}\left(S_1(x), S_2(x + z)\right)\] where $x$ and $x+z$ are integer values appropriately defined within the domains of $S_1$ and $S_2$ and  $z$ is the lag between the two sequences.
+The lag $z$ used to achieve maximum $CC$ is used to determine the best alignment; the maximal value of $CC$ is  used to assess the match strength between the two sequences.
+When signatures contain missing values, pairwise complete observations are used to calculate the cross-correlation.
 
@@
 ## Identification Errors and Algorithm Evaluation
-When evaluating algorithm performance, it is useful to \svp{enumerate the set of possible outcomes}.
-If ground truth and the algorithm's prediction match, we have either a correct identification or a correct exclusion.
+When evaluating algorithm performance, it is useful to \hh{systematically assess }\svp{ the set of possible outcomes} \hh{first}: if ground truth and the algorithm's prediction match, we have either a correct identification or a correct exclusion.
 If ground truth and the algorithm's prediction do not match, we distinguish between two types of errors, which we will refer to as \textbf{wrong identification} and \textbf{missed identification}.
@@
 Missed identifications are those in which two bullets from the same source are determined to be from different sources.
-\svp{The full range of possible outcomes is shown in \autoref{tab:possible-outcomes}.}
+The full range of possible outcomes is shown in \autoref{tab:possible-outcomes}.
 \begin{table}
@@
 \svp{When evaluating the \citet{aoas} algorithm, we will describe errors as a percentage of correct evaluations.
-Some machine learning tools, such as ROC curves and AUC, will be presented in the decimal (rate) form because they do not easily translate into a percent-form interpretation.
+Statistics that do not translate to percent-form interpretation, such as area under the curve (AUC), will be presented as a decimal.
 }
@@
 A perfect classifier would have 100\% correct identifications and 0\% wrong identifications, e.g.
-the ROC curve would be a right angle at (0, 100).
+the ROC curve would be a right angle at $(0, 100)$.
 Classifiers with better performance will be closer to this corner of the plot.
-A random classifier would have an ROC curve that was a straight diagonal line through (0,0) and (100,100).
+A random classifier would have an ROC curve that was a straight diagonal line through $(0,0)$ and $(100,100)$.
 
@@
   geom_roc(aes(m = score, d = as.numeric(sameland)), 
-           labels = FALSE, pointsize = 0, alpha = 0.8, size = .75) +
+           labels = FALSE, pointsize = 0, alpha = 0.8, size = 1) +
   theme_bw() +
@@
   scale_linetype_manual("Measure", values=c(1,3,2)) +
-  scale_x_continuous(breaks=seq(0,1, by=0.25), labels=c("0", "25\\%", "50\\%", "75\\%", "100\\%")) +
-  scale_y_continuous(breaks=seq(0,1, by=0.25), labels=c("0", "25\\%", "50\\%", "75\\%", "100\\%"))
+  scale_x_continuous(breaks=seq(0,1, by=0.25), labels=c("0", "25%", "50%", "75%", "100%")) +
+  scale_y_continuous(breaks=seq(0,1, by=0.25), labels=c("0", "25%", "50%", "75%", "100%"))
 
@@
   xlim(c(0.5, 1)) +
-  scale_y_continuous("", breaks = unique(aucs$measurenum), labels = unique(aucs$measurechar)) + 
+  scale_y_continuous("", breaks = unique(aucs$measurenum), labels = unique(aucs$measurechar), limits=c(0.5, 3.5)) + 
   scale_colour_manual("Study", values = c("#fe9929", "#aa0000",
@@
 
-```{r roc-auc, echo = FALSE, out.width = c("0.49\\textwidth", ".45\\textwidth"), out.height = c("0.6\\textwidth", "0.6\\textwidth"), fig.width = c(5, 6), fig.height = c(7.5, 7.5), fig.subcap=c("ROC curves", "Area under the curve (AUC)"), fig.cap="ROC curves and AUC for each (set) of the studies for the random forest score (RF), cross-correlation (CC) and consecutively matching striae (CMS).
+```{r roc-auc, echo = FALSE, out.width = c("0.51\\textwidth", ".4725\\textwidth"), out.height = c("0.63\\textwidth", "0.63\\textwidth"), fig.width = c(5, 6), fig.height = c(7.5, 7.5), fig.subcap=c("ROC curves", "Area under the curve (AUC)"), fig.cap="ROC curves and AUC for each (set) of the studies for the random forest score (RF), cross-correlation (CC) and consecutively matching striae (CMS).
 As seen in the ROC curves, the algorithm performs the least well on set 2 of the Houston FSI study.
@@
 Interestingly, for consecutively matching striae, equal errors are reached for just two matching striae, which from a practical point of view is not sustainable.
-A  CMS of  6 or above is an established cutoff used in identifications.
-This leads to 21 wrong identifications (out of 49286; for an error of 4 in 10,000 -- or 0.04\%) but only identifies 139 out of  715 known same-source lands, for an error of 80.6\%.
+\svp{A  CMS of  6 or above is an established cutoff used in identifications; when applied to the land-to-land comparisons, it is much more likely to result in a missed identification than a wrong identification.
+A threshold of 6 leads to 21 wrong identifications out of 49,286 different-source comparisons (0.04\% errors, or about 4 errors for each 10,000 evaluations); when applied to same-source comparisons, 517 of 715 same-source lands were incorrectly classified as exclusions, for an error rate of 80.6\%.
+}
+<!-- This leads to 21 wrong identifications (out of 49286; for an error of 4 in 10,000 -- or 0.04\%) but only identifies 139 out of 715 known same-source lands, for an error of 80.6\%.  -->
 
@@
 \svp{The random forest matching algorithm presented in \citet{aoas} performs well on three different external test sets.
-While a single cutoff value cannot be used to distinguish matches and non-matches across the different test sets, the algorithm makes no errors when a set-specific cutoff value is used on bullet-to-bullet scores aggregated using sequence average maximum.
-XXX Need more here XXX}
+While a single cutoff value cannot be used to distinguish matches and non-matches across the different test sets, the algorithm makes no errors when a set-specific cutoff value is used on bullet-to-bullet scores aggregated using sequence average maximum.} \svp{XXX more here XXX}
 

### 2018-11-20 index.Rmd
reorder

@@
 The cross-correlation (CC) function between two signatures $S_1(x)$ and $S_2(x)$ is defined as \[CC(S_1, S_2, z) = \text{cor}\left(S_1(x), S_2(x + z)\right)\] where \hh{$x$ and $x+z$ are integer values appropriately defined within the domains of $S_1$ and $S_2$ and  $z$ is the lag between the two sequences.
-The lag $z$ used to achieve maximum $CC$ is used to determine the best alignment; the maximal value of $CC$ is  used to assess the match strength between the two sequences.} When signatures contain missing values, pairwise complete observations are used to calculate the cross-correlation.
+When signatures contain missing values, pairwise complete observations are used to calculate the cross-correlation.
+The lag $z$ used to achieve maximum $CC$ is used to determine the best alignment.}
 
 ## Statistics for Matching Aligned Signatures
-
-<!-- \hh{XXX define signatures mathematically here - as random spatial process; extracted from crosscut by removing big structure first and then applying (cross-validated, \citet{aoas}) 'roughness' filter.
-'roughness' defined by NIST paper.
-} -->
 
@@
 One of the measures firearms examiners use to describe the strength of a match quantitatively is the number of consecutive matching striae (CMS) as defined by \citet{biasotti}.
+\svp{In order to evaluate CMS, we must first identify peaks and valleys in each of the signatures, then determine whether these peaks and valleys overlap sufficiently.
+Features which firearms examiners assess visually depend on the identification of these extrema; numerical features such as cross-correlation can be computed from the aligned signatures alone.
+}
+
 If there are at least 6 consecutive matching striae, the bullets are considered to be a good match \citep{biasotti_firearms_1997,nichols_scientific_2006}. 
@@
 
+; the maximal value of $CC$ is  used to assess the match strength between the two sequences
+
 The random forest presented in \citet{aoas} is based on a combination of multiple characteristics, such as maximized cross-correlation, number of matching striae, and number of consecutively matching striae among others.
@@
 
-### Land-to-Land Comparisons
+### Bullet Structure
 
@@
 
-\svp{XXX This is a bit of an incomplete thought. I think I know where you were going, but can you expand on it a bit so that I'm not putting words in your mouth? XXX}
-\citet{Morrison:2018km}  stress the need to additionally include the typicality of a signature.
-
 ## Digging deeper: comparison of land-to-land scores

### 2018-11-20 index.Rmd

--- 2018-11-20 index.Rmd
@@
 The cross-correlation (CC) function between two signatures $S_1(x)$ and $S_2(x)$ is defined as \[CC(S_1, S_2, z) = \text{cor}\left(S_1(x), S_2(x + z)\right)\] where $x$ and $x+z$ are integer values appropriately defined within the domains of $S_1$ and $S_2$ and  $z$ is the lag between the two sequences.
-The lag $z$ used to achieve maximum $CC$ is used to determine the best alignment; the maximal value of $CC$ is  used to assess the match strength between the two sequences.
 When signatures contain missing values, pairwise complete observations are used to calculate the cross-correlation.
+The lag $z$ used to achieve maximum $CC$ is used to determine the best alignment.
 
 ## Statistics for Matching Aligned Signatures
-
-<!-- \hh{XXX define signatures mathematically here - as random spatial process; extracted from crosscut by removing big structure first and then applying (cross-validated, \citet{aoas}) 'roughness' filter.
-'roughness' defined by NIST paper.
-} -->
 

### 2018-11-20 index.Rmd
restructuring

@@
 CMS in this setting are counted as the number of consecutive peaks in signatures of two lands that match, in part because it is difficult to visually assess valleys except as relative to peaks. 
-Note that because signatures are moved into phase based on the maximum cross-correlation rather than matching striation marks, it is possible for two lands to have a consecutive matching striae of 0.
-CMS are generally considered to be a very conservative criteria, in that bullets fired through the same barrel may not have a high number of consecutively matching striae; in this sense, CMS may be more likely to produce an inconclusive result.
+Note that \svp{with automatic alignment }based on the maximum cross-correlation rather than matching striation marks, it is possible for two lands to have a consecutive matching striae of 0. 
+<!-- CMS are generally considered to be a very conservative criteria, in that bullets fired through the same barrel may not have a high number of consecutively matching striae; in this sense, CMS may be more likely to produce an inconclusive result.
+-->
 <!--\hh{XXX for CMS we need to define peaks, and matching peaks, and for that we need to define the additional filter/smooth.}\svp{XXX Can't we just refer to the aoas paper?
 It seems to be defined in there relatively well.}-->
-\svp{XXX Feels like we need to mention something about ccf here just because these are the 3 main features we're using to evaluate the algorithm...?}
-
-; the maximal value of $CC$ is  used to assess the match strength between the two sequences
-
-The random forest presented in \citet{aoas} is based on a combination of multiple characteristics, such as maximized cross-correlation, number of matching striae, and number of consecutively matching striae among others.
+<!-- \svp{XXX Feels like we need to mention something about ccf here just because these are the 3 main features we're using to evaluate the algorithm...?} -->
+
+\svp{Cross correlation can also be used to assess match strength between two aligned signatures; this cross-correlation value corresponds to the maximal cross-correlation over all valid lag values $z$.
+Cross correlation varies between -1 and 1, but as we are taking the maximal value, in practice the cross-correlations for aligned signatures are generally positive.
+Cross-correlation values which are close to 1 indicate signatures which are similar and may originate from the same source.
+Cross-correlation values which are close to 0 indicate signatures which may originate from different sources.}
+
+The random forest presented in \citet{aoas} is based on a combination of multiple characteristics, including cross-correlation, number of matching striae, and number of consecutively matching striae.
 The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where scores close to 1 indicate that the algorithm believes it to be highly likely that the aligned signatures come from the same source, while low scores indicate that the aligned signatures are highly likely to originate from different lands.
 
-## Assessing Match Strength
-
-Each of the features described in the previous section, as well as the random forest score, describe comparisons on a land-to-land basis.
-While these comparisons are useful, \hh{the question of interest typically involves }  the entire physical object (e.g.
-all lands on a single bullet), \hh{and a conclusion on same or different source should be reached based on the evidence of all lands of the bullets}.
-<!--to describe the correspondence between two pieces of physical evidence.-->
-<!-- \svp{In order to aggregate land-to-land match scores, we must consider the physical structure of the evidence; that is, that the sequence of lands stems from spatial information, but the numbering of each land represents the scan order and start point rather than encoding useful information.
-} --><!--This last bit might be redundant-->
-
-```{r, echo=FALSE, results = 'hide', warning = FALSE, message = FALSE}
-features <- read.csv("data/fsi-features.csv.gz")
-features <- features %>% mutate(
-  bullet1 = gsub("ullet ", "", bullet1),
-  bullet2 = gsub("ullet ", "", bullet2)
-)
-b12 <- features %>% filter(
-  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
-  group2 == "G1", barrel2 == "KA", bullet2 == "B2"
-)
-p1 <- b12 %>%
-  ggplot(aes(x = land1, y = land2, fill = rfscore)) +
-  geom_tile() + ggtitle("Same barrel") +
-  scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
-  theme_bw() +
-  xlab("Bullet 2") + ylab("Bullet 1") +
-  theme(legend.position = "bottom") + coord_equal()
-
-bab <- features %>% filter(
-  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
-  group2 == "G1", barrel2 == "KB", bullet2 == "B1"
-)
-p2 <- bab %>%
-  ggplot(aes(x = land1, y = land2, fill = rfscore)) +
-  geom_tile() + ggtitle("Different barrel") +
-  scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
-  theme_bw() +
-  xlab("Bullet 2") + ylab("Bullet 1") +
-  theme(legend.position = "bottom") + coord_equal()
-```
-
-### Bullet Structure
-
-Land-to-land comparisons lead to a whole set of scores for bullet-to-bullet comparisons.
-\autoref{fig:b2b} shows two matrices of scores for \hh{two pairs} of bullet-to-bullet matches.
-On the left, a matrix is shown that is typical for scores from two bullets from the same barrel.
-On the right, values for a pair of known non-matching bullets are shown.
-
-```{r b2b, echo = FALSE, fig.width = 9, fig.height = 5,  out.width='.6\\textwidth', fig.cap="Overview of land-land matching scores for two pairs of bullet-bullet comparisons.
-On the left the two bullets are known to come from the same source (barrel) on the right, the bullets are from two different sources (barrels)"}
-grid.arrange(p1, p2, ncol = 2, padding = 2)
-```
-
-When imaging bullets, operators scan one land at a time in a clockwise (left twisted rifling) or anti-clockwise (right twisted rifling) sequence.
-The order in which scans are acquired is kept as meta-information.
-Let us assume that lands on a bullet are labelled $\ell_i$ with $i = 1, .., p$, \hh{where $p$ indicates the number of lands a bullet has, as determined by the rifling of a barrel.
-For all of the bullets considered here the number of lands $p$ is 6}.
-A match between  \hh{a pair of lands from two} bullets therefore results in an expected additional $p-1$ matches between pairs of lands.
-These lands are also expected to be in a sequence, i.e.
-if there is a match between lands $\ell_i$ on bullet 1 and $\ell_j$ on bullet 2, we also expect lands $\ell_{i\oplus s}$ and $\ell_{j\oplus s}$ to match for all integers $s$, where $\oplus$ is defined as $a \oplus b \equiv \left((a + b - 1)\mod p\right) + 1$.
-This relationship gives rise to the sequence average maximum (SAM) to quantify a bullet-to-bullet match.
-
-### Bullet-to-Bullet Comparisons
-
-\begin{definition}{Sequence Average and its Maximum}
-Let $A$ be a square real-valued matrix of dimensions $p \times p$. For the purpose of this paper, $A$ consists of scores describing the similarity between two sets of land engraved areas.
-The $k$th \emph{sequence average} $SA(A, k)$ for $k = 0, ... p-1$ is defined  as 
-$$SA(A, k) = \frac{1}{p} \sum_{i=1}^{p} a_{i,i \oplus k}, \text{ where } i\oplus k := \left((i + k - 1)\mod p\right) + 1.$$
-The \emph{Sequence Average Maximum} \citep[SAM, ][]{sam} of square matrix $A$ of scores is defined as
-$$SAM (A) = \max_{k = 1}^{p} SA(A, k).$$
-\end{definition}
-
-Note that the sequence average maximum of the correlation between lands is used in SensoComp to capture the similarity between bullets.
-The correlation based SAM score has also been called the 'average correlation calculated at the max phase' in \citet{chuPilotStudyAutomated2010}.
-
-```{r sam-sketch, out.width='\\textwidth', fig.width=12, fig.height = 2.5, fig.cap="Sketch of all six land-to-land sequences between two bullets with six lands. "}
-df2 <- data.frame(expand.grid(x = 1:6, y = 1:6))
-df2$case1 <- df2$x == ((df2$y + 5) %% 6) + 1
-df2$case2 <- df2$x == ((df2$y + 0) %% 6) + 1
-df2$case3 <- df2$x == ((df2$y + 1) %% 6) + 1
-df2$case4 <- df2$x == ((df2$y + 2) %% 6) + 1
-df2$case5 <- df2$x == ((df2$y + 3) %% 6) + 1
-df2$case6 <- df2$x == ((df2$y + 4) %% 6) + 1
-
-df2 %>%
-  gather("case", "values", starts_with("case")) %>%
-  mutate(case = gsub("case", "", case)) %>%
-  mutate(case = paste0("k = ", as.numeric(case) - 1)) %>%
-  ggplot(aes(x = factor(x), y = factor(y, levels = 1:6), fill = values)) +
-  geom_tile(colour = "grey20") +
-  facet_wrap(~case, ncol = 6) +
-  xlab("Bullet 2") + ylab("Bullet 1") +
-  scale_x_discrete(labels = paste0("L", 1:6)) +
-  scale_y_discrete(labels = paste0("L", 1:6)) +
-  scale_fill_manual("Cell included in SA(A,k)", values = c("white", "grey20")) +
-  coord_equal() +
-  theme(legend.position = "bottom")
-```
-
-Looking back at \autoref{fig:b2b}, we see that for the two bullets from the same barrel, the sequence average for 
-$k=2$ is higher than the other sequence averages, and also higher than the sequence averages for the other pair of bullets shown on the right of the figure.
-
-SAM scores allow us to define a single quantity for each pair of bullets that describes the similarity between these two bullets. 
-This leads us to two minimal requirements the \hh{similarity} scores should fulfill:
+## Interpreting the Algorithm Score
+
+\svp{For any of the features described above, high values over the expected range indicate that the signatures are more similar; low values indicate that the signatures are less similar.
+At some point along the range of possible values, an automated algorithm would produce a threshold - below that score, the signatures would be categorized as an exclusion; above it, the signatures would be categorized as an identification.
+}
+<!-- Firearms examiners typically set the threshold for CMS at 6; this is a threshold which is conservative, as it is much more likely to exclude same-source comparisons than it is to identify different-source comparisons.
+-->
+
+We can enumerate characteristics of an ideal similarity scoring mechanism: 
 \begin{description}
@@
 
+\svp{In particular, requirement (R2) would imply that the same threshold value would be used for all comparisons of a certain scoring mechanism.
+If the same-source and different-source similarity score distributions overlap, setting a threshold value will introduce classification errors.}
+
+\svp{This threshold system introduces a binary classification: identification or exclusion.
+This} \hh{definition of error is more stringent than the AFTE Theory of Identification, in that it does not allow for inconclusive results.
+} \svp{This increased rigor may increase the error rate of the model when compared to examiner error rates, but we expect that the increased information available from the 3D measurement data will compensate for this loss.
+A binary decision model is also easier to interpret and provides more clear-cut, definitive results.}
+<!-- \svp{It is likely that matches which were deemed  inconclusive would instead be marked as exclusions by the algorithm.
+XXX I don't know if we can make this statement, but I think it would help examiners understand how the line gets drawn if we could.
+I have exactly no evidence to back that up (except maybe the Hamby 44 tank rash results?)} -->
+
 ## Identification Errors and Algorithm Evaluation
-When evaluating algorithm performance, it is useful to \hh{systematically assess }\svp{ the set of possible outcomes} \hh{first}: if ground truth and the algorithm's prediction match, we have either a correct identification or a correct exclusion.
+
+When evaluating algorithm performance, it is useful to \hh{systematically assess }\svp{ the set of possible outcomes}.
+If ground truth and the algorithm's prediction match, we have either a correct identification or a correct exclusion.
 If ground truth and the algorithm's prediction do not match, we distinguish between two types of errors, which we will refer to as \textbf{wrong identification} and \textbf{missed identification}.
@@
 \end{table}
-\hh{Note that this definition of error is more stringent than the AFTE Theory of Identification, in that it does not allow for inconclusive results, but forces a decision based on a cutoff value.
-} \svp{This increased rigor may increase the error rate of the model when compared to examiner error rates, but we expect that the increased information available from the 3D measurement data will compensate for this loss.
-A binary decision model is also easier to interpret and provides more clear-cut, definitive results.}
-
-\svp{It is likely that matches which were deemed  inconclusive would instead be marked as exclusions by the algorithm.
-XXX I don't know if we can make this statement, but I think it would help examiners understand how the line gets drawn if we could.
-I have exactly no evidence to back that up (except maybe the Hamby 44 tank rash results?)}
 
@@
 
-\svp{XXX Should we have some sort of header here to differentiate it from the error section?}
+## Assessing Match Strength
+
+\svp{Each of the scoring methods we have described is computed on a land-to-land basis.} While these comparisons are useful, \hh{the question of interest typically involves }  the entire physical object (e.g.
+all lands on a single bullet), \hh{and a conclusion on same or different source should be reached based on the evidence of all lands of the bullets}.
+<!--to describe the correspondence between two pieces of physical evidence.-->
+<!-- \svp{In order to aggregate land-to-land match scores, we must consider the physical structure of the evidence; that is, that the sequence of lands stems from spatial information, but the numbering of each land represents the scan order and start point rather than encoding useful information.
+} --><!--This last bit might be redundant-->
+
+```{r, echo=FALSE, results = 'hide', warning = FALSE, message = FALSE}
+features <- read.csv("data/fsi-features.csv.gz")
+features <- features %>% mutate(
+  bullet1 = gsub("ullet ", "", bullet1),
+  bullet2 = gsub("ullet ", "", bullet2)
+)
+b12 <- features %>% filter(
+  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
+  group2 == "G1", barrel2 == "KA", bullet2 == "B2"
+)
+p1 <- b12 %>%
+  ggplot(aes(x = land1, y = land2, fill = rfscore)) +
+  geom_tile() + ggtitle("Same barrel") +
+  scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
+  theme_bw() +
+  xlab("Bullet 2") + ylab("Bullet 1") +
+  theme(legend.position = "bottom") + coord_equal()
+
+bab <- features %>% filter(
+  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
+  group2 == "G1", barrel2 == "KB", bullet2 == "B1"
+)
+p2 <- bab %>%
+  ggplot(aes(x = land1, y = land2, fill = rfscore)) +
+  geom_tile() + ggtitle("Different barrel") +
+  scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
+  theme_bw() +
+  xlab("Bullet 2") + ylab("Bullet 1") +
+  theme(legend.position = "bottom") + coord_equal()
+```
+
+Land-to-land comparisons lead to a whole set of scores for bullet-to-bullet comparisons.
+\autoref{fig:b2b} shows two matrices of scores for \hh{two pairs} of bullet-to-bullet matches.
+On the left, a matrix is shown that is typical for scores from two bullets from the same barrel.
+On the right, values for a pair of known non-matching bullets are shown.
+
+```{r b2b, echo = FALSE, fig.width = 9, fig.height = 5,  out.width='.6\\textwidth', fig.cap="Overview of land-land matching scores for two pairs of bullet-bullet comparisons.
+On the left the two bullets are known to come from the same source (barrel) on the right, the bullets are from two different sources (barrels)"}
+grid.arrange(p1, p2, ncol = 2, padding = 2)
+```
+
+When imaging bullets, operators scan one land at a time in a clockwise (left twisted rifling) or anti-clockwise (right twisted rifling) sequence.
+The order in which scans are acquired is kept as meta-information.
+Let us assume that lands on a bullet are labelled $\ell_i$ with $i = 1, .., p$, \hh{where $p$ indicates the number of lands a bullet has, as determined by the rifling of a barrel.
+For all of the bullets considered here the number of lands $p$ is 6}.
+A match between  \hh{a pair of lands from two} bullets therefore results in an expected additional $p-1$ matches between pairs of lands.
+These lands are also expected to be in a sequence, i.e.
+if there is a match between lands $\ell_i$ on bullet 1 and $\ell_j$ on bullet 2, we also expect lands $\ell_{i\oplus s}$ and $\ell_{j\oplus s}$ to match for all integers $s$, where $\oplus$ is defined as $a \oplus b \equiv \left((a + b - 1)\mod p\right) + 1$.
+This relationship gives rise to the sequence average maximum (SAM) to quantify a bullet-to-bullet match.
+
+\begin{definition}{Sequence Average and its Maximum}
+Let $A$ be a square real-valued matrix of dimensions $p \times p$. For the purpose of this paper, $A$ consists of scores describing the similarity between two sets of land engraved areas.
+The $k$th \emph{sequence average} $SA(A, k)$ for $k = 0, ... p-1$ is defined  as 
+$$SA(A, k) = \frac{1}{p} \sum_{i=1}^{p} a_{i,i \oplus k}, \text{ where } i\oplus k := \left((i + k - 1)\mod p\right) + 1.$$
+The \emph{Sequence Average Maximum} \citep[SAM, ][]{sam} of square matrix $A$ of scores is defined as
+$$SAM (A) = \max_{k = 1}^{p} SA(A, k).$$
+\end{definition}
+
+Note that the sequence average maximum of the correlation between lands is used in SensoComp to capture the similarity between bullets.
+The correlation based SAM score has also been called the 'average correlation calculated at the max phase' in \citet{chuPilotStudyAutomated2010}.
+
+```{r sam-sketch, out.width='\\textwidth', fig.width=12, fig.height = 2.5, fig.cap="Sketch of all six land-to-land sequences between two bullets with six lands. "}
+df2 <- data.frame(expand.grid(x = 1:6, y = 1:6))
+df2$case1 <- df2$x == ((df2$y + 5) %% 6) + 1
+df2$case2 <- df2$x == ((df2$y + 0) %% 6) + 1
+df2$case3 <- df2$x == ((df2$y + 1) %% 6) + 1
+df2$case4 <- df2$x == ((df2$y + 2) %% 6) + 1
+df2$case5 <- df2$x == ((df2$y + 3) %% 6) + 1
+df2$case6 <- df2$x == ((df2$y + 4) %% 6) + 1
+
+df2 %>%
+  gather("case", "values", starts_with("case")) %>%
+  mutate(case = gsub("case", "", case)) %>%
+  mutate(case = paste0("k = ", as.numeric(case) - 1)) %>%
+  ggplot(aes(x = factor(x), y = factor(y, levels = 1:6), fill = values)) +
+  geom_tile(colour = "grey20") +
+  facet_wrap(~case, ncol = 6) +
+  xlab("Bullet 2") + ylab("Bullet 1") +
+  scale_x_discrete(labels = paste0("L", 1:6)) +
+  scale_y_discrete(labels = paste0("L", 1:6)) +
+  scale_fill_manual("Cell included in SA(A,k)", values = c("white", "grey20")) +
+  coord_equal() +
+  theme(legend.position = "bottom")
+```
+
+Looking back at \autoref{fig:b2b}, we see that for the two bullets from the same barrel, the sequence average for 
+$k=2$ is higher than the other sequence averages, and also higher than the sequence averages for the other pair of bullets shown on the right of the figure.
+
+SAM scores allow us to define a single quantity for each pair of bullets that describes the similarity between these two bullets.
+
+\svp{XXX Should we have some sort of header here to differentiate it from the previous section?}
 
@@
 
+Several lands in Hamby set 44 have major deficiencies,  such as 'tank rash' (a collision of the bullet after exiting the barrel with a surface causing markings on top of the striations from the barrel) or extreme pitting (holes caused by direct contact with burning gun powder).
+\autoref{fig:tankrash} gives an overview of all lands of affected bullets in the Hamby-44 set.
+These issues affect the algorithm's ability to identify a stable signature, which impacts the subsequent match scores \svp{ at the land-to-land and bullet-to-bullet level.}
+```{r tankrash, fig.cap="Overview of bullet lands with prominent deficiency such as tank rash or extreme pitting in the Hamby-44 study.", fig.width=4, fig.height = 2, out.width='47.5%', fig.subcap=c("Br1-B1-L6 (tank rask)", "Br2-B2-L5 (tank rash)", "Br3-B1-L5 (tank rash)", "Br8-B1-L6 (tank rash)", "Br8-B2-L2 (pitting)", "Br8-B2-L6 (tank rash)", "Unk-E-L6 (tank rash)", "Unk-I-L6 (tank rash and break-off)"), fig.align="middle", fig.ncol=2}
+knitr::include_graphics(
+  c(
+    "./images/B1-B1-L6.png", "./images/B2-B2-L5.png",
+    "./images/B3-B1-L5.png", "./images/B8-B1-L6.png",
+    "./images/B8-B2-L2.png", "./images/B8-B2-L6.png",
+    "./images/NA-BE-L6.png", "./images/NA-BI-L6.png"
+  )
+)
+```
+
+Three lands from bullets known to be fired from barrel 8 are affected, producing low scores for barrel 8 in \autoref{fig:h44}.
+We also see that questioned bullet 'I' is affected, explaining some of the low similarity scores for this bullet.
+\hh{None of the lands of any of the bullets in the other studies are affected by tank rash in a similar manner.}
+
 ```{r, cache=FALSE}
@@
 
-What is quite surprising is that while we anticipated Hamby 44 would be the easiest set to work with, it turned out to be the hardest to solve.
-This is likely due to several lands in Hamby set 44 that have major deficiencies,  such as 'tank rash' (a collision of the bullet after exiting the barrel with a surface causing markings on top of the striations from the barrel) or extreme pitting (holes caused by direct contact with burning gun powder).
-\autoref{fig:tankrash} gives an overview of all lands of affected bullets in the Hamby-44 set.
-These issues affect the algorithm's ability to identify a stable signature, which impacts the subsequent match scores.
-```{r tankrash, fig.cap="Overview of bullet lands with prominent deficiency such as tank rash or extreme pitting in the Hamby-44 study.", fig.width=4, fig.height = 2, out.width='47.5%', fig.subcap=c("Br1-B1-L6 (tank rask)", "Br2-B2-L5 (tank rash)", "Br3-B1-L5 (tank rash)", "Br8-B1-L6 (tank rash)", "Br8-B2-L2 (pitting)", "Br8-B2-L6 (tank rash)", "Unk-E-L6 (tank rash)", "Unk-I-L6 (tank rash and break-off)"), fig.align="middle", fig.ncol=2}
-knitr::include_graphics(
-  c(
-    "./images/B1-B1-L6.png", "./images/B2-B2-L5.png",
-    "./images/B3-B1-L5.png", "./images/B8-B1-L6.png",
-    "./images/B8-B2-L2.png", "./images/B8-B2-L6.png",
-    "./images/NA-BE-L6.png", "./images/NA-BI-L6.png"
-  )
-)
-```
-
-Three lands from bullets known to be fired from barrel 8 are affected, producing low scores for barrel 8 in \autoref{fig:h44}.
-We also see that questioned bullet 'I' is affected, explaining some of the low similarity scores for this bullet.
-\hh{None of the lands of any of the bullets in the other studies are affected by tank rash in a similar manner.}
-
-## Limitations
+What is quite surprising is that while we anticipated Hamby 44 would be the easiest set to work with, it turned out to be the hardest to solve, in part because of damage to the bullets that obscure LEA striae.
+It has been shown \citet{lpr} that parts of lands can be used for successful identifications.
+The algorithm proposed by \citet{aoas} is not capable of  automatically detecting parts of lands with well expressed striae.
+It may be useful to couple the \citet{aoas} algorithm with an algorithm which assesses the quality of the input data and determines which portions of the data to use for comparison.
+This would emulate the process used by examiners, who first assess the quality of the evidence and whether there is enough information present to attempt a match.
+\svp{One major disadvantage of an automated algorithm is that all of the decisions humans make (recognizing degraded land areas, excluding those areas from consideration, matching only the remaining areas) must be explicitly characterized; however, this explicit characterization means that the process can be scientifically validated to a much higher degree than the human perceptual process.
+}
 
@@
 
-As seen in the earlier discussion, similarity scores are affected by deficiencies on lands.
-It has been shown \citet{lpr} that parts of lands can be used for successful identifications.
-The algorithm proposed by \citet{aoas} is not capable of  automatically detecting parts of lands with well expressed striae.
-It may be useful to couple the \citet{aoas} algorithm with an algorithm which assesses the quality of the input data and determines which portions of the data to use for comparison.
-This would emulate the process used by examiners, who first assess the quality of the evidence and whether there is enough information present to attempt a match.
-\svp{One major disadvantage of an automated algorithm is that all of the decisions humans make without further thought (recognizing degraded land areas, excluding those areas from consideration, matching only the remaining areas) must be explicitly characterized; however, this explicit characterization means that the process can be scientifically validated to a much higher degree than the human perceptual process
-. }
-
-## Conclusion
-
-\svp{The random forest matching algorithm presented in \citet{aoas} performs well on three different external test sets.
-While a single cutoff value cannot be used to distinguish matches and non-matches across the different test sets, the algorithm makes no errors when a set-specific cutoff value is used on bullet-to-bullet scores aggregated using sequence average maximum.} \svp{XXX more here XXX}
+\svp{The random forest matching algorithm presented in \citet{aoas} does perform well on three different external test sets.
+While a single cutoff value cannot be used to distinguish matches and non-matches across the different test sets, the algorithm makes no errors when a set-specific cutoff value is used on bullet-to-bullet scores aggregated using sequence average maximum.} \svp{XXX comparison to FTEs XXX} \svp{By validating the algorithm on external test sets, we have demonstrated that the method can be generalized to different types of ammunition and is not overly sensitive to small differences in rifling procedure.
+Future studies can generalize this to a wider range of external test sets to establish the limits of the algorithm's generalizability.}
 

### 2018-11-21 index.Rmd
JFS submission guidelines, keywords
todo

@@
 # to produce blinded version set to 1
-blinded: 0
+blinded: 1
 
@@
 keywords:
-- 3 to 6 keywords
-- that do not appear in the title
+- random forest
+- 3d microscopy
+- toolmark
 
@@
 \begin{itemize}
-\item define false positives and false negatives once and stick to those definitions.
+\item \sout{define false positives and false negatives once and stick to those definitions.
 We should avoid false positive and false negatives, because that only makes sense in a test setting, which we have not defined formally.
-Let's use instead wrong identification and missed identification. 
-\item error rates: we are using rates exchangeably with percentages. We need to fix that. I think percentages might be easier understood. XXX Should be fixed now
+Let's use instead wrong identification and missed identification. }
+\item \sout{error rates: we are using rates exchangeably with percentages. We need to fix that. I think percentages might be easier understood. XXX Should be fixed now}
+\end{itemize}
+}
+\svp{
+Outstanding XXX items:
+\begin{itemize}
+\item thanks - scanning and hamby advice
+\item comparison to FTEs in conclusion
 \end{itemize}

### 2018-11-21 index.Rmd
remove unnecessary comment
compile separate title page when document is compiled

@@
 library(kableExtra)
+```
+
+```{r child-doc-compile, include = F}
+library(stringr)
+thisdoc <- readLines("index.Rmd")
+idxs <- which(thisdoc == "---")
+if (length(idxs) >= 2) {
+  # Get yaml metadata, set to nonblinded, write to title-page.Rmd
+  thisdoc[idxs[1]:idxs[2]] %>%
+    str_replace("^blinded: 1", "blinded: 0") %>%
+    writeLines(con = "title-page.Rmd")
+}
+
+rmarkdown::render("title-page.Rmd", clean = T)
 ```
@@
 SAM scores allow us to define a single quantity for each pair of bullets that describes the similarity between these two bullets.
-
-\svp{XXX Should we have some sort of header here to differentiate it from the previous section?}
 

### 2018-11-21 index.Rmd
JFS author guidelines specify highest degree must be listed

@@
 authors: 
-- name: Heike Hofmann
+- name: Heike Hofmann, Ph.D.
   thanks: This work was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.
@@
 
-- name: Susan VanderPlas
+- name: Susan VanderPlas, Ph.D.
   affiliation: Department of Statistics, Iowa State University

### 2018-11-21 index.Rmd
New "thanks" paragraph for technical help

@@
 - name: Heike Hofmann, Ph.D.
-  thanks: This work was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.
-The authors wish to thank Jim Hamby, Tylor Klep (Phoenix PD), and Melissa Nally and Kasi Kirksey (FSI Houston) for access to test set bullets.
-XXX Scanning.
-Hamby advice.
   affiliation: Department of Statistics, Iowa State University
+  thanks: 
+    - This work was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.
+    - The authors wish to thank Jim Hamby, Tylor Klep (Phoenix PD), and Melissa Nally and Kasi Kirksey (FSI Houston) for access to test set bullets. XXX Scanning. Hamby advice.
 

### 2018-11-21 index.Rmd
remove abstract from title page

@@
     str_replace("^blinded: 1", "blinded: 0") %>%
+    str_replace("^abstract:.*", "") %>%
     writeLines(con = "title-page.Rmd")

### 2018-11-21 index.Rmd
other JFS formatting guidelines

@@
   pdf_document:
+    number_sections: false
     citation_package: natbib
@@
 \begin{itemize}
+\item ``Authors are responsible for obtaining written permission from persons acknowledged by name, because readers may infer their endorsement of the data and conclusions."
 \item thanks - scanning and hamby advice
@@
   fig.width = 6,
+  dpi = 300,
   fig.align = "center",

### 2018-11-21 index.Rmd
Fixing headings for intro, Methods sections
template update and yaml changes

@@
 
+affiliation: 
+- Department of Statistics, Iowa State University
+- Center for Statistical Applications in Forensic Evidence (CSAFE), Iowa State University
+
 authors: 
-- name: Heike Hofmann, Ph.D.
-  affiliation: Department of Statistics, Iowa State University
+- name: Heike Hofmann
+  degree: Doctor of Philosophy (Ph.D.)
+- name: Susan VanderPlas
+  degree: Doctor of Philosophy (Ph.D.)
+
+funding: This work was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.
+
 thanks: 
-    - This work was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.
-    - The authors wish to thank Jim Hamby, Tylor Klep (Phoenix PD), and Melissa Nally and Kasi Kirksey (FSI Houston) for access to test set bullets. XXX Scanning. Hamby advice.
-
-- name: Susan VanderPlas, Ph.D.
-  affiliation: Department of Statistics, Iowa State University
+- The authors wish to thank Jim Hamby, Tylor Klep (Phoenix PD), and Melissa Nally and Kasi Kirksey (FSI Houston) for access to test set bullets. 
+- XXX Scanning. 
+- Thanks to Jim Hamby for his advice on XXX.
+
+presentedat: This work was presented in the form of a poster at Forensics\@NIST on November 7, 2018 in Gaithersburg, Maryland
 
 keywords:
+- forensic science
+- toolmark
+- cross-correlation
 - random forest
 - 3d microscopy
-- toolmark
+- land engraved areas (LEAs)
 
@@
 bibliography: bibliography
-biblio-style: apsr
+biblio-style: jfs
+
+numsecdepth: 1
 
@@
   pdf_document:
-    number_sections: false
     citation_package: natbib
@@
 
-# Introduction and Background
-
 In current practice, firearms and toolmark examiners (FTE) evaluate the similarity of striae on bullets by placing the evidence together with another bullet under a comparison microscope.
@@
 
-## Automated Processing of 3D Scans
+# Methods
+
+<!-- ## Automated Processing of 3D Scans -->
 
@@
 
-## Statistics for Matching Aligned Signatures
+<!-- ## Statistics for Matching Aligned Signatures -->
 
@@
 
-## Interpreting the Algorithm Score
+<!-- ## Interpreting the Algorithm Score -->
 
@@
 
-## Identification Errors and Algorithm Evaluation
+<!-- ## Identification Errors and Algorithm Evaluation -->
 
@@
 
-## Assessing Match Strength
+<!-- ## Assessing Match Strength -->
 

### 2018-11-21 index.Rmd
tweaks due to new style file, output all pictures separately and automatically

@@
 
+library(stringr)
+
 library(tidyverse)
@@
 
-```{r child-doc-compile, include = F}
-library(stringr)
+```{r child-doc-compile, include = F, cache = F}
 thisdoc <- readLines("index.Rmd")
-idxs <- which(thisdoc == "---")
+idxs <- which(str_detect(thisdoc, "^---$"))
 if (length(idxs) >= 2) {
@@
 rmarkdown::render("title-page.Rmd", clean = T)
+file.remove("title-page.tex", "title-page.Rmd")
 ```
@@
 
-## Identification Using 3D Imaging Technology
+<!-- ## Identification Using 3D Imaging Technology -->
 
@@
 \multicolumn{2}{c}{} & \multicolumn{1}{|c}{Same Source} & \multicolumn{1}{|c|}{Different Source} \\\cline{2-4}
-\multirow{8}{*}{\rotatebox[origin=c]{90}{\hfill \textbf{Results}\hspace{6mm}\hfill}} & 
-  \multirow{4}{*}{\rotatebox[origin=c]{90}{\hspace{1mm}Exclusion\hspace{1mm}}} & \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\
+\multirow{8}{*}{\rotatebox[origin=c]{90}{\hfill \textbf{Results}\hspace{5mm}\hfill}} & 
+  \multirow{6}{*}{\rotatebox[origin=c]{90}{\hspace{1mm}Exclusion\hspace{1mm}}} & \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\
+  & &   \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\
   & &  \multicolumn{1}{|c}{Missed} & \multicolumn{1}{|c|}{Correct}\\
   & &  \multicolumn{1}{|c}{Identification} & \multicolumn{1}{|c|}{Exclusion}\\
+  & &   \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\
   & &  \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{} \\\cline{2-4}
-& \multirow{4}{*}{\rotatebox[origin=c]{90}{\hspace{1mm}Identification\hspace{1mm}}}  & \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\
+& \multirow{6}{*}{\rotatebox[origin=c]{90}{\hspace{1mm}Identification\hspace{1mm}}}  & \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\
+  & &  \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\
   & &  \multicolumn{1}{|c}{Correct} & \multicolumn{1}{|c|}{Wrong}\\
   & &  \multicolumn{1}{|c}{Identification} & \multicolumn{1}{|c|}{Identification}\\
+  & &  \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\
   & &  \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{} \\\cline{2-4}
@@
 
-```{r rocs, fig.cap="ROC curves for all studies based on land-to-land comparisons of the random forest score."}
+```{r rocs}
 library(plotROC)
@@
 
-```{r aucs-pic, fig.cap=""}
+```{r aucs-pic}
 eer <- function(roc){
@@
 
+```{r output-all-figs, include = F}
+thisdoc <- readLines("index.Rmd")
+
+# First, get figures that aren't in code chunks
+# Haven't really tested this part - regex may need work
+raw_figures <- data_frame(
+  line_start = which(str_detect(thisdoc, "\\begin\\{figure\\}")), 
+  line_end = which(str_detect(thisdoc, "\\end\\{figure\\}")),
+  subfloats = purrr::map2(line_start, line_end, 
+                          ~str_extract(thisdoc[.x:.y], "\\includegraphics(?:\\[.*\\])?\\{(.*)\\}") %>%
+                            str_replace("\\includegraphics(?:\\[.*\\])?\\{(.*)\\}", "\\1"))
+)
+
+# Code chunk figures
+chunk_figures <- data_frame(
+  line_start = which(str_detect(thisdoc, "^```\\{r.*\\}$")),
+  line_end = which(str_detect(thisdoc, "^``` *$")),
+  chunk_name = str_replace(thisdoc[line_start], "(```\\{r ?)([0-9A-z-]*)(,?.*\\})", "\\2")
+) %>% 
+  mutate(
+    chunk_name = ifelse(chunk_name == "", "unnamed-chunk", chunk_name),
+  ) %>% 
+  mutate(
+    subfloats = purrr::map2(line_start, chunk_name, ~thisdoc[.x] %>%
+                              str_extract("fig\\.subcap ?= ?.*?\\),") %>%
+                              str_replace("fig\\.subcap ?= ?", "") %>%
+                              str_replace(",$", "") %>%
+                              parse(text=.) %>% eval %>% length() %>% 
+                              seq(1, ., by = 1) %>%
+                              sprintf("figures/%s-%d.pdf", .y, .))
+  )
+
+# Knitr included graphics figures
+chunk_incl_figures <- data_frame(
+  line_start = which(str_detect(thisdoc, "^```\\{r.*\\}$")),
+  line_end = which(str_detect(thisdoc, "^``` *$")),
+  chunk_name = str_replace(thisdoc[line_start], "(```\\{r ?)([0-9A-z-]*)(,?.*\\})", "\\2")
+) %>% 
+  mutate(
+    chunk_name = ifelse(chunk_name == "", "unnamed-chunk", chunk_name),
+  ) %>% 
+  mutate(
+    subfloats = purrr::map2(line_start, line_end, function(.x, .y) {
+      thisdoc[.x:.y] %>%
+        paste(collapse = " ") %>%
+        str_extract(., "include_graphics\\(c?\\(?.*?\\)?\\)") %>%
+        str_replace("include_graphics\\(", "") %>%
+        str_extract_all('\\"[\\.A-z0-9 /-]*?\\"') %>%
+        parse(text = .) %>% eval() %>%
+        str_remove_all("\\\\|\"")
+    })
+  )
+
+# All figures
+figs <- bind_rows(raw_figures, chunk_figures, chunk_incl_figures) %>%
+  arrange(line_start) %>%
+  mutate(subfloats = purrr::map(subfloats, function(x) data_frame(file = x, exists = file.exists(x)))) %>%
+  tidyr::unnest() %>%
+  filter(exists) %>%
+  mutate(fig_num = group_indices(., line_start, line_end)) %>%
+  group_by(fig_num) %>%
+  mutate(rn = 1:n(),
+         maxrn = n(),
+         fig_label = letters[rn],
+         fig_label = ifelse(maxrn > 1, fig_label, ""),
+         fileext = tools::file_ext(file),
+         fig_label = sprintf("Figure_%d%s.%s", fig_num, fig_label, fileext))
+
+if (!dir.exists("separate_fig_files")) dir.create("separate_fig_files")
+file.copy(figs$file, file.path("separate_fig_files", figs$fig_label))
+
+```
 # References

### 2018-11-21 index.Rmd
Not working yet, but try to fix draft behavior AND title page behavior

@@
 
-# to produce blinded version set to 1
-blinded: 1
+blinded: 1 # to produce blinded version set to 1
+titlepage: 0 # This gets changed automatically when the title page is compiled in the chunk child-doc-compile
+draft: 0 # Set to 1 if figures should be included with the text
 
@@
     str_replace("^blinded: 1", "blinded: 0") %>%
+    str_replace("^titlepage: 0", "titlepage: 1") %>%
     str_replace("^abstract:.*", "") %>%
@@
 if (!dir.exists("separate_fig_files")) dir.create("separate_fig_files")
-file.copy(figs$file, file.path("separate_fig_files", figs$fig_label))
+file.copy(figs$file, file.path("separate_fig_files", figs$fig_label), overwrite = T)
 

### 2018-11-26 index.Rmd
phrasing change

@@
 In many cases, these approaches also provide some level of automation of the comparison process, with the goal of reducing human biases by augmenting the visual information with 3D measurements.
-Utilizing the 3D measurements allows for examination of both peaks and valleys in LEAs, \hh{it also allows us to take the depth of striae into account, something that is difficult, if not impossible, for visual examination.}
+Utilizing the 3D measurements allows for examination of both peaks and valleys in LEAs, \hh{it also allows us to take the depth of striae into account, something that is difficult, if not impossible,} \svp{from} \hh{visual examination.}
 

### 2018-11-26 index.Rmd
proof-read

--- 2018-11-21 index.Rmd
@@
 - random forest
-- 3d microscopy
+- 3D microscopy
 - land engraved areas (LEAs)
 
-abstract: Recent advances in microscopy have made it possible to collect 3d topographic data, enabling virtual comparisons based on the collected 3D data next to traditional comparison microscopy.
+abstract: Recent advances in microscopy have made it possible to collect 3D topographic data, enabling virtual comparisons based on the collected 3D data next to traditional comparison microscopy.
 Automatic matching algorithms have been introduced for various scenarios, such as matching cartridge cases (Tai and Eddy 2018) or matching bullet striae (Hare et al.
@@
 
-```{r child-doc-compile, include = F, cache = F}
+```{r child-doc-compile, eval=FALSE, include = F, cache = F}
 thisdoc <- readLines("index.Rmd")
-idxs <- which(str_detect(thisdoc, "^---$"))
+idxs <- which(stringr::str_detect(thisdoc, "^---$"))
 if (length(idxs) >= 2) {
@@
 file.remove("title-page.tex", "title-page.Rmd")
-```
-
-```{r functions, eval=FALSE}
-estimateBeta <- function(x, method = "mm", probs = c(0.25, 0.75)) {
-  if (method == "mm") {
-    mu <- mean(x, na.rm = TRUE)
-    sd <- sd(x, na.rm = TRUE)
-    lambda <- mu / sd * (1 - mu) / sd - 1
-    alpha <- mu * lambda
-    beta <- (1 - mu) * lambda
-  }
-  if (method == "LB") {
-    quantile1 <- list(p = probs[1], x = quantile(probs = probs[1], x))
-    quantile2 <- list(p = probs[2], x = quantile(probs = probs[2], x))
-    bs <- LearnBayes::beta.select(quantile1, quantile2)
-    alpha <- bs[1]
-    beta <- bs[2]
-  }
-  return(c(alpha = alpha, beta = beta))
-}
 ```
@@
 
-More explicit characterization of bullet surfaces has been discussed since at least 1958 (\citeauthor{davis1968introduction}), but at the time technology was not sufficiently advanced to make an analysis based on 3D measurements or surface traces a viable option.
+More explicit characterization of bullet surfaces has been discussed since at least 1958 by \citeauthor{davis1968introduction} \cite{davis1968introduction}, but at the time technology was not sufficiently advanced to make an analysis based on 3D measurements or surface traces a viable option.
 \citet{ComputerIdentificationBullets1978} demonstrated use of a scanning electron microscope to quantitatively examine and compare bullet striae.
-More recently, approaches using 3D measurement data were explicitly described in 1999 (\citeauthor{dekinderAutomatedComparisonsBullet1999}), and have been further developed by \citet{bachrachDevelopment3DbasedAutomated2002,xieAutomatedBulletidentificationSystem2009, chuPilotStudyAutomated2010}.
+More recently, approaches using 3D measurement data were explicitly described in 1999 (\citeauthor{dekinderAutomatedComparisonsBullet1999}), and have been further developed in \citep{bachrachDevelopment3DbasedAutomated2002,xieAutomatedBulletidentificationSystem2009, chuPilotStudyAutomated2010}.
 These approaches utilize 3D surface measurements directly to characterize the topology of land engraved areas (LEAs), rather than using visual or image comparison techniques.
 In many cases, these approaches also provide some level of automation of the comparison process, with the goal of reducing human biases by augmenting the visual information with 3D measurements.
-Utilizing the 3D measurements allows for examination of both peaks and valleys in LEAs, \hh{it also allows us to take the depth of striae into account, something that is difficult, if not impossible, for visual examination.}
+Utilizing the 3D measurements allows for examination of both peaks and valleys in LEAs, it also allows us to take the depth of striae into account, something that is difficult, if not impossible, from visual inspection.
 
@@
 
-In the matching algorithm proposed in \citet{aoas}, several of the features described above are combined using a random forest \citep{breiman} to produce a similarity score based on 3d topographic scans of land engraved areas (LEAs).
+In the matching algorithm proposed in \citet{aoas}, several of the features described above are combined using a random forest \citep{breiman} to produce a similarity score based on 3D topographic scans of land engraved areas (LEAs).
 From each land engraved area a signature is derived using the process described in detail in \citet{aoas}:

### 2018-11-26 index.Rmd
up to page 9

@@
 \begin{itemize}
-\item[(a)] Identify an area on the LEA with expressed striae. Locate a stable crosscut (see \citet{aoas}) in the identified area.
+\item[(a)] Identify an area on the LEA with expressed striae. Locate a stable crosscut in the identified area.
 \item[(b)] Discard extraneous/contaminated data, such as data from groove engraved area and areas affected by break-off or contact with objects after the bullet exited the barrel  (``tank rash").
@@
 \autoref{fig:signatures} shows a set of six signatures corresponding to the six land engraved areas of bullets 1 and 2 from barrel 1 of set 44 of the Hamby study \citep{hamby}.
-```{r signatures, fig.cap="Signatures of all six lands of bullets 1 and 2 from barrel 1 of the Hamby set 44.", fig.height=4, fig.width=10}
+```{r signatures, fig.cap="Signatures of all six lands of bullets 1 and 2 from barrel 1 of the Hamby set 44.
+Matching lands between the two bullets have been placed above each other.", fig.height=4, fig.width=10}
 # signatures for one bullet
@@
 
-The cross-correlation (CC) function between two signatures $S_1(x)$ and $S_2(x)$ is defined as \[CC(S_1, S_2, z) = \text{cor}\left(S_1(x), S_2(x + z)\right)\] where $x$ and $x+z$ are integer values appropriately defined within the domains of $S_1$ and $S_2$ and  $z$ is the lag between the two sequences.
-When signatures contain missing values, pairwise complete observations are used to calculate the cross-correlation.
-The lag $z$ used to achieve maximum $CC$ is used to determine the best alignment.
+The cross-correlation (CC) function between two signatures $S_1(x)$ and $S_2(x)$ is defined as 
+\[
+CC(S_1, S_2, z) = \text{cor}\left(S_1(x), S_2(x + z)\right)
+\] 
+where $x$ and $x+z$ are integer values appropriately defined within the domains of $S_1$ and $S_2$ and  $z$ is the lag between the two sequences.
+\hh{$\text{cor}(.,.)$ is the Pearson correlation coefficient,}
+when signatures contain missing values, pairwise complete observations are used to calculate the cross-correlation. The lag $z$ used to achieve maximum $CC$ is used to determine the best alignment.
+
+\svp{Cross correlation can also be used to assess match strength between two aligned signatures.
+Cross correlation varies between -1 and 1, but as we are taking the maximal value, in practice the cross-correlations for aligned signatures are generally positive.
+Cross-correlation values which are close to 1 indicate similarity between signatures and with it may indicate that the two signatures come from the same source.
+Cross-correlation values which are close to 0 indicate signatures which may originate from different sources.}
 
@@
 
-Once two signatures are aligned, quantitative features describing the similarity of two signature are extracted.
-Features include characteristics of the signature that firearms and toolmark examiners base their visual assessment on, such as striae depth, number of matching striae, and the number of consecutively matching striae (CMS).
-Statistical features such as cross-correlation and Euclidean distance between the signatures are also extracted and used to assess the strength of the correspondence between the aligned signatures.
-
-One of the measures firearms examiners use to describe the strength of a match quantitatively is the number of consecutive matching striae (CMS) as defined by \citet{biasotti}.
+\hh{Using the lag determined for aligning two signatures} other quantitative features describing the similarity of the two signatures can be extracted.
+
+Numerical features such as cross-correlation  and Euclidean distance
+can be computed from the aligned signatures alone.
+
+\hh{Other} features include characteristics of the signature that firearms and toolmark examiners base their visual assessment on, such as striae depth, number of matching striae, and the number of consecutively matching striae (CMS), \hh{one of the measures firearms examiners use to describe the strength of a match quantitatively.}
+\sout{is the number of consecutive matching striae (CMS) as defined by} \citet{biasotti}.
+
 \svp{In order to evaluate CMS, we must first identify peaks and valleys in each of the signatures, then determine whether these peaks and valleys overlap sufficiently.
-Features which firearms examiners assess visually depend on the identification of these extrema; numerical features such as cross-correlation can be computed from the aligned signatures alone.
+Features which firearms examiners assess visually depend on the identification of these extrema.
 }
 
-If there are at least 6 consecutive matching striae, the bullets are considered to be a good match \citep{biasotti_firearms_1997,nichols_scientific_2006}. 
-CMS in this setting are counted as the number of consecutive peaks in signatures of two lands that match, in part because it is difficult to visually assess valleys except as relative to peaks. 
-Note that \svp{with automatic alignment }based on the maximum cross-correlation rather than matching striation marks, it is possible for two lands to have a consecutive matching striae of 0. 
-<!-- CMS are generally considered to be a very conservative criteria, in that bullets fired through the same barrel may not have a high number of consecutively matching striae; in this sense, CMS may be more likely to produce an inconclusive result.
--->
-<!--\hh{XXX for CMS we need to define peaks, and matching peaks, and for that we need to define the additional filter/smooth.}\svp{XXX Can't we just refer to the aoas paper?
-It seems to be defined in there relatively well.}-->
-<!-- \svp{XXX Feels like we need to mention something about ccf here just because these are the 3 main features we're using to evaluate the algorithm...?} -->
-
-\svp{Cross correlation can also be used to assess match strength between two aligned signatures; this cross-correlation value corresponds to the maximal cross-correlation over all valid lag values $z$.
-Cross correlation varies between -1 and 1, but as we are taking the maximal value, in practice the cross-correlations for aligned signatures are generally positive.
-Cross-correlation values which are close to 1 indicate signatures which are similar and may originate from the same source.
-Cross-correlation values which are close to 0 indicate signatures which may originate from different sources.}
+If there are at least six consecutive matching striae, the bullets are considered to be a good match in practice \citep{biasotti_firearms_1997,nichols_scientific_2006}. 
+CMS in this setting are counted as the number of consecutively matching peaks in signatures of two aligned lands.
+Peaks -- rather than peaks and valleys -- are used for this determination in part because it is difficult to visually assess valleys except as relative to peaks.
+Note that with automatic alignment based on the maximum cross-correlation rather than matching striation marks, it is possible for two lands to have a consecutive matching striae of zero.
 
 The random forest presented in \citet{aoas} is based on a combination of multiple characteristics, including cross-correlation, number of matching striae, and number of consecutively matching striae.
-The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where scores close to 1 indicate that the algorithm believes it to be highly likely that the aligned signatures come from the same source, while low scores indicate that the aligned signatures are highly likely to originate from different lands.
+The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where scores close to 1 indicate \sout{that the algorithm believes} \hh{XXX can an algorithm have a belief?
+XXX} it to be highly likely that the aligned signatures come from the same source, while low scores indicate that the aligned signatures are highly likely to originate from different lands.
 
@@
 
-\svp{For any of the features described above, high values over the expected range indicate that the signatures are more similar; low values indicate that the signatures are less similar.
+\hh{Any of the above described features are supposed to separate between signatures from same source and different source, e.g.\ high cross-correlation values are indicative of same sourceness of two signatures, whereas  low values are more indicate of different sources.
+An automated algorithm determines a cutoff-value along the range of possible values that introduces a barrier between identification and exclusion.
+}
+
+\sout{\svp{For any of the features described above, \hh{XXX not for Euclidean distance XXX} high values over the expected range indicate that the signatures are more similar; low values indicate that the signatures are less similar.
 At some point along the range of possible values, an automated algorithm would produce a threshold - below that score, the signatures would be categorized as an exclusion; above it, the signatures would be categorized as an identification.
-}
+}}
 <!-- Firearms examiners typically set the threshold for CMS at 6; this is a threshold which is conservative, as it is much more likely to exclude same-source comparisons than it is to identify different-source comparisons.
 -->
+This threshold system introduces a binary classification: identification or exclusion.
+This definition of error is more stringent than the AFTE Theory of Identification, in that it does not allow for inconclusive results.
+This increased rigor may increase the error rate of the model when compared to examiner error rates, but we expect that the increased information available from the 3D measurement data will compensate for this loss.
+A binary decision model is also easier to interpret and provides more clear-cut, definitive results.
 
@@
 
-\svp{In particular, requirement (R2) would imply that the same threshold value would be used for all comparisons of a certain scoring mechanism.
-If the same-source and different-source similarity score distributions overlap, setting a threshold value will introduce classification errors.}
-
-\svp{This threshold system introduces a binary classification: identification or exclusion.
-This} \hh{definition of error is more stringent than the AFTE Theory of Identification, in that it does not allow for inconclusive results.
-} \svp{This increased rigor may increase the error rate of the model when compared to examiner error rates, but we expect that the increased information available from the 3D measurement data will compensate for this loss.
-A binary decision model is also easier to interpret and provides more clear-cut, definitive results.}
+In particular, requirement (R2) would imply that the same threshold value would be used for all comparisons of a certain scoring mechanism.
+If the same-source and different-source similarity score distributions overlap, setting a threshold value will introduce classification errors.
+
 <!-- \svp{It is likely that matches which were deemed  inconclusive would instead be marked as exclusions by the algorithm.
@@
 
-When evaluating algorithm performance, it is useful to \hh{systematically assess }\svp{ the set of possible outcomes}.
+When evaluating algorithm performance, it is useful to systematically assess  the set of possible outcomes.
 If ground truth and the algorithm's prediction match, we have either a correct identification or a correct exclusion.
@@
 The full range of possible outcomes is shown in \autoref{tab:possible-outcomes}.
+\hh{XXX I know that you have spent a ton of time on this table - but could you make it less tall?
+It's taking a ton of space.
+}
 \begin{table}
@@
 
-\svp{In machine learning and statistics, it is common to evaluate algorithms using sensitivity and specificity; these concepts are related to the error rate.
+In machine learning and statistics, it is common to evaluate algorithms using sensitivity and specificity; these concepts are related to the error rate.
 \textbf{Sensitivity} is defined as the proportion of actual positives that are correctly identified; that is, when ground truth is same source, the sensitivity is the proportion of correct identifications.
@@
 The combination of sensitivity and specificity is sufficient to describe the reliability of an algorithm.
-}
-
-\svp{When evaluating the \citet{aoas} algorithm, we will describe errors as a percentage of correct evaluations.
+
+When evaluating the \citet{aoas} algorithm, we will describe errors as a percentage of correct evaluations.
 Statistics that do not translate to percent-form interpretation, such as area under the curve (AUC), will be presented as a decimal.
-}
 
@@
 
-\svp{Each of the scoring methods we have described is computed on a land-to-land basis.} While these comparisons are useful, \hh{the question of interest typically involves }  the entire physical object (e.g.
-all lands on a single bullet), \hh{and a conclusion on same or different source should be reached based on the evidence of all lands of the bullets}.
+Each of the scoring methods we have described is computed on a land-to-land basis.
+While these comparisons are useful, the question of interest typically involves   the entire physical object (e.g.
+all lands on a single bullet), and a conclusion on same or different source should be reached based on the evidence of all lands of the bullets.
 <!--to describe the correspondence between two pieces of physical evidence.-->
@@
 Land-to-land comparisons lead to a whole set of scores for bullet-to-bullet comparisons.
-\autoref{fig:b2b} shows two matrices of scores for \hh{two pairs} of bullet-to-bullet matches.
+\autoref{fig:b2b} shows two matrices of scores for two pairs of bullet-to-bullet matches.
 On the left, a matrix is shown that is typical for scores from two bullets from the same barrel.
@@
 
-Note that the sequence average maximum of the correlation between lands is used in SensoComp to capture the similarity between bullets.
+Note that the sequence average maximum of the correlation between lands is used in SensoComp \hh{citation for that?} to capture the similarity between bullets.
 The correlation based SAM score has also been called the 'average correlation calculated at the max phase' in \citet{chuPilotStudyAutomated2010}.

### 2018-11-26 index.Rmd
Fixing draft options?

--- 2018-11-26 index.Rmd
@@
 
-blinded: 1 # to produce blinded version set to 1
-titlepage: 0 # This gets changed automatically when the title page is compiled in the chunk child-doc-compile
-draft: 0 # Set to 1 if figures should be included with the text
+# to produce blinded version set to 1
+blinded: 1
+# This gets changed automatically when the title page is compiled in the chunk child-doc-compile
+titlepage: 0
+# Set to 1 if figures should be included with the text
+draft: 1
 

### 2018-11-26 index.Rmd
finished proof-read

@@
 The order in which scans are acquired is kept as meta-information.
-Let us assume that lands on a bullet are labelled $\ell_i$ with $i = 1, .., p$, \hh{where $p$ indicates the number of lands a bullet has, as determined by the rifling of a barrel.
-For all of the bullets considered here the number of lands $p$ is 6}.
-A match between  \hh{a pair of lands from two} bullets therefore results in an expected additional $p-1$ matches between pairs of lands.
+Let us assume that lands on a bullet are labelled $\ell_i$ with $i = 1, .., p$, where $p$ indicates the number of lands a bullet has, as determined by the rifling of a barrel.
+For all of the bullets considered here the number of lands $p$ is 6.
+A match between  a pair of lands from two bullets therefore results in an expected additional $p-1$ matches between pairs of lands.
 These lands are also expected to be in a sequence, i.e.
@@
 Ten questioned bullets were provided (B, E, H, J, K N, Q, T, Y, Z).
-This set is an open set; that is,\ it is not known in advance whether all (or any) of the questioned bullets are fired from the known barrels.
-\hh{In fact, the results will show that three of the questioned bullets were fired from three different barrels not included in the knowns.
-These three barrels correspond to an eleventh Ruger P-95, a Ruger P-95C and a Ruger P-85.} \svp{All bullets fired for this study are American Eagle 9mm  copper jacketed Luger.}
+This set is an open set; that is, it is not known in advance whether all (or any) of the questioned bullets are fired from the known barrels.
+In fact, the results will show that three of the questioned bullets were fired from three different barrels not included in the knowns.
+These three barrels correspond to an eleventh Ruger P-95, a Ruger P-95C and a Ruger P-85.
+All bullets fired for this study are American Eagle 9mm  copper jacketed Luger.
 Land engraved areas for each of the six lands of each bullet were scanned by Bill Henderson (Sensofar).
@@
 The ammunition used in both test fires and the questioned bullets were Remington UMC 9mm Luger Full Metal Jackets.
-All three of the test sets are open; that is,\ not every one of the questioned bullets is fired from the five known barrels in each of the test set.
+All three of the test sets are open; that is, not every one of the questioned bullets is fired from the five known barrels in each of the test set.
 The structure of these three sets is similar to a forthcoming study by Nally and Kirksey, but the results here come from preliminary test sets made available to us.
@@
 similar to the Ruger P-85, but are treated to a secondary round of heating after rifling.
-This second round of heat exposure has the potential to introduce some subclass characteristics, i.e.
-anomalies that are shared between different barrels manufactured in this manner.
+This second round of heat exposure has the potential to introduce some subclass characteristics, i.e.\ anomalies that are shared between different barrels manufactured in this manner.
 This should make the automatic matching harder, and in particular, is expected to complicate the classification as different-source land engraved areas/bullets.
@@
 These strips correspond to known barrels 1 through 10.
-\svp{Each barrel was test-fired twice, so each of the questioned bullets (shown along the $y$ axis) matches two bullets fired from a known barrel.
+Each barrel was test-fired twice, so each of the questioned bullets (shown along the $y$ axis) matches two bullets fired from a known barrel.
 Colored tiles are used to show the strength of the comparison; light grey colors correspond to low similarity scores, dark colors correspond to high similarity scores.
-squares with a colored frame indicate ground truth.
-} Ground truth is encoded in this figure as a thin, dark, colored frame for all pairs of same-source bullets.
-\svp{We want to see one barrel with two dark-filled, dark-framed tiles for each questioned bullet, and light grey tiles for all other barrels, indicating a match between a questioned bullet and a single barrel.} This expectation is met for all questioned bullets, i.e.
-the automatic matching  identifies the correct barrel for all questioned bullets.
+Ground truth is encoded in this figure as a thin, dark, colored frame for all pairs of same-source bullets.
+Ideally, we want to see one barrel with two dark-filled, dark-framed tiles for each questioned bullet, and light grey tiles for all other barrels, indicating a match between a questioned bullet and a single barrel.
+This expectation is met for all questioned bullets, i.e.\ the automatic matching  identifies the correct barrel for all questioned bullets.
 For two of the questioned bullets, 'I' and 'F', the similarity scores to the matching barrels are considerably smaller than for the other questioned bullets.
@@
 None of the questioned bullets in set 1 match bullets fired from barrels KC or KD.
-Additionally, two of the questioned bullets which do not match any of the barrels in the set match each other.
-In set 2, four of the questioned bullets can be matched to known bullets in the set; of the additional four questioned bullets which do not match any known bullets, two were fired from the same barrel.
-In set 3, six of the questioned bullets match known bullets, and the remaining two questioned bullets match each other but do not match any known bullets in the set.
+Additionally, two of the questioned bullets which do not match any of the barrels in the set match each other (U28 and U37).
+In set 2, four of the questioned bullets can be matched to known bullets in the set; of the additional four questioned bullets which do not match any known bullets, two were fired from the same barrel (bullets U34 and U73).
+In set 3, six of the questioned bullets match known bullets, and the remaining two questioned bullets match each other (bullets U14 and U45) but do not match any known bullets in the set.
 <!--Thus, we can conclude that the algorithm performs well on open set studies, as it can identify matches among questioned bullets which do not match any known bullets. -->
-While all bullet-to-bullet matches are correctly identified, and no known matches are missed, it is obvious from the generally lighter shades of the tiles corresponding to matching bullets in \autoref{fig:hou}\svp{(b) that the algorithm is not performing as well on set 2 as it performs on sets 1 and 3.}
+While all bullet-to-bullet matches are correctly identified, and no known matches are missed, it is obvious from the generally lighter shades of the tiles corresponding to matching bullets in \autoref{fig:hou}(b) that the algorithm is not performing as well on set 2 as it performs on sets 1 and 3.
 ```{r hou, fig.width=9, fig.height = 3, fig.cap="Overview of matching scores for all pairs of questioned bullets to known bullets from five barrels (on the left) and questioned bullets to themselves (tiles on the right).", fig.subcap=c("Set 1", "Set 2", "Set 3"), fig.align="middle", fig.ncol=1}
@@
 # Evaluating the Random Forest Algorithm
-\svp{The random forest SAM scores correctly separate known matches from known non-matches, using SAM scores to aggregate land-to-land scores into a bullet-to-bullet comparison.
-With no errors at the bullet-to-bullet level, we now need to evaluate the scores in light of (R1) and (R2).
+The random forest SAM scores correctly separate known matches from known non-matches, using SAM scores to aggregate land-to-land scores into a bullet-to-bullet comparison.
+With no errors at the bullet-to-bullet level, we can now evaluate the scores in light of (R1) and (R2).
 In this section, we will compare the random forest scores to cross-correlation scores; cross-correlation is one of the components of the random forest, but has also been used to quantify the strength of a match in its own right \citep{chuPilotStudyAutomated2010}.
-In addition, both scoring methods use the same scale (0 to 1) and are continuous along that interval; other evaluation methods, such as consecutive matching striae are discrete and more difficult to directly compare to continuous measurements.}
-
-<!-- \svp{XXX We may want to add in SAM scores for CMS as well, but that's going to make the the scales on the density plots (and the density plots themselves) a bit wonky. Thoughts?} -->
+In addition, both scoring methods use the same scale (0 to 1) and are continuous along that interval; other evaluation methods, such as consecutive matching striae are discrete and more difficult to directly compare to continuous measurements.
 
@@
 Small vertical lines below the $x$-axis indicate observed scores in a **rug-plot**.
-The plot shows that for all three case studies and both continuous measures, i.e.
-SAM scores based on cross-correlation as well as SAM scores based on the random forest scores, all similarity scores are higher for pairs of bullets from the same source, i.e.
-bullets fired through the same barrel, than pairs of bullets from different sources (fired through different barrels).
+The plot shows that for all three case studies and both continuous measures, i.e.\ SAM scores based on cross-correlation as well as SAM scores based on the random forest scores, all similarity scores are higher for pairs of bullets from the same source, i.e.\ bullets fired through the same barrel, than pairs of bullets from different sources (fired through different barrels).
 
@@
 
-\svp{According to \autoref{fig:compare}, neither cross-correlation nor random forest score} fulfill both of the minimal requirements laid out \svp{in the introduction}.
-Both measures fulfill (R1), i.e.
-all  scores  of  different-source pairs are lower than scores of same-source pairs for SAM scores based on cross-correlation and RF scores.
+According to \autoref{fig:compare}, neither cross-correlation nor random forest score fulfill both of the minimal requirements laid out in the introduction.
+Both measures fulfill (R1), i.e.\ all  scores  of  different-source pairs are lower than scores of same-source pairs for SAM scores based on cross-correlation and RF scores.
 However, SAM scores based on the multivariate random forest score show better separation between same-source and different-source scores than SAM scores based on cross-correlation.
@@
 There are several approaches for potential improvements: we can try to fine tune the matching algorithm to take make and model of the firearm and ammunition used into account or expand the training base of the algorithm to include a wider variety of makes and models.
-The downside of \svp{either of these options} is that we \svp{would} need to considerably expand the database used for training.
-Another solution would be to \svp{augment} the aggregation used to get from land-to-land scores to bullet scores: SAM scores only take the scores of the maximum sequence into account and ignore scores associated with pairs of land engraved areas from different sources.
+The downside of either of these options is that we would need to considerably expand the database used for training.
+Another solution would be to augment the aggregation used to get from land-to-land scores to bullet scores: SAM scores only take the scores of the maximum sequence into account and ignore scores associated with pairs of land engraved areas from different sources.
 Those scores might be useful in establishing a baseline that better expresses similarity between pairs of bullets.
@@
 
-\svp{Examining land-to-land scores provides more details about the matching algorithm's performance.
+Examining land-to-land scores provides more details about the matching algorithm's performance.
 \autoref{fig:compare-land-to-land} shows density curves for each study.
 There is some overlap between known matching and known non-matching land-to-land scores; in many cases the random forest scores for these overlapping values are more extreme than the corresponding cross-correlation scores.
-This suggests that the matching algorithm is sensitive to the presence of LEAs which have low land-to-land match scores, such as the matches of bullets `I' and `F' in the Hamby 44 study.
+This suggests that the matching algorithm is sensitive to the presence of LEAs which have low land-to-land match scores, such as the matches of bullets 'I' and 'F' in the Hamby 44 study.
 Comparing this to the bullet-to-bullet score density plot shown in \autoref{fig:compare}, we see that there is much greater separation between the same-source and different-source densities for each study in the bullet-to-bullet comparisons than in the land-to-land comparisons, that is, one or two poorly matching lands does not prevent the bullet from showing a matching score, though if there are weak matches on several lands, such as the marked Hamby 44 comparisons in \autoref{fig:compare}, the overall score may be affected.
-}
 
@@
 \autoref{fig:tankrash} gives an overview of all lands of affected bullets in the Hamby-44 set.
-These issues affect the algorithm's ability to identify a stable signature, which impacts the subsequent match scores \svp{ at the land-to-land and bullet-to-bullet level.}
+These issues affect the algorithm's ability to identify a stable signature, which impacts the subsequent match scores  at the land-to-land and bullet-to-bullet level.
 ```{r tankrash, fig.cap="Overview of bullet lands with prominent deficiency such as tank rash or extreme pitting in the Hamby-44 study.", fig.width=4, fig.height = 2, out.width='47.5%', fig.subcap=c("Br1-B1-L6 (tank rask)", "Br2-B2-L5 (tank rash)", "Br3-B1-L5 (tank rash)", "Br8-B1-L6 (tank rash)", "Br8-B2-L2 (pitting)", "Br8-B2-L6 (tank rash)", "Unk-E-L6 (tank rash)", "Unk-I-L6 (tank rash and break-off)"), fig.align="middle", fig.ncol=2}
@@
 We also see that questioned bullet 'I' is affected, explaining some of the low similarity scores for this bullet.
-\hh{None of the lands of any of the bullets in the other studies are affected by tank rash in a similar manner.}
+None of the lands of any of the bullets in the other studies are affected by tank rash in a similar manner.
 
@@
 \autoref{fig:eer} shows an overview of Equal Error and their thresholds based on the ROC curves of \autoref{fig:roc-auc}.
-Equal Errors are errors when sensitivity and specificity of a test are equal, i.e.
-we see the same percentage of missed and wrong identifications for a land-to-land comparison.
+Equal Errors are errors when sensitivity and specificity of a test are equal, i.e.\ we see the same percentage of missed and wrong identifications for a land-to-land comparison.
 Equal errors based on CMS of two lands are significantly higher than equal errors based on cross-correlation or the RF score.
@@
 Interestingly, for consecutively matching striae, equal errors are reached for just two matching striae, which from a practical point of view is not sustainable.
-\svp{A  CMS of  6 or above is an established cutoff used in identifications; when applied to the land-to-land comparisons, it is much more likely to result in a missed identification than a wrong identification.
+A  CMS of  6 or above is an established cutoff used in identifications; when applied to the land-to-land comparisons, it is much more likely to result in a missed identification than a wrong identification.
 A threshold of 6 leads to 21 wrong identifications out of 49,286 different-source comparisons (0.04\% errors, or about 4 errors for each 10,000 evaluations); when applied to same-source comparisons, 517 of 715 same-source lands were incorrectly classified as exclusions, for an error rate of 80.6\%.
-}
-<!-- This leads to 21 wrong identifications (out of 49286; for an error of 4 in 10,000 -- or 0.04\%) but only identifies 139 out of 715 known same-source lands, for an error of 80.6\%.  -->
 
@@
 This would emulate the process used by examiners, who first assess the quality of the evidence and whether there is enough information present to attempt a match.
-\svp{One major disadvantage of an automated algorithm is that all of the decisions humans make (recognizing degraded land areas, excluding those areas from consideration, matching only the remaining areas) must be explicitly characterized; however, this explicit characterization means that the process can be scientifically validated to a much higher degree than the human perceptual process.
-}
+One major disadvantage of an automated algorithm is that all of the decisions humans make (recognizing degraded land areas, excluding those areas from consideration, matching only the remaining areas) must be explicitly characterized; however, this explicit characterization means that the process can be scientifically validated to a much higher degree than the human perceptual process.
 
@@
 
-\svp{The random forest matching algorithm presented in \citet{aoas} does perform well on three different external test sets.
-While a single cutoff value cannot be used to distinguish matches and non-matches across the different test sets, the algorithm makes no errors when a set-specific cutoff value is used on bullet-to-bullet scores aggregated using sequence average maximum.} \svp{XXX comparison to FTEs XXX} \svp{By validating the algorithm on external test sets, we have demonstrated that the method can be generalized to different types of ammunition and is not overly sensitive to small differences in rifling procedure.
-Future studies can generalize this to a wider range of external test sets to establish the limits of the algorithm's generalizability.}
+The random forest matching algorithm presented in \citet{aoas} does perform well on three different external test sets.
+While a single cutoff value cannot be used to distinguish matches and non-matches across the different test sets, the algorithm makes no errors when a set-specific cutoff value is used on bullet-to-bullet scores aggregated using sequence average maximum.
+\hh{This performance is on par with the performance of forensic examiners.
+For a future round of the Houston study, we are planning to compare the algorithmic performance directly with scores given by forensic toolmark examiners.
+XXX need to ask XXX However, the intent is not to replace a fornesic examiner.
+The decision on matching should not be taken away from the human assessment.
+Instead, the algorithm's score is meant as a complement to an examiner's testimony as a way to measure and quantify the strength of an identification.
+}
+By validating the algorithm on external test sets, we have demonstrated that the method can be generalized to different types of ammunition and is not overly sensitive to small differences in rifling procedure.
+Future studies can \hh{and should} generalize this to a wider range of external test sets to establish the limits of the algorithm's generalizability.
 
@@
 # References
+
+<!--include marginal distributions of training data and compare to test data-->
+
+```{r, echo=FALSE, eval=FALSE}
+training <- read.csv("data/features-hamby.csv")
+gtrain <- training %>% 
+  select(profile1_id, profile2_id, study1, study2, match, ccf:sum_peaks) %>%
+  gather(feature, value, ccf:sum_peaks)
+# now compare to test data
+```

### 2018-11-27 index.Rmd
tweaks. Still need to do a final read through.

@@
 \] 
-where $x$ and $x+z$ are integer values appropriately defined within the domains of $S_1$ and $S_2$ and  $z$ is the lag between the two sequences.
-\hh{$\text{cor}(.,.)$ is the Pearson correlation coefficient,}
-when signatures contain missing values, pairwise complete observations are used to calculate the cross-correlation. The lag $z$ used to achieve maximum $CC$ is used to determine the best alignment.
+where $x$ and $x+z$ are integer values appropriately defined within the domains of $S_1$ and $S_2$, $z$ is the lag between the two sequences, \svp{and }\hh{$\text{cor}(.,.)$ is the Pearson correlation coefficient.}
+When signatures contain missing values, pairwise complete observations are used to calculate the cross-correlation. The lag $z$ used to achieve maximum $CC$ is used to determine the best alignment.
 
@@
 
-\hh{Using the lag determined for aligning two signatures} other quantitative features describing the similarity of the two signatures can be extracted.
-
+\hh{Using the lag determined for aligning two signatures,} other quantitative features describing the similarity of the two signatures can be extracted.
 Numerical features such as cross-correlation and Euclidean distance
 can be computed from the aligned signatures alone.
-
 \hh{Other} features include characteristics of the signature that firearms and toolmark examiners base their visual assessment on, such as striae depth, number of matching striae, and the number of consecutively matching striae (CMS), \hh{one of the measures firearms examiners use to describe the strength of a match quantitatively.}
-\sout{is the number of consecutive matching striae (CMS) as defined by} \citet{biasotti}.
 
@@
 Features which firearms examiners assess visually depend on the identification of these extrema.
-}
-
-If there are at least six consecutive matching striae, the bullets are considered to be a good match in practice \citep{biasotti_firearms_1997,nichols_scientific_2006}. 
+} If there are at least six consecutive matching striae, the bullets are considered to be a good match in practice \citep{biasotti_firearms_1997,nichols_scientific_2006}.
 CMS in this setting are counted as the number of consecutively matching peaks in signatures of two aligned lands.
@@
 The random forest presented in \citet{aoas} is based on a combination of multiple characteristics, including cross-correlation, number of matching striae, and number of consecutively matching striae.
-The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where scores close to 1 indicate \sout{that the algorithm believes} \hh{XXX can an algorithm have a belief?
-XXX} it to be highly likely that the aligned signatures come from the same source, while low scores indicate that the aligned signatures are highly likely to originate from different lands.
+The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where scores close to 1 indicate it to be highly likely that the aligned signatures come from the same source, while low scores indicate that the aligned signatures are highly likely to originate from different lands.
 
@@
 
-\sout{\svp{For any of the features described above, \hh{XXX not for Euclidean distance XXX} high values over the expected range indicate that the signatures are more similar; low values indicate that the signatures are less similar.
-At some point along the range of possible values, an automated algorithm would produce a threshold - below that score, the signatures would be categorized as an exclusion; above it, the signatures would be categorized as an identification.
-}}
 <!-- Firearms examiners typically set the threshold for CMS at 6; this is a threshold which is conservative, as it is much more likely to exclude same-source comparisons than it is to identify different-source comparisons.
@@
 
-Note that the sequence average maximum of the correlation between lands is used in SensoComp \hh{citation for that?} to capture the similarity between bullets.
+Note that the sequence average maximum of the correlation between lands is used in SensoComp \hh{citation for that?}\svp{It's your claim...
+I looked for a citation and didn't find one.
+Closed-source proprietary = no references?} to capture the similarity between bullets.
 The correlation based SAM score has also been called the 'average correlation calculated at the max phase' in \citet{chuPilotStudyAutomated2010}.
@@
 For a future round of the Houston study, we are planning to compare the algorithmic performance directly with scores given by forensic toolmark examiners.
-XXX need to ask XXX However, the intent is not to replace a fornesic examiner.
+XXX need to ask XXX \sout{However, the intent is not to replace a fornesic examiner.
 The decision on matching should not be taken away from the human assessment.
-Instead, the algorithm's score is meant as a complement to an examiner's testimony as a way to measure and quantify the strength of an identification.
+Instead, the algorithm's score is meant as a complement to an examiner's testimony as a way to measure and quantify the strength of an identification.}  }\svp{The algorithm's score is not intended to replace an examiner's judgment; instead, it is a tool to measure and quantify the strength of an identification.
 }

### 2018-11-27 index.Rmd
more tweaks

@@
 # Set to 1 if figures should be included with the text
-draft: 1
+draft: 0
 
@@
 \multirow{8}{*}{\rotatebox[origin=c]{90}{\hfill \textbf{Results}\hspace{5mm}\hfill}} & 
-  \multirow{6}{*}{\rotatebox[origin=c]{90}{\hspace{1mm}Exclusion\hspace{1mm}}} & \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\
-  & &   \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\
+  \multirow{4}{*}{\rotatebox[origin=c]{90}{\hspace{1mm}Exclusion\hspace{1mm}}} & \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\
+  <!-- & &   \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\ -->
   & &  \multicolumn{1}{|c}{Missed} & \multicolumn{1}{|c|}{Correct}\\
   & &  \multicolumn{1}{|c}{Identification} & \multicolumn{1}{|c|}{Exclusion}\\
-  & &   \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\
+  <!-- & &   \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\ -->
   & &  \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{} \\\cline{2-4}
-& \multirow{6}{*}{\rotatebox[origin=c]{90}{\hspace{1mm}Identification\hspace{1mm}}}  & \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\
-  & &  \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\
+& \multirow{4}{*}{\rotatebox[origin=c]{90}{\hspace{1mm}Identification\hspace{1mm}}}  & \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\
+  <!-- & &  \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\ -->
   & &  \multicolumn{1}{|c}{Correct} & \multicolumn{1}{|c|}{Wrong}\\
   & &  \multicolumn{1}{|c}{Identification} & \multicolumn{1}{|c|}{Identification}\\
-  & &  \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\
+  <!-- & &  \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\ -->
   & &  \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{} \\\cline{2-4}

### 2018-11-27 index.Rmd
fix table... again. Swear I fixed that before....

@@
   \multirow{4}{*}{\rotatebox[origin=c]{90}{\hspace{1mm}Exclusion\hspace{1mm}}} & \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\
-  <!-- & &   \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\ -->
+  % & &   \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\ 
   & &  \multicolumn{1}{|c}{Missed} & \multicolumn{1}{|c|}{Correct}\\
   & &  \multicolumn{1}{|c}{Identification} & \multicolumn{1}{|c|}{Exclusion}\\
-  <!-- & &   \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\ -->
+  % & &   \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\ 
   & &  \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{} \\\cline{2-4}
 & \multirow{4}{*}{\rotatebox[origin=c]{90}{\hspace{1mm}Identification\hspace{1mm}}}  & \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\
-  <!-- & &  \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\ -->
+  % & &  \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\ 
   & &  \multicolumn{1}{|c}{Correct} & \multicolumn{1}{|c|}{Wrong}\\
   & &  \multicolumn{1}{|c}{Identification} & \multicolumn{1}{|c|}{Identification}\\
-  <!-- & &  \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\ -->
+  % & &  \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\ 
   & &  \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{} \\\cline{2-4}

### 2018-11-27 index.Rmd
Fix figure text size

--- 2018-11-27 index.Rmd
@@
 \] 
-where $x$ and $x+z$ are integer values appropriately defined within the domains of $S_1$ and $S_2$, $z$ is the lag between the two sequences, \svp{and }\hh{$\text{cor}(.,.)$ is the Pearson correlation coefficient.}
-When signatures contain missing values, pairwise complete observations are used to calculate the cross-correlation. The lag $z$ used to achieve maximum $CC$ is used to determine the best alignment.
+where $x$ and $x+z$ are integer values appropriately defined within the domains of $S_1$ and $S_2$ and  $z$ is the lag between the two sequences.
+\hh{$\text{cor}(.,.)$ is the Pearson correlation coefficient,}
+when signatures contain missing values, pairwise complete observations are used to calculate the cross-correlation. The lag $z$ used to achieve maximum $CC$ is used to determine the best alignment.
 
@@
 
-\hh{Using the lag determined for aligning two signatures,} other quantitative features describing the similarity of the two signatures can be extracted.
+\hh{Using the lag determined for aligning two signatures} other quantitative features describing the similarity of the two signatures can be extracted.
+
 Numerical features such as cross-correlation  and Euclidean distance
 can be computed from the aligned signatures alone.
+
 \hh{Other} features include characteristics of the signature that firearms and toolmark examiners base their visual assessment on, such as striae depth, number of matching striae, and the number of consecutively matching striae (CMS), \hh{one of the measures firearms examiners use to describe the strength of a match quantitatively.}
+\sout{is the number of consecutive matching striae (CMS) as defined by} \citet{biasotti}.
 
@@
 Features which firearms examiners assess visually depend on the identification of these extrema.
-} If there are at least six consecutive matching striae, the bullets are considered to be a good match in practice \citep{biasotti_firearms_1997,nichols_scientific_2006}.
+}
+
+If there are at least six consecutive matching striae, the bullets are considered to be a good match in practice \citep{biasotti_firearms_1997,nichols_scientific_2006}. 
 CMS in this setting are counted as the number of consecutively matching peaks in signatures of two aligned lands.
@@
 The random forest presented in \citet{aoas} is based on a combination of multiple characteristics, including cross-correlation, number of matching striae, and number of consecutively matching striae.
-The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where scores close to 1 indicate it to be highly likely that the aligned signatures come from the same source, while low scores indicate that the aligned signatures are highly likely to originate from different lands.
+The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where scores close to 1 indicate \sout{that the algorithm believes} \hh{XXX can an algorithm have a belief?
+XXX} it to be highly likely that the aligned signatures come from the same source, while low scores indicate that the aligned signatures are highly likely to originate from different lands.
 
@@
 
+\sout{\svp{For any of the features described above, \hh{XXX not for Euclidean distance XXX} high values over the expected range indicate that the signatures are more similar; low values indicate that the signatures are less similar.
+At some point along the range of possible values, an automated algorithm would produce a threshold - below that score, the signatures would be categorized as an exclusion; above it, the signatures would be categorized as an identification.
+}}
 <!-- Firearms examiners typically set the threshold for CMS at 6; this is a threshold which is conservative, as it is much more likely to exclude same-source comparisons than it is to identify different-source comparisons.
@@
 
-```{r b2b, echo = FALSE, fig.width = 9, fig.height = 5,  out.width='.6\\textwidth', fig.cap="Overview of land-land matching scores for two pairs of bullet-bullet comparisons.
+```{r b2b, echo = FALSE, fig.width = 5.4, fig.height = 3, out.width = ".6\\textwidth", fig.cap="Overview of land-land matching scores for two pairs of bullet-bullet comparisons.
 On the left the two bullets are known to come from the same source (barrel) on the right, the bullets are from two different sources (barrels)"}
@@
 The order in which scans are acquired is kept as meta-information.
-Let us assume that lands on a bullet are labelled $\ell_i$ with $i = 1, .., p$, where $p$ indicates the number of lands a bullet has, as determined by the rifling of a barrel.
-For all of the bullets considered here the number of lands $p$ is 6.
-A match between  a pair of lands from two bullets therefore results in an expected additional $p-1$ matches between pairs of lands.
+Let us assume that lands on a bullet are labelled $\ell_i$ with $i = 1, .., p$, \hh{where $p$ indicates the number of lands a bullet has, as determined by the rifling of a barrel.
+For all of the bullets considered here the number of lands $p$ is 6}.
+A match between  \hh{a pair of lands from two} bullets therefore results in an expected additional $p-1$ matches between pairs of lands.
 These lands are also expected to be in a sequence, i.e.
@@
 
-Note that the sequence average maximum of the correlation between lands is used in SensoComp \hh{citation for that?}\svp{It's your claim...
-I looked for a citation and didn't find one.
-Closed-source proprietary = no references?} to capture the similarity between bullets.
+Note that the sequence average maximum of the correlation between lands is used in SensoComp \hh{citation for that?} to capture the similarity between bullets.
 The correlation based SAM score has also been called the 'average correlation calculated at the max phase' in \citet{chuPilotStudyAutomated2010}.
@@
 Ten questioned bullets were provided (B, E, H, J, K N, Q, T, Y, Z).
-This set is an open set; that is, it is not known in advance whether all (or any) of the questioned bullets are fired from the known barrels.
-In fact, the results will show that three of the questioned bullets were fired from three different barrels not included in the knowns.
-These three barrels correspond to an eleventh Ruger P-95, a Ruger P-95C and a Ruger P-85.
-All bullets fired for this study are American Eagle 9mm  copper jacketed Luger.
+This set is an open set; that is,\ it is not known in advance whether all (or any) of the questioned bullets are fired from the known barrels.
+\hh{In fact, the results will show that three of the questioned bullets were fired from three different barrels not included in the knowns.
+These three barrels correspond to an eleventh Ruger P-95, a Ruger P-95C and a Ruger P-85.} \svp{All bullets fired for this study are American Eagle 9mm  copper jacketed Luger.}
 Land engraved areas for each of the six lands of each bullet were scanned by Bill Henderson (Sensofar).
@@
 The ammunition used in both test fires and the questioned bullets were Remington UMC 9mm Luger Full Metal Jackets.
-All three of the test sets are open; that is, not every one of the questioned bullets is fired from the five known barrels in each of the test set.
+All three of the test sets are open; that is,\ not every one of the questioned bullets is fired from the five known barrels in each of the test set.
 The structure of these three sets is similar to a forthcoming study by Nally and Kirksey, but the results here come from preliminary test sets made available to us.
@@
 similar to the Ruger P-85, but are treated to a secondary round of heating after rifling.
-This second round of heat exposure has the potential to introduce some subclass characteristics, i.e.\ anomalies that are shared between different barrels manufactured in this manner.
+This second round of heat exposure has the potential to introduce some subclass characteristics, i.e.
+anomalies that are shared between different barrels manufactured in this manner.
 This should make the automatic matching harder, and in particular, is expected to complicate the classification as different-source land engraved areas/bullets.
@@
 These strips correspond to known barrels 1 through 10.
-Each barrel was test-fired twice, so each of the questioned bullets (shown along the $y$ axis) matches two bullets fired from a known barrel.
+\svp{Each barrel was test-fired twice, so each of the questioned bullets (shown along the $y$ axis) matches two bullets fired from a known barrel.
 Colored tiles are used to show the strength of the comparison; light grey colors correspond to low similarity scores, dark colors correspond to high similarity scores.
-Ground truth is encoded in this figure as a thin, dark, colored frame for all pairs of same-source bullets.
-Ideally, we want to see one barrel with two dark-filled, dark-framed tiles for each questioned bullet, and light grey tiles for all other barrels, indicating a match between a questioned bullet and a single barrel.
-This expectation is met for all questioned bullets, i.e.\ the automatic matching  identifies the correct barrel for all questioned bullets.
+squares with a colored frame indicate ground truth.
+} Ground truth is encoded in this figure as a thin, dark, colored frame for all pairs of same-source bullets.
+\svp{We want to see one barrel with two dark-filled, dark-framed tiles for each questioned bullet, and light grey tiles for all other barrels, indicating a match between a questioned bullet and a single barrel.} This expectation is met for all questioned bullets, i.e.
+the automatic matching  identifies the correct barrel for all questioned bullets.
 For two of the questioned bullets, 'I' and 'F', the similarity scores to the matching barrels are considerably smaller than for the other questioned bullets.
@@
 None of the questioned bullets in set 1 match bullets fired from barrels KC or KD.
-Additionally, two of the questioned bullets which do not match any of the barrels in the set match each other (U28 and U37).
-In set 2, four of the questioned bullets can be matched to known bullets in the set; of the additional four questioned bullets which do not match any known bullets, two were fired from the same barrel (bullets U34 and U73).
-In set 3, six of the questioned bullets match known bullets, and the remaining two questioned bullets match each other (bullets U14 and U45) but do not match any known bullets in the set.
+Additionally, two of the questioned bullets which do not match any of the barrels in the set match each other.
+In set 2, four of the questioned bullets can be matched to known bullets in the set; of the additional four questioned bullets which do not match any known bullets, two were fired from the same barrel.
+In set 3, six of the questioned bullets match known bullets, and the remaining two questioned bullets match each other but do not match any known bullets in the set.
 <!--Thus, we can conclude that the algorithm performs well on open set studies, as it can identify matches among questioned bullets which do not match any known bullets. -->
-While all bullet-to-bullet matches are correctly identified, and no known matches are missed, it is obvious from the generally lighter shades of the tiles corresponding to matching bullets in \autoref{fig:hou}(b) that the algorithm is not performing as well on set 2 as it performs on sets 1 and 3.
+While all bullet-to-bullet matches are correctly identified, and no known matches are missed, it is obvious from the generally lighter shades of the tiles corresponding to matching bullets in \autoref{fig:hou}\svp{(b) that the algorithm is not performing as well on set 2 as it performs on sets 1 and 3.}
 ```{r hou, fig.width=9, fig.height = 3, fig.cap="Overview of matching scores for all pairs of questioned bullets to known bullets from five barrels (on the left) and questioned bullets to themselves (tiles on the right).", fig.subcap=c("Set 1", "Set 2", "Set 3"), fig.align="middle", fig.ncol=1}
@@
 # Evaluating the Random Forest Algorithm
-The random forest SAM scores correctly separate known matches from known non-matches, using SAM scores to aggregate land-to-land scores into a bullet-to-bullet comparison.
-With no errors at the bullet-to-bullet level, we can now evaluate the scores in light of (R1) and (R2).
+\svp{The random forest SAM scores correctly separate known matches from known non-matches, using SAM scores to aggregate land-to-land scores into a bullet-to-bullet comparison.
+With no errors at the bullet-to-bullet level, we now need to evaluate the scores in light of (R1) and (R2).
 In this section, we will compare the random forest scores to cross-correlation scores; cross-correlation is one of the components of the random forest, but has also been used to quantify the strength of a match in its own right \citep{chuPilotStudyAutomated2010}.
-In addition, both scoring methods use the same scale (0 to 1) and are continuous along that interval; other evaluation methods, such as consecutive matching striae are discrete and more difficult to directly compare to continuous measurements.
+In addition, both scoring methods use the same scale (0 to 1) and are continuous along that interval; other evaluation methods, such as consecutive matching striae are discrete and more difficult to directly compare to continuous measurements.}
+
+<!-- \svp{XXX We may want to add in SAM scores for CMS as well, but that's going to make the the scales on the density plots (and the density plots themselves) a bit wonky. Thoughts?} -->
 
@@
 Small vertical lines below the $x$-axis indicate observed scores in a **rug-plot**.
-The plot shows that for all three case studies and both continuous measures, i.e.\ SAM scores based on cross-correlation as well as SAM scores based on the random forest scores, all similarity scores are higher for pairs of bullets from the same source, i.e.\ bullets fired through the same barrel, than pairs of bullets from different sources (fired through different barrels).
+The plot shows that for all three case studies and both continuous measures, i.e.
+SAM scores based on cross-correlation as well as SAM scores based on the random forest scores, all similarity scores are higher for pairs of bullets from the same source, i.e.
+bullets fired through the same barrel, than pairs of bullets from different sources (fired through different barrels).
 
@@
 
-According to \autoref{fig:compare}, neither cross-correlation nor random forest score fulfill both of the minimal requirements laid out in the introduction.
-Both measures fulfill (R1), i.e.\ all  scores  of  different-source pairs are lower than scores of same-source pairs for SAM scores based on cross-correlation and RF scores.
+\svp{According to \autoref{fig:compare}, neither cross-correlation nor random forest score} fulfill both of the minimal requirements laid out \svp{in the introduction}.
+Both measures fulfill (R1), i.e.
+all  scores  of  different-source pairs are lower than scores of same-source pairs for SAM scores based on cross-correlation and RF scores.
 However, SAM scores based on the multivariate random forest score show better separation between same-source and different-source scores than SAM scores based on cross-correlation.
@@
 There are several approaches for potential improvements: we can try to fine tune the matching algorithm to take make and model of the firearm and ammunition used into account or expand the training base of the algorithm to include a wider variety of makes and models.
-The downside of either of these options is that we would need to considerably expand the database used for training.
-Another solution would be to augment the aggregation used to get from land-to-land scores to bullet scores: SAM scores only take the scores of the maximum sequence into account and ignore scores associated with pairs of land engraved areas from different sources.
+The downside of \svp{either of these options} is that we \svp{would} need to considerably expand the database used for training.
+Another solution would be to \svp{augment} the aggregation used to get from land-to-land scores to bullet scores: SAM scores only take the scores of the maximum sequence into account and ignore scores associated with pairs of land engraved areas from different sources.
 Those scores might be useful in establishing a baseline that better expresses similarity between pairs of bullets.
@@
 
-Examining land-to-land scores provides more details about the matching algorithm's performance.
+\svp{Examining land-to-land scores provides more details about the matching algorithm's performance.
 \autoref{fig:compare-land-to-land} shows density curves for each study.
 There is some overlap between known matching and known non-matching land-to-land scores; in many cases the random forest scores for these overlapping values are more extreme than the corresponding cross-correlation scores.
-This suggests that the matching algorithm is sensitive to the presence of LEAs which have low land-to-land match scores, such as the matches of bullets 'I' and 'F' in the Hamby 44 study.
+This suggests that the matching algorithm is sensitive to the presence of LEAs which have low land-to-land match scores, such as the matches of bullets `I' and `F' in the Hamby 44 study.
 Comparing this to the bullet-to-bullet score density plot shown in \autoref{fig:compare}, we see that there is much greater separation between the same-source and different-source densities for each study in the bullet-to-bullet comparisons than in the land-to-land comparisons, that is, one or two poorly matching lands does not prevent the bullet from showing a matching score, though if there are weak matches on several lands, such as the marked Hamby 44 comparisons in \autoref{fig:compare}, the overall score may be affected.
+}
 
@@
 \autoref{fig:tankrash} gives an overview of all lands of affected bullets in the Hamby-44 set.
-These issues affect the algorithm's ability to identify a stable signature, which impacts the subsequent match scores  at the land-to-land and bullet-to-bullet level.
+These issues affect the algorithm's ability to identify a stable signature, which impacts the subsequent match scores \svp{ at the land-to-land and bullet-to-bullet level.}
 ```{r tankrash, fig.cap="Overview of bullet lands with prominent deficiency such as tank rash or extreme pitting in the Hamby-44 study.", fig.width=4, fig.height = 2, out.width='47.5%', fig.subcap=c("Br1-B1-L6 (tank rask)", "Br2-B2-L5 (tank rash)", "Br3-B1-L5 (tank rash)", "Br8-B1-L6 (tank rash)", "Br8-B2-L2 (pitting)", "Br8-B2-L6 (tank rash)", "Unk-E-L6 (tank rash)", "Unk-I-L6 (tank rash and break-off)"), fig.align="middle", fig.ncol=2}
@@
 We also see that questioned bullet 'I' is affected, explaining some of the low similarity scores for this bullet.
-None of the lands of any of the bullets in the other studies are affected by tank rash in a similar manner.
+\hh{None of the lands of any of the bullets in the other studies are affected by tank rash in a similar manner.}
 
@@
 \autoref{fig:eer} shows an overview of Equal Error and their thresholds based on the ROC curves of \autoref{fig:roc-auc}.
-Equal Errors are errors when sensitivity and specificity of a test are equal, i.e.\ we see the same percentage of missed and wrong identifications for a land-to-land comparison.
+Equal Errors are errors when sensitivity and specificity of a test are equal, i.e.
+we see the same percentage of missed and wrong identifications for a land-to-land comparison.
 Equal errors based on CMS of two lands are significantly higher than equal errors based on cross-correlation or the RF score.
@@
 Interestingly, for consecutively matching striae, equal errors are reached for just two matching striae, which from a practical point of view is not sustainable.
-A  CMS of  6 or above is an established cutoff used in identifications; when applied to the land-to-land comparisons, it is much more likely to result in a missed identification than a wrong identification.
+\svp{A  CMS of  6 or above is an established cutoff used in identifications; when applied to the land-to-land comparisons, it is much more likely to result in a missed identification than a wrong identification.
 A threshold of 6 leads to 21 wrong identifications out of 49,286 different-source comparisons (0.04\% errors, or about 4 errors for each 10,000 evaluations); when applied to same-source comparisons, 517 of 715 same-source lands were incorrectly classified as exclusions, for an error rate of 80.6\%.
+}
+<!-- This leads to 21 wrong identifications (out of 49286; for an error of 4 in 10,000 -- or 0.04\%) but only identifies 139 out of 715 known same-source lands, for an error of 80.6\%.  -->
 
@@
 This would emulate the process used by examiners, who first assess the quality of the evidence and whether there is enough information present to attempt a match.
-One major disadvantage of an automated algorithm is that all of the decisions humans make (recognizing degraded land areas, excluding those areas from consideration, matching only the remaining areas) must be explicitly characterized; however, this explicit characterization means that the process can be scientifically validated to a much higher degree than the human perceptual process.
+\svp{One major disadvantage of an automated algorithm is that all of the decisions humans make (recognizing degraded land areas, excluding those areas from consideration, matching only the remaining areas) must be explicitly characterized; however, this explicit characterization means that the process can be scientifically validated to a much higher degree than the human perceptual process.
+}
 
@@
 
-The random forest matching algorithm presented in \citet{aoas} does perform well on three different external test sets.
-While a single cutoff value cannot be used to distinguish matches and non-matches across the different test sets, the algorithm makes no errors when a set-specific cutoff value is used on bullet-to-bullet scores aggregated using sequence average maximum.
-\hh{This performance is on par with the performance of forensic examiners.
-For a future round of the Houston study, we are planning to compare the algorithmic performance directly with scores given by forensic toolmark examiners.
-XXX need to ask XXX \sout{However, the intent is not to replace a fornesic examiner.
-The decision on matching should not be taken away from the human assessment.
-Instead, the algorithm's score is meant as a complement to an examiner's testimony as a way to measure and quantify the strength of an identification.}  }\svp{The algorithm's score is not intended to replace an examiner's judgment; instead, it is a tool to measure and quantify the strength of an identification.
-}
-By validating the algorithm on external test sets, we have demonstrated that the method can be generalized to different types of ammunition and is not overly sensitive to small differences in rifling procedure.
-Future studies can \hh{and should} generalize this to a wider range of external test sets to establish the limits of the algorithm's generalizability.
+\svp{The random forest matching algorithm presented in \citet{aoas} does perform well on three different external test sets.
+While a single cutoff value cannot be used to distinguish matches and non-matches across the different test sets, the algorithm makes no errors when a set-specific cutoff value is used on bullet-to-bullet scores aggregated using sequence average maximum.} \svp{XXX comparison to FTEs XXX} \svp{By validating the algorithm on external test sets, we have demonstrated that the method can be generalized to different types of ammunition and is not overly sensitive to small differences in rifling procedure.
+Future studies can generalize this to a wider range of external test sets to establish the limits of the algorithm's generalizability.}
 
@@
 # References
-
-<!--include marginal distributions of training data and compare to test data-->
-
-```{r, echo=FALSE, eval=FALSE}
-training <- read.csv("data/features-hamby.csv")
-gtrain <- training %>% 
-  select(profile1_id, profile2_id, study1, study2, match, ccf:sum_peaks) %>%
-  gather(feature, value, ccf:sum_peaks)
-# now compare to test data
-```

### 2018-11-27 index.Rmd
remove comments and accept changes, fix table

@@
 In virtually all labs, individual characteristics used to identify matching bullets are derived from visual assessment; some class characteristics may be directly measured, but these are not sufficient for individualization.
-<!-- Want to make it clear that individual characteristics are visually derived, rather than numerically derived...
-hopefully managed that without being insulting.
-HH: reads well -->
 
@@
 
-<!-- XXXX get from area to crosscut XXXX -->
-<!-- \hh{ -->
-<!-- Let $z(t)$ describe a digitized toolmark for the striation marks on a land engraved area. -->
-<!-- $z(t)$ is then a spatial process indexed by $t$, where $t, t = 1,...,T$ denotes equally spaced pixel locations for a total length of $T$. } -->
 \begin{itemize}
@@
 \] 
-where $x$ and $x+z$ are integer values appropriately defined within the domains of $S_1$ and $S_2$, $z$ is the lag between the two sequences, \svp{and }\hh{$\text{cor}(.,.)$ is the Pearson correlation coefficient.}
+where $x$ and $x+z$ are integer values appropriately defined within the domains of $S_1$ and $S_2$, $z$ is the lag between the two sequences, \svp{and }$\text{cor}(.,.)$ is the Pearson correlation coefficient.
 When signatures contain missing values, pairwise complete observations are used to calculate the cross-correlation. The lag $z$ used to achieve maximum $CC$ is used to determine the best alignment.
@@
 
-\hh{Using the lag determined for aligning two signatures,} other quantitative features describing the similarity of the two signatures can be extracted.
+Using the lag determined for aligning two signatures, other quantitative features describing the similarity of the two signatures can be extracted.
 Numerical features such as cross-correlation and Euclidean distance
 can be computed from the aligned signatures alone.
-\hh{Other} features include characteristics of the signature that firearms and toolmark examiners base their visual assessment on, such as striae depth, number of matching striae, and the number of consecutively matching striae (CMS), \hh{one of the measures firearms examiners use to describe the strength of a match quantitatively.}
-
-\svp{In order to evaluate CMS, we must first identify peaks and valleys in each of the signatures, then determine whether these peaks and valleys overlap sufficiently.
+Other features include characteristics of the signature that firearms and toolmark examiners base their visual assessment on, such as striae depth, number of matching striae, and the number of consecutively matching striae (CMS), one of the measures firearms examiners use to describe the strength of a match quantitatively.
+
+In order to evaluate CMS, we must first identify peaks and valleys in each of the signatures, then determine whether these peaks and valleys overlap sufficiently.
 Features which firearms examiners assess visually depend on the identification of these extrema.
-} If there are at least six consecutive matching striae, the bullets are considered to be a good match in practice \citep{biasotti_firearms_1997,nichols_scientific_2006}.
+If there are at least six consecutive matching striae, the bullets are considered to be a good match in practice \citep{biasotti_firearms_1997,nichols_scientific_2006}.
 CMS in this setting are counted as the number of consecutively matching peaks in signatures of two aligned lands.
@@
 
-<!-- \svp{It is likely that matches which were deemed  inconclusive would instead be marked as exclusions by the algorithm.
-XXX I don't know if we can make this statement, but I think it would help examiners understand how the line gets drawn if we could.
-I have exactly no evidence to back that up (except maybe the Hamby 44 tank rash results?)} -->
-
 <!-- ## Identification Errors and Algorithm Evaluation -->
@@
 The full range of possible outcomes is shown in \autoref{tab:possible-outcomes}.
-\hh{XXX I know that you have spent a ton of time on this table - but could you make it less tall?
-It's taking a ton of space.
-}
+
 \begin{table}
 \centering
-\begin{tabular}{c|ccc}
-\multicolumn{2}{c}{} & \multicolumn{2}{c}{\textbf{Ground Truth}}\\\cline{3-4}
-\multicolumn{2}{c}{} & \multicolumn{1}{|c}{Same Source} & \multicolumn{1}{|c|}{Different Source} \\\cline{2-4}
-\multirow{8}{*}{\rotatebox[origin=c]{90}{\hfill \textbf{Results}\hspace{5mm}\hfill}} & 
-  \multirow{4}{*}{\rotatebox[origin=c]{90}{\hspace{1mm}Exclusion\hspace{1mm}}} & \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\
-  % & &   \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\ 
-  & &  \multicolumn{1}{|c}{Missed} & \multicolumn{1}{|c|}{Correct}\\
-  & &  \multicolumn{1}{|c}{Identification} & \multicolumn{1}{|c|}{Exclusion}\\
-  % & &   \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\ 
-  & &  \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{} \\\cline{2-4}
-& \multirow{4}{*}{\rotatebox[origin=c]{90}{\hspace{1mm}Identification\hspace{1mm}}}  & \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\
-  % & &  \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\ 
-  & &  \multicolumn{1}{|c}{Correct} & \multicolumn{1}{|c|}{Wrong}\\
-  & &  \multicolumn{1}{|c}{Identification} & \multicolumn{1}{|c|}{Identification}\\
-  % & &  \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{}\\ 
-  & &  \multicolumn{1}{|c}{} & \multicolumn{1}{|c|}{} \\\cline{2-4}
+\bgroup
+\def\arraystretch{.8}%  1 is the default, change whatever you need
+\begin{tabular}{|ccc}
+\multicolumn{1}{c}{} & \multicolumn{2}{c}{\textbf{Ground Truth}}\\\cline{2-3}
+\multicolumn{1}{c}{\textbf{Results}} & \multicolumn{1}{|c}{Same Source} & \multicolumn{1}{|c|}{Different Source} \\\hline
+\multirow{2}{*}{Exclusion} & \multicolumn{1}{|c}{Missed} & \multicolumn{1}{|c|}{Correct}\\
+   &  \multicolumn{1}{|c}{Identification} & \multicolumn{1}{|c|}{Exclusion}\\\hline
+\multirow{2}{*}{Identification}  &  \multicolumn{1}{|c}{Correct} & \multicolumn{1}{|c|}{Wrong}\\
+   &  \multicolumn{1}{|c}{Identification} & \multicolumn{1}{|c|}{Identification}\\\hline
 \end{tabular}
+\egroup
 \caption{Possible outcomes of an examination of two pieces of evidence.
@@
 While a single cutoff value cannot be used to distinguish matches and non-matches across the different test sets, the algorithm makes no errors when a set-specific cutoff value is used on bullet-to-bullet scores aggregated using sequence average maximum.
-\hh{This performance is on par with the performance of forensic examiners.
+This performance is on par with the performance of forensic examiners.
 For a future round of the Houston study, we are planning to compare the algorithmic performance directly with scores given by forensic toolmark examiners.
-XXX need to ask XXX \sout{However, the intent is not to replace a fornesic examiner.
+\hh{XXX need to ask XXX \sout{However, the intent is not to replace a fornesic examiner.
 The decision on matching should not be taken away from the human assessment.
@@
 By validating the algorithm on external test sets, we have demonstrated that the method can be generalized to different types of ammunition and is not overly sensitive to small differences in rifling procedure.
-Future studies can \hh{and should} generalize this to a wider range of external test sets to establish the limits of the algorithm's generalizability.
+Future studies can and should generalize this to a wider range of external test sets to establish the limits of the algorithm's generalizability.
 

### 2018-11-27 index.Rmd
Cite press release for sensofar

@@
 
-Note that the sequence average maximum of the correlation between lands is used in SensoComp \hh{citation for that?}\svp{It's your claim...
-I looked for a citation and didn't find one.
-Closed-source proprietary = no references?} to capture the similarity between bullets.
+Note that the sequence average maximum of the correlation between lands is used in SensoComp \citep{sensofar-sam} to capture the similarity between bullets.
 The correlation based SAM score has also been called the 'average correlation calculated at the max phase' in \citet{chuPilotStudyAutomated2010}.

### 2018-11-27 index.Rmd
Add subheadings back in, slightly rearrange and add "flow" text

@@
 
-<!-- ## Automated Processing of 3D Scans -->
-
-In the matching algorithm proposed in \citet{aoas}, several of the features described above are combined using a random forest \citep{breiman} to produce a similarity score based on 3D topographic scans of land engraved areas (LEAs).
+In the matching algorithm proposed in \citet{aoas}, several features are combined using a random forest \citep{breiman} to produce a similarity score based on 3D topographic scans of land engraved areas (LEAs).
+\svp{In order to generate these features, some pre-processing is necessary in order to transform the 3D surface measurements into `signatures' which can be compared.}
+
+## Automated Processing of 3D Scans
 From each land engraved area a signature is derived using the process described in detail in \citet{aoas}:
@@
 
-\svp{Cross correlation can also be used to assess match strength between two aligned signatures.
-Cross correlation varies between -1 and 1, but as we are taking the maximal value, in practice the cross-correlations for aligned signatures are generally positive.
-Cross-correlation values which are close to 1 indicate similarity between signatures and with it may indicate that the two signatures come from the same source.
-Cross-correlation values which are close to 0 indicate signatures which may originate from different sources.}
-
-<!-- ## Statistics for Matching Aligned Signatures -->
+## Statistics for Matching Aligned Signatures
 
@@
 
+<!-- I really think this belongs in this section: it's discussing cross-correlation as a scoring statistic, rather than as a convenient way to align signatures so that scoring stats can be calculated.
+-->
+
+\svp{Cross correlation can also be used to assess match strength between two aligned signatures.
+Cross correlation varies between -1 and 1, but as we are taking the maximal value, in practice the cross-correlations for aligned signatures are generally positive.
+Cross-correlation values which are close to 1 indicate similarity between signatures and with it may indicate that the two signatures come from the same source.
+Cross-correlation values which are close to 0 indicate signatures which may originate from different sources.}
+
 The random forest presented in \citet{aoas} is based on a combination of multiple characteristics, including cross-correlation, number of matching striae, and number of consecutively matching striae.
-The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where scores close to 1 indicate it to be highly likely that the aligned signatures come from the same source, while low scores indicate that the aligned signatures are highly likely to originate from different lands.
-
-<!-- ## Interpreting the Algorithm Score -->
+The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where scores close to 1 indicate it to be highly likely that the aligned signatures come from the same source, while scores close to 0 indicate that the aligned signatures are highly likely to originate from different sources.
+
+## Interpreting the Algorithm Score
+
+\svp{The features described above are each individually designed to separate same-source and different-source signatures: for instance, same-source signatures would have high cross-correlation values and different-source signatures would have low cross-correlation values.
+An automated algorithm selects a cutoff value, introducing a barrier between identification and exclusion.}
+<!-- rephrased? trying to get rid of "supposed" and "same sourceness" :-p
 
@@
 }
+-->
 
@@
 -->
+
 This threshold system introduces a binary classification: identification or exclusion.
@@
 
-<!-- ## Identification Errors and Algorithm Evaluation -->
+## Identification Errors and Algorithm Evaluation
 
@@
 
-<!-- ## Assessing Match Strength -->
+## Assessing Bullet-to-Bullet Match Strength
 
@@
 all lands on a single bullet), and a conclusion on same or different source should be reached based on the evidence of all lands of the bullets.
-<!--to describe the correspondence between two pieces of physical evidence.-->
-<!-- \svp{In order to aggregate land-to-land match scores, we must consider the physical structure of the evidence; that is, that the sequence of lands stems from spatial information, but the numbering of each land represents the scan order and start point rather than encoding useful information.
-} --><!--This last bit might be redundant-->
 
@@
 \end{definition}
-
-Note that the sequence average maximum of the correlation between lands is used in SensoComp \citep{sensofar-sam} to capture the similarity between bullets.
+\svp{XXX Rearranged this slightly XXX}
+Looking back at \autoref{fig:b2b}, we see that for the two bullets from the same barrel, the sequence average for 
+$k=2$ is higher than the other sequence averages, and also higher than the sequence averages for the other pair of bullets shown on the right of the figure.
+SAM scores allow us to define a single quantity for each pair of bullets that describes the similarity between these two bullets.
+
+The sequence average maximum of the correlation between lands is used in SensoComp \citep{sensofar-sam} to capture the similarity between bullets.
 The correlation based SAM score has also been called the 'average correlation calculated at the max phase' in \citet{chuPilotStudyAutomated2010}.
@@
 ```
-
-Looking back at \autoref{fig:b2b}, we see that for the two bullets from the same barrel, the sequence average for 
-$k=2$ is higher than the other sequence averages, and also higher than the sequence averages for the other pair of bullets shown on the right of the figure.
-
-SAM scores allow us to define a single quantity for each pair of bullets that describes the similarity between these two bullets.
 

### 2018-11-27 index.Rmd
phrasing

@@
 
-What is quite surprising is that while we anticipated Hamby 44 would be the easiest set to work with, it turned out to be the hardest to solve, in part because of damage to the bullets that obscure LEA striae.
+\svp{At the beginning of this study,} we anticipated Hamby 44 would be the easiest set to work with \svp{because of its similarity to the sets used to train the random forest algorithm.
+In a surprise turn of events, it was the}  hardest to solve, in part because of damage to the bullets that obscured LEA striae.
 It has been shown \citet{lpr} that parts of lands can be used for successful identifications.

### 2018-11-27 index.Rmd
dpi fix

@@
   fig.width = 6,
-  dpi = 300,
+  dpi = 600,
   fig.align = "center",

### 2018-11-30 index.Rmd
couple more edits

@@
 # to produce blinded version set to 1
-blinded: 1
+blinded: 0
 # This gets changed automatically when the title page is compiled in the chunk child-doc-compile
-titlepage: 0
+titlepage: 1
 # Set to 1 if figures should be included with the text
-draft: 0
+draft: 1
 
@@
 thanks: 
-- The authors wish to thank Jim Hamby, Tylor Klep (Phoenix PD), and Melissa Nally and Kasi Kirksey (FSI Houston) for access to test set bullets. 
-- XXX Scanning. 
-- Thanks to Jim Hamby for his advice on XXX.
+- The authors wish to thank Alan Zheng (NIST), Tylor Klep (Phoenix PD), and Melissa Nally and Kasi Kirksey (FSI Houston) for access to test set bullets. 
+- We would like to thank the efforts of the Roy J Carver High Resolution lab in scanning the bullet sets and providing the scans to us. 
+- Thanks to Prof James Hamby for his advice on XXX.
 
@@
 - toolmark
-- cross-correlation
+- cross correlation
 - random forest
@@
 2017b) in three studies.
-We are considering matching performance based on the Random forest score, cross-correlation, and consecutive matching striae (CMS) at the land-to-land level and, using Sequential Average Maxima scores, also at the bullet-to bullet level.
-Cross-correlation and Random Forest scores both result in perfect discrimination of same-source and different-source bullets.
+We are considering matching performance based on the Random forest score, cross corrleation, and consecutive matching striae (CMS) at the land-to-land level and, using Sequential Average Maxima scores, also at the bullet-to bullet level.
+Cross corrleation and Random Forest scores both result in perfect discrimination of same-source and different-source bullets.
 At the land-to-land level, discrimination (based on area under the curve, AUC) is excellent ($ge$ 0.90).
@@
 \citet{ComputerIdentificationBullets1978} demonstrated use of a scanning electron microscope to quantitatively examine and compare bullet striae.
-More recently, approaches using 3D measurement data were explicitly described in 1999 (\citeauthor{dekinderAutomatedComparisonsBullet1999}), and have been further developed in \citep{bachrachDevelopment3DbasedAutomated2002,xieAutomatedBulletidentificationSystem2009, chuPilotStudyAutomated2010}.
+More recently, approaches using 3D measurement data were explicitly described in 1999 by \citeauthor{dekinderAutomatedComparisonsBullet1999} \cite{dekinderAutomatedComparisonsBullet1999}, and have been further developed in \citep{bachrachDevelopment3DbasedAutomated2002,xieAutomatedBulletidentificationSystem2009, chuPilotStudyAutomated2010}.
 These approaches utilize 3D surface measurements directly to characterize the topology of land engraved areas (LEAs), rather than using visual or image comparison techniques.
@@
 Individual characteristics, which are not shared by all members of a class, can also be automatically assessed from 3D surface measurements.
-One of the most common features used to assess the strength of a match based on individual characteristics is the cross-correlation function, which is utilized in several studies \citep{maNISTBulletSignature2004, vorburgerApplicationsCrosscorrelationFunctions2011, chuPilotStudyAutomated2010}.
+One of the most common features used to assess the strength of a match based on individual characteristics is the cross corrleation function, which is utilized in several studies \citep{maNISTBulletSignature2004, vorburgerApplicationsCrosscorrelationFunctions2011, chuPilotStudyAutomated2010}.
 Additional features proposed for automatically assessing match strength also include signature distance \citep{maNISTBulletSignature2004}, striae depth and width \citep{ComputerIdentificationBullets1978}, and consecutive matching striae (CMS) \citep{chuAutomaticIdentificationBullet2013}.
@@
 In the matching algorithm proposed in \citet{aoas}, several features are combined using a random forest \citep{breiman} to produce a similarity score based on 3D topographic scans of land engraved areas (LEAs).
-\svp{In order to generate these features, some pre-processing is necessary in order to transform the 3D surface measurements into `signatures' which can be compared.}
+In order to generate these features, some pre-processing is necessary in order to transform the 3D surface measurements into `signatures' which can be compared.
 
@@
 Once signatures have been extracted from stable regions with expressed striae, they must be aligned in order to assess their match strength.
-Maximized cross-correlation is used to pair-wise align signatures \citep{aoas,vorburgerApplicationsCrosscorrelationFunctions2011}.
-
-The cross-correlation (CC) function between two signatures $S_1(x)$ and $S_2(x)$ is defined as 
+Maximized cross corrleation is used to pair-wise align signatures \citep{aoas,vorburgerApplicationsCrosscorrelationFunctions2011}.
+
+The cross corrleation (CC) function between two signatures $S_1(x)$ and $S_2(x)$ is defined as 
 \[
@@
 \] 
-where $x$ and $x+z$ are integer values appropriately defined within the domains of $S_1$ and $S_2$, $z$ is the lag between the two sequences, \svp{and }$\text{cor}(.,.)$ is the Pearson correlation coefficient.
-When signatures contain missing values, pairwise complete observations are used to calculate the cross-correlation. The lag $z$ used to achieve maximum $CC$ is used to determine the best alignment.
+where $x$ and $x+z$ are integer values appropriately defined within the domains of $S_1$ and $S_2$, $z$ is the lag between the two sequences, and $\text{cor}(.,.)$ is the Pearson correlation coefficient.
+When signatures contain missing values, pairwise complete observations are used to calculate the cross corrleation. The lag $z$ used to achieve maximum $CC$ is used to determine the best alignment.
 
@@
 Using the lag determined for aligning two signatures, other quantitative features describing the similarity of the two signatures can be extracted.
-Numerical features such as cross-correlation and Euclidean distance
+Numerical features such as cross corrleation and Euclidean distance
 can be computed from the aligned signatures alone.
@@
 Peaks -- rather than peaks and valleys -- are used for this determination in part because it is difficult to visually assess valleys except as relative to peaks.
-Note that with automatic alignment based on the maximum cross-correlation rather than matching striation marks, it is possible for two lands to have a consecutive matching striae of zero.
-
-<!-- I really think this belongs in this section: it's discussing cross-correlation as a scoring statistic, rather than as a convenient way to align signatures so that scoring stats can be calculated.
+Note that with automatic alignment based on the maximum cross corrleation rather than matching striation marks, it is possible for two lands to have a consecutive matching striae of zero.
+
+<!-- I really think this belongs in this section: it's discussing cross corrleation as a scoring statistic, rather than as a convenient way to align signatures so that scoring stats can be calculated.
 -->
 
-\svp{Cross correlation can also be used to assess match strength between two aligned signatures.
-Cross correlation varies between -1 and 1, but as we are taking the maximal value, in practice the cross-correlations for aligned signatures are generally positive.
-Cross-correlation values which are close to 1 indicate similarity between signatures and with it may indicate that the two signatures come from the same source.
-Cross-correlation values which are close to 0 indicate signatures which may originate from different sources.}
-
-The random forest presented in \citet{aoas} is based on a combination of multiple characteristics, including cross-correlation, number of matching striae, and number of consecutively matching striae.
+Cross correlation can also be used to assess match strength between two aligned signatures.
+Cross correlation varies between -1 and 1, but as the alignment is based on the maximized cross correlation value, in practice the cross correlations for aligned signatures are generally positive.
+Cross correlation values which are close to 1 indicate similarity between signatures and with it may indicate that the two signatures come from the same source.
+Low cross corrleation values indicate signatures which may originate from different sources.
+
+The random forest presented in \citet{aoas} is based on a combination of multiple characteristics, including cross corrleation, number of matching striae, and number of consecutively matching striae.
 The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where scores close to 1 indicate it to be highly likely that the aligned signatures come from the same source, while scores close to 0 indicate that the aligned signatures are highly likely to originate from different sources.
@@
 
-\svp{The features described above are each individually designed to separate same-source and different-source signatures: for instance, same-source signatures would have high cross-correlation values and different-source signatures would have low cross-correlation values.
-An automated algorithm selects a cutoff value, introducing a barrier between identification and exclusion.}
-<!-- rephrased? trying to get rid of "supposed" and "same sourceness" :-p
-
-\hh{Any of the above described features are supposed to separate between signatures from same source and different source, e.g.\ high cross-correlation values are indicative of same sourceness of two signatures, whereas  low values are more indicate of different sources.
-An automated algorithm determines a cutoff-value along the range of possible values that introduces a barrier between identification and exclusion.
-}
--->
-
-<!-- Firearms examiners typically set the threshold for CMS at 6; this is a threshold which is conservative, as it is much more likely to exclude same-source comparisons than it is to identify different-source comparisons.
--->
+The features described above are each individually designed to separate same-source and different-source signatures: for instance, same-source signatures would have high cross corrleation values and different-source signatures would have low cross corrleation values.
+An automated algorithm selects a cutoff value, introducing a barrier between identification and exclusion.
 
@@
 \end{definition}
-\svp{XXX Rearranged this slightly XXX}
 Looking back at \autoref{fig:b2b}, we see that for the two bullets from the same barrel, the sequence average for 
@@
 We assess the results on each set and evaluate whether the minimal requirements described above are fulfilled for bullet-to-bullet SAM scores.
-We compare the random forest algorithm's performance to the performance of other suggested measures for quantitative assessment of match strength, such as cross-correlation and consecutive matching striae.
+We compare the random forest algorithm's performance to the performance of other suggested measures for quantitative assessment of match strength, such as cross corrleation and consecutive matching striae.
 
@@
 With no errors at the bullet-to-bullet level, we can now evaluate the scores in light of (R1) and (R2).
-In this section, we will compare the random forest scores to cross-correlation scores; cross-correlation is one of the components of the random forest, but has also been used to quantify the strength of a match in its own right \citep{chuPilotStudyAutomated2010}.
+In this section, we will compare the random forest scores to cross corrleation scores; cross corrleation is one of the components of the random forest, but has also been used to quantify the strength of a match in its own right \citep{chuPilotStudyAutomated2010}.
 In addition, both scoring methods use the same scale (0 to 1) and are continuous along that interval; other evaluation methods, such as consecutive matching striae are discrete and more difficult to directly compare to continuous measurements.
@@
 ## Comparison of SAM scores
-\autoref{fig:compare} shows density curves for the scores bullet-to-bullet matches of each of the studies, comparing the SAM scores based on cross-correlation (top) and the random forest score (bottom).
+\autoref{fig:compare} shows density curves for the scores bullet-to-bullet matches of each of the studies, comparing the SAM scores based on cross corrleation (top) and the random forest score (bottom).
 Color indicates ground truth.
 Small vertical lines below the $x$-axis indicate observed scores in a **rug-plot**.
-The plot shows that for all three case studies and both continuous measures, i.e.\ SAM scores based on cross-correlation as well as SAM scores based on the random forest scores, all similarity scores are higher for pairs of bullets from the same source, i.e.\ bullets fired through the same barrel, than pairs of bullets from different sources (fired through different barrels).
-
-```{r compare, fig.width = 8, fig.height = 5, fig.cap="Density curves of similarity scores from cross-correlation (top) and RF scores (bottom).
+The plot shows that for all three case studies and both continuous measures, i.e.\ SAM scores based on cross corrleation as well as SAM scores based on the random forest scores, all similarity scores are higher for pairs of bullets from the same source, i.e.\ bullets fired through the same barrel, than pairs of bullets from different sources (fired through different barrels).
+
+```{r compare, fig.width = 8, fig.height = 5, fig.cap="Density curves of similarity scores from cross corrleation (top) and RF scores (bottom).
 Different colors indicate same source versus different source.
@@
 nestslong$measure <- factor(nestslong$measure)
-levels(nestslong$measure) <- c("Cross-correlation",  "RF score")
+levels(nestslong$measure) <- c("Cross corrleation",  "RF score")
 
@@
 
-According to \autoref{fig:compare}, neither cross-correlation nor random forest score fulfill both of the minimal requirements laid out in the introduction.
-Both measures fulfill (R1), i.e.\ all  scores  of  different-source pairs are lower than scores of same-source pairs for SAM scores based on cross-correlation and RF scores.
-However, SAM scores based on the multivariate random forest score show better separation between same-source and different-source scores than SAM scores based on cross-correlation.
-This can be seen from the larger horizontal distance between the modes (peaks) of the density curves of RF based SAM scores compared to the cross-correlation SAM scores.
+According to \autoref{fig:compare}, neither cross corrleation nor random forest score fulfill both of the minimal requirements laid out in the introduction.
+Both measures fulfill (R1), i.e.\ all  scores  of  different-source pairs are lower than scores of same-source pairs for SAM scores based on cross corrleation and RF scores.
+However, SAM scores based on the multivariate random forest score show better separation between same-source and different-source scores than SAM scores based on cross corrleation.
+This can be seen from the larger horizontal distance between the modes (peaks) of the density curves of RF based SAM scores compared to the cross corrleation SAM scores.
 
@@
 \autoref{fig:compare-land-to-land} shows density curves for each study.
-There is some overlap between known matching and known non-matching land-to-land scores; in many cases the random forest scores for these overlapping values are more extreme than the corresponding cross-correlation scores.
+There is some overlap between known matching and known non-matching land-to-land scores; in many cases the random forest scores for these overlapping values are more extreme than the corresponding cross corrleation scores.
 This suggests that the matching algorithm is sensitive to the presence of LEAs which have low land-to-land match scores, such as the matches of bullets 'I' and 'F' in the Hamby 44 study.
@@
 
-```{r compare-land-to-land, fig.width = 8, fig.height = 4, fig.cap="Density curves of land-to-land similarity scores from cross-correlation (top) and RF scores (bottom).
+```{r compare-land-to-land, fig.width = 8, fig.height = 4, fig.cap="Density curves of land-to-land similarity scores from cross corrleation (top) and RF scores (bottom).
 Different colors indicate same source versus different source for each land.
@@
   mutate(measure = factor(measure, levels = c("ccf", "rfscore"),
-                          labels = c("Cross-correlation", "RF score")),
+                          labels = c("Cross corrleation", "RF score")),
          study = paste(study, group1) %>% str_trim() %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"))) %>%
@@
 
-```{r roc-auc, echo = FALSE, out.width = c("0.51\\textwidth", ".4725\\textwidth"), out.height = c("0.63\\textwidth", "0.63\\textwidth"), fig.width = c(5, 6), fig.height = c(7.5, 7.5), fig.subcap=c("ROC curves", "Area under the curve (AUC)"), fig.cap="ROC curves and AUC for each (set) of the studies for the random forest score (RF), cross-correlation (CC) and consecutively matching striae (CMS).
+```{r roc-auc, echo = FALSE, out.width = c("0.51\\textwidth", ".4725\\textwidth"), out.height = c("0.63\\textwidth", "0.63\\textwidth"), fig.width = c(5, 6), fig.height = c(7.5, 7.5), fig.subcap=c("ROC curves", "Area under the curve (AUC)"), fig.cap="ROC curves and AUC for each (set) of the studies for the random forest score (RF), cross corrleation (CC) and consecutively matching striae (CMS).
 As seen in the ROC curves, the algorithm performs the least well on set 2 of the Houston FSI study.
-Based on AUC, the overall performance on all sets is very good to excellent for cross-correlation and the random forest score.
-At the land level there is no significant difference in prediction power between RF score and cross-correlation.", cache=FALSE}
+Based on AUC, the overall performance on all sets is very good to excellent for cross corrleation and the random forest score.
+At the land level there is no significant difference in prediction power between RF score and cross corrleation.", cache=FALSE}
   rocpic
@@
 AUC values are useful for differentiating between poor, good, and excellent model performance, but are not particularly useful when determining which of several models with approximately the same level of performance should be used \citep{marzban_roc_2004}.
-Hamby 44 and Phoenix PD are the only studies with a somewhat noticable difference between the AUC score for RF compared to cross-correlation, but even those differences are not statistically significant.
+Hamby 44 and Phoenix PD are the only studies with a somewhat noticable difference between the AUC score for RF compared to cross corrleation, but even those differences are not statistically significant.
 
@@
 Equal Errors are errors when sensitivity and specificity of a test are equal, i.e.\ we see the same percentage of missed and wrong identifications for a land-to-land comparison.
-Equal errors based on CMS of two lands are significantly higher than equal errors based on cross-correlation or the RF score.
-At best, RF score and cross-correlation have an equal error of around 5\% for land-to-land comparisons.
+Equal errors based on CMS of two lands are significantly higher than equal errors based on cross corrleation or the RF score.
+At best, RF score and cross corrleation have an equal error of around 5\% for land-to-land comparisons.
 Interestingly, for consecutively matching striae, equal errors are reached for just two matching striae, which from a practical point of view is not sustainable.
@@
 
-```{r eer, fig.width=7, fig.height = 4, fig.cap="Equal Error (in Percent)  and corresponding thresholds for all measures (RF score, Cross-correlation and CMS) under all five (sub-)studies.
+```{r eer, fig.width=7, fig.height = 4, fig.cap="Equal Error (in Percent)  and corresponding thresholds for all measures (RF score, Cross corrleation and CMS) under all five (sub-)studies.
 CMS has an equal error threshold of 2 for all studies.
-CMS is signifcantly worse than both Cross-correlation and the RF score.
-At best, Cross-correlation and RF score have equal errors for land-to-land comparison of around 5\\%.", out.width = '.8\\textwidth'}
+CMS is signifcantly worse than both Cross corrleation and the RF score.
+At best, Cross corrleation and RF score have equal errors for land-to-land comparison of around 5\\%.", out.width = '.8\\textwidth'}
 aucs %>% 
@@
 
-\svp{At the beginning of this study,} we anticipated Hamby 44 would be the easiest set to work with \svp{because of its similarity to the sets used to train the random forest algorithm.
+At the beginning of this study, we anticipated Hamby 44 would be the easiest set to work with \svp{because of its similarity to the sets used to train the random forest algorithm.
 In a surprise turn of events, it was the}  hardest to solve, in part because of damage to the bullets that obscured LEA striae.
@@
 For a future round of the Houston study, we are planning to compare the algorithmic performance directly with scores given by forensic toolmark examiners.
-\hh{XXX need to ask XXX \sout{However, the intent is not to replace a fornesic examiner.
-The decision on matching should not be taken away from the human assessment.
-Instead, the algorithm's score is meant as a complement to an examiner's testimony as a way to measure and quantify the strength of an identification.}  }\svp{The algorithm's score is not intended to replace an examiner's judgment; instead, it is a tool to measure and quantify the strength of an identification.
-}
+\hh{XXX need to ask XXX  } The algorithm's score is not intended to replace an examiner's judgment; instead, it provides a complementary tool to measure and quantify the strength of an identification.
 By validating the algorithm on external test sets, we have demonstrated that the method can be generalized to different types of ammunition and is not overly sensitive to small differences in rifling procedure.

### 2018-12-11 index.Rmd
spelling issues, comment out extraneous stuff

@@
 # to produce blinded version set to 1
-blinded: 0
+blinded: 1
 # This gets changed automatically when the title page is compiled in the chunk child-doc-compile
@@
 2017b) in three studies.
-We are considering matching performance based on the Random forest score, cross corrleation, and consecutive matching striae (CMS) at the land-to-land level and, using Sequential Average Maxima scores, also at the bullet-to bullet level.
-Cross corrleation and Random Forest scores both result in perfect discrimination of same-source and different-source bullets.
+We are considering matching performance based on the Random forest score, cross correlation, and consecutive matching striae (CMS) at the land-to-land level and, using Sequential Average Maxima scores, also at the bullet-to bullet level.
+Cross correlation and Random Forest scores both result in perfect discrimination of same-source and different-source bullets.
 At the land-to-land level, discrimination (based on area under the curve, AUC) is excellent ($ge$ 0.90).
@@
 
+<!--
 \hh{
@@
 Individual characteristics, which are not shared by all members of a class, can also be automatically assessed from 3D surface measurements.
-One of the most common features used to assess the strength of a match based on individual characteristics is the cross corrleation function, which is utilized in several studies \citep{maNISTBulletSignature2004, vorburgerApplicationsCrosscorrelationFunctions2011, chuPilotStudyAutomated2010}.
+One of the most common features used to assess the strength of a match based on individual characteristics is the cross correlation function, which is utilized in several studies \citep{maNISTBulletSignature2004, vorburgerApplicationsCrosscorrelationFunctions2011, chuPilotStudyAutomated2010}.
 Additional features proposed for automatically assessing match strength also include signature distance \citep{maNISTBulletSignature2004}, striae depth and width \citep{ComputerIdentificationBullets1978}, and consecutive matching striae (CMS) \citep{chuAutomaticIdentificationBullet2013}.
@@
 Once signatures have been extracted from stable regions with expressed striae, they must be aligned in order to assess their match strength.
-Maximized cross corrleation is used to pair-wise align signatures \citep{aoas,vorburgerApplicationsCrosscorrelationFunctions2011}.
-
-The cross corrleation (CC) function between two signatures $S_1(x)$ and $S_2(x)$ is defined as 
+Maximized cross correlation is used to pair-wise align signatures \citep{aoas,vorburgerApplicationsCrosscorrelationFunctions2011}.
+
+The cross correlation (CC) function between two signatures $S_1(x)$ and $S_2(x)$ is defined as 
 \[
@@
 where $x$ and $x+z$ are integer values appropriately defined within the domains of $S_1$ and $S_2$, $z$ is the lag between the two sequences, and $\text{cor}(.,.)$ is the Pearson correlation coefficient.
-When signatures contain missing values, pairwise complete observations are used to calculate the cross corrleation. The lag $z$ used to achieve maximum $CC$ is used to determine the best alignment.
+When signatures contain missing values, pairwise complete observations are used to calculate the cross correlation. The lag $z$ used to achieve maximum $CC$ is used to determine the best alignment.
 
@@
 Using the lag determined for aligning two signatures, other quantitative features describing the similarity of the two signatures can be extracted.
-Numerical features such as cross corrleation and Euclidean distance
+Numerical features such as cross correlation and Euclidean distance
 can be computed from the aligned signatures alone.
@@
 Peaks -- rather than peaks and valleys -- are used for this determination in part because it is difficult to visually assess valleys except as relative to peaks.
-Note that with automatic alignment based on the maximum cross corrleation rather than matching striation marks, it is possible for two lands to have a consecutive matching striae of zero.
-
-<!-- I really think this belongs in this section: it's discussing cross corrleation as a scoring statistic, rather than as a convenient way to align signatures so that scoring stats can be calculated.
+Note that with automatic alignment based on the maximum cross correlation rather than matching striation marks, it is possible for two lands to have a consecutive matching striae of zero.
+
+<!-- I really think this belongs in this section: it's discussing cross correlation as a scoring statistic, rather than as a convenient way to align signatures so that scoring stats can be calculated.
 -->
@@
 Cross correlation values which are close to 1 indicate similarity between signatures and with it may indicate that the two signatures come from the same source.
-Low cross corrleation values indicate signatures which may originate from different sources.
-
-The random forest presented in \citet{aoas} is based on a combination of multiple characteristics, including cross corrleation, number of matching striae, and number of consecutively matching striae.
+Low cross correlation values indicate signatures which may originate from different sources.
+
+The random forest presented in \citet{aoas} is based on a combination of multiple characteristics, including cross correlation, number of matching striae, and number of consecutively matching striae.
 The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where scores close to 1 indicate it to be highly likely that the aligned signatures come from the same source, while scores close to 0 indicate that the aligned signatures are highly likely to originate from different sources.
@@
 
-The features described above are each individually designed to separate same-source and different-source signatures: for instance, same-source signatures would have high cross corrleation values and different-source signatures would have low cross corrleation values.
+The features described above are each individually designed to separate same-source and different-source signatures: for instance, same-source signatures would have high cross correlation values and different-source signatures would have low cross correlation values.
 An automated algorithm selects a cutoff value, introducing a barrier between identification and exclusion.
@@
 We assess the results on each set and evaluate whether the minimal requirements described above are fulfilled for bullet-to-bullet SAM scores.
-We compare the random forest algorithm's performance to the performance of other suggested measures for quantitative assessment of match strength, such as cross corrleation and consecutive matching striae.
+We compare the random forest algorithm's performance to the performance of other suggested measures for quantitative assessment of match strength, such as cross correlation and consecutive matching striae.
 
@@
 With no errors at the bullet-to-bullet level, we can now evaluate the scores in light of (R1) and (R2).
-In this section, we will compare the random forest scores to cross corrleation scores; cross corrleation is one of the components of the random forest, but has also been used to quantify the strength of a match in its own right \citep{chuPilotStudyAutomated2010}.
+In this section, we will compare the random forest scores to cross correlation scores; cross correlation is one of the components of the random forest, but has also been used to quantify the strength of a match in its own right \citep{chuPilotStudyAutomated2010}.
 In addition, both scoring methods use the same scale (0 to 1) and are continuous along that interval; other evaluation methods, such as consecutive matching striae are discrete and more difficult to directly compare to continuous measurements.
@@
 ## Comparison of SAM scores
-\autoref{fig:compare} shows density curves for the scores bullet-to-bullet matches of each of the studies, comparing the SAM scores based on cross corrleation (top) and the random forest score (bottom).
+\autoref{fig:compare} shows density curves for the scores bullet-to-bullet matches of each of the studies, comparing the SAM scores based on cross correlation (top) and the random forest score (bottom).
 Color indicates ground truth.
 Small vertical lines below the $x$-axis indicate observed scores in a **rug-plot**.
-The plot shows that for all three case studies and both continuous measures, i.e.\ SAM scores based on cross corrleation as well as SAM scores based on the random forest scores, all similarity scores are higher for pairs of bullets from the same source, i.e.\ bullets fired through the same barrel, than pairs of bullets from different sources (fired through different barrels).
-
-```{r compare, fig.width = 8, fig.height = 5, fig.cap="Density curves of similarity scores from cross corrleation (top) and RF scores (bottom).
+The plot shows that for all three case studies and both continuous measures, i.e.\ SAM scores based on cross correlation as well as SAM scores based on the random forest scores, all similarity scores are higher for pairs of bullets from the same source, i.e.\ bullets fired through the same barrel, than pairs of bullets from different sources (fired through different barrels).
+
+```{r compare, fig.width = 8, fig.height = 5, fig.cap="Density curves of similarity scores from cross correlation (top) and RF scores (bottom).
 Different colors indicate same source versus different source.
@@
 nestslong$measure <- factor(nestslong$measure)
-levels(nestslong$measure) <- c("Cross corrleation",  "RF score")
+levels(nestslong$measure) <- c("Cross correlation",  "RF score")
 
@@
 
-According to \autoref{fig:compare}, neither cross corrleation nor random forest score fulfill both of the minimal requirements laid out in the introduction.
-Both measures fulfill (R1), i.e.\ all  scores  of  different-source pairs are lower than scores of same-source pairs for SAM scores based on cross corrleation and RF scores.
-However, SAM scores based on the multivariate random forest score show better separation between same-source and different-source scores than SAM scores based on cross corrleation.
-This can be seen from the larger horizontal distance between the modes (peaks) of the density curves of RF based SAM scores compared to the cross corrleation SAM scores.
+According to \autoref{fig:compare}, neither cross correlation nor random forest score fulfill both of the minimal requirements laid out in the introduction.
+Both measures fulfill (R1), i.e.\ all  scores  of  different-source pairs are lower than scores of same-source pairs for SAM scores based on cross correlation and RF scores.
+However, SAM scores based on the multivariate random forest score show better separation between same-source and different-source scores than SAM scores based on cross correlation.
+This can be seen from the larger horizontal distance between the modes (peaks) of the density curves of RF based SAM scores compared to the cross correlation SAM scores.
 
@@
 \autoref{fig:compare-land-to-land} shows density curves for each study.
-There is some overlap between known matching and known non-matching land-to-land scores; in many cases the random forest scores for these overlapping values are more extreme than the corresponding cross corrleation scores.
+There is some overlap between known matching and known non-matching land-to-land scores; in many cases the random forest scores for these overlapping values are more extreme than the corresponding cross correlation scores.
 This suggests that the matching algorithm is sensitive to the presence of LEAs which have low land-to-land match scores, such as the matches of bullets 'I' and 'F' in the Hamby 44 study.
@@
 
-```{r compare-land-to-land, fig.width = 8, fig.height = 4, fig.cap="Density curves of land-to-land similarity scores from cross corrleation (top) and RF scores (bottom).
+```{r compare-land-to-land, fig.width = 8, fig.height = 4, fig.cap="Density curves of land-to-land similarity scores from cross correlation (top) and RF scores (bottom).
 Different colors indicate same source versus different source for each land.
@@
   mutate(measure = factor(measure, levels = c("ccf", "rfscore"),
-                          labels = c("Cross corrleation", "RF score")),
+                          labels = c("Cross correlation", "RF score")),
          study = paste(study, group1) %>% str_trim() %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"))) %>%
@@
 
-```{r roc-auc, echo = FALSE, out.width = c("0.51\\textwidth", ".4725\\textwidth"), out.height = c("0.63\\textwidth", "0.63\\textwidth"), fig.width = c(5, 6), fig.height = c(7.5, 7.5), fig.subcap=c("ROC curves", "Area under the curve (AUC)"), fig.cap="ROC curves and AUC for each (set) of the studies for the random forest score (RF), cross corrleation (CC) and consecutively matching striae (CMS).
+```{r roc-auc, echo = FALSE, out.width = c("0.51\\textwidth", ".4725\\textwidth"), out.height = c("0.63\\textwidth", "0.63\\textwidth"), fig.width = c(5, 6), fig.height = c(7.5, 7.5), fig.subcap=c("ROC curves", "Area under the curve (AUC)"), fig.cap="ROC curves and AUC for each (set) of the studies for the random forest score (RF), cross correlation (CC) and consecutively matching striae (CMS).
 As seen in the ROC curves, the algorithm performs the least well on set 2 of the Houston FSI study.
-Based on AUC, the overall performance on all sets is very good to excellent for cross corrleation and the random forest score.
-At the land level there is no significant difference in prediction power between RF score and cross corrleation.", cache=FALSE}
+Based on AUC, the overall performance on all sets is very good to excellent for cross correlation and the random forest score.
+At the land level there is no significant difference in prediction power between RF score and cross correlation.", cache=FALSE}
   rocpic
@@
 AUC values are useful for differentiating between poor, good, and excellent model performance, but are not particularly useful when determining which of several models with approximately the same level of performance should be used \citep{marzban_roc_2004}.
-Hamby 44 and Phoenix PD are the only studies with a somewhat noticable difference between the AUC score for RF compared to cross corrleation, but even those differences are not statistically significant.
+Hamby 44 and Phoenix PD are the only studies with a somewhat noticeable difference between the AUC score for RF compared to cross correlation, but even those differences are not statistically significant.
 
@@
 Equal Errors are errors when sensitivity and specificity of a test are equal, i.e.\ we see the same percentage of missed and wrong identifications for a land-to-land comparison.
-Equal errors based on CMS of two lands are significantly higher than equal errors based on cross corrleation or the RF score.
-At best, RF score and cross corrleation have an equal error of around 5\% for land-to-land comparisons.
+Equal errors based on CMS of two lands are significantly higher than equal errors based on cross correlation or the RF score.
+At best, RF score and cross correlation have an equal error of around 5\% for land-to-land comparisons.
 Interestingly, for consecutively matching striae, equal errors are reached for just two matching striae, which from a practical point of view is not sustainable.
@@
 
-```{r eer, fig.width=7, fig.height = 4, fig.cap="Equal Error (in Percent)  and corresponding thresholds for all measures (RF score, Cross corrleation and CMS) under all five (sub-)studies.
+```{r eer, fig.width=7, fig.height = 4, fig.cap="Equal Error (in Percent)  and corresponding thresholds for all measures (RF score, Cross correlation and CMS) under all five (sub-)studies.
 CMS has an equal error threshold of 2 for all studies.
-CMS is signifcantly worse than both Cross corrleation and the RF score.
-At best, Cross corrleation and RF score have equal errors for land-to-land comparison of around 5\\%.", out.width = '.8\\textwidth'}
+CMS is signifcantly worse than both Cross correlation and the RF score.
+At best, Cross correlation and RF score have equal errors for land-to-land comparison of around 5\\%.", out.width = '.8\\textwidth'}
 aucs %>% 
@@
 
-At the beginning of this study, we anticipated Hamby 44 would be the easiest set to work with \svp{because of its similarity to the sets used to train the random forest algorithm.
-In a surprise turn of events, it was the}  hardest to solve, in part because of damage to the bullets that obscured LEA striae.
+At the beginning of this study, we anticipated Hamby 44 would be the easiest set to work with because of its similarity to the sets used to train the random forest algorithm.
+In a surprise turn of events, it was the hardest to solve, in part because of damage to the bullets that obscured LEA striae.
 It has been shown \citet{lpr} that parts of lands can be used for successful identifications.
-The algorithm proposed by \citet{aoas} is not capable of  automatically detecting parts of lands with well expressed striae.
+The algorithm proposed by \citet{aoas} is not capable of  automatically detecting parts of lands with well-expressed striae.
 It may be useful to couple the \citet{aoas} algorithm with an algorithm which assesses the quality of the input data and determines which portions of the data to use for comparison.

### 2019-04-29 index.Rmd
fix bullets

--- 2018-11-30 index.Rmd
@@
 
-h44nest <- h44nest %>% mutate(
+h44nest <- h44nest %>% 
+  mutate(bullet1 = as.character(bullet1), bullet2 = as.character(bullet2)) %>%
+  mutate(
   sam_ccf = data %>% purrr::map_dbl(.f = function(d) {

### 2019-04-29 index.Rnw

--- 2018-12-11 index.Rmd
@@
----
-title: Case study validations of automatic bullet matching
-
-# to produce blinded version set to 1
-blinded: 1
-# This gets changed automatically when the title page is compiled in the chunk child-doc-compile
-titlepage: 1
-# Set to 1 if figures should be included with the text
-draft: 1
-
-affiliation: 
-- Department of Statistics, Iowa State University
-- Center for Statistical Applications in Forensic Evidence (CSAFE), Iowa State University
-
-authors: 
-- name: Heike Hofmann
-  degree: Doctor of Philosophy (Ph.D.)
-- name: Susan VanderPlas
-  degree: Doctor of Philosophy (Ph.D.)
-
-funding: This work was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.
-
-thanks: 
-- The authors wish to thank Alan Zheng (NIST), Tylor Klep (Phoenix PD), and Melissa Nally and Kasi Kirksey (FSI Houston) for access to test set bullets. 
-- We would like to thank the efforts of the Roy J Carver High Resolution lab in scanning the bullet sets and providing the scans to us. 
-- Thanks to Prof James Hamby for his advice on XXX.
-
-presentedat: This work was presented in the form of a poster at Forensics\@NIST on November 7, 2018 in Gaithersburg, Maryland
-
-keywords:
-- forensic science
-- toolmark
-- cross correlation
-- random forest
-- 3D microscopy
-- land engraved areas (LEAs)
-
-abstract: Recent advances in microscopy have made it possible to collect 3D topographic data, enabling virtual comparisons based on the collected 3D data next to traditional comparison microscopy.
+\documentclass[review]{elsarticle}
+
+\usepackage{lineno,hyperref}
+\modulolinenumbers[5]
+
+\usepackage{amsmath}
+\usepackage{amsthm}
+\usepackage{graphicx,psfrag,epsf}
+\usepackage{enumerate}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{adjustbox}
+\usepackage[table]{xcolor}
+\usepackage{wrapfig}
+\usepackage{subfig}
+\usepackage{float}
+% \usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{url} % not crucial - just used below for the URL
+\usepackage{ifdraft}
+
+\journal{Forensic Science International}
+%%%%%%%%%%%%%%%%%%%%%%%
+%% Elsevier bibliography styles
+%%%%%%%%%%%%%%%%%%%%%%%
+%% To change the style, put a % in front of the second line of the current style and
+%% remove the % from the second line of the style you would like to use.
+%%%%%%%%%%%%%%%%%%%%%%%
+
+%% Numbered
+%\bibliographystyle{model1-num-names}
+
+%% Numbered without titles
+%\bibliographystyle{model1a-num-names}
+
+%% Harvard
+%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
+
+%% Vancouver numbered
+%\usepackage{numcompress}\bibliographystyle{model3-num-names}
+
+%% Vancouver name/year
+%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
+
+%% APA style
+%\bibliographystyle{model5-names}\biboptions{authoryear}
+
+%% AMA style
+%\usepackage{numcompress}\bibliographystyle{model6-num-names}
+
+%% `Elsevier LaTeX' style
+\bibliographystyle{elsarticle-num}
+%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{document}
+
+\begin{frontmatter}
+
+\title{Case study validations of automatic bullet matching\tnoteref{t1}\tnoteref{t2}}
+
+\tnotetext[t1]{This work was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.}
+
+\tnotetext[t2]{The authors wish to thank Alan Zheng (NIST), Tylor Klep (Phoenix PD), and Melissa Nally and Kasi Kirksey (FSI Houston) for access to test set bullets.
+We would also like to thank the efforts of the Roy J Carver High Resolution lab in scanning the bullet sets and providing the scans to us.
+Finally, thanks to Prof James Hamby for his advice on XXX.}
+
+%% Group authors per affiliation:
+% \author{Elsevier\fnref{myfootnote}}
+% \address{Radarweg 29, Amsterdam}
+% \fntext[myfootnote]{Since 1880.}
+
+%% or include affiliations in footnotes:
+% \author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
+% \ead[url]{www.elsevier.com}
+% 
+% \author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
+% \cortext[mycorrespondingauthor]{Corresponding author}
+% \ead{support@elsevier.com}
+% 
+% \address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
+% \address[mysecondaryaddress]{360 Park Avenue South, New York}
+
+\author[isu,csafe]{Heike Hofmann\corref{corauthor}}
+\cortext[corauthor]{Corresponding author}\ead{hofmann@iastate.edu}
+\author[isu,csafe]{Susan Vanderplas}\ead{srvander@iastate.edu}
+
+\address[iastate]{Statistics Department, Iowa State University\\2438 Osborne Dr, Ames, IA 50011}
+\address[csafe]{Center for Statistical Applications in Forensic Evidence, Iowa State University\\613 Morrill Rd, Ames, IA 50011}
+
+\begin{abstract}
+Recent advances in microscopy have made it possible to collect 3D topographic data, enabling virtual comparisons based on the collected 3D data next to traditional comparison microscopy.
 Automatic matching algorithms have been introduced for various scenarios, such as matching cartridge cases (Tai and Eddy 2018) or matching bullet striae (Hare et al.
@@
 At the land-to-land level, discrimination (based on area under the curve, AUC) is excellent ($ge$ 0.90).
-
-bibliography: bibliography
-biblio-style: jfs
-
-numsecdepth: 1
-
-output: 
-  pdf_document:
-    citation_package: natbib
-    keep_tex: true
-    fig_caption: true
-    latex_engine: pdflatex
-    template: template.tex
----
+\end{abstract}
+
+\begin{keyword}
+forensic science, toolmark, cross correlation, random forest, 3D microscopy, land engraved areas (LEAs)
+\end{keyword}
+
+\end{frontmatter}
 
@@
 
-<!--
-\hh{
-Cleanup TO-DO
-\begin{itemize}
-\item \sout{define false positives and false negatives once and stick to those definitions.
+% \hh{
+% Cleanup TO-DO
+% \begin{itemize}
+% \item \sout{define false positives and false negatives once and stick to those definitions.
 We should avoid false positive and false negatives, because that only makes sense in a test setting, which we have not defined formally.
-Let's use instead wrong identification and missed identification. }
-\item \sout{error rates: we are using rates exchangeably with percentages. We need to fix that. I think percentages might be easier understood. XXX Should be fixed now}
-\end{itemize}
-}
-\svp{
-Outstanding XXX items:
-\begin{itemize}
-\item ``Authors are responsible for obtaining written permission from persons acknowledged by name, because readers may infer their endorsement of the data and conclusions."
-\item thanks - scanning and hamby advice
-\item comparison to FTEs in conclusion
-\end{itemize}
-}
--->
-
-```{r, echo = FALSE, message = FALSE, warning = FALSE}
+% Let's use instead wrong identification and missed identification. }
+% \item \sout{error rates: we are using rates exchangeably with percentages. We need to fix that. I think percentages might be easier understood. XXX Should be fixed now}
+% \end{itemize}
+% }
+% \svp{
+% Outstanding XXX items:
+% \begin{itemize}
+% \item ``Authors are responsible for obtaining written permission from persons acknowledged by name, because readers may infer their endorsement of the data and conclusions."
+% \item thanks - scanning and hamby advice
+% \item comparison to FTEs in conclusion
+% \end{itemize}
+% }
+
+<<echo = FALSE, message = FALSE, warning = FALSE>>=
 library(knitr)
@@
 library(kableExtra)
-```
-
-```{r child-doc-compile, eval=FALSE, include = F, cache = F}
-thisdoc <- readLines("index.Rmd")
-idxs <- which(stringr::str_detect(thisdoc, "^---$"))
-if (length(idxs) >= 2) {
-  # Get yaml metadata, set to nonblinded, write to title-page.Rmd
-  thisdoc[idxs[1]:idxs[2]] %>%
-    str_replace("^blinded: 1", "blinded: 0") %>%
-    str_replace("^titlepage: 0", "titlepage: 1") %>%
-    str_replace("^abstract:.*", "") %>%
-    writeLines(con = "title-page.Rmd")
-}
-
-rmarkdown::render("title-page.Rmd", clean = T)
-file.remove("title-page.tex", "title-page.Rmd")
-```
+@
 
@@
 
-<!-- ## Identification Using 3D Imaging Technology -->
+% Identification using 3D Scanning Technology
 
@@
 
-# Methods
+\section{Methods}
 
@@
 
-## Automated Processing of 3D Scans
+\subsection{Automated Processing of 3D Scans}
+
 From each land engraved area a signature is derived using the process described in detail in \citet{aoas}:
@@
 \autoref{fig:signatures} shows a set of six signatures corresponding to the six land engraved areas of bullets 1 and 2 from barrel 1 of set 44 of the Hamby study \citep{hamby}.
-```{r signatures, fig.cap="Signatures of all six lands of bullets 1 and 2 from barrel 1 of the Hamby set 44.
-Matching lands between the two bullets have been placed above each other.", fig.height=4, fig.width=10}
+<<signatures, fig.cap="Signatures of all six lands of bullets 1 and 2 from barrel 1 of the Hamby set 44.
+Matching lands between the two bullets have been placed above each other.", fig.height=4, fig.width=10>>=
 # signatures for one bullet
@@
   xlab("Relative Location (in millimeters)")
-```
+@
 
@@
 
-## Statistics for Matching Aligned Signatures
+\subsection{Statistics for Matching Aligned Signatures}
 
@@
 
-<!-- I really think this belongs in this section: it's discussing cross correlation as a scoring statistic, rather than as a convenient way to align signatures so that scoring stats can be calculated.
+% <!-- I really think this belongs in this section: it's discussing cross correlation as a scoring statistic, rather than as a convenient way to align signatures so that scoring stats can be calculated.
 -->
@@
 
-## Interpreting the Algorithm Score
+\subsection{Interpreting the Algorithm Score}
 
@@
 
-## Identification Errors and Algorithm Evaluation
+\subsection{Identification Errors and Algorithm Evaluation}
 
@@
 
-## Assessing Bullet-to-Bullet Match Strength
+\subsection{Assessing Bullet-to-Bullet Match Strength}
 
@@
 
-```{r, echo=FALSE, results = 'hide', warning = FALSE, message = FALSE}
+<<echo=FALSE, results = 'hide', warning = FALSE, message = FALSE>>=
 features <- read.csv("data/fsi-features.csv.gz")
@@
   theme(legend.position = "bottom") + coord_equal()
-```
+@
 
@@
 
-```{r b2b, echo = FALSE, fig.width = 5.4, fig.height = 3, out.width = ".6\\textwidth", fig.cap="Overview of land-land matching scores for two pairs of bullet-bullet comparisons.
-On the left the two bullets are known to come from the same source (barrel) on the right, the bullets are from two different sources (barrels)"}
+<<b2b, echo = FALSE, fig.width = 5.4, fig.height = 3, out.width = ".6\\textwidth", fig.cap="Overview of land-land matching scores for two pairs of bullet-bullet comparisons.
+On the left the two bullets are known to come from the same source (barrel) on the right, the bullets are from two different sources (barrels).">>=
 grid.arrange(p1, p2, ncol = 2, padding = 2)
-```
+@
 
@@
 
-```{r sam-sketch, out.width='\\textwidth', fig.width=12, fig.height = 2.5, fig.cap="Sketch of all six land-to-land sequences between two bullets with six lands. "}
+<<sam-sketch, out.width='\\textwidth', fig.width=12, fig.height = 2.5, fig.cap="Sketch of all six land-to-land sequences between two bullets with six lands.">>=
 df2 <- data.frame(expand.grid(x = 1:6, y = 1:6))
@@
   theme(legend.position = "bottom")
-```
+@
 
@@
 
-# Validation Sets
+\section{Validation Sets}
 
@@
 
-# Results
-
-## Hamby Set 44
-
-```{r h44, echo=FALSE, fig.width = 10, fig.height=5, fig.cap="Hamby set 44: overview of all bullet-to-bullet matches between all pairs of questioned bullets ($y$-axis) and known test fires from ten barrels ($x$ axis)."}
+\section{Results}
+
+\subsection{Hamby Set 44}
+
+<<h44, echo=FALSE, fig.width = 10, fig.height=5, fig.cap="Hamby set 44: overview of all bullet-to-bullet matches between all pairs of questioned bullets (y-axis) and known test fires from ten barrels (x axis).">>=
 h44features <- read.csv("data/h44-features.csv")
@@
 h44
-```
+@
 
@@
 
-```{r close, eval=FALSE, fig.width=11.25, fig.height=4.2, fig.cap="Examples of land-to-land comparisons from the Hamby-44 set.
-Along the top are the three bullets from different source with the highest RF scores, along the bottom are the three bullets from same source with the lowest RF score.", fig.subcap=c("Ground truth: different source", "Ground truth: same-source"), fig.align="middle", fig.ncol=1}
+<<close, eval=FALSE, fig.width=11.25, fig.height=4.2, fig.cap="Examples of land-to-land comparisons from the Hamby-44 set.
+Along the top are the three bullets from different source with the highest RF scores, along the bottom are the three bullets from same source with the lowest RF score.", fig.subcap=c("Ground truth: different source", "Ground truth: same-source"), fig.align="middle", fig.ncol=1>>=
 # ex2 <- h44nest %>% filter(samesource == FALSE & sam_rf >= 0.287)
@@
 It is obvious that not all of the scores are as high as we would usually see in same-source comparisons.
-```
-
-## Phoenix PD
-
-```{r pd, echo=FALSE, fig.width = 9, fig.height=4, fig.cap="Phoenix study: overview of all bullet-to-bullet matches between all pairs of questioned bullets ($y$-axis) and known test fires from ten barrels ($x$ axis).
-The order of the bullets on the $y$ axis is determined by matching barrel."}
+@
+
+\subsection{Phoenix PD}
+
+<<pd, echo=FALSE, fig.width = 9, fig.height=4, fig.cap="Phoenix study: overview of all bullet-to-bullet matches between all pairs of questioned bullets (y-axis) and known test fires from ten barrels (x axis).
+The order of the bullets on the y axis is determined by matching barrel.">>=
 features <- read.csv("data/pd-features.csv.gz")
@@
 p3 + ggtitle("Phoenix PD set")
-```
+@
 
@@
 
-## Houston FSI
+\subsection{Houston FSI}
 
@@
 In set 3, six of the questioned bullets match known bullets, and the remaining two questioned bullets match each other (bullets U14 and U45) but do not match any known bullets in the set.
-<!--Thus, we can conclude that the algorithm performs well on open set studies, as it can identify matches among questioned bullets which do not match any known bullets. -->
+% <!--Thus, we can conclude that the algorithm performs well on open set studies, as it can identify matches among questioned bullets which do not match any known bullets. -->
 While all bullet-to-bullet matches are correctly identified, and no known matches are missed, it is obvious from the generally lighter shades of the tiles corresponding to matching bullets in \autoref{fig:hou}(b) that the algorithm is not performing as well on set 2 as it performs on sets 1 and 3.
-```{r hou, fig.width=9, fig.height = 3, fig.cap="Overview of matching scores for all pairs of questioned bullets to known bullets from five barrels (on the left) and questioned bullets to themselves (tiles on the right).", fig.subcap=c("Set 1", "Set 2", "Set 3"), fig.align="middle", fig.ncol=1}
+
+<<hou, fig.width=9, fig.height = 3, fig.cap="Overview of matching scores for all pairs of questioned bullets to known bullets from five barrels (on the left) and questioned bullets to themselves (tiles on the right).", fig.subcap=c("Set 1", "Set 2", "Set 3"), fig.align="middle", fig.ncol=1>>=
 features <- read.csv("data/fsi-features.csv.gz")
@@
 hou3
-```
-
-# Evaluating the Random Forest Algorithm
+@
+
+\section{Evaluating the Random Forest Algorithm}
 The random forest SAM scores correctly separate known matches from known non-matches, using SAM scores to aggregate land-to-land scores into a bullet-to-bullet comparison.
@@
 
-## Comparison of SAM scores
+\subsection{Comparison of SAM scores}
 \autoref{fig:compare} shows density curves for the scores bullet-to-bullet matches of each of the studies, comparing the SAM scores based on cross correlation (top) and the random forest score (bottom).
@@
 
-```{r compare, fig.width = 8, fig.height = 5, fig.cap="Density curves of similarity scores from cross correlation (top) and RF scores (bottom).
+<<compare, fig.width = 8, fig.height = 5, fig.cap="Density curves of similarity scores from cross correlation (top) and RF scores (bottom).
 Different colors indicate same source versus different source.
-Ideally all scores of different source pairs should be much lower than scores for same source pairs."}
+Ideally all scores of different source pairs should be much lower than scores for same source pairs.">>=
 nests <- rbind(
@@
   )
-```
+@
 
@@
 
-## Digging deeper: comparison of land-to-land scores
+\subsection{Digging deeper: comparison of land-to-land scores}
 
@@
 These issues affect the algorithm's ability to identify a stable signature, which impacts the subsequent match scores  at the land-to-land and bullet-to-bullet level.
-```{r tankrash, fig.cap="Overview of bullet lands with prominent deficiency such as tank rash or extreme pitting in the Hamby-44 study.", fig.width=4, fig.height = 2, out.width='47.5%', fig.subcap=c("Br1-B1-L6 (tank rask)", "Br2-B2-L5 (tank rash)", "Br3-B1-L5 (tank rash)", "Br8-B1-L6 (tank rash)", "Br8-B2-L2 (pitting)", "Br8-B2-L6 (tank rash)", "Unk-E-L6 (tank rash)", "Unk-I-L6 (tank rash and break-off)"), fig.align="middle", fig.ncol=2}
+<<tankrash, fig.cap="Overview of bullet lands with prominent deficiency such as tank rash or extreme pitting in the Hamby-44 study.", fig.width=4, fig.height = 2, out.width='47.5%', fig.subcap=c("Br1-B1-L6 (tank rask)", "Br2-B2-L5 (tank rash)", "Br3-B1-L5 (tank rash)", "Br8-B1-L6 (tank rash)", "Br8-B2-L2 (pitting)", "Br8-B2-L6 (tank rash)", "Unk-E-L6 (tank rash)", "Unk-I-L6 (tank rash and break-off)"), fig.align="middle", fig.ncol=2>>=
 knitr::include_graphics(
@@
 )
-```
+@
 
@@
 
-```{r, cache=FALSE}
+<<cache=FALSE>>=
 hounest <- hounest %>% mutate(
@@
   )
-```
-
-```{r check44, eval=FALSE}
+@
+
+<<check44, eval=FALSE>>=
 h44lands <- h44lands %>%
@@
 
-```
-
-```{r checkpd, eval=FALSE}
+@
+
+<<checkpd, eval=FALSE>>=
 pdlands <- pdlands %>%
@@
 
-```
-
-```{r checkhou, eval=FALSE}
+@
+
+<<checkhou, eval=FALSE>>=
 houlands <- houlands %>%
@@
   facet_grid(bullet1~barrel2+bullet2)  
-```
-
-```{r compare-land-to-land, fig.width = 8, fig.height = 4, fig.cap="Density curves of land-to-land similarity scores from cross correlation (top) and RF scores (bottom).
+@
+
+<<compare-land-to-land, fig.width = 8, fig.height = 4, fig.cap="Density curves of land-to-land similarity scores from cross correlation (top) and RF scores (bottom).
 Different colors indicate same source versus different source for each land.
 RF scores for different source comparisons are generally well below 0.5.
-Some same-source comparisons are also below 0.5 for RF scores, indicating potential problems with at least one of the lands involved in the comparison."}
+Some same-source comparisons are also below 0.5 for RF scores, indicating potential problems with at least one of the lands involved in the comparison.">>=
 all_lands %>% 
@@
   geom_hline(yintercept = 0, show.legend = F)
-```
+@
 
@@
 
-```{r rocs}
+<<rocs>>=
 library(plotROC)
@@
 
-```
-
-```{r aucs-pic}
+@
+
+<<aucs-pic>>=
 eer <- function(roc){
@@
 
-```
-
-```{r roc-auc, echo = FALSE, out.width = c("0.51\\textwidth", ".4725\\textwidth"), out.height = c("0.63\\textwidth", "0.63\\textwidth"), fig.width = c(5, 6), fig.height = c(7.5, 7.5), fig.subcap=c("ROC curves", "Area under the curve (AUC)"), fig.cap="ROC curves and AUC for each (set) of the studies for the random forest score (RF), cross correlation (CC) and consecutively matching striae (CMS).
+@
+
+<<roc-auc, echo = FALSE, out.width = c("0.51\\textwidth", ".4725\\textwidth"), out.height = c("0.63\\textwidth", "0.63\\textwidth"), fig.width = c(5, 6), fig.height = c(7.5, 7.5), fig.subcap=c("ROC curves", "Area under the curve (AUC)"), fig.cap="ROC curves and AUC for each (set) of the studies for the random forest score (RF), cross correlation (CC) and consecutively matching striae (CMS).
 As seen in the ROC curves, the algorithm performs the least well on set 2 of the Houston FSI study.
 Based on AUC, the overall performance on all sets is very good to excellent for cross correlation and the random forest score.
-At the land level there is no significant difference in prediction power between RF score and cross correlation.", cache=FALSE}
+At the land level there is no significant difference in prediction power between RF score and cross correlation.", cache=FALSE>>=
   rocpic
   aucpic
-```
+@
 
@@
 
-```{r eval=FALSE}
+<<eval=FALSE>>=
 xtabs(~sameland+I(round(cms)>=6), data = all_lands)
-```
-
-```{r eer, fig.width=7, fig.height = 4, fig.cap="Equal Error (in Percent)  and corresponding thresholds for all measures (RF score, Cross correlation and CMS) under all five (sub-)studies.
+@
+
+<<eer, fig.width=7, fig.height = 4, fig.cap="Equal Error (in Percent)  and corresponding thresholds for all measures (RF score, Cross correlation and CMS) under all five (sub-)studies.
 CMS has an equal error threshold of 2 for all studies.
 CMS is signifcantly worse than both Cross correlation and the RF score.
-At best, Cross correlation and RF score have equal errors for land-to-land comparison of around 5\\%.", out.width = '.8\\textwidth'}
+At best, Cross correlation and RF score have equal errors for land-to-land comparison of around 5\\%.", out.width = '.8\\textwidth'>>=
 aucs %>% 
@@
   xlim(c(0, NA))
-```
-
-# Discussion and Conclusions
+@
+
+\section{Discussion and Conclusions}
 
@@
 
-```{r output-all-figs, include = F}
+<<output-all-figs, include = F>>=
 thisdoc <- readLines("index.Rmd")
@@
 
-```
-# References
-
-<!--include marginal distributions of training data and compare to test data-->
-
-```{r, echo=FALSE, eval=FALSE}
+@
+\section{References}
+
+\bibliography{bibliography}
+
+% <!--include marginal distributions of training data and compare to test data-->
+
+<< echo=FALSE, eval=FALSE>>=
 training <- read.csv("data/features-hamby.csv")
@@
 # now compare to test data
-```
+@

### 2019-04-29 index.Rnw
Switch to FSI latex template from Rmarkdown

@@
 
-\usepackage{amsmath}
-\usepackage{amsthm}
+\usepackage{amsmath,amsthm}
 \usepackage{graphicx,psfrag,epsf}
@@
 \usepackage{threeparttable}
-\usepackage{url} % not crucial - just used below for the URL
-\usepackage{ifdraft}
+
+\usepackage[numbers]{natbib}
+
+\newenvironment{definition}[1]% environment name 
+{% begin code 
+  \par\vspace{.75\baselineskip}\noindent 
+  \textbf{Definition (#1)}\begin{itshape}% 
+  \par\vspace{.5\baselineskip}\noindent\ignorespaces 
+}% 
+{% end code 
+  \end{itshape}\ignorespacesafterend 
+}
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
@@
 %% Harvard
-%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
+% \bibliographystyle{model2-names.bst}\biboptions{authoryear}
 
@@
 %% APA style
-%\bibliographystyle{model5-names}\biboptions{authoryear}
+% \bibliographystyle{model5-names}\biboptions{authoryear}
 
@@
 %% `Elsevier LaTeX' style
-\bibliographystyle{elsarticle-num}
+\bibliographystyle{elsarticle-num-names}
 %%%%%%%%%%%%%%%%%%%%%%%
@@
 \cortext[corauthor]{Corresponding author}\ead{hofmann@iastate.edu}
-\author[isu,csafe]{Susan Vanderplas}\ead{srvander@iastate.edu}
-
-\address[iastate]{Statistics Department, Iowa State University\\2438 Osborne Dr, Ames, IA 50011}
+\author[isu,csafe]{Susan Vanderplas}
+\address[isu]{Statistics Department, Iowa State University\\2438 Osborne Dr, Ames, IA 50011}
 \address[csafe]{Center for Statistical Applications in Forensic Evidence, Iowa State University\\613 Morrill Rd, Ames, IA 50011}
@@
 We see some variations in scores across different models of Rugers, and it is probable that we will see even bigger variations with different brands of firearms that vary in number of lands and land length.
-We also know that different firearms and ammunition combinations mark differently well \citep{bolton-king_preventing_2016, bonfanti_influence_1999}.
+We also know that different firearms and ammunition combinations mark differently well \citep{boltonkingPreventingMiscarriagesJustice2016, bonfantiInfluenc1999}.
 Rugers are among the firearms that mark very well, so we would correspondingly expect to see lowered similarity scores for other firearms.
@@
 
-\bibliography{bibliography}
+\bibliography{bibfile}
+
+\end{document}
 

### 2019-04-29 index.Rnw
fix typos and image sizing issues with new template

@@
 Cross correlation and Random Forest scores both result in perfect discrimination of same-source and different-source bullets.
-At the land-to-land level, discrimination (based on area under the curve, AUC) is excellent ($ge$ 0.90).
+At the land-to-land level, discrimination (based on area under the curve, AUC) is excellent ($\ge$ 0.90).
 \end{abstract}
@@
 
-<<sam-sketch, out.width='\\textwidth', fig.width=12, fig.height = 2.5, fig.cap="Sketch of all six land-to-land sequences between two bullets with six lands.">>=
+<<sam-sketch, out.width='\\textwidth', fig.width=8, fig.height = 3, fig.cap="Sketch of all six land-to-land sequences between two bullets with six lands.">>=
 df2 <- data.frame(expand.grid(x = 1:6, y = 1:6))
@@
 
-<<h44, echo=FALSE, fig.width = 10, fig.height=5, fig.cap="Hamby set 44: overview of all bullet-to-bullet matches between all pairs of questioned bullets (y-axis) and known test fires from ten barrels (x axis).">>=
+<<h44, echo=FALSE, fig.width = 8, fig.height=4, fig.cap="Hamby set 44: overview of all bullet-to-bullet matches between all pairs of questioned bullets (y-axis) and known test fires from ten barrels (x axis).">>=
 h44features <- read.csv("data/h44-features.csv")
@@
 
-<<pd, echo=FALSE, fig.width = 9, fig.height=4, fig.cap="Phoenix study: overview of all bullet-to-bullet matches between all pairs of questioned bullets (y-axis) and known test fires from ten barrels (x axis).
+<<pd, echo=FALSE, fig.width = 8, fig.height=3.56, fig.cap="Phoenix study: overview of all bullet-to-bullet matches between all pairs of questioned bullets (y-axis) and known test fires from ten barrels (x axis).
 The order of the bullets on the y axis is determined by matching barrel.">>=
@@
 
-<<hou, fig.width=9, fig.height = 3, fig.cap="Overview of matching scores for all pairs of questioned bullets to known bullets from five barrels (on the left) and questioned bullets to themselves (tiles on the right).", fig.subcap=c("Set 1", "Set 2", "Set 3"), fig.align="middle", fig.ncol=1>>=
+<<hou, fig.width=8, fig.height = 2.67, fig.cap="Overview of matching scores for all pairs of questioned bullets to known bullets from five barrels (on the left) and questioned bullets to themselves (tiles on the right).", fig.subcap=c("Set 1", "Set 2", "Set 3"), fig.align="middle", fig.ncol=1>>=
 features <- read.csv("data/fsi-features.csv.gz")

### 2019-08-06 index.Rnw
Tylor's changes

@@
 \item {\bf Phoenix PD}
-Tylor Klep from Phoenix PD provided sets of known test fires and questioned bullets: the set of known bullets consists of three test fires (B1, B2, B3) from each of eight different, consecutively manufactured Ruger P-95 barrels (A9, C8, F6, L5, M2, P7, R3, U10).
+Tylor Klep from Phoenix PD provided sets of known test fires and questioned bullets: the set of known bullets consists of three test fires (B1, B2, B3) from each of eight different, consecutively rifled Ruger P-95 barrels (A9, C8, F6, L5, M2, P7, R3, U10).
 Ten questioned bullets were provided (B, E, H, J, K N, Q, T, Y, Z).
@@
 In fact, the results will show that three of the questioned bullets were fired from three different barrels not included in the knowns.
-These three barrels correspond to an eleventh Ruger P-95, a Ruger P-95C and a Ruger P-85.
-All bullets fired for this study are American Eagle 9mm  copper jacketed Luger.
+These three barrels correspond to an eleventh Ruger P-95 barrel, a Ruger P-95C barrel and a Ruger P-85 barrel.
+All bullets fired for this study are American Eagle 9mm  Luger full metal jackets.
 Land engraved areas for each of the six lands of each bullet were scanned by Bill Henderson (Sensofar).

### 2019-08-07 acknowledgements.tex
submission files

\documentclass{article}
\usepackage{fullpage}
\author{Acknowledgements}
\title{Case study validations of automatic bullet matching}
\date{}
\begin{document}
\maketitle

This work was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.\vspace{1cm}

The authors wish to thank Alan Zheng (NIST), Tylor Klep (Phoenix PD), and Melissa Nally and Kasi Kirksey (FSI Houston) for access to test set bullets. We would also like to thank the efforts of the Roy J Carver High Resolution lab in scanning the bullet sets and providing the scans to us. 
\end{document}

### 2019-08-07 cover_letter.tex
submission files

\documentclass[11pt]{article}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{float}
% \setlength{\textheight}{26cm}
% \setlength{\textwidth}{18cm}
\topmargin0in

\oddsidemargin-1cm
\evensidemargin0in
\textheight9in
\textwidth7in
\setlength{\parindent}{0em}

\pagestyle{fancyplain}
\addtolength{\headheight}{\baselineskip}
% \addtolength{\headheight}{43pt}
\rhead{}
\cfoot{}
\markright{{\hfill\large \sc Center for Statistics and Applications in Forensic Evidence}\\{\hfill\sc Iowa State University} \\{\hfill 195 Durham Center, 613 Morill Rd., Ames, IA 50011}}

\begin{document}
C. Jackowski \\
Institute of Forensic Medicine\vspace{.5em}\\
C. Cattaneo \\
Università degli Studi di Milano \vspace{2em}

\today\\\vspace{1em}

{\bf Re: Submission to Forensic Science International}\\\vspace{2em}

Dear Professor Jackowski and Professor Cattaneo, \vspace{1em}

Please find attached a manuscript, ``Case study validations of automatic bullet matching" intended for submission to FSI. \vspace{1em}

The manuscript describes a validation of an automatic bullet matching algorithm on several sets of bullets used to test examiner proficiency. This algorithm uses a scheme similar to that used by footwear forensic analysts, unlike other convolutional neural network based analyses in footwear forensics. We describe the model's structure, capabilities, and overall performance; we also explore situations where the model's automatic recognition process does not succeed and provide some possible solutions for these issues going forward. We believe this work lays the groundwork to move beyond theoretical applications of machine learning in toolmark forensics toward implementation of these algorithms in the crime lab. \vspace{1em}

I hope that you find the manuscript acceptable for publication in FSI. If there is any additional information I can provide, please feel free to contact us directly. \vspace{0.25in}

Sincerely,\\
\begin{minipage}{.25\linewidth}
\vspace*{.5cm}
\includegraphics[width=2.5in,keepaspectratio=TRUE]{signature}
\vspace*{.5cm}
\end{minipage}

Susan Vanderplas\\
Center for Statistics and Applications in Forensic Evidence\\
Iowa State University

\end{document}

### 2019-08-07 index-blinded.Rnw
submission files

--- 2019-08-06 index.Rnw
@@
-\documentclass[review]{elsarticle}
+\documentclass[doubleblind]{elsarticle}
 
@@
 
-\title{Case study validations of automatic bullet matching\tnoteref{t1}\tnoteref{t2}}
-
-\tnotetext[t1]{This work was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.}
-
-\tnotetext[t2]{The authors wish to thank Alan Zheng (NIST), Tylor Klep (Phoenix PD), and Melissa Nally and Kasi Kirksey (FSI Houston) for access to test set bullets.
+\title{Case study validations of automatic bullet matching}%\tnoteref{t1}\tnoteref{t2}}
+
+% \tnotetext[t1]{This work was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.}
+% 
+% \tnotetext[t2]{The authors wish to thank Alan Zheng (NIST), Tylor Klep (Phoenix PD), and Melissa Nally and Kasi Kirksey (FSI Houston) for access to test set bullets.
 We would also like to thank the efforts of the Roy J Carver High Resolution lab in scanning the bullet sets and providing the scans to us.
@@
 
-\author[isu,csafe]{Heike Hofmann\corref{corauthor}}
-\cortext[corauthor]{Corresponding author}\ead{hofmann@iastate.edu}
-\author[isu,csafe]{Susan Vanderplas}
+\author[isu,csafe]{Heike Hofmann}
+\cortext[corauthor]{Corresponding author}\ead{srvander@iastate.edu}
+\author[isu,csafe]{Susan Vanderplas\corref{corauthor}}
 \address[isu]{Statistics Department, Iowa State University\\2438 Osborne Dr, Ames, IA 50011}
@@
 Recent advances in microscopy have made it possible to collect 3D topographic data, enabling virtual comparisons based on the collected 3D data next to traditional comparison microscopy.
-Automatic matching algorithms have been introduced for various scenarios, such as matching cartridge cases (Tai and Eddy 2018) or matching bullet striae (Hare et al.
-2017b,  Chu et al 2013,  De Kinder and Bonfanti 1999).
+Automatic matching algorithms have been introduced for various scenarios, such as matching cartridge cases \citep{tai2018fully} or matching bullet striae \citep{aoas, chuAutomaticIdentificationBullet2013, dekinderAutomatedComparisonsBullet1999}.
 One key aspect of validating automatic matching algorithms is to evaluate the performance of the algorithm on external tests.
-Here, we are presenting a discussion of the performance of the matching algorithm (Hare et al.
-2017b) in three studies.
+Here, we are presenting a discussion of the performance of the matching algorithm \citep{aoas} in three studies.
 We are considering matching performance based on the Random forest score, cross correlation, and consecutive matching striae (CMS) at the land-to-land level and, using Sequential Average Maxima scores, also at the bullet-to bullet level.
@@
 Exact guidelines for this classification vary from lab to lab; some labs will exclude only on the basis of non-matching class characteristics, such as direction of the twist in rifling, land length or number of lands, or type of rifling.
-In other labs, CMS (consecutively matching striae) as defined by \citet{biasotti} is used as a measure to quantify the strength of a match.
+In other labs, CMS (consecutively matching striae) as defined by \citeauthor{biasotti} \citep{biasotti} is used as a measure to quantify the strength of a match.
 In virtually all labs, individual characteristics used to identify matching bullets are derived from visual assessment; some class characteristics may be directly measured, but these are not sufficient for individualization.
@@
 Scans of all land engraved areas for the validation data were taken on a Sensofar Confocal Light microscope at 20x magnification resulting in a resolution of 0.645 microns per pixel.
-If not indicated otherwise, scans were taken at the Roy J Carver high resolution microscopy lab at Iowa State University.
+If not indicated otherwise, scans were taken at \texttt{Redacted}.
+% the Roy J Carver high resolution microscopy lab at Iowa State University.
 
@@
 For a future round of the Houston study, we are planning to compare the algorithmic performance directly with scores given by forensic toolmark examiners.
-\hh{XXX need to ask XXX  } The algorithm's score is not intended to replace an examiner's judgment; instead, it provides a complementary tool to measure and quantify the strength of an identification.
+The algorithm's score is not intended to replace an examiner's judgment; instead, it provides a complementary tool to measure and quantify the strength of an identification.
 By validating the algorithm on external test sets, we have demonstrated that the method can be generalized to different types of ammunition and is not overly sensitive to small differences in rifling procedure.

### 2019-08-07 title_page.tex
submission files

\documentclass[titlepage]{article}
\title{Case study validations of automatic bullet matching}
\author{Susan Vanderplas and Heike Hofmann}
\date{Center for Statistics and Applications in Forensic Evidence, Iowa State University. 195 Durham Center, 613 Morill Rd., Ames, IA 50011}

\begin{document}
\maketitle

\end{document}

### 2019-11-10 index.Rnw
Abstract changes

@@
 \tnotetext[t2]{The authors wish to thank Alan Zheng (NIST), Tylor Klep (Phoenix PD), and Melissa Nally and Kasi Kirksey (FSI Houston) for access to test set bullets.
-We would also like to thank the efforts of the Roy J Carver High Resolution lab in scanning the bullet sets and providing the scans to us.
-Finally, thanks to Prof James Hamby for his advice on XXX.}
+We would also like to thank the efforts of the Roy J Carver High Resolution lab in scanning the bullet sets and providing the scans to us.}
 
@@
 \begin{abstract}
-Recent advances in microscopy have made it possible to collect 3D topographic data, enabling virtual comparisons based on the collected 3D data next to traditional comparison microscopy.
-Automatic matching algorithms have been introduced for various scenarios, such as matching cartridge cases (Tai and Eddy 2018) or matching bullet striae (Hare et al.
-2017b,  Chu et al 2013,  De Kinder and Bonfanti 1999).
-One key aspect of validating automatic matching algorithms is to evaluate the performance of the algorithm on external tests.
-Here, we are presenting a discussion of the performance of the matching algorithm (Hare et al.
-2017b) in three studies.
-We are considering matching performance based on the Random forest score, cross correlation, and consecutive matching striae (CMS) at the land-to-land level and, using Sequential Average Maxima scores, also at the bullet-to bullet level.
-Cross correlation and Random Forest scores both result in perfect discrimination of same-source and different-source bullets.
-At the land-to-land level, discrimination (based on area under the curve, AUC) is excellent ($\ge$ 0.90).
+Recent advances in microscopy have made it possible to collect 3D topographic data, enabling virtual comparisons based on the collected 3D data as a supplement to traditional comparison microscopy and 2D photography.
+Automatic matching algorithms have been introduced for various scenarios, such as matching cartridge cases \citep{tai2018fully,song3DTopographyMeasurements2014} or matching bullet striae \citep{aoas2, chuAutomaticIdentificationBullet2013, dekinderAutomatedComparisonsBullet1999}.
+One key aspect of validating automatic matching algorithms is to evaluate the performance of the algorithm on external tests, that is, using data which were not used to train the algorithm.
+Here, we present a discussion of the performance of the matching algorithm \citep{aoas2} in three studies conducted using different Ruger weapons.
+We consider the performance of three scoring measures: random forest score, cross correlation, and consecutive matching striae (CMS) at the land-to-land level and, using Sequential Average Maxima scores, also at the bullet-to bullet level.
+Cross correlation and random forest scores both result in perfect discrimination of same-source and different-source bullets.
+At the land-to-land level, discrimination for both cross correlation and random forest scores (based on area under the curve, AUC) is excellent ($\ge$ 0.90).
 \end{abstract}
@@
 
-In the matching algorithm proposed in \citet{aoas}, several features are combined using a random forest \citep{breiman} to produce a similarity score based on 3D topographic scans of land engraved areas (LEAs).
+In the matching algorithm proposed in \citet{aoas2}, several features are combined using a random forest \citep{breiman} to produce a similarity score based on 3D topographic scans of land engraved areas (LEAs).
 In order to generate these features, some pre-processing is necessary in order to transform the 3D surface measurements into `signatures' which can be compared.
@@
 
-From each land engraved area a signature is derived using the process described in detail in \citet{aoas}:
+From each land engraved area a signature is derived using the process described in detail in \citet{aoas2}:
 
@@
 Once signatures have been extracted from stable regions with expressed striae, they must be aligned in order to assess their match strength.
-Maximized cross correlation is used to pair-wise align signatures \citep{aoas,vorburgerApplicationsCrosscorrelationFunctions2011}.
+Maximized cross correlation is used to pair-wise align signatures \citep{aoas2,vorburgerApplicationsCrosscorrelationFunctions2011}.
 
@@
 
-The random forest presented in \citet{aoas} is based on a combination of multiple characteristics, including cross correlation, number of matching striae, and number of consecutively matching striae.
+The random forest presented in \citet{aoas2} is based on a combination of multiple characteristics, including cross correlation, number of matching striae, and number of consecutively matching striae.
 The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where scores close to 1 indicate it to be highly likely that the aligned signatures come from the same source, while scores close to 0 indicate that the aligned signatures are highly likely to originate from different sources.
@@
 
-When evaluating the \citet{aoas} algorithm, we will describe errors as a percentage of correct evaluations.
+When evaluating the \citet{aoas2} algorithm, we will describe errors as a percentage of correct evaluations.
 Statistics that do not translate to percent-form interpretation, such as area under the curve (AUC), will be presented as a decimal.
@@
 
-In this paper, we validate the algorithm described in \citet{aoas} on three external test sets.
+In this paper, we validate the algorithm described in \citet{aoas2} on three external test sets.
 We assess the results on each set and evaluate whether the minimal requirements described above are fulfilled for bullet-to-bullet SAM scores.
@@
 
-The algorithm in \citet{aoas} is trained on scans from Hamby sets 252 and 173 made available through \citet{nist}.
+The algorithm in \citet{aoas2} is trained on scans from Hamby sets 252 and 173 made available through \citet{nist}.
 Set 173 was originally published as Hamby set 44.
@@
 Case studies were chosen such that difficulty for the matching algorithm increases with each case study: Hamby set 44 is part of the Hamby study.
-The algorithm in \citet{aoas} was trained on sets 173 and 252, so the bullets in Hamby 44 are of the same type of ammunition and are fired through the same barrels as the bullets in the training set.
+The algorithm in \citet{aoas2} was trained on sets 173 and 252, so the bullets in Hamby 44 are of the same type of ammunition and are fired through the same barrels as the bullets in the training set.
 The Phoenix PD set uses Ruger P-95 barrels, which are different from the barrels in the Hamby sets used for training the algorithm, but the barrels are rifled similarly to the Ruger P-85 barrels.
@@
 This should make the automatic matching harder, and in particular, is expected to complicate the classification as different-source land engraved areas/bullets.
-In all three studies, the scan resolution is much higher than the scans used to train the algorithm in \citet{aoas}; this difference also provides a test of the algorithm's ability to generalize to scans taken at different resolutions.
+In all three studies, the scan resolution is much higher than the scans used to train the algorithm in \citet{aoas2}; this difference also provides a test of the algorithm's ability to generalize to scans taken at different resolutions.
 
@@
 In a surprise turn of events, it was the hardest to solve, in part because of damage to the bullets that obscured LEA striae.
-It has been shown \citet{lpr} that parts of lands can be used for successful identifications.
-The algorithm proposed by \citet{aoas} is not capable of  automatically detecting parts of lands with well-expressed striae.
-It may be useful to couple the \citet{aoas} algorithm with an algorithm which assesses the quality of the input data and determines which portions of the data to use for comparison.
+It has been shown \citet{lpr2} that parts of lands can be used for successful identifications.
+The algorithm proposed by \citet{aoas2} is not capable of  automatically detecting parts of lands with well-expressed striae.
+It may be useful to couple the \citet{aoas2} algorithm with an algorithm which assesses the quality of the input data and determines which portions of the data to use for comparison.
 This would emulate the process used by examiners, who first assess the quality of the evidence and whether there is enough information present to attempt a match.
@@
 
-The random forest matching algorithm presented in \citet{aoas} does perform well on three different external test sets.
+The random forest matching algorithm presented in \citet{aoas2} does perform well on three different external test sets.
 While a single cutoff value cannot be used to distinguish matches and non-matches across the different test sets, the algorithm makes no errors when a set-specific cutoff value is used on bullet-to-bullet scores aggregated using sequence average maximum.

### 2019-11-10 index.Rnw
Fixing so that it'll compile?

@@
 )
-pdlands <- pdnest %>% unnest(data)
+pdlands <- pdnest %>%
+  rename(bullet_KM = KM) %>% 
+  unnest(data)
 

### 2019-11-10 index.Rnw
match strength -> similarity

@@
 Recent advances in microscopy have made it possible to collect 3D topographic data, enabling virtual comparisons based on the collected 3D data as a supplement to traditional comparison microscopy and 2D photography.
-Automatic matching algorithms have been introduced for various scenarios, such as matching cartridge cases \citep{tai2018fully,song3DTopographyMeasurements2014} or matching bullet striae \citep{aoas2, chuAutomaticIdentificationBullet2013, dekinderAutomatedComparisonsBullet1999}.
-One key aspect of validating automatic matching algorithms is to evaluate the performance of the algorithm on external tests, that is, using data which were not used to train the algorithm.
+Automatic comparison algorithms have been introduced for various scenarios, such as matching cartridge cases \citep{tai2018fully,song3DTopographyMeasurements2014} or matching bullet striae \citep{aoas2, chuAutomaticIdentificationBullet2013, dekinderAutomatedComparisonsBullet1999}.
+One key aspect of validating these automatic comparison algorithms is to evaluate the performance of the algorithm on external tests, that is, using data which were not used to train the algorithm.
 Here, we present a discussion of the performance of the matching algorithm \citep{aoas2} in three studies conducted using different Ruger weapons.
@@
 
-Once signatures have been extracted from stable regions with expressed striae, they must be aligned in order to assess their match strength.
+Once signatures have been extracted from stable regions with expressed striae, they must be aligned in order to assess their similarity, just as examiners would manually align two bullets under a comparison microscope.
 Maximized cross correlation is used to pair-wise align signatures \citep{aoas2,vorburgerApplicationsCrosscorrelationFunctions2011}.
@@
 can be computed from the aligned signatures alone.
-Other features include characteristics of the signature that firearms and toolmark examiners base their visual assessment on, such as striae depth, number of matching striae, and the number of consecutively matching striae (CMS), one of the measures firearms examiners use to describe the strength of a match quantitatively.
+Other features include characteristics of the signature that firearms and toolmark examiners base their visual assessment on, such as striae depth, number of matching striae, and the number of consecutively matching striae (CMS), one of the measures firearms examiners use to describe the similarity of two bullets quantitatively.
 
@@
 Features which firearms examiners assess visually depend on the identification of these extrema.
-If there are at least six consecutive matching striae, the bullets are considered to be a good match in practice \citep{biasotti_firearms_1997,nichols_scientific_2006}.
+If there are at least six consecutive matching striae, the bullets are considered to be similar in practice \citep{biasotti_firearms_1997,nichols_scientific_2006}.
 CMS in this setting are counted as the number of consecutively matching peaks in signatures of two aligned lands.
 Peaks -- rather than peaks and valleys -- are used for this determination in part because it is difficult to visually assess valleys except as relative to peaks.
-Note that with automatic alignment based on the maximum cross correlation rather than matching striation marks, it is possible for two lands to have a consecutive matching striae of zero.
+Utilizing automatic alignment based on maximum cross correlation (as opposed to matching striation marks) allows the possibility that two lands will have zero consecutive matching striae; this would not occur with manual alignment based on striation marks.
 
@@
 
-Cross correlation can also be used to assess match strength between two aligned signatures.
+Cross correlation can also be used to assess similarity between two aligned signatures.
 Cross correlation varies between -1 and 1, but as the alignment is based on the maximized cross correlation value, in practice the cross correlations for aligned signatures are generally positive.
-Cross correlation values which are close to 1 indicate similarity between signatures and with it may indicate that the two signatures come from the same source.
+Cross correlation values which are close to 1 indicate similarity between signatures, and may consequentially indicate that the two signatures come from the same source.
 Low cross correlation values indicate signatures which may originate from different sources.
@@
 The random forest presented in \citet{aoas2} is based on a combination of multiple characteristics, including cross correlation, number of matching striae, and number of consecutively matching striae.
-The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where scores close to 1 indicate it to be highly likely that the aligned signatures come from the same source, while scores close to 0 indicate that the aligned signatures are highly likely to originate from different sources.
+The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the similarity of the two signatures, where scores close to 1 indicate it to be highly likely that the aligned signatures come from the same source, while scores close to 0 indicate that the aligned signatures are highly likely to originate from different sources.
 
@@
 
-\subsection{Assessing Bullet-to-Bullet Match Strength}
+\subsection{Assessing Bullet-to-Bullet Similarity}
 
@@
 
-When imaging bullets, operators scan one land at a time in a clockwise (left twisted rifling) or anti-clockwise (right twisted rifling) sequence.
+When imaging bullets using a confocal light microscope, operators scan one land at a time in a clockwise (left twisted rifling) or anti-clockwise (right twisted rifling) sequence (Other imaging systems, such as BulletTrax 3D\footnote{https://www.ultra-forensictechnology.com/en/our-products/ballistic-identification/bullettrax/} or BalScan\footnote{https://www.forensic.cz/en/products/balscan}, may capture lower-resolution scans of an entire bullet at one time).
 The order in which scans are acquired is kept as meta-information.
@@
 We assess the results on each set and evaluate whether the minimal requirements described above are fulfilled for bullet-to-bullet SAM scores.
-We compare the random forest algorithm's performance to the performance of other suggested measures for quantitative assessment of match strength, such as cross correlation and consecutive matching striae.
+We compare the random forest algorithm's performance to the performance of other suggested measures for quantitative assessment of bullet similarity, such as cross correlation and consecutive matching striae.
 
@@
 In this section, we will compare the random forest scores to cross correlation scores; cross correlation is one of the components of the random forest, but has also been used to quantify the strength of a match in its own right \citep{chuPilotStudyAutomated2010}.
-In addition, both scoring methods use the same scale (0 to 1) and are continuous along that interval; other evaluation methods, such as consecutive matching striae are discrete and more difficult to directly compare to continuous measurements.
+In addition, both scoring methods use the same scale (0 to 1) and are continuous along that interval; other evaluation methods, such as consecutive matching striae, are discrete and more difficult to directly compare to continuous measurements.
 
@@
 There is some overlap between known matching and known non-matching land-to-land scores; in many cases the random forest scores for these overlapping values are more extreme than the corresponding cross correlation scores.
-This suggests that the matching algorithm is sensitive to the presence of LEAs which have low land-to-land match scores, such as the matches of bullets 'I' and 'F' in the Hamby 44 study.
+This suggests that the matching algorithm is sensitive to the presence of LEAs which have low land-to-land similarity scores, such as the matches of bullets 'I' and 'F' in the Hamby 44 study.
 Comparing this to the bullet-to-bullet score density plot shown in \autoref{fig:compare}, we see that there is much greater separation between the same-source and different-source densities for each study in the bullet-to-bullet comparisons than in the land-to-land comparisons, that is, one or two poorly matching lands does not prevent the bullet from showing a matching score, though if there are weak matches on several lands, such as the marked Hamby 44 comparisons in \autoref{fig:compare}, the overall score may be affected.
@@
 It may be useful to couple the \citet{aoas2} algorithm with an algorithm which assesses the quality of the input data and determines which portions of the data to use for comparison.
-This would emulate the process used by examiners, who first assess the quality of the evidence and whether there is enough information present to attempt a match.
+This would emulate the process used by examiners, who first assess the quality of the evidence and whether there is enough information present to attempt a comparison.
 One major disadvantage of an automated algorithm is that all of the decisions humans make (recognizing degraded land areas, excluding those areas from consideration, matching only the remaining areas) must be explicitly characterized; however, this explicit characterization means that the process can be scientifically validated to a much higher degree than the human perceptual process.
@@
 Rugers are among the firearms that mark very well, so we would correspondingly expect to see lowered similarity scores for other firearms.
-The matching algorithm works on matching lands.
-Barrels with polygonal rifling, such as Glocks, do not introduce well defined lands on bullets, which makes an application of the matching algorithm impossible.
+The matching algorithm works on matching lands, so barrels with polygonal rifling, such as Glocks, which do not introduce well defined lands on bullets, cannot be compared using the matching algorithm examined here.
 

### 2019-11-10 index.Rnw
Stable crosscut, some modification to binary classification explanation.

@@
 \begin{itemize}
-\item[(a)] Identify an area on the LEA with expressed striae. Locate a stable crosscut in the identified area.
+\item[(a)] Identify an area on the LEA with expressed striae. Locate a stable crosscut in the identified area; that is, an area where the striae are similar in the region above and below the crosscut.
 \item[(b)] Discard extraneous/contaminated data, such as data from groove engraved area and areas affected by break-off or contact with objects after the bullet exited the barrel  (``tank rash").
@@
 This increased rigor may increase the error rate of the model when compared to examiner error rates, but we expect that the increased information available from the 3D measurement data will compensate for this loss.
-A binary decision model is also easier to interpret and provides more clear-cut, definitive results.
-
-We can enumerate characteristics of an ideal similarity scoring mechanism: 
+A binary decision model is also easier to interpret and provides more clear-cut, definitive results than the AFTE Theory of Identification system used in most forensic laboratories in the United States.
+Of course, in jurisdictions which utilize score-based likelihood ratios, this threshold system is not necessary because likelihood ratios are evaluated on a continuous spectrum.
+Working under the binary decision model, we can enumerate characteristics of an ideal similarity scoring mechanism:
+
 \begin{description}

### 2019-11-10 index.Rnw
Clean up figure 2

@@
 <<echo=FALSE, results = 'hide', warning = FALSE, message = FALSE>>=
-features <- read.csv("data/fsi-features.csv.gz")
-features <- features %>% mutate(
-  bullet1 = gsub("ullet ", "", bullet1),
-  bullet2 = gsub("ullet ", "", bullet2)
-)
-b12 <- features %>% filter(
-  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
-  group2 == "G1", barrel2 == "KA", bullet2 == "B2"
-)
-p1 <- b12 %>%
-  ggplot(aes(x = land1, y = land2, fill = rfscore)) +
-  geom_tile() + ggtitle("Same barrel") +
-  scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
-  theme_bw() +
-  xlab("Bullet 2") + ylab("Bullet 1") +
-  theme(legend.position = "bottom") + coord_equal()
-
-bab <- features %>% filter(
-  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
-  group2 == "G1", barrel2 == "KB", bullet2 == "B1"
-)
-p2 <- bab %>%
-  ggplot(aes(x = land1, y = land2, fill = rfscore)) +
-  geom_tile() + ggtitle("Different barrel") +
-  scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
-  theme_bw() +
-  xlab("Bullet 2") + ylab("Bullet 1") +
-  theme(legend.position = "bottom") + coord_equal()
+# features <- read.csv("data/fsi-features.csv.gz")
+# features <- features %>% mutate(
+#   bullet1 = gsub("ullet ", "", bullet1),
+#   bullet2 = gsub("ullet ", "", bullet2)
+# )
+# b12 <- features %>% filter(
+#   group1 == "G1", barrel1 == "KA", bullet1 == "B1",
+#   group2 == "G1", barrel2 == "KA", bullet2 == "B2"
+# )
+# 
+# p1 <- b12 %>%
+#   ggplot(aes(x = land1, y = land2, fill = rfscore)) +
+#   geom_tile() + ggtitle("Same barrel") +
+#   scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
+#   theme_bw() +
+#   xlab("Bullet 2") + ylab("Bullet 1") +
+#   theme(legend.position = "bottom") + coord_equal()
+# 
+# bab <- features %>% filter(
+#   group1 == "G1", barrel1 == "KA", bullet1 == "B1",
+#   group2 == "G1", barrel2 == "KB", bullet2 == "B1"
+# )
+# p2 <- bab %>%
+#   ggplot(aes(x = land1, y = land2, fill = rfscore)) +
+#   geom_tile() + ggtitle("Different barrel") +
+#   scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
+#   theme_bw() +
+#   xlab("Bullet 2") + ylab("Bullet 1") +
+#   theme(legend.position = "bottom") + coord_equal()
+
+l2lscores <- bind_rows(
+  mutate(b12, type = "Same barrel"),
+  mutate(bab, type = "Different barrel")
+) %>% 
+  mutate(type = factor(type, levels = c("Same barrel", "Different barrel"), ordered = T))
+
 @
@@
 On the left the two bullets are known to come from the same source (barrel) on the right, the bullets are from two different sources (barrels).">>=
-grid.arrange(p1, p2, ncol = 2, padding = 2)
-@
-
-When imaging bullets using a confocal light microscope, operators scan one land at a time in a clockwise (left twisted rifling) or anti-clockwise (right twisted rifling) sequence (Other imaging systems, such as BulletTrax 3D\footnote{https://www.ultra-forensictechnology.com/en/our-products/ballistic-identification/bullettrax/} or BalScan\footnote{https://www.forensic.cz/en/products/balscan}, may capture lower-resolution scans of an entire bullet at one time).
+# grid.arrange(p1, p2, ncol = 2, padding = 2)
+ggplot(data = l2lscores, aes(x = land1, y = land2, fill = rfscore)) +
+  geom_tile() + 
+  scale_fill_gradient2("Random Forest (RF) score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
+  theme_bw() +
+  xlab("Bullet 2") + ylab("Bullet 1") +
+  theme(legend.position = "bottom") + coord_equal() + facet_wrap(~type)
+@
+
+While some imaging systems, such as BulletTrax 3D\footnote{https://www.ultra-forensictechnology.com/en/our-products/ballistic-identification/bullettrax/} or BalScan\footnote{https://www.forensic.cz/en/products/balscan}, may capture lower-resolution scans of an entire bullet at one time, confocal light microscopes do not have 360$^\circ$ capability.
+Thus, when imaging bullets using a confocal light microscope, operators scan one land at a time in a clockwise (left twisted rifling) or anti-clockwise (right twisted rifling) sequence.
 The order in which scans are acquired is kept as meta-information.
+
 Let us assume that lands on a bullet are labelled $\ell_i$ with $i = 1, .., p$, where $p$ indicates the number of lands a bullet has, as determined by the rifling of a barrel.

### 2019-11-10 index.Rnw
same and different source as separate facets with RF score on the side, not bottom.

@@
 
-<<echo=FALSE, results = 'hide', warning = FALSE, message = FALSE>>=
-# features <- read.csv("data/fsi-features.csv.gz")
-# features <- features %>% mutate(
-#   bullet1 = gsub("ullet ", "", bullet1),
-#   bullet2 = gsub("ullet ", "", bullet2)
-# )
-# b12 <- features %>% filter(
-#   group1 == "G1", barrel1 == "KA", bullet1 == "B1",
-#   group2 == "G1", barrel2 == "KA", bullet2 == "B2"
-# )
+<<demo-l2lscores, echo=FALSE, results = 'hide', warning = FALSE, message = FALSE>>=
+features <- read.csv("data/fsi-features.csv.gz")
+features <- features %>% mutate(
+  bullet1 = gsub("ullet ", "", bullet1),
+  bullet2 = gsub("ullet ", "", bullet2)
+)
+b12 <- features %>% filter(
+  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
+  group2 == "G1", barrel2 == "KA", bullet2 == "B2"
+)
 # 
@@
 # 
-# bab <- features %>% filter(
-#   group1 == "G1", barrel1 == "KA", bullet1 == "B1",
-#   group2 == "G1", barrel2 == "KB", bullet2 == "B1"
-# )
+bab <- features %>% filter(
+  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
+  group2 == "G1", barrel2 == "KB", bullet2 == "B1"
+)
 # p2 <- bab %>%
@@
   geom_tile() + 
-  scale_fill_gradient2("Random Forest (RF) score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
+  scale_fill_gradient2("Random Forest\n(RF) score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
   theme_bw() +
   xlab("Bullet 2") + ylab("Bullet 1") +
-  theme(legend.position = "bottom") + coord_equal() + facet_wrap(~type)
+  theme(legend.position = "right") + coord_equal() + facet_wrap(~type)
 @
@@
 \begin{definition}{Sequence Average and its Maximum}
-Let $A$ be a square real-valued matrix of dimensions $p \times p$. For the purpose of this paper, $A$ consists of scores describing the similarity between two sets of land engraved areas.
+Let $A$ be a square real-valued matrix of dimensions $p \times p$.
+For the purpose of this paper, $A$ consists of scores describing the similarity between two sets of land engraved areas.
+The entries in $A$ are represented as $a_{i,j}$, where $i$ and $j$ are row and column indexes.
+
 The $k$th \emph{sequence average} $SA(A, k)$ for $k = 0, ... p-1$ is defined  as 

### 2019-11-10 index.Rnw
Fix resolution emphasis thing

@@
 The scans used to train the model were taken at 20 fold magnification for a resolution of 1.5625 microns per pixel.
-While magnification is generally of interest in microscopy, **resolution** -- usually measured in microns per pixel -- is of more interest for 3D topographic measurements, as it determines the operative level of available data.
+While magnification is generally of interest in microscopy, \emph{resolution} -- usually measured in microns per pixel -- is of more interest for 3D topographic measurements, as it determines the operative level of available data.
 

### 2019-11-10 index.Rnw
Add CMS to the comparison of scores figures

@@
 
-<<compare, fig.width = 8, fig.height = 5, fig.cap="Density curves of similarity scores from cross correlation (top) and RF scores (bottom).
+<<compare, fig.width = 8, fig.height = 7, fig.cap="Density curves of similarity scores from random forest scores (top), cross correlation (middle), and consecutive matching striae (bottom).
 Different colors indicate same source versus different source.
@@
   h44nest %>%
-    select(sam_rf, sam_ccf, samesource) %>%
+    select(sam_rf, sam_ccf, sam_cms, samesource) %>%
     mutate(study = "Hamby 44", group1 = NA),
   pdnest %>%
-    select(sam_rf, sam_ccf, samesource) %>%
+    select(sam_rf, sam_ccf, sam_cms, samesource) %>%
     mutate(study = "Phoenix PD", group1 = NA),
   hounest %>%
-    select(sam_rf, sam_ccf,  samesource, group1) %>%
+    select(sam_rf, sam_ccf, sam_cms, samesource, group1) %>%
     mutate(study = "Houston FSI")
@@
 nests$study <- factor(nests$study, levels = c("Hamby 44", "Phoenix PD", "Houston FSI"))
-nestslong <- nests %>% gather(key = "measure", value = "value", sam_rf, sam_ccf)
+nestslong <- nests %>% gather(key = "measure", value = "value", sam_rf, sam_ccf, sam_cms)
 nestslong$measure <- factor(nestslong$measure)
-levels(nestslong$measure) <- c("Cross correlation",  "RF score")
+levels(nestslong$measure) <- c("Cross correlation", "CMS", "RF score")
 
@@
 labels$y <- 5 + 4:1
-nestslong %>%
+p1 <- nestslong %>%
+  filter(measure %in% c("RF score")) %>%
   ggplot(aes(
@@
   geom_density(alpha = 0.6) +
-  geom_rug(alpha = 0.5, show.legend = F) +
+  geom_rug(alpha = 0.3, show.legend = F) +
   facet_grid(measure ~ study) +
   theme_bw() +
-  scale_fill_manual("Same source", values = c("darkgrey", "darkorange")) +
-  scale_colour_manual("Same source", values = c("darkgrey", "darkorange")) +
+  scale_fill_manual("Same source", values = c("darkgrey", "darkorange"), guide = F) +
+  scale_colour_manual("Same source", values = c("darkgrey", "darkorange"), guide = F) +
   scale_x_continuous("SAM scores", limits=c(0,1)) +
@@
                data = labels
-  )
+  ) + 
+  theme(axis.title.x = element_blank(), axis.text.y = element_text(hjust = .5, angle = 90))
+
+p2 <- nestslong %>%
+  filter(measure %in% c("Cross correlation")) %>% 
+  ggplot(aes(
+    x = value, fill = samesource,
+    colour = samesource
+  )) +
+  geom_density(alpha = 0.6) +
+  geom_rug(alpha = 0.3, show.legend = F) +
+  facet_grid(measure ~ study) +
+  theme_bw() +
+  theme(legend.position = "bottom") +
+  scale_fill_manual("Same source", values = c("darkgrey", "darkorange"), guide = F) +
+  scale_colour_manual("Same source", values = c("darkgrey", "darkorange"), guide = F) +
+  scale_x_continuous("SAM scores") + 
+  theme(axis.title.x = element_blank(), axis.text.y = element_text(hjust = .5, angle = 90))
+
+p3 <- nestslong %>%
+  filter(measure %in% c("CMS")) %>% 
+  ggplot(aes(
+    x = value, fill = samesource,
+    colour = samesource
+  )) +
+  geom_density(alpha = 0.6) +
+  geom_rug(aes(x = jitter(value)), alpha = 0.3, show.legend = F) +
+  facet_grid(measure ~ study) +
+  theme_bw() +
+  theme(legend.position = "bottom") +
+  scale_fill_manual("Same source", values = c("darkgrey", "darkorange")) +
+  scale_colour_manual("Same source", values = c("darkgrey", "darkorange")) +
+  scale_x_continuous("SAM scores")  + 
+  theme(axis.text.y = element_text(hjust = .5, angle = 90))
+
+grid.arrange(p1, p2, p3, nrow = 3, heights = c(1, 1, 1.4))
 @
@@
 
-<<compare-land-to-land, fig.width = 8, fig.height = 4, fig.cap="Density curves of land-to-land similarity scores from cross correlation (top) and RF scores (bottom).
+<<compare-land-to-land, fig.width = 8, fig.height = 6.5, fig.cap="Density curves of land-to-land similarity scores from RF scores (top), cross-correlation (middle), and consecutive matching striae (CMS)(bottom).
 Different colors indicate same source versus different source for each land.
@@
 Some same-source comparisons are also below 0.5 for RF scores, indicating potential problems with at least one of the lands involved in the comparison.">>=
-all_lands %>% 
-  gather(key = "measure", value = "value", ccf, rfscore) %>%
-  mutate(measure = factor(measure, levels = c("ccf", "rfscore"),
-                          labels = c("Cross correlation", "RF score")),
+# all_lands %>% 
+#   gather(key = "measure", value = "value", ccf, rfscore) %>%
+#   mutate(measure = factor(measure, levels = c("ccf", "rfscore"),
+#                           labels = c("Cross correlation", "RF score")),
+#          study = paste(study, group1) %>% str_trim() %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"))) %>%
+#   ggplot(aes(
+#     x = value, fill = sameland,
+#     colour = sameland
+#   )) +
+#   geom_density(alpha = 0.6) +
+# #  geom_rug(alpha = 0.5/12, show.legend = F) +
+#   facet_grid(measure ~ study) +
+#   theme_bw() +
+#   scale_fill_manual("Same source", values = c("darkgrey", "darkorange")) +
+#   scale_colour_manual("Same source", values = c("darkgrey", "darkorange")) +
+#   scale_x_continuous("Land-to-Land scores", limits = c(0, 1), labels = c(0, 0.25, 0.5, 0.75, 1)) +
+#   theme(legend.position = "bottom") +
+#   geom_hline(yintercept = 0, show.legend = F)
+p1 <- all_lands %>% 
+  gather(key = "measure", value = "value", ccf, cms, rfscore) %>%
+  mutate(measure = factor(measure, levels = c("ccf", "cms", "rfscore"),
+                          labels = c("Cross correlation", "CMS", "RF score")),
          study = paste(study, group1) %>% str_trim() %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"))) %>%
+  filter(measure == "RF score") %>%
+  ggplot(aes(
+    x = value, fill = sameland,
+    colour = sameland
+  )) +
+  geom_density(alpha = 0.6) +
+#  geom_rug(alpha = 0.5/12, show.legend = F) +
+  facet_grid(measure ~ study) +
+  theme_bw() +
+  scale_fill_manual("Same source", values = c("darkgrey", "darkorange"), guide = F) +
+  scale_colour_manual("Same source", values = c("darkgrey", "darkorange"), guide = F) +
+  scale_x_continuous("Land-to-Land scores", limits = c(0, 1), labels = c(0, 0.25, 0.5, 0.75, 1)) +
+  theme(legend.position = "bottom", axis.title.x = element_blank(), axis.text.y = element_text(hjust = .5, angle = 90)) +
+  geom_hline(yintercept = 0, show.legend = F)
+
+p2 <- all_lands %>% 
+  gather(key = "measure", value = "value", ccf, cms, rfscore) %>%
+  mutate(measure = factor(measure, levels = c("ccf", "cms", "rfscore"),
+                          labels = c("Cross correlation", "CMS", "RF score")),
+         study = paste(study, group1) %>% str_trim() %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"))) %>%
+  filter(measure == "Cross correlation") %>%
+  ggplot(aes(
+    x = value, fill = sameland,
+    colour = sameland
+  )) +
+  geom_density(alpha = 0.6) +
+#  geom_rug(alpha = 0.5/12, show.legend = F) +
+  facet_grid(measure ~ study) +
+  theme_bw() +
+  scale_fill_manual("Same source", values = c("darkgrey", "darkorange"), guide = F) +
+  scale_colour_manual("Same source", values = c("darkgrey", "darkorange"), guide = F) +
+  scale_x_continuous("Land-to-Land scores", limits = c(0, 1), labels = c(0, 0.25, 0.5, 0.75, 1)) +
+  theme(legend.position = "bottom", axis.title.x = element_blank(), axis.text.y = element_text(hjust = .5, angle = 90)) +
+  geom_hline(yintercept = 0, show.legend = F)
+
+p3 <- all_lands %>% 
+  gather(key = "measure", value = "value", ccf, cms, rfscore) %>%
+  mutate(measure = factor(measure, levels = c("ccf", "cms", "rfscore"),
+                          labels = c("Cross correlation", "CMS", "RF score")),
+         study = paste(study, group1) %>% str_trim() %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"))) %>%
+  filter(measure == "CMS") %>%
   ggplot(aes(
@@
   scale_colour_manual("Same source", values = c("darkgrey", "darkorange")) +
-  scale_x_continuous("Land-to-Land scores", limits = c(0, 1), labels = c(0, 0.25, 0.5, 0.75, 1)) +
-  theme(legend.position = "bottom") +
+  scale_x_continuous("Land-to-Land scores") +
+  theme(legend.position = "bottom", axis.text.y = element_text(hjust = .5, angle = 90)) +
   geom_hline(yintercept = 0, show.legend = F)
+
+grid.arrange(p1, p2, p3, nrow = 3, heights = c(1, 1, 1.4))
 @

### 2019-11-10 index.Rnw
Addressing conclusion and plot meddling.

@@
   ) %>%
-  ggplot(aes(linetype = measure, colour = studyg)) +
+  ggplot(aes(linetype = measure, colour = studyg, shape = measure)) +
   geom_roc(aes(m = score, d = as.numeric(sameland)), 
            labels = FALSE, pointsize = 0, alpha = 0.8, size = 1) +
-  theme_bw() +
+  theme_bw(base_size = 16) +
   # coord_fixed() + 
-  facet_wrap(~studyg, nrow=3) +
+  facet_wrap(~studyg, nrow = 3) +
   xlab("Percent Wrong Identifications") +
@@
   theme(legend.position= c(1, 0), legend.justification = c(1, 0), legend.direction = "vertical",
-        legend.key.width = unit(1, "cm")) +
-  guides(linetype = guide_legend(override.aes = list(shape = NA, size = .5))) + 
+        legend.key.width = unit(1, "cm"), axis.text.y = element_text(angle = 90, hjust = .5)) +
+  guides(linetype = guide_legend(override.aes = c(listsize = .5))) + 
   scale_linetype_manual("Measure", values=c(1,3,2)) +
+  scale_shape_discrete("Measure") +
   scale_x_continuous(breaks=seq(0,1, by=0.25), labels=c("0", "25%", "50%", "75%", "100%")) +
@@
 
-<<aucs-pic>>=
+<<aucs-pic, include = F>>=
 eer <- function(roc){
@@
   ) %>% 
-  mutate(
-    studyg = paste(study, group1) %>% str_trim %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3")),
+  ungroup() %>% 
+  mutate(
+    studyg = paste(study, group1) %>% str_trim %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"),
+                                                          labels = c("Hamby 44", "Phoenix PD", "Houston FSI\nG1", "Houston FSI\nG2", "Houston FSI\nG3")),
     measure = factor(measure, levels = c("cms", "ccf", "rfscore"), labels = c("CMS", "CC", "RF")),
@@
   geom_point(size=3.5, aes(shape = measure)) +
-  theme_bw() +
+  theme_bw(base_size = 16) +
   xlab("AUC and 95% confidence intervals") +
@@
                         guide = FALSE) +
+  theme(axis.text.y = element_text(angle = 90, hjust = 0.5)) + 
   # theme(legend.position = "bottom",
@@
 
-<<eer, fig.width=7, fig.height = 4, fig.cap="Equal Error (in Percent)  and corresponding thresholds for all measures (RF score, Cross correlation and CMS) under all five (sub-)studies.
-CMS has an equal error threshold of 2 for all studies.
-CMS is signifcantly worse than both Cross correlation and the RF score.
-At best, Cross correlation and RF score have equal errors for land-to-land comparison of around 5\\%.", out.width = '.8\\textwidth'>>=
+<<eer, fig.width=6, fig.height = 2.6, fig.cap="Equal error (in Percent)  and corresponding thresholds for all measures (random forest score, cross correlation and CMS) under all five (sub-)studies.
+CMS has an equal error threshold of 2 for all studies, and perform signifcantly worse than both cross correlation and the random forest score.
+At best, Cross correlation and RF score have equal errors for land-to-land comparisons of around 5\\%.
+The performance of the random forest score is slightly better than the cross-correlation on sets which have higher equal error rates (Houston FSI G2, Hamby 44).", out.width = '\\textwidth'>>=
 aucs %>% 
   mutate(
-    eer = 100*eer
+    eer = 100*eer,
+    studyg = factor(studyg, labels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"))
   ) %>%
@@
   theme_bw() +
-  xlab("Equal Error (in Percent, left) and corresponding thresholds (right)") +
+  xlab("Equal Error % (left) and corresponding thresholds (right)") +
   ylab("") +
-  scale_color_brewer(palette = "Set2") +
-  scale_shape_discrete(solid = F) + 
-  theme(legend.position = "bottom") +
+  scale_color_brewer("Measure", palette = "Set2") +
+  scale_shape_discrete("Measure", solid = F) + 
+  theme(legend.position = "right") +
   facet_grid(.~eer_thresh, scales="free") +
@@
 In a surprise turn of events, it was the hardest to solve, in part because of damage to the bullets that obscured LEA striae.
-It has been shown \citet{lpr2} that parts of lands can be used for successful identifications.
-The algorithm proposed by \citet{aoas2} is not capable of  automatically detecting parts of lands with well-expressed striae.
+It has been shown \citet{lpr2} that parts of lands can be used for successful identifications, if at least 50\% of the land is present (using full-length scores as the reference distribution).
+The random forest algorithm proposed in \citet{aoas2} is not capable of automatically detecting parts of lands with well-expressed striae.
 It may be useful to couple the \citet{aoas2} algorithm with an algorithm which assesses the quality of the input data and determines which portions of the data to use for comparison.
@@
 Rugers are among the firearms that mark very well, so we would correspondingly expect to see lowered similarity scores for other firearms.
-The matching algorithm works on matching lands, so barrels with polygonal rifling, such as Glocks, which do not introduce well defined lands on bullets, cannot be compared using the matching algorithm examined here.
+While the performance of cross-correlation and random forest scores are similar when evaluated on Rugers, it is likely that the additional information used to generate the random forest score may be useful when comparing bullets from barrels which produce fewer distinctive marks.
+A future study will compare the performance of the random forest score, cross-correlation score, and CMS on non-Ruger barrels.
+The random forest algorithm is trained on scans taken at the land level, so barrels with polygonal rifling, such as Glocks, which do not introduce well defined lands on bullets, cannot be compared using the random forest algorithm examined here.
 

### 2019-12-06 index.Rnw
Tweaks to title, phrasing, etc.

@@
 
-\title{Case study validations of automatic bullet matching\tnoteref{t1}\tnoteref{t2}}
+\title{Comparison of three similarity scores for bullet LEA matching\tnoteref{t1}\tnoteref{t2}}
 
@@
 Exact guidelines for this classification vary from lab to lab; some labs will exclude only on the basis of non-matching class characteristics, such as direction of the twist in rifling, land length or number of lands, or type of rifling.
-In other labs, CMS (consecutively matching striae) as defined by \citeauthor{biasotti} \citep{biasotti} is used as a measure to quantify the strength of a match.
+In other labs, CMS (consecutively matching striae) as defined by \citeauthor{biasotti} \citep{biasotti} is used as a measure to quantify the similarity of two lands.
 In virtually all labs, individual characteristics used to identify matching bullets are derived from visual assessment; some class characteristics may be directly measured, but these are not sufficient for individualization.
@@
 \item[(R1)] {\bf Monotonicity:} a higher score is indicative of higher similarity between a pair of bullets, in particular, similarity scores of same-source pairs of bullets are higher than different-source pairs.
-\item[(R2)] {\bf Stability:} the same score leads to the same conclusion. 
+\item[(R2)] {\bf Stability:} the same score leads to the same conclusion in all situations and under separately assembled reference distributions. 
 \end{description}

### 2019-12-06 index.Rnw
LR fixes

@@
 A binary decision model is also easier to interpret and provides more clear-cut, definitive results than the AFTE Theory of Identification system used in most forensic laboratories in the United States.
-Of course, in jurisdictions which utilize score-based likelihood ratios, this threshold system is not necessary because likelihood ratios are evaluated on a continuous spectrum.
-Working under the binary decision model, we can enumerate characteristics of an ideal similarity scoring mechanism:
+Of course, in jurisdictions which utilize score-based likelihood ratios, this threshold system is not necessary because likelihood ratios are evaluated on a continuous spectrum (though, there is a natural threshold at 1, and many times a more flexible threshold is used to determine the level of significance of a result).
+Working under the binary decision model common in US jurisdictions, we can enumerate characteristics of an ideal similarity scoring mechanism:
 

### 2019-12-06 Reviewer_response.Rmd
Reviewer response -- start -- this is a private repo, so we can have this here.

---
title: "Response"
author: "Susan Vanderplas"
date: "12/6/2019"
output: html_document
---

We thank the reviewer for his or her careful and constructive comments on our article, which we have retitled "Comparison of three similarity scores for bullet LEA matching". We have made several changes to the document in response to the suggestions made by the reviewer, and have responses to the comments which could not be easily addressed in this paper as it stands.

The figures have been adjusted to have a larger font size for labels and titles; the caption size is set by the LaTeX template and is thus not easily editable by the authors. We expect that the captions will be sized appropriately in the final product; at the moment, we do not have control over the sizing of the figure captions.

The choice to exclude score-based likelihood ratios was a conscious choice because the land-to-land comparisons result in density estimates with different variability between known matches and known non-matches. As a result, in these cases, the SLR is nonmonotonic and produces unexpected and potentially misleading results at the extremes of the distribution. As a result, we would like to defer the explicit use of SLRs in this paper; we have added a paragraph on the potential of using likelihood ratios with the scoring methods described in the paper. The threshold-based model we used in this paper is more consistent with the model used by examiners in our jurisdiction: decisions include, broadly, identification (the bullets match), exclusion (the bullets do not match), inconclusive, and insufficient evidence to make a conclusion. Our computational model does not require the last two categories, reducing this problem to a binary decision with a single threshold. We would also note that using likelihood ratios does not remove the need for a threshold; rather, it moves (implicitly or explicitly) to a threshold of 1, or 10, or whatever the criteria for "enough evidence" is set to.

The reviewer suggests using multiplication to combine land-to-land SLRs into a bullet-to-bullet SLR; we cannot recommend this course of action in part because the land to land scores are not independent, rather, they result from pairwise comparisons that are inherently dependent. Further, we can't assume that striae observed on different lands from the same physical bullet are independent, because they were all part of the same firing event. While this assumption is often made for practical purposes, the assumption is highly suspect and should not be further promoted without considering the covariance structures within the data. For this reason, we feel that the use of the sequence average match method for combining scores is more reasonable because it accounts for the spatial relationships between the set of lands belonging to the same bullet, without requiring assumptions of independence.

We appreciate the reviewer's discussion about the Ruger vs. non-Ruger bullets; we have included some of the suggestions in the paper, but do not currently have sufficient data to include a fourth case study in this paper. We strongly agree that this would be an interesting test of the random forest score compared to the cross correlation and are planning several future studies which explore the random forest score's use on non-Ruger bullets. In addition, the Houston FSI study was also used to conduct examiner studies; we are preparing a paper which compares the algorithm's performance using the decision criteria shown in this paper to the performance of firearm examiners and expect to submit this paper for review in mid 2020.

We are also working on gaining approval to publish a data in brief paper with the data used in this paper, so that we can make the land scans publicly available for additional analysis; we strongly agree with the reviewer that the community would benefit from access to these scans. Two of the three sets have been uploaded to NBTRD at a lower resolution than the resolution used in our calculations.

### 2019-12-07 Reviewer_response.Rmd
More reviewer responses

@@
 non-Ruger bullets; we have included some of the suggestions in the paper, but do not currently have sufficient data to include a fourth case study in this paper.
-We strongly agree that this would be an interesting test of the random forest score compared to the cross correlation and are planning several future studies which explore the random forest score's use on non-Ruger bullets.
+We agree that this would be an interesting test of the random forest score compared to the cross correlation and are planning several future studies which explore the random forest score's use on non-Ruger bullets.
+As this paper is already fairly lengthy, as noted by the reviewer, we feel that the addition of a non-Ruger case study is best reserved for one of these future papers.
 In addition, the Houston FSI study was also used to conduct examiner studies; we are preparing a paper which compares the algorithm's performance using the decision criteria shown in this paper to the performance of firearm examiners and expect to submit this paper for review in mid 2020.
+
+The final broad XXX comment addresses the matching of partial signatures, specifically suggesting that these signatures should be compared to a reference distribution assembled from partial signature comparisons.
+Hare, Hofmann, and Carriquiry (2017) showed how to use the random forest score to match degraded and partial land impressions; the random forest score remained fairly stable up to 50% signature length, suggesting that the random forest score, at least, is relatively robust to degradation and partial mark size.
+For the reasons stated above, we do not carry this conclusion all the way to the calculation of SLRs, but given the previously cited paper, we do not expect to require a separate comparison distribution when the partial LEA includes at least 50% of the area.
 
@@
 Two of the three sets have been uploaded to NBTRD at a lower resolution than the resolution used in our calculations.
+
+We have addressed many of the other minor comments, adding clarifications to address the specific issues with phrasing and technical corrections.
+We created a replacement for Figure 1 with the signatures overlaid, but found it difficult to read because the sequences are not (at this point in the algorithm) aligned, so much of the similarity is difficult to see when the unaligned details are overlaid.
+We include the modified Figure 1 here as a demonstration, but ultimately feel that the original Fig 1 (with labels of increased size) allows for easier visual comparison of the peaks in the sequence.
+
+![](fig1-alt.png)
+
+More specifically, we have addressed the corrections to the abstract, the phrasing of match strength vs.
+similarity, ambiguous terms such as stable crosscut, and the definition of stability.
+We appreciate these corrections, and hope that this revised manuscript is more clear.
+% need to talk through cutoff, imaging, training, peak/valley detection.

### 2019-12-09 index.Rnw
Fixing match strength, zero cms

@@
 Individual characteristics, which are not shared by all members of a class, can also be automatically assessed from 3D surface measurements.
-One of the most common features used to assess the strength of a match based on individual characteristics is the cross correlation function, which is utilized in several studies \citep{maNISTBulletSignature2004, vorburgerApplicationsCrosscorrelationFunctions2011, chuPilotStudyAutomated2010}.
-Additional features proposed for automatically assessing match strength also include signature distance \citep{maNISTBulletSignature2004}, striae depth and width \citep{ComputerIdentificationBullets1978}, and consecutive matching striae (CMS) \citep{chuAutomaticIdentificationBullet2013}.
+One of the most common features used to describe the similarity of two surfaces is the cross correlation function, which is utilized in several studies \citep{maNISTBulletSignature2004, vorburgerApplicationsCrosscorrelationFunctions2011, chuPilotStudyAutomated2010}.
+Additional features proposed for automatically assessing similarity also include signature distance \citep{maNISTBulletSignature2004}, striae depth and width \citep{ComputerIdentificationBullets1978}, and consecutive matching striae (CMS) \citep{chuAutomaticIdentificationBullet2013}.
 
@@
 Peaks -- rather than peaks and valleys -- are used for this determination in part because it is difficult to visually assess valleys except as relative to peaks.
-Utilizing automatic alignment based on maximum cross correlation (as opposed to matching striation marks) allows the possibility that two lands will have zero consecutive matching striae; this would not occur with manual alignment based on striation marks.
+Utilizing automatic alignment based on maximum cross correlation (as opposed to matching striation marks) allows the possibility that two lands will have zero consecutive matching striae; this would not occur with manual alignment based on striation marks, as the examiner typically lines up two distinctive markings, resulting in at least one matching striation mark.
 
@@
 \autoref{fig:tankrash} gives an overview of all lands of affected bullets in the Hamby-44 set.
-These issues affect the algorithm's ability to identify a stable signature, which impacts the subsequent match scores  at the land-to-land and bullet-to-bullet level.
+These issues affect the algorithm's ability to identify a stable signature, which impacts the subsequent similarity scores  at the land-to-land and bullet-to-bullet level.
 <<tankrash, fig.cap="Overview of bullet lands with prominent deficiency such as tank rash or extreme pitting in the Hamby-44 study.", fig.width=4, fig.height = 2, out.width='47.5%', fig.subcap=c("Br1-B1-L6 (tank rask)", "Br2-B2-L5 (tank rash)", "Br3-B1-L5 (tank rash)", "Br8-B1-L6 (tank rash)", "Br8-B2-L2 (pitting)", "Br8-B2-L6 (tank rash)", "Unk-E-L6 (tank rash)", "Unk-I-L6 (tank rash and break-off)"), fig.align="middle", fig.ncol=2>>=
@@
 For a future round of the Houston study, we are planning to compare the algorithmic performance directly with scores given by forensic toolmark examiners.
-The algorithm's score is not intended to replace an examiner's judgment; instead, it provides a complementary tool to measure and quantify the strength of an identification.
+The algorithm's score is not intended to replace an examiner's judgment; instead, it provides a complementary tool to measure, quantify, and compare the similarity of two bullets in an identification.
+With more research into the behavior of score-based likelihood ratios, this method may also be used to compute
 By validating the algorithm on external test sets, we have demonstrated that the method can be generalized to different types of ammunition and is not overly sensitive to small differences in rifling procedure.

### 2019-12-09 index.Rnw
Binary decision model

@@
 A binary decision model is also easier to interpret and provides more clear-cut, definitive results than the AFTE Theory of Identification system used in most forensic laboratories in the United States.
-Of course, in jurisdictions which utilize score-based likelihood ratios, this threshold system is not necessary because likelihood ratios are evaluated on a continuous spectrum (though, there is a natural threshold at 1, and many times a more flexible threshold is used to determine the level of significance of a result).
-Working under the binary decision model common in US jurisdictions, we can enumerate characteristics of an ideal similarity scoring mechanism:
+Of course, in jurisdictions which utilize score-based likelihood ratios, this threshold system is not necessary because likelihood ratios are continuous.
+Even with likelihood ratios, however, there is a natural threshold at 1, and many times a more flexible threshold (100, 1000) is used to determine the weight given to a result.
+
+The remainder of this paper will utilize the binary decision model common in US jurisdictions. With this model, we can enumerate characteristics of an ideal similarity scoring mechanism:
 

### 2019-12-09 index.Rnw
More explicit citation to aoas paper

@@
 Utilizing automatic alignment based on maximum cross correlation (as opposed to matching striation marks) allows the possibility that two lands will have zero consecutive matching striae; this would not occur with manual alignment based on striation marks, as the examiner typically lines up two distinctive markings, resulting in at least one matching striation mark.
+The algorithms used to identify peaks and valleys in each signature and determine CMS from the peak and valley identifications are described in more detail in \citep[p.
+2340, step 3 -- 4]{aoas2}.
 
@@
 Each barrel was test-fired twice, so each of the questioned bullets (shown along the $y$ axis) matches two bullets fired from a known barrel.
-Colored tiles are used to show the strength of the comparison; light grey colors correspond to low similarity scores, dark colors correspond to high similarity scores.
+Colored tiles are used to show the similarity score: light grey colors correspond to low similarity scores, dark colors correspond to high similarity scores.
 Ground truth is encoded in this figure as a thin, dark, colored frame for all pairs of same-source bullets.
@@
 With no errors at the bullet-to-bullet level, we can now evaluate the scores in light of (R1) and (R2).
-In this section, we will compare the random forest scores to cross correlation scores; cross correlation is one of the components of the random forest, but has also been used to quantify the strength of a match in its own right \citep{chuPilotStudyAutomated2010}.
+In this section, we will compare the random forest scores to cross correlation scores; cross correlation is one of the components of the random forest, but has also been used to quantify the similarity of two signatures in its own right \citep{chuPilotStudyAutomated2010}.
 In addition, both scoring methods use the same scale (0 to 1) and are continuous along that interval; other evaluation methods, such as consecutive matching striae, are discrete and more difficult to directly compare to continuous measurements.

### 2019-12-09 index.Rnw
Additional attempts to address reviewer responses - brevity, stability, etc.

@@
 \begin{itemize}
-\item[(a)] Identify an area on the LEA with expressed striae. Locate a stable crosscut in the identified area; that is, an area where the striae are similar in the region above and below the crosscut.
+\item[(a)] Identify an area on the LEA with expressed striae.
+Locate a stable crosscut in the identified area; that is, an area where the striae are similar in the region above and below the crosscut.
+This excludes regions with extreme pitting, breakoff, tank rash, and other flaws that would interfere with similarity score calculation.
 \item[(b)] Discard extraneous/contaminated data, such as data from groove engraved area and areas affected by break-off or contact with objects after the bullet exited the barrel  (``tank rash").
@@
 Using the lag determined for aligning two signatures, other quantitative features describing the similarity of the two signatures can be extracted.
-Numerical features such as cross correlation and Euclidean distance
-can be computed from the aligned signatures alone.
-Other features include characteristics of the signature that firearms and toolmark examiners base their visual assessment on, such as striae depth, number of matching striae, and the number of consecutively matching striae (CMS), one of the measures firearms examiners use to describe the similarity of two bullets quantitatively.
+Numerical features such as cross correlation and Euclidean distance can be computed from the aligned signatures alone.
+Additional features are modeled after visual assessment methods used by examiners, including striae depth, total number of matching striae, and the number of consecutively matching striae (CMS).
 
 In order to evaluate CMS, we must first identify peaks and valleys in each of the signatures, then determine whether these peaks and valleys overlap sufficiently.
-Features which firearms examiners assess visually depend on the identification of these extrema.
+Features which firearms examiners assess visually generally depend on the identification of these extrema, as striation marks and the corresponding peaks are the most salient feature when viewing a bullet using a microscope.
 If there are at least six consecutive matching striae, the bullets are considered to be similar in practice \citep{biasotti_firearms_1997,nichols_scientific_2006}.
-CMS in this setting are counted as the number of consecutively matching peaks in signatures of two aligned lands.
-Peaks -- rather than peaks and valleys -- are used for this determination in part because it is difficult to visually assess valleys except as relative to peaks.
-Utilizing automatic alignment based on maximum cross correlation (as opposed to matching striation marks) allows the possibility that two lands will have zero consecutive matching striae; this would not occur with manual alignment based on striation marks, as the examiner typically lines up two distinctive markings, resulting in at least one matching striation mark.
+CMS, as measured by examiners, is the number of consecutively matching peaks in signatures of two aligned lands, and, because examiner typically lines up two distinctive markings, resulting in at least one matching striation mark.
+Peaks -- rather than peaks and valleys -- are used in part because it is difficult to visually assess valleys except as relative to peaks.
+In contrast, the matching algorithm can identify both peaks and valleys, and signatures are aligned based on maximum cross-correlation (as opposed to matching striation marks), there is the possibility that two lands will have zero consecutive matching striae.
+Thus, the matching algorithm uses a slightly different measure of CMS than examiners, but the underlying principle is the same.
 The algorithms used to identify peaks and valleys in each signature and determine CMS from the peak and valley identifications are described in more detail in \citep[p.
@@
 The features described above are each individually designed to separate same-source and different-source signatures: for instance, same-source signatures would have high cross correlation values and different-source signatures would have low cross correlation values.
-An automated algorithm selects a cutoff value, introducing a barrier between identification and exclusion.
-
+For each feature, or an aggregate score composed of many features, the end result for any two signatures is a number which must be compared to a distribution of other scores from similar situations.
+
+Using this distribution of scores and training data, an automated algorithm selects a cutoff value, introducing a barrier between identification and exclusion.
 This threshold system introduces a binary classification: identification or exclusion.
@@
 Of course, in jurisdictions which utilize score-based likelihood ratios, this threshold system is not necessary because likelihood ratios are continuous.
-Even with likelihood ratios, however, there is a natural threshold at 1, and many times a more flexible threshold (100, 1000) is used to determine the weight given to a result.
-
-The remainder of this paper will utilize the binary decision model common in US jurisdictions. With this model, we can enumerate characteristics of an ideal similarity scoring mechanism:
+Even with likelihood ratios, however, there is a natural threshold at 1 which functions similarly to the selected threshold in the binary decision case.
+
+The remainder of this paper will utilize the binary decision model which is compatible with the legal framework commonly used in US jurisdictions.
+With this model, we can enumerate characteristics of an ideal similarity scoring mechanism:
 
@@
 While these comparisons are useful, the question of interest typically involves the entire physical object (e.g.
-all lands on a single bullet), and a conclusion on same or different source should be reached based on the evidence of all lands of the bullets.
+all lands on a single bullet), and a conclusion on same or different source should be reached based on the evidence of all lands of the two bullets.
 
@@
 
-Land-to-land comparisons lead to a whole set of scores for bullet-to-bullet comparisons.
+Land-to-land comparisons lead to a whole set of scores to evaluate for bullet-to-bullet comparisons.
 \autoref{fig:b2b} shows two matrices of scores for two pairs of bullet-to-bullet matches.
@@
 Let us assume that lands on a bullet are labelled $\ell_i$ with $i = 1, .., p$, where $p$ indicates the number of lands a bullet has, as determined by the rifling of a barrel.
-For all of the bullets considered here the number of lands $p$ is 6.
+For all of the bullets considered here the number of lands, $p$, is 6.
 A match between  a pair of lands from two bullets therefore results in an expected additional $p-1$ matches between pairs of lands.
@@
 
-In this paper, we validate the algorithm described in \citet{aoas2} on three external test sets.
-We assess the results on each set and evaluate whether the minimal requirements described above are fulfilled for bullet-to-bullet SAM scores.
-We compare the random forest algorithm's performance to the performance of other suggested measures for quantitative assessment of bullet similarity, such as cross correlation and consecutive matching striae.
-
-\section{Validation Sets}
-
 Random forests \citep{breiman} have a built-in internal testing mechanism to prevent potential overfitting.
-Errors reported by random forest algorithms are based on these internal test sets.
-However, there is some discussion on whether or in which situations these errors are biased \citep{rfOver2, rfOver}.
+Errors reported by random forest algorithms are based on these internal test sets, though there is some debate about the bias of these errors \citep{rfOver2, rfOver}.
 Neither of these papers addresses another issue with internal test sets: internal test sets are constructed to have the same distribution as the training data (apart from sampling variability).
@@
 
+In this paper, we validate the algorithm described in \citet{aoas2} on three external test sets.
+We assess the results on each set and evaluate whether the minimal requirements described above are fulfilled for bullet-to-bullet SAM scores.
+We compare the random forest algorithm's performance to the performance of other suggested measures for quantitative assessment of bullet similarity, such as cross correlation and consecutive matching striae.
+
+\section{Validation Sets}
+
 The algorithm in \citet{aoas2} is trained on scans from Hamby sets 252 and 173 made available through \citet{nist}.
-Set 173 was originally published as Hamby set 44.
-This mis-labeling has been corrected now.
+Set 173 was originally published as Hamby set 44, but the mis-labeling has been recently corrected.
 The scans used to train the model were taken at 20 fold magnification for a resolution of 1.5625 microns per pixel.
@@
 
-Case studies were chosen such that difficulty for the matching algorithm increases with each case study: Hamby set 44 is part of the Hamby study.
+Case studies were chosen such that theoretical difficulty for the matching algorithm increases with each case study: Hamby set 44 is part of the Hamby study.
 The algorithm in \citet{aoas2} was trained on sets 173 and 252, so the bullets in Hamby 44 are of the same type of ammunition and are fired through the same barrels as the bullets in the training set.
@@
 These barrels are first rifled traditionally, i.e.
-similar to the Ruger P-85, but are treated to a secondary round of heating after rifling.
-This second round of heat exposure has the potential to introduce some subclass characteristics, i.e.\ anomalies that are shared between different barrels manufactured in this manner.
+similar to the Ruger P-85, but are treated with a secondary round of heating after rifling, which may introduce some subclass characteristics.
 This should make the automatic matching harder, and in particular, is expected to complicate the classification as different-source land engraved areas/bullets.

### 2019-12-09 index.Rnw
Bandwidth fix

@@
 
-p3 <- all_lands %>% 
+p3data <- all_lands %>% 
   gather(key = "measure", value = "value", ccf, cms, rfscore) %>%
@@
          study = paste(study, group1) %>% str_trim() %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"))) %>%
-  filter(measure == "CMS") %>%
-  ggplot(aes(
+  filter(measure == "CMS")
+p3bw <- bw.nrd0(p3data$value[p3data$sameland & p3data$study == "Houston FSI G1"])
+p3 <- ggplot(p3data, aes(
     x = value, fill = sameland,
@@
   )) +
-  geom_density(alpha = 0.6) +
-#  geom_rug(alpha = 0.5/12, show.legend = F) +
+  geom_density(alpha = 0.6, bw = p3bw) +
   facet_grid(measure ~ study) +

### 2019-12-09 index.Rnw
Add CMS explanations, brevity

@@
 \section{Evaluating the Random Forest Algorithm}
-The random forest SAM scores correctly separate known matches from known non-matches, using SAM scores to aggregate land-to-land scores into a bullet-to-bullet comparison.
+The random forest scores correctly separate known matches from known non-matches, using SAM scores to aggregate land-to-land scores into a bullet-to-bullet comparison.
 With no errors at the bullet-to-bullet level, we can now evaluate the scores in light of (R1) and (R2).
-In this section, we will compare the random forest scores to cross correlation scores; cross correlation is one of the components of the random forest, but has also been used to quantify the similarity of two signatures in its own right \citep{chuPilotStudyAutomated2010}.
+In this section, we will primarily compare the random forest scores to cross correlation scores; cross correlation is one of the components of the random forest, but has also been used to quantify the similarity of two signatures in its own right \citep{chuPilotStudyAutomated2010}.
 In addition, both scoring methods use the same scale (0 to 1) and are continuous along that interval; other evaluation methods, such as consecutive matching striae, are discrete and more difficult to directly compare to continuous measurements.
@@
 Color indicates ground truth.
-Small vertical lines below the $x$-axis indicate observed scores in a **rug-plot**.
+Small vertical lines below the $x$-axis indicate observed scores in a \emph{rug-plot}.
 The plot shows that for all three case studies and both continuous measures, i.e.\ SAM scores based on cross correlation as well as SAM scores based on the random forest scores, all similarity scores are higher for pairs of bullets from the same source, i.e.\ bullets fired through the same barrel, than pairs of bullets from different sources (fired through different barrels).
+There is considerably more ambiguity in the SAM CMS scores, where the same-source distribution significantly overlaps the different-source distribution.
 
@@
 
-According to \autoref{fig:compare}, neither cross correlation nor random forest score fulfill both of the minimal requirements laid out in the introduction.
+According to \autoref{fig:compare}, neither cross correlation nor random forest (RF) score fulfill both of the minimal requirements laid out in the introduction.
 Both measures fulfill (R1), i.e.\ all  scores  of  different-source pairs are lower than scores of same-source pairs for SAM scores based on cross correlation and RF scores.
@@
 This can be seen from the larger horizontal distance between the modes (peaks) of the density curves of RF based SAM scores compared to the cross correlation SAM scores.
-
-Regarding requirement (R2), \autoref{fig:compare} shows that a single cutoff value does not exist that separates same-source scores from different-source scores across all three studies without introducing errors.
+Given the overlap between the distributions of SAM CMS scores, the scores fail to meet (R1) in that there is not complete separation.
+
+Regarding requirement (R2), \autoref{fig:compare} shows that a single cutoff value does not exist that separates same-source scores from different-source scores across all three studies and all three measures without introducing errors.
 
@@
 The downside of either of these options is that we would need to considerably expand the database used for training.
-Another solution would be to augment the aggregation used to get from land-to-land scores to bullet scores: SAM scores only take the scores of the maximum sequence into account and ignore scores associated with pairs of land engraved areas from different sources.
+Another solution would be to augment the aggregation used to get from land-to-land scores to bullet scores: SAM scores only take the scores of the maximum sequence into account and ignore scores associated with pairs of land engraved areas which are off the matching sequence.
 Those scores might be useful in establishing a baseline that better expresses similarity between pairs of bullets.

### 2019-12-09 index.Rnw
tightening up language and fixing dangling sentences.

@@
 In many cases, these approaches also provide some level of automation of the comparison process, with the goal of reducing human biases by augmenting the visual information with 3D measurements.
-Utilizing the 3D measurements allows for examination of both peaks and valleys in LEAs, it also allows us to take the depth of striae into account, something that is difficult, if not impossible, from visual inspection.
+Utilizing the 3D measurements allows for examination of both peaks and valleys in LEAs.
+It also allows us to take the depth of striae into account; something that is difficult, if not impossible, from visual inspection.
 
@@
 This can be seen from the larger horizontal distance between the modes (peaks) of the density curves of RF based SAM scores compared to the cross correlation SAM scores.
-Given the overlap between the distributions of SAM CMS scores, the scores fail to meet (R1) in that there is not complete separation.
-
-Regarding requirement (R2), \autoref{fig:compare} shows that a single cutoff value does not exist that separates same-source scores from different-source scores across all three studies and all three measures without introducing errors.
+
+Regarding requirement (R2), \autoref{fig:compare} shows that a single cutoff value does not exist that separates same-source scores from different-source scores across all three studies and all three scoring measures.
 
@@
 
-Several lands in Hamby set 44 have major deficiencies,  such as 'tank rash' (a collision of the bullet after exiting the barrel with a surface causing markings on top of the striations from the barrel) or extreme pitting (holes caused by direct contact with burning gun powder).
-\autoref{fig:tankrash} gives an overview of all lands of affected bullets in the Hamby-44 set.
+Several lands in Hamby set 44 have major deficiencies,  such as `tank rash' (a collision of the bullet after exiting the barrel with a surface causing markings on top of the striations from the barrel) or extreme pitting (holes caused by direct contact with burning gun powder).
+\autoref{fig:tankrash} shows rendered scans of all affected lands in the Hamby-44 set.
 These issues affect the algorithm's ability to identify a stable signature, which impacts the subsequent similarity scores at the land-to-land and bullet-to-bullet level.
@@
 When evaluating classifier performance, it is common to use a receiver operating characteristic (ROC) curve, which plots the percent of wrong identifications against the percent correct identifications for each possible value of the cutoff between the two classes (in this case, known match and known non-match).
-Land-to-land scores are not perfectly separated, we can therefore use ROC curves to distinguish between the performance of the different methods and studies.
+As the land-to-land scores are not perfectly separated, we can use ROC curves to distinguish between the performance of the different methods and studies.
 The ROC curves for the land-to-land scores are shown in \autoref{fig:roc-auc} (a).
@@
 As seen in the ROC curves, the algorithm performs the least well on set 2 of the Houston FSI study.
-Based on AUC, the overall performance on all sets is very good to excellent for cross correlation and the random forest score.
+Based on AUC, the overall performance on all sets is very good to excellent for both the cross correlation and the random forest score, and moderate for CMS scores.
 At the land level there is no significant difference in prediction power between RF score and cross correlation.", cache=FALSE>>=
@@
 AUC values are useful for differentiating between poor, good, and excellent model performance, but are not particularly useful when determining which of several models with approximately the same level of performance should be used \citep{marzban_roc_2004}.
-Hamby 44 and Phoenix PD are the only studies with a somewhat noticeable difference between the AUC score for RF compared to cross correlation, but even those differences are not statistically significant.
+%Hamby 44 and Phoenix PD are the only studies with a somewhat noticeable difference between the AUC score for RF compared to cross correlation, but even those differences are not statistically significant.
 
 \autoref{fig:eer} shows an overview of Equal Error and their thresholds based on the ROC curves of \autoref{fig:roc-auc}.
-Equal Errors are errors when sensitivity and specificity of a test are equal, i.e.\ we see the same percentage of missed and wrong identifications for a land-to-land comparison.
+Equal error rates are the error rates when the sensitivity and specificity of a test are equal, i.e.\ we see the same percentage of missed and wrong identifications for a land-to-land comparison.
 Equal errors based on CMS of two lands are significantly higher than equal errors based on cross correlation or the RF score.
 At best, RF score and cross correlation have an equal error of around 5\% for land-to-land comparisons.
-Interestingly, for consecutively matching striae, equal errors are reached for just two matching striae, which from a practical point of view is not sustainable.
+% Interestingly, for consecutively matching striae, equal errors are reached for just two matching striae, which from a practical point of view is not sustainable.
 A  CMS of 6 or above is an established cutoff used in identifications; when applied to the land-to-land comparisons, it is much more likely to result in a missed identification than a wrong identification.
@@
 
-At the beginning of this study, we anticipated Hamby 44 would be the easiest set to work with because of its similarity to the sets used to train the random forest algorithm.
-In a surprise turn of events, it was the hardest to solve, in part because of damage to the bullets that obscured LEA striae.
+At the beginning of this study, we anticipated Hamby 44 would be the easiest set to evaluate because of its similarity to the sets used to train the random forest algorithm.
+In a surprise turn of events, it was the hardest, in part because of damage to the bullets that obscured LEA striae.
 It has been shown \citet{lpr2} that parts of lands can be used for successful identifications, if at least 50\% of the land is present (using full-length scores as the reference distribution).
@@
 The random forest algorithm is trained on scans taken at the land level, so barrels with polygonal rifling, such as Glocks, which do not introduce well defined lands on bullets, cannot be compared using the random forest algorithm examined here.
+In addition, we plan to re-fit the random forest using a wider set of scans taken at different resolutions, with different ammunition, fired from barrels of different models.
+We expect that a random forest trained on a wider set of input data will be more likely to produce scores which meet the criteria for monotonicity and stability.
 
@@
 The algorithm's score is not intended to replace an examiner's judgment; instead, it provides a complementary tool to measure, quantify, and compare the similarity of two bullets in an identification.
-With more research into the behavior of score-based likelihood ratios, this method may also be used to compute
+With more research into the behavior of score-based likelihood ratios, the random forest score may also be used to compute a SLR to explicitly quantify the strength of the match between two bullets relative to other matches in the population.
 By validating the algorithm on external test sets, we have demonstrated that the method can be generalized to different types of ammunition and is not overly sensitive to small differences in rifling procedure.

### 2019-12-09 Reviewer_response.Rmd
Updated reviewer response...

--- 2019-12-06 Reviewer_response.Rmd
@@
 We thank the reviewer for his or her careful and constructive comments on our article, which we have retitled "Comparison of three similarity scores for bullet LEA matching".
-We have made several changes to the document in response to the suggestions made by the reviewer, and have responses to the comments which could not be easily addressed in this paper as it stands.
+We have made several changes to the document in response to the suggestions made by the reviewer, and have responses to the comments which we could not address according to the suggestions.
+We will first address the minor changes and then discuss some of the broader requested changes.
 
-The figures have been adjusted to have a larger font size for labels and titles; the caption size is set by the LaTeX template and is thus not easily editable by the authors.
+The figures have been adjusted to have a larger font size for labels and titles.
+The caption size is set by the LaTeX template and is thus not easily editable by the authors.
 We expect that the captions will be sized appropriately in the final product; at the moment, we do not have control over the sizing of the figure captions.
+We have added CMS to Figures 7 and 9; the ommission was an oversight that we are glad to have a chance to remedy.
 
-The choice to exclude score-based likelihood ratios was a conscious choice because the land-to-land comparisons result in density estimates with different variability between known matches and known non-matches.
+We have modified the abstract in accordance with the reviewer's suggestions.
+We have also (throughout the paper) changed the language to distinguish between similarity metrics and match strength; we appreciate the importance of this clarification.
+
+We created a replacement for Figure 1 with the signatures overlaid, but found it difficult to read because the sequences are not (at this stage of the algorithm) aligned, so much of the similarity is difficult to see when the unaligned details are overlaid.
+We include the modified Figure 1 here as a demonstration, but ultimately feel that the original Fig 1 (with labels of increased size) allows for easier visual comparison of the peaks in the sequence.
+
+![](fig1-alt.png)
+
+As requested, we have added clarifying language and descriptions when referencing the matching algorithm, and in some cases made the references more explicit, such as when describing peak identification for CMS calculation.
+We have included references to BulletTrax and other full-bullet scanning systems.
+In addition, we have made efforts to reduce the wordiness of some explanations while maintaining sufficient detail to describe the process used to automatically match two bullets.
+
+The suggestions addressed thus far required relatively minor changes to the manuscript; the suggestions and comments we address next are somewhat more substantive.
+We have addressed these comments but would like to provide additional context for the way we have approached each suggestion.
+
+The choice to exclude score-based likelihood ratios from the scope of this article was a conscious decision, as the land-to-land comparisons produce density estimates with different variability between known matches and known non-matches.
 As a result, in these cases, the SLR is nonmonotonic and produces unexpected and potentially misleading results at the extremes of the distribution.
-As a result, we would like to defer the explicit use of SLRs in this paper; we have added a paragraph on the potential of using likelihood ratios with the scoring methods described in the paper.
+We would like to defer the explicit use of SLRs; we have added a paragraph on the potential of using likelihood ratios with the scoring methods described in the paper, as well as several mentions of how this method could be used with LRs in theory.
 The threshold-based model we used in this paper is more consistent with the model used by examiners in our jurisdiction: decisions include, broadly, identification (the bullets match), exclusion (the bullets do not match), inconclusive, and insufficient evidence to make a conclusion.
 Our computational model does not require the last two categories, reducing this problem to a binary decision with a single threshold.
-We would also note that using likelihood ratios does not remove the need for a threshold; rather, it moves (implicitly or explicitly) to a threshold of 1, or 10, or whatever the criteria for "enough evidence" is set to.
+We would also note that using likelihood ratios does not remove the need for a threshold; rather, it moves (implicitly or explicitly) to a threshold of 1, where scores greater than 1 indicate more evidence of a match than not, and scores less than 1 indicate more evidence of a non-match than a match.
 
-The reviewer suggests using multiplication to combine land-to-land SLRs into a bullet-to-bullet SLR; we cannot recommend this course of action in part because the land to land scores are not independent, rather, they result from pairwise comparisons that are inherently dependent.
+As to the suggestion to use multiplication to combine land-to-land SLRs into a bullet-to-bullet SLR; we cannot recommend this course of action in part because the land to land scores are not independent, rather, they result from pairwise comparisons that are inherently dependent.
 Further, we can't assume that striae observed on different lands from the same physical bullet are independent, because they were all part of the same firing event.
@@
 non-Ruger bullets; we have included some of the suggestions in the paper, but do not currently have sufficient data to include a fourth case study in this paper.
-We strongly agree that this would be an interesting test of the random forest score compared to the cross correlation and are planning several future studies which explore the random forest score's use on non-Ruger bullets.
-In addition, the Houston FSI study was also used to conduct examiner studies; we are preparing a paper which compares the algorithm's performance using the decision criteria shown in this paper to the performance of firearm examiners and expect to submit this paper for review in mid 2020.
+We agree that this would be an interesting test of the random forest score compared to the cross correlation and are planning several future studies which explore the random forest score's use on non-Ruger bullets.
+As this paper is already fairly lengthy, as noted by the reviewer, we feel that the addition of a non-Ruger case study is best reserved for one of these future papers.
+Regarding the suggestion to compare decisions made by the algorithm to decisions made by the examiner, the Houston FSI study data shown here were part of a larger examiner study; we are preparing a paper which compares the algorithm's performance using the decision criteria shown in this paper to the performance of firearm examiners and expect to submit this paper for review in mid 2020.
 
-We are also working on gaining approval to publish a data in brief paper with the data used in this paper, so that we can make the land scans publicly available for additional analysis; we strongly agree with the reviewer that the community would benefit from access to these scans.
-Two of the three sets have been uploaded to NBTRD at a lower resolution than the resolution used in our calculations.
+The final wider-scope comment addresses the matching of partial signatures, specifically suggesting that these signatures should be compared to a reference distribution assembled from partial signature comparisons.
+Hare, Hofmann, and Carriquiry (2017) showed how to use the random forest score to match degraded and partial land impressions; the random forest score remained fairly stable up to 50% signature length, suggesting that the random forest score, at least, is relatively robust to degradation and partial mark size.
+For the reasons stated above, we do not carry this conclusion all the way to the calculation of SLRs, but given the previously cited paper, we do not expect to require a separate comparison distribution when the partial LEA includes at least 50% of the length of a full LEA signature.
+
+We are working on gaining approval to publish a data in brief paper with the data used in this paper, to make the land scans publicly available for additional analysis; we strongly agree with the reviewer that the community would benefit from access to these scans.
+Two of the three sets have been uploaded to NBTRD at a lower resolution than the resolution used in our calculations, but before we provide bulk downloads of the higher-resolution data, we must consult with CSAFE collaborators at NIST.

### 2019-12-09 Reviewer_response.Rmd
Response to error rate

@@
 
+Finally, we have emphasized that at the bullet-to-bullet level, the distributions for each study are perfectly separated, indicating that the error rate would be nil.
+It is only when we proceed to analyze the land-to-land comparisons that we can do any sort of assessment of the error rate.
+<!-- I actually didn't change anything, but if we suggest we emphasized it, I think it will make it more obvious to the reviewer without calling it out directly  -->
+
 The suggestions addressed thus far required relatively minor changes to the manuscript; the suggestions and comments we address next are somewhat more substantive.

### 2019-12-13 title_page.tex
Update authors on title page

@@
-\documentclass[titlepage]{article}
+\documentclass{article}
+\usepackage{fullpage}
+
 \title{Case study validations of automatic bullet matching}
-\author{Susan Vanderplas and Heike Hofmann}
-\date{Center for Statistics and Applications in Forensic Evidence, Iowa State University. 195 Durham Center, 613 Morill Rd., Ames, IA 50011}
+\author{Susan Vanderplas$^\ast$$^1$, Melissa Nally$^2$, Tylor Klep$^3$, Cristina Cadevall$^4$, Heike Hofmann$^1$}
+
+\date{$^\ast$ Corresponding author. email: srvander@iastate.edu\\[1em]
+$^1$Center for Statistics and Applications in Forensic Evidence\\Iowa State University\\
+195 Durham Center, 613 Morill Rd., Ames, IA 50011\\[1em]
+$^2$Houston Forensic Science Center\\500 Jefferson St 13th Floor, Houston, TX 77002\\[1em]
+$^3$Phoenix Police Department\\Laboratory Services Bureau\\621 W Washington St, Phoenix, AZ 85003\\[1em]
+$^4$Sensofar Group\\Parc Audiovisual de Catalunya Ctra, BV-1274, Km 1, 08225 Terrassa, Barcelona, Spain}
 

### 2019-12-13 Reviewer_response.Rmd
Danica changes

@@
 Our computational model does not require the last two categories, reducing this problem to a binary decision with a single threshold.
-We would also note that using likelihood ratios does not remove the need for a threshold; rather, it moves (implicitly or explicitly) to a threshold of 1, where scores greater than 1 indicate more evidence of a match than not, and scores less than 1 indicate more evidence of a non-match than a match.
+<!-- We would also note that using likelihood ratios does not remove the need for a threshold; rather, it moves (implicitly or explicitly) to a threshold of 1, where scores greater than 1 indicate more evidence of a match than not, and scores less than 1 indicate more evidence of a non-match than a match.
+-->
 
@@
 As this paper is already fairly lengthy, as noted by the reviewer, we feel that the addition of a non-Ruger case study is best reserved for one of these future papers.
-Regarding the suggestion to compare decisions made by the algorithm to decisions made by the examiner, the Houston FSI study data shown here were part of a larger examiner study; we are preparing a paper which compares the algorithm's performance using the decision criteria shown in this paper to the performance of firearm examiners and expect to submit this paper for review in mid 2020.
+Regarding the suggestion to compare decisions made by the algorithm to decisions made by the examiner, the Houston FSC study data shown here were part of a larger examiner study; we are preparing a paper which compares the algorithm's performance using the decision criteria shown in this paper to the performance of firearm examiners and expect to submit this paper for review in mid 2020.
 

### 2019-12-13 Reviewer_response.Rmd
Updates and figshare links

@@
 We thank the reviewer for his or her careful and constructive comments on our article, which we have retitled "Comparison of three similarity scores for bullet LEA matching".
-We have made several changes to the document in response to the suggestions made by the reviewer, and have responses to the comments which we could not address according to the suggestions.
-We will first address the minor changes and then discuss some of the broader requested changes.
+We have made numerous changes to the document in response to the suggestions made by the reviewer, and have responses to the comments which we could not address according to the suggestions.
+Rather than address each point in turn, we have summarized the requested changes, as some of the issues appear in several separate comments.
+We first address the minor changes and then discuss some of the broader, more conceptual comments.
+We would be happy to provide a point by point response to the reviewer's suggestions, if this format does not work.
 
+\paragraph{Figures}
 The figures have been adjusted to have a larger font size for labels and titles.
@@
 We have added CMS to Figures 7 and 9; the ommission was an oversight that we are glad to have a chance to remedy.
-
-We have modified the abstract in accordance with the reviewer's suggestions.
-We have also (throughout the paper) changed the language to distinguish between similarity metrics and match strength; we appreciate the importance of this clarification.
 
@@
 
+\paragraph{Minor Narrative Changes}
+
+We have modified the abstract in accordance with the reviewer's suggestions.
+We have also (throughout the paper) changed the language to distinguish between similarity metrics and match strength; we appreciate the importance of this clarification.
 As requested, we have added clarifying language and descriptions when referencing the matching algorithm, and in some cases made the references more explicit, such as when describing peak identification for CMS calculation.
@@
 
+\paragraph{Likelihood Ratios}
 The choice to exclude score-based likelihood ratios from the scope of this article was a conscious decision, as the land-to-land comparisons produce density estimates with different variability between known matches and known non-matches.
-As a result, in these cases, the SLR is nonmonotonic and produces unexpected and potentially misleading results at the extremes of the distribution.
+As a result, in these cases, the SLR is nonmonotonic and produces unexpected and potentially misleading results at the extremes of the distribution (see Fig.
+4 in Neumann and Ausdemore, https://arxiv.org/pdf/1910.05240.pdf).
 We would like to defer the explicit use of SLRs; we have added a paragraph on the potential of using likelihood ratios with the scoring methods described in the paper, as well as several mentions of how this method could be used with LRs in theory.
-The threshold-based model we used in this paper is more consistent with the model used by examiners in our jurisdiction: decisions include, broadly, identification (the bullets match), exclusion (the bullets do not match), inconclusive, and insufficient evidence to make a conclusion.
+<!-- The threshold-based model we used in this paper is more consistent with the model used by examiners in our jurisdiction: decisions include, broadly, identification (the bullets match), exclusion (the bullets do not match), inconclusive, and insufficient evidence to make a conclusion.
 Our computational model does not require the last two categories, reducing this problem to a binary decision with a single threshold.
+-->
 <!-- We would also note that using likelihood ratios does not remove the need for a threshold; rather, it moves (implicitly or explicitly) to a threshold of 1, where scores greater than 1 indicate more evidence of a match than not, and scores less than 1 indicate more evidence of a non-match than a match.
@@
 
+\paragraph{Suggestions for Expanded Scope}
 We appreciate the reviewer's discussion about the Ruger vs.
@@
 
+\paragraph{Data in Brief}
 We are working on gaining approval to publish a data in brief paper with the data used in this paper, to make the land scans publicly available for additional analysis; we strongly agree with the reviewer that the community would benefit from access to these scans.
-Two of the three sets have been uploaded to NBTRD at a lower resolution than the resolution used in our calculations, but before we provide bulk downloads of the higher-resolution data, we must consult with CSAFE collaborators at NIST.
+At the moment, two of the three sets have been uploaded to NBTRD at a lower resolution than the resolution used in our calculations, but before we provide bulk downloads of the higher-resolution data, we must consult with CSAFE collaborators at NIST and other institutions.
+We are in the process of creating a [metadata-only figshare collection](https://doi.org/10.25380/iastate.c.4781447) with contact information to request the data (it will be available at once it is released by our librarians, but can be seen [here](https://figshare.com/s/bb22f76c9cb196a4f9d3) until that approval is granted).
+Once we have collaborator approval, we hope to submit a Data in Brief paper and upload the full datasets for public consumption.

### 2019-12-24 index-blinded.Rnw
Required documents for upload

@@
 
-\title{Case study validations of automatic bullet matching}%\tnoteref{t1}\tnoteref{t2}}
+\title{Comparison of three similarity scores for bullet LEA matching
+%\tnoteref{t1}\tnoteref{t2}
+}
 
@@
 % \tnotetext[t2]{The authors wish to thank Alan Zheng (NIST), Tylor Klep (Phoenix PD), and Melissa Nally and Kasi Kirksey (FSI Houston) for access to test set bullets.
-We would also like to thank the efforts of the Roy J Carver High Resolution lab in scanning the bullet sets and providing the scans to us.
-Finally, thanks to Prof James Hamby for his advice on XXX.}
+We would also like to thank the efforts of the Roy J Carver High Resolution lab in scanning the bullet sets and providing the scans to us.}
 
@@
 
-\author[isu,csafe]{Heike Hofmann}
-\cortext[corauthor]{Corresponding author}\ead{srvander@iastate.edu}
-\author[isu,csafe]{Susan Vanderplas\corref{corauthor}}
-\address[isu]{Statistics Department, Iowa State University\\2438 Osborne Dr, Ames, IA 50011}
-\address[csafe]{Center for Statistical Applications in Forensic Evidence, Iowa State University\\613 Morrill Rd, Ames, IA 50011}
+% \author[isu,csafe]{Heike Hofmann}
+% \cortext[corauthor]{Corresponding author}\ead{srvander@iastate.edu}
+% \author[isu,csafe]{Susan Vanderplas\corref{corauthor}}
+% \address[isu]{Statistics Department, Iowa State University\\2438 Osborne Dr, Ames, IA 50011}
+% \address[csafe]{Center for Statistical Applications in Forensic Evidence, Iowa State University\\613 Morrill Rd, Ames, IA 50011}
+%
 
 \begin{abstract}
-Recent advances in microscopy have made it possible to collect 3D topographic data, enabling virtual comparisons based on the collected 3D data next to traditional comparison microscopy.
-Automatic matching algorithms have been introduced for various scenarios, such as matching cartridge cases \citep{tai2018fully} or matching bullet striae \citep{aoas, chuAutomaticIdentificationBullet2013, dekinderAutomatedComparisonsBullet1999}.
-One key aspect of validating automatic matching algorithms is to evaluate the performance of the algorithm on external tests.
-Here, we are presenting a discussion of the performance of the matching algorithm \citep{aoas} in three studies.
-We are considering matching performance based on the Random forest score, cross correlation, and consecutive matching striae (CMS) at the land-to-land level and, using Sequential Average Maxima scores, also at the bullet-to bullet level.
-Cross correlation and Random Forest scores both result in perfect discrimination of same-source and different-source bullets.
-At the land-to-land level, discrimination (based on area under the curve, AUC) is excellent ($\ge$ 0.90).
+Recent advances in microscopy have made it possible to collect 3D topographic data, enabling virtual comparisons based on the collected 3D data as a supplement to traditional comparison microscopy and 2D photography.
+Automatic comparison algorithms have been introduced for various scenarios, such as matching cartridge cases \citep{tai2018fully,song3DTopographyMeasurements2014} or matching bullet striae \citep{aoas2, chuAutomaticIdentificationBullet2013, dekinderAutomatedComparisonsBullet1999}.
+One key aspect of validating these automatic comparison algorithms is to evaluate the performance of the algorithm on external tests, that is, using data which were not used to train the algorithm.
+Here, we present a discussion of the performance of the matching algorithm \citep{aoas} in three studies conducted using different Ruger weapons.
+We consider the performance of three scoring measures: random forest score, cross correlation, and consecutive matching striae (CMS) at the land-to-land level and, using Sequential Average Maxima scores, also at the bullet-to bullet level.
+Cross correlation and random forest scores both result in perfect discrimination of same-source and different-source bullets.
+At the land-to-land level, discrimination for both cross correlation and random forest scores (based on area under the curve, AUC) is excellent ($\ge$ 0.90).
 \end{abstract}
@@
 \begin{itemize}
-\item[(a)] Identify an area on the LEA with expressed striae. Locate a stable crosscut in the identified area.
+\item[(a)] Identify an area on the LEA with expressed striae.
+Locate a stable crosscut in the identified area; that is, an area where the striae are similar in the region above and below the crosscut.
+This excludes regions with extreme pitting, breakoff, tank rash, and other flaws that would interfere with similarity score calculation.
 \item[(b)] Discard extraneous/contaminated data, such as data from groove engraved area and areas affected by break-off or contact with objects after the bullet exited the barrel  (``tank rash").
@@
 
-Once signatures have been extracted from stable regions with expressed striae, they must be aligned in order to assess their match strength.
-Maximized cross correlation is used to pair-wise align signatures \citep{aoas,vorburgerApplicationsCrosscorrelationFunctions2011}.
+Once signatures have been extracted from stable regions with expressed striae, they must be aligned in order to assess their similarity, just as examiners would manually align two bullets under a comparison microscope.
+Maximized cross correlation is used to pair-wise align signatures \citep{aoas2,vorburgerApplicationsCrosscorrelationFunctions2011}.
 
@@
 Using the lag determined for aligning two signatures, other quantitative features describing the similarity of the two signatures can be extracted.
-Numerical features such as cross correlation and Euclidean distance
-can be computed from the aligned signatures alone.
-Other features include characteristics of the signature that firearms and toolmark examiners base their visual assessment on, such as striae depth, number of matching striae, and the number of consecutively matching striae (CMS), one of the measures firearms examiners use to describe the strength of a match quantitatively.
+Numerical features such as cross correlation and Euclidean distance can be computed from the aligned signatures alone.
+Additional features are modeled after visual assessment methods used by examiners, including striae depth, total number of matching striae, and the number of consecutively matching striae (CMS).
 
 In order to evaluate CMS, we must first identify peaks and valleys in each of the signatures, then determine whether these peaks and valleys overlap sufficiently.
-Features which firearms examiners assess visually depend on the identification of these extrema.
-If there are at least six consecutive matching striae, the bullets are considered to be a good match in practice \citep{biasotti_firearms_1997,nichols_scientific_2006}.
-CMS in this setting are counted as the number of consecutively matching peaks in signatures of two aligned lands.
-Peaks -- rather than peaks and valleys -- are used for this determination in part because it is difficult to visually assess valleys except as relative to peaks.
-Note that with automatic alignment based on the maximum cross correlation rather than matching striation marks, it is possible for two lands to have a consecutive matching striae of zero.
+Features which firearms examiners assess visually generally depend on the identification of these extrema, as striation marks and the corresponding peaks are the most salient feature when viewing a bullet using a microscope.
+If there are at least six consecutive matching striae, the bullets are considered to be similar in practice \citep{biasotti_firearms_1997,nichols_scientific_2006}.
+CMS, as measured by examiners, is the number of consecutively matching peaks in signatures of two aligned lands, and, because examiner typically lines up two distinctive markings, resulting in at least one matching striation mark.
+Peaks -- rather than peaks and valleys -- are used in part because it is difficult to visually assess valleys except as relative to peaks.
+In contrast, the matching algorithm can identify both peaks and valleys, and signatures are aligned based on maximum cross-correlation (as opposed to matching striation marks), there is the possibility that two lands will have zero consecutive matching striae.
+Thus, the matching algorithm uses a slightly different measure of CMS than examiners, but the underlying principle is the same.
+The algorithms used to identify peaks and valleys in each signature and determine CMS from the peak and valley identifications are described in more detail in \citep[p.
+2340, step 3 -- 4]{aoas}.
 
@@
 The random forest presented in \citet{aoas} is based on a combination of multiple characteristics, including cross correlation, number of matching striae, and number of consecutively matching striae.
-The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where scores close to 1 indicate it to be highly likely that the aligned signatures come from the same source, while scores close to 0 indicate that the aligned signatures are highly likely to originate from different sources.
+The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the similarity of the two signatures, where scores close to 1 indicate it to be highly likely that the aligned signatures come from the same source, while scores close to 0 indicate that the aligned signatures are highly likely to originate from different sources.
 
@@
 This increased rigor may increase the error rate of the model when compared to examiner error rates, but we expect that the increased information available from the 3D measurement data will compensate for this loss.
-A binary decision model is also easier to interpret and provides more clear-cut, definitive results.
-
-We can enumerate characteristics of an ideal similarity scoring mechanism: 
+A binary decision model is also easier to interpret and provides more clear-cut, definitive results than the AFTE Theory of Identification system used in most forensic laboratories in the United States.
+Of course, in jurisdictions which utilize score-based likelihood ratios, this threshold system is not necessary because likelihood ratios are continuous.
+Even with likelihood ratios, however, there is a natural threshold at 1 which functions similarly to the selected threshold in the binary decision case.
+
+The remainder of this paper will utilize the binary decision model which is compatible with the legal framework commonly used in US jurisdictions.
+With this model, we can enumerate characteristics of an ideal similarity scoring mechanism:
+
 \begin{description}
 \item[(R1)] {\bf Monotonicity:} a higher score is indicative of higher similarity between a pair of bullets, in particular, similarity scores of same-source pairs of bullets are higher than different-source pairs.
-\item[(R2)] {\bf Stability:} the same score leads to the same conclusion. 
+\item[(R2)] {\bf Stability:} the same score leads to the same conclusion in all situations and under separately assembled reference distributions. 
 \end{description}
@@
 While these comparisons are useful, the question of interest typically involves the entire physical object (e.g.
-all lands on a single bullet), and a conclusion on same or different source should be reached based on the evidence of all lands of the bullets.
-
-<<echo=FALSE, results = 'hide', warning = FALSE, message = FALSE>>=
+all lands on a single bullet), and a conclusion on same or different source should be reached based on the evidence of all lands of the two bullets.
+
+<<demo-l2lscores, echo=FALSE, results = 'hide', warning = FALSE, message = FALSE>>=
 features <- read.csv("data/fsi-features.csv.gz")
@@
 )
-p1 <- b12 %>%
-  ggplot(aes(x = land1, y = land2, fill = rfscore)) +
-  geom_tile() + ggtitle("Same barrel") +
-  scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
-  theme_bw() +
-  xlab("Bullet 2") + ylab("Bullet 1") +
-  theme(legend.position = "bottom") + coord_equal()
-
+# 
+# p1 <- b12 %>%
+#   ggplot(aes(x = land1, y = land2, fill = rfscore)) +
+#   geom_tile() + ggtitle("Same barrel") +
+#   scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
+#   theme_bw() +
+#   xlab("Bullet 2") + ylab("Bullet 1") +
+#   theme(legend.position = "bottom") + coord_equal()
+# 
 bab <- features %>% filter(
@@
 )
-p2 <- bab %>%
-  ggplot(aes(x = land1, y = land2, fill = rfscore)) +
-  geom_tile() + ggtitle("Different barrel") +
-  scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
-  theme_bw() +
-  xlab("Bullet 2") + ylab("Bullet 1") +
-  theme(legend.position = "bottom") + coord_equal()
-@
-
-Land-to-land comparisons lead to a whole set of scores for bullet-to-bullet comparisons.
+# p2 <- bab %>%
+#   ggplot(aes(x = land1, y = land2, fill = rfscore)) +
+#   geom_tile() + ggtitle("Different barrel") +
+#   scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
+#   theme_bw() +
+#   xlab("Bullet 2") + ylab("Bullet 1") +
+#   theme(legend.position = "bottom") + coord_equal()
+
+l2lscores <- bind_rows(
+  mutate(b12, type = "Same barrel"),
+  mutate(bab, type = "Different barrel")
+) %>% 
+  mutate(type = factor(type, levels = c("Same barrel", "Different barrel"), ordered = T))
+
+@
+
+Land-to-land comparisons lead to a whole set of scores to evaluate for bullet-to-bullet comparisons.
 \autoref{fig:b2b} shows two matrices of scores for two pairs of bullet-to-bullet matches.
@@
 On the left the two bullets are known to come from the same source (barrel) on the right, the bullets are from two different sources (barrels).">>=
-grid.arrange(p1, p2, ncol = 2, padding = 2)
-@
-
-When imaging bullets, operators scan one land at a time in a clockwise (left twisted rifling) or anti-clockwise (right twisted rifling) sequence.
+# grid.arrange(p1, p2, ncol = 2, padding = 2)
+ggplot(data = l2lscores, aes(x = land1, y = land2, fill = rfscore)) +
+  geom_tile() + 
+  scale_fill_gradient2("Random Forest\n(RF) score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
+  theme_bw() +
+  xlab("Bullet 2") + ylab("Bullet 1") +
+  theme(legend.position = "right") + coord_equal() + facet_wrap(~type)
+@
+
+While some imaging systems, such as BulletTrax 3D\footnote{https://www.ultra-forensictechnology.com/en/our-products/ballistic-identification/bullettrax/} or BalScan\footnote{https://www.forensic.cz/en/products/balscan}, may capture lower-resolution scans of an entire bullet at one time, confocal light microscopes do not have 360$^\circ$ capability.
+Thus, when imaging bullets using a confocal light microscope, operators scan one land at a time in a clockwise (left twisted rifling) or anti-clockwise (right twisted rifling) sequence.
 The order in which scans are acquired is kept as meta-information.
+
 Let us assume that lands on a bullet are labelled $\ell_i$ with $i = 1, .., p$, where $p$ indicates the number of lands a bullet has, as determined by the rifling of a barrel.
-For all of the bullets considered here the number of lands $p$ is 6.
+For all of the bullets considered here the number of lands, $p$, is 6.
 A match between  a pair of lands from two bullets therefore results in an expected additional $p-1$ matches between pairs of lands.
@@
 
-In this paper, we validate the algorithm described in \citet{aoas} on three external test sets.
-We assess the results on each set and evaluate whether the minimal requirements described above are fulfilled for bullet-to-bullet SAM scores.
-We compare the random forest algorithm's performance to the performance of other suggested measures for quantitative assessment of match strength, such as cross correlation and consecutive matching striae.
-
-\section{Validation Sets}
-
 Random forests \citep{breiman} have a built-in internal testing mechanism to prevent potential overfitting.
-Errors reported by random forest algorithms are based on these internal test sets.
-However, there is some discussion on whether or in which situations these errors are biased \citep{rfOver2, rfOver}.
+Errors reported by random forest algorithms are based on these internal test sets, though there is some debate about the bias of these errors \citep{rfOver2, rfOver}.
 Neither of these papers addresses another issue with internal test sets: internal test sets are constructed to have the same distribution as the training data (apart from sampling variability).
@@
 
+In this paper, we validate the algorithm described in \citet{aoas} on three external test sets.
+We assess the results on each set and evaluate whether the minimal requirements described above are fulfilled for bullet-to-bullet SAM scores.
+We compare the random forest algorithm's performance to the performance of other suggested measures for quantitative assessment of bullet similarity, such as cross correlation and consecutive matching striae.
+
+\section{Validation Sets}
+
 The algorithm in \citet{aoas} is trained on scans from Hamby sets 252 and 173 made available through \citet{nist}.
-Set 173 was originally published as Hamby set 44.
-This mis-labeling has been corrected now.
+Set 173 was originally published as Hamby set 44, but the mis-labeling has been recently corrected.
 The scans used to train the model were taken at 20 fold magnification for a resolution of 1.5625 microns per pixel.
-While magnification is generally of interest in microscopy, **resolution** -- usually measured in microns per pixel -- is of more interest for 3D topographic measurements, as it determines the operative level of available data.
+While magnification is generally of interest in microscopy, \emph{resolution} -- usually measured in microns per pixel -- is of more interest for 3D topographic measurements, as it determines the operative level of available data.
 
@@
 Scans of all land engraved areas for the validation data were taken on a Sensofar Confocal Light microscope at 20x magnification resulting in a resolution of 0.645 microns per pixel.
-If not indicated otherwise, scans were taken at \texttt{Redacted}.
-% the Roy J Carver high resolution microscopy lab at Iowa State University.
-
-Case studies were chosen such that difficulty for the matching algorithm increases with each case study: Hamby set 44 is part of the Hamby study.
+If not indicated otherwise, scans were taken at the Roy J Carver high resolution microscopy lab at Iowa State University.
+
+Case studies were chosen such that theoretical difficulty for the matching algorithm increases with each case study: Hamby set 44 is part of the Hamby study.
 The algorithm in \citet{aoas} was trained on sets 173 and 252, so the bullets in Hamby 44 are of the same type of ammunition and are fired through the same barrels as the bullets in the training set.
@@
 h44nest <- h44nest %>%
+  ungroup() %>%
   mutate(
@@
 unknowns <- unknowns %>% filter(bullet1 != bullet2)
-unknowns <- unknowns %>% mutate(
+unknowns <- unknowns %>% ungroup() %>% mutate(
   bullet2 = gsub("Unknown 1-", "", bullet2),
@@
 
-pdnest <- pdnest %>%
+pdnest <- pdnest %>% ungroup() %>% 
   mutate(
@@
 
-hou1nest <- hounest %>%
+hou1nest <- hounest %>%ungroup() %>% 
   filter(group1 == "G1") %>%
@@
 
-hou2nest <- hounest %>%
+hou2nest <- hounest %>%ungroup() %>% 
   filter(group1 == "G2") %>%
@@
 
-hou3nest <- hounest %>%
+hou3nest <- hounest %>%ungroup() %>% 
   filter(group1 == "G3") %>%
@@
 \section{Evaluating the Random Forest Algorithm}
-The random forest SAM scores correctly separate known matches from known non-matches, using SAM scores to aggregate land-to-land scores into a bullet-to-bullet comparison.
+The random forest scores correctly separate known matches from known non-matches, using SAM scores to aggregate land-to-land scores into a bullet-to-bullet comparison.
 With no errors at the bullet-to-bullet level, we can now evaluate the scores in light of (R1) and (R2).
-In this section, we will compare the random forest scores to cross correlation scores; cross correlation is one of the components of the random forest, but has also been used to quantify the strength of a match in its own right \citep{chuPilotStudyAutomated2010}.
-In addition, both scoring methods use the same scale (0 to 1) and are continuous along that interval; other evaluation methods, such as consecutive matching striae are discrete and more difficult to directly compare to continuous measurements.
+In this section, we will primarily compare the random forest scores to cross correlation scores; cross correlation is one of the components of the random forest, but has also been used to quantify the similarity of two signatures in its own right \citep{chuPilotStudyAutomated2010}.
+In addition, both scoring methods use the same scale (0 to 1) and are continuous along that interval; other evaluation methods, such as consecutive matching striae, are discrete and more difficult to directly compare to continuous measurements.
 
@@
 Color indicates ground truth.
-Small vertical lines below the $x$-axis indicate observed scores in a **rug-plot**.
+Small vertical lines below the $x$-axis indicate observed scores in a \emph{rug-plot}.
 The plot shows that for all three case studies and both continuous measures, i.e.\ SAM scores based on cross correlation as well as SAM scores based on the random forest scores, all similarity scores are higher for pairs of bullets from the same source, i.e.\ bullets fired through the same barrel, than pairs of bullets from different sources (fired through different barrels).
-
-<<compare, fig.width = 8, fig.height = 5, fig.cap="Density curves of similarity scores from cross correlation (top) and RF scores (bottom).
+There is considerably more ambiguity in the SAM CMS scores, where the same-source distribution significantly overlaps the different-source distribution.
+
+<<compare, fig.width = 8, fig.height = 7, fig.cap="Density curves of similarity scores from random forest scores (top), cross correlation (middle), and consecutive matching striae (bottom).
 Different colors indicate same source versus different source.
@@
 nests <- rbind(
-  h44nest %>%
-    select(sam_rf, sam_ccf, samesource) %>%
+  h44nest %>% ungroup() %>% 
+    select(sam_rf, sam_ccf, sam_cms, samesource) %>%
     mutate(study = "Hamby 44", group1 = NA),
-  pdnest %>%
-    select(sam_rf, sam_ccf, samesource) %>%
+  pdnest %>%ungroup() %>% 
+    select(sam_rf, sam_ccf, sam_cms, samesource) %>%
     mutate(study = "Phoenix PD", group1 = NA),
-  hounest %>%
-    select(sam_rf, sam_ccf,  samesource, group1) %>%
+  hounest %>%ungroup() %>% 
+    select(sam_rf, sam_ccf, sam_cms, samesource, group1) %>%
     mutate(study = "Houston FSI")
@@
                data = labels
-  )
-@
-
-According to \autoref{fig:compare}, neither cross correlation nor random forest score fulfill both of the minimal requirements laid out in the introduction.
+  ) + 
+  theme(axis.title.x = element_blank(), axis.text.y = element_text(hjust = .5, angle = 90))
+
+p2 <- nestslong %>%
+  filter(measure %in% c("Cross correlation")) %>% 
+  ggplot(aes(
+    x = value, fill = samesource,
+    colour = samesource
+  )) +
+  geom_density(alpha = 0.6) +
+  geom_rug(alpha = 0.3, show.legend = F) +
+  facet_grid(measure ~ study) +
+  theme_bw() +
+  theme(legend.position = "bottom") +
+  scale_fill_manual("Same source", values = c("darkgrey", "darkorange"), guide = F) +
+  scale_colour_manual("Same source", values = c("darkgrey", "darkorange"), guide = F) +
+  scale_x_continuous("SAM scores") + 
+  theme(axis.title.x = element_blank(), axis.text.y = element_text(hjust = .5, angle = 90))
+
+p3 <- nestslong %>%
+  filter(measure %in% c("CMS")) %>% 
+  ggplot(aes(
+    x = value, fill = samesource,
+    colour = samesource
+  )) +
+  geom_density(alpha = 0.6) +
+  geom_rug(aes(x = jitter(value)), alpha = 0.3, show.legend = F) +
+  facet_grid(measure ~ study) +
+  theme_bw() +
+  theme(legend.position = "bottom") +
+  scale_fill_manual("Same source", values = c("darkgrey", "darkorange")) +
+  scale_colour_manual("Same source", values = c("darkgrey", "darkorange")) +
+  scale_x_continuous("SAM scores")  + 
+  theme(axis.text.y = element_text(hjust = .5, angle = 90))
+
+grid.arrange(p1, p2, p3, nrow = 3, heights = c(1, 1, 1.4))
+@
+
+According to \autoref{fig:compare}, neither cross correlation nor random forest (RF) score fulfill both of the minimal requirements laid out in the introduction.
 Both measures fulfill (R1), i.e.\ all  scores  of  different-source pairs are lower than scores of same-source pairs for SAM scores based on cross correlation and RF scores.
@@
 
-Regarding requirement (R2), \autoref{fig:compare} shows that a single cutoff value does not exist that separates same-source scores from different-source scores across all three studies without introducing errors.
+Regarding requirement (R2), \autoref{fig:compare} shows that a single cutoff value does not exist that separates same-source scores from different-source scores across all three studies and all three scoring measures.
 
@@
 There is some overlap between known matching and known non-matching land-to-land scores; in many cases the random forest scores for these overlapping values are more extreme than the corresponding cross correlation scores.
-This suggests that the matching algorithm is sensitive to the presence of LEAs which have low land-to-land match scores, such as the matches of bullets 'I' and 'F' in the Hamby 44 study.
+This suggests that the matching algorithm is sensitive to the presence of LEAs which have low land-to-land similarity scores, such as the matches of bullets 'I' and 'F' in the Hamby 44 study.
 Comparing this to the bullet-to-bullet score density plot shown in \autoref{fig:compare}, we see that there is much greater separation between the same-source and different-source densities for each study in the bullet-to-bullet comparisons than in the land-to-land comparisons, that is, one or two poorly matching lands does not prevent the bullet from showing a matching score, though if there are weak matches on several lands, such as the marked Hamby 44 comparisons in \autoref{fig:compare}, the overall score may be affected.
 
-Several lands in Hamby set 44 have major deficiencies,  such as 'tank rash' (a collision of the bullet after exiting the barrel with a surface causing markings on top of the striations from the barrel) or extreme pitting (holes caused by direct contact with burning gun powder).
-\autoref{fig:tankrash} gives an overview of all lands of affected bullets in the Hamby-44 set.
-These issues affect the algorithm's ability to identify a stable signature, which impacts the subsequent match scores  at the land-to-land and bullet-to-bullet level.
+Several lands in Hamby set 44 have major deficiencies,  such as `tank rash' (a collision of the bullet after exiting the barrel with a surface causing markings on top of the striations from the barrel) or extreme pitting (holes caused by direct contact with burning gun powder).
+\autoref{fig:tankrash} shows rendered scans of all affected lands in the Hamby-44 set.
+These issues affect the algorithm's ability to identify a stable signature, which impacts the subsequent similarity scores at the land-to-land and bullet-to-bullet level.
 <<tankrash, fig.cap="Overview of bullet lands with prominent deficiency such as tank rash or extreme pitting in the Hamby-44 study.", fig.width=4, fig.height = 2, out.width='47.5%', fig.subcap=c("Br1-B1-L6 (tank rask)", "Br2-B2-L5 (tank rash)", "Br3-B1-L5 (tank rash)", "Br8-B1-L6 (tank rash)", "Br8-B2-L2 (pitting)", "Br8-B2-L6 (tank rash)", "Unk-E-L6 (tank rash)", "Unk-I-L6 (tank rash and break-off)"), fig.align="middle", fig.ncol=2>>=
@@
 )
-pdlands <- pdnest %>% unnest(data)
-
-all_lands <- rbind(
+pdlands <- pdnest %>%
+  rename(bullet_KM = KM) %>% 
+  unnest(data)
+
+all_lands <- bind_rows(
   h44lands %>%
@@
 Some same-source comparisons are also below 0.5 for RF scores, indicating potential problems with at least one of the lands involved in the comparison.">>=
-all_lands %>% 
-  gather(key = "measure", value = "value", ccf, rfscore) %>%
-  mutate(measure = factor(measure, levels = c("ccf", "rfscore"),
-                          labels = c("Cross correlation", "RF score")),
+# all_lands %>% 
+#   gather(key = "measure", value = "value", ccf, rfscore) %>%
+#   mutate(measure = factor(measure, levels = c("ccf", "rfscore"),
+#                           labels = c("Cross correlation", "RF score")),
+#          study = paste(study, group1) %>% str_trim() %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"))) %>%
+#   ggplot(aes(
+#     x = value, fill = sameland,
+#     colour = sameland
+#   )) +
+#   geom_density(alpha = 0.6) +
+# #  geom_rug(alpha = 0.5/12, show.legend = F) +
+#   facet_grid(measure ~ study) +
+#   theme_bw() +
+#   scale_fill_manual("Same source", values = c("darkgrey", "darkorange")) +
+#   scale_colour_manual("Same source", values = c("darkgrey", "darkorange")) +
+#   scale_x_continuous("Land-to-Land scores", limits = c(0, 1), labels = c(0, 0.25, 0.5, 0.75, 1)) +
+#   theme(legend.position = "bottom") +
+#   geom_hline(yintercept = 0, show.legend = F)
+p1 <- all_lands %>% 
+  gather(key = "measure", value = "value", ccf, cms, rfscore) %>%
+  mutate(measure = factor(measure, levels = c("ccf", "cms", "rfscore"),
+                          labels = c("Cross correlation", "CMS", "RF score")),
          study = paste(study, group1) %>% str_trim() %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"))) %>%
+  filter(measure == "RF score") %>%
   ggplot(aes(
@@
   theme_bw() +
+  scale_fill_manual("Same source", values = c("darkgrey", "darkorange"), guide = F) +
+  scale_colour_manual("Same source", values = c("darkgrey", "darkorange"), guide = F) +
+  scale_x_continuous("Land-to-Land scores", limits = c(0, 1), labels = c(0, 0.25, 0.5, 0.75, 1)) +
+  theme(legend.position = "bottom", axis.title.x = element_blank(), axis.text.y = element_text(hjust = .5, angle = 90)) +
+  geom_hline(yintercept = 0, show.legend = F)
+
+p2 <- all_lands %>% 
+  gather(key = "measure", value = "value", ccf, cms, rfscore) %>%
+  mutate(measure = factor(measure, levels = c("ccf", "cms", "rfscore"),
+                          labels = c("Cross correlation", "CMS", "RF score")),
+         study = paste(study, group1) %>% str_trim() %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"))) %>%
+  filter(measure == "Cross correlation") %>%
+  ggplot(aes(
+    x = value, fill = sameland,
+    colour = sameland
+  )) +
+  geom_density(alpha = 0.6) +
+#  geom_rug(alpha = 0.5/12, show.legend = F) +
+  facet_grid(measure ~ study) +
+  theme_bw() +
+  scale_fill_manual("Same source", values = c("darkgrey", "darkorange"), guide = F) +
+  scale_colour_manual("Same source", values = c("darkgrey", "darkorange"), guide = F) +
+  scale_x_continuous("Land-to-Land scores", limits = c(0, 1), labels = c(0, 0.25, 0.5, 0.75, 1)) +
+  theme(legend.position = "bottom", axis.title.x = element_blank(), axis.text.y = element_text(hjust = .5, angle = 90)) +
+  geom_hline(yintercept = 0, show.legend = F)
+
+p3data <- all_lands %>% 
+  gather(key = "measure", value = "value", ccf, cms, rfscore) %>%
+  mutate(measure = factor(measure, levels = c("ccf", "cms", "rfscore"),
+                          labels = c("Cross correlation", "CMS", "RF score")),
+         study = paste(study, group1) %>% str_trim() %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"))) %>%
+  filter(measure == "CMS")
+p3bw <- bw.nrd0(p3data$value[p3data$sameland & p3data$study == "Houston FSI G1"])
+p3 <- ggplot(p3data, aes(
+    x = value, fill = sameland,
+    colour = sameland
+  )) +
+  geom_density(alpha = 0.6, bw = p3bw) +
+  facet_grid(measure ~ study) +
+  theme_bw() +
   scale_fill_manual("Same source", values = c("darkgrey", "darkorange")) +
   scale_colour_manual("Same source", values = c("darkgrey", "darkorange")) +
-  scale_x_continuous("Land-to-Land scores", limits = c(0, 1), labels = c(0, 0.25, 0.5, 0.75, 1)) +
-  theme(legend.position = "bottom") +
+  scale_x_continuous("Land-to-Land scores") +
+  theme(legend.position = "bottom", axis.text.y = element_text(hjust = .5, angle = 90)) +
   geom_hline(yintercept = 0, show.legend = F)
+
+grid.arrange(p1, p2, p3, nrow = 3, heights = c(1, 1, 1.4))
 @
@@
 
-At the beginning of this study, we anticipated Hamby 44 would be the easiest set to work with because of its similarity to the sets used to train the random forest algorithm.
-In a surprise turn of events, it was the hardest to solve, in part because of damage to the bullets that obscured LEA striae.
-It has been shown \citet{lpr} that parts of lands can be used for successful identifications.
-The algorithm proposed by \citet{aoas} is not capable of  automatically detecting parts of lands with well-expressed striae.
+At the beginning of this study, we anticipated Hamby 44 would be the easiest set to evaluate because of its similarity to the sets used to train the random forest algorithm.
+In a surprise turn of events, it was the hardest, in part because of damage to the bullets that obscured LEA striae.
+It has been shown \citet{lpr} that parts of lands can be used for successful identifications, if at least 50\% of the land is present (using full-length scores as the reference distribution).
+The random forest algorithm proposed in \citet{aoas} is not capable of automatically detecting parts of lands with well-expressed striae.
 It may be useful to couple the \citet{aoas} algorithm with an algorithm which assesses the quality of the input data and determines which portions of the data to use for comparison.
-This would emulate the process used by examiners, who first assess the quality of the evidence and whether there is enough information present to attempt a match.
+This would emulate the process used by examiners, who first assess the quality of the evidence and whether there is enough information present to attempt a comparison.
 One major disadvantage of an automated algorithm is that all of the decisions humans make (recognizing degraded land areas, excluding those areas from consideration, matching only the remaining areas) must be explicitly characterized; however, this explicit characterization means that the process can be scientifically validated to a much higher degree than the human perceptual process.
@@
 Rugers are among the firearms that mark very well, so we would correspondingly expect to see lowered similarity scores for other firearms.
-The matching algorithm works on matching lands.
-Barrels with polygonal rifling, such as Glocks, do not introduce well defined lands on bullets, which makes an application of the matching algorithm impossible.
+While the performance of cross-correlation and random forest scores are similar when evaluated on Rugers, it is likely that the additional information used to generate the random forest score may be useful when comparing bullets from barrels which produce fewer distinctive marks.
+A future study will compare the performance of the random forest score, cross-correlation score, and CMS on non-Ruger barrels.
+The random forest algorithm is trained on scans taken at the land level, so barrels with polygonal rifling, such as Glocks, which do not introduce well defined lands on bullets, cannot be compared using the random forest algorithm examined here.
+In addition, we plan to re-fit the random forest using a wider set of scans taken at different resolutions, with different ammunition, fired from barrels of different models.
+We expect that a random forest trained on a wider set of input data will be more likely to produce scores which meet the criteria for monotonicity and stability.
 
@@
 For a future round of the Houston study, we are planning to compare the algorithmic performance directly with scores given by forensic toolmark examiners.
-The algorithm's score is not intended to replace an examiner's judgment; instead, it provides a complementary tool to measure and quantify the strength of an identification.
+The algorithm's score is not intended to replace an examiner's judgment; instead, it provides a complementary tool to measure, quantify, and compare the similarity of two bullets in an identification.
+With more research into the behavior of score-based likelihood ratios, the random forest score may also be used to compute a SLR to explicitly quantify the strength of the match between two bullets relative to other matches in the population.
 By validating the algorithm on external test sets, we have demonstrated that the method can be generalized to different types of ammunition and is not overly sensitive to small differences in rifling procedure.
@@
 
-<<output-all-figs, include = F>>=
-thisdoc <- readLines("index.Rmd")
+<<output-all-figs,eval = F, include = F>>=
+thisdoc <- readLines("index.Rnw")
 
@@
 # Haven't really tested this part - regex may need work
-raw_figures <- data_frame(
+raw_figures <- tibble(
   line_start = which(str_detect(thisdoc, "\\begin\\{figure\\}")), 
@@
 # Code chunk figures
-chunk_figures <- data_frame(
+chunk_figures <- tibble(
   line_start = which(str_detect(thisdoc, "^```\\{r.*\\}$")),
@@
 # Knitr included graphics figures
-chunk_incl_figures <- data_frame(
+chunk_incl_figures <- tibble(
   line_start = which(str_detect(thisdoc, "^```\\{r.*\\}$")),

### 2019-12-24 index.Rnw
Required documents for upload

@@
 )
-h44nest <- h44nest %>%
+h44nest <- h44nest %>%ungroup() %>% 
   mutate(
@@
 
-pdnest <- pdnest %>%
+pdnest <- pdnest %>%ungroup() %>% 
   mutate(
@@
 nests <- rbind(
-  h44nest %>%
+  h44nest %>%ungroup() %>% 
     select(sam_rf, sam_ccf, sam_cms, samesource) %>%
     mutate(study = "Hamby 44", group1 = NA),
-  pdnest %>%
+  pdnest %>%ungroup() %>% 
     select(sam_rf, sam_ccf, sam_cms, samesource) %>%
     mutate(study = "Phoenix PD", group1 = NA),
-  hounest %>%
+  hounest %>%ungroup() %>% 
     select(sam_rf, sam_ccf, sam_cms, samesource, group1) %>%
@@
 
-all_lands <- rbind(
+all_lands <- bind_rows(
   h44lands %>%
@@
 
-<<output-all-figs, include = F>>=
-thisdoc <- readLines("index.Rmd")
+<<output-all-figs, eval = F, include = F>>=
+thisdoc <- readLines("index.Rnw")
 
@@
 # Code chunk figures
-chunk_figures <- data_frame(
-  line_start = which(str_detect(thisdoc, "^```\\{r.*\\}$")),
-  line_end = which(str_detect(thisdoc, "^``` *$")),
-  chunk_name = str_replace(thisdoc[line_start], "(```\\{r ?)([0-9A-z-]*)(,?.*\\})", "\\2")
+chunk_figures <- tibble(
+  line_start = which(str_detect(thisdoc, "^<<.*>>=$")),
+  line_end = which(str_detect(thisdoc, "^@$")),
+  chunk_name = str_replace(thisdoc[line_start], "<<([0-9A-z-]*?),?((?:.*=.*){1,})>>=", "\\1")
 ) %>% 
@@
 # Knitr included graphics figures
-chunk_incl_figures <- data_frame(
-  line_start = which(str_detect(thisdoc, "^```\\{r.*\\}$")),
-  line_end = which(str_detect(thisdoc, "^``` *$")),
-  chunk_name = str_replace(thisdoc[line_start], "(```\\{r ?)([0-9A-z-]*)(,?.*\\})", "\\2")
+chunk_incl_figures <- tibble(
+  line_start = which(str_detect(thisdoc, "^<<.*>>=$")),
+  line_end = which(str_detect(thisdoc, "^@$")),
+  chunk_name = str_replace(thisdoc[line_start], "<<([0-9A-z-]*?),?((?:.*=.*){1,})>>=", "\\1")
 ) %>% 
@@
   group_by(fig_num) %>%
-  mutate(rn = 1:n(),
+  mutate(rn = row_number(),
          maxrn = n(),

### 2019-12-24 title_page.tex
Required documents for upload

--- 2019-08-07 title_page.tex
@@
-\documentclass[titlepage]{article}
-\title{Case study validations of automatic bullet matching}
-\author{Susan Vanderplas and Heike Hofmann}
-\date{Center for Statistics and Applications in Forensic Evidence, Iowa State University. 195 Durham Center, 613 Morill Rd., Ames, IA 50011}
+\documentclass{article}
+\usepackage{fullpage}
+
+\title{Comparison of three similarity scores for bullet LEA matching}
+\author{Susan Vanderplas$^\ast$$^{1,2}$, Melissa Nally$^3$, Tylor Klep$^4$, Cristina Cadevall$^5$, Heike Hofmann$^1$}
+
+\date{$^\ast$ Corresponding author. email: susan.vanderplas@unl.edu\\[1em]
+$^1$Center for Statistics and Applications in Forensic Evidence, Iowa State University\\
+195 Durham Center, 613 Morill Rd., Ames, IA 50011\\[1em]
+$^2$Statistics Department, University of Nebraska Lincoln\\340 Hardin Hall North Wing, Lincoln, NE 68583-0963\\[1em]
+$^2$Houston Forensic Science Center\\500 Jefferson St 13th Floor, Houston, TX 77002\\[1em]
+$^3$Phoenix Police Department\\Laboratory Services Bureau\\621 W Washington St, Phoenix, AZ 85003\\[1em]
+$^4$Sensofar Group\\Parc Audiovisual de Catalunya Ctra, BV-1274, Km 1, 08225 Terrassa, Barcelona, Spain}
 

### 2020-01-17 Point_by_point_response.Rmd
updated filfes, changes, etc.

---
title: "Point by Point Reviewer Response"
date: "2020-01-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We thank the reviewer for his or her careful and constructive comments on our article, which we have retitled "Comparison of three similarity scores for bullet LEA matching". We have made numerous changes to the document in response to the suggestions made by the reviewer, and have responses to the comments which we could not address according to the suggestions. An included document shows the differences between the text in the original version and the text in the resubmitted version.

## Summary
The aim of this article is to compare three different measures of LEA similarity, the author developed random forest score, the cross correlation and consecutive matching striae, determined from three Ruger bullet test sets.

## General comments
> In general, the article is clearly structured and well written. Some of the figures are a bit small, particularly the captions. More details are given below. One thing that I'm missing in the document is the step from score distribution determination to the calculation of the strength of the evidence, e.g. in the form of a score based likelihood ratio or calculate identification error rates. I strongly encourage the authors to add this to the document. In addition I think that the authors did a great evaluation using three different test sets. However, they only include Ruger bullets, which are known to mark very well and in fact, based on the results presented in the article, the added value of using the random forest score instead of cross correlation is not justified, in contrary. It would also be very interesting to see the performance on less well marking firearms as I think for those cases the random forest score should indeed perform better than the cross correlation. Finally I think that the document is pretty lengthy and detailed, given what you are studying, and I suggest to shorten some paragraphs.

Response:

- We have addressed the figure legend/label size throughought the document. 
- The choice to exclude score-based likelihood ratios from the scope of this article was a conscious decision, as the land-to-land comparisons produce density estimates with different variability between known matches and known non-matches. As a result, in these cases, the SLR is nonmonotonic and produces unexpected and potentially misleading results at the extremes of the distribution (see Fig. 4 in Neumann and Ausdemore, https://arxiv.org/pdf/1910.05240.pdf). We would like to defer the explicit use of SLRs; we have added a paragraph on the potential of using likelihood ratios with the scoring methods described in the paper, as well as several mentions of how this method could be used with LRs in theory. 
- Error rates are present in the ROC curves presented in Figure 10 and are  presented more explicitly in Figure 11, which shows the equal errorr rates and thresholds for each land-to-land score. We have emphasized that at the bullet-to-bullet level, the distributions for each study are perfectly separated, indicating that the error rate would be nil. It is only when we proceed to analyze the land-to-land comparisons that we can do any sort of assessment of the error rate. 
- We appreciate the reviewer's discussion about the Ruger vs. non-Ruger bullets; we have included some of the suggestions in the paper, but do not currently have sufficient data to include a fourth case study in this paper. We agree that this would be an interesting test of the random forest score compared to the cross correlation and are planning several future studies which explore the random forest score's use on non-Ruger bullets. As this paper is already fairly lengthy, as noted by the reviewer, we feel that the addition of a non-Ruger case study is best reserved for one of these future papers. 
- We have (throughout the document) tightened up language to make the paper more succinct.

## Further comments

- Title: I think the title is not covering the content very well. How about "Comparison of three different similarity scores for bullet LEA matching" or something like that.

> We have modified the title according to the reviewer's suggestion

- Abstract: I think it's odd to refer to the article by Tai et al. as an example for an automated impression mark comparison algorithm that works on 2D images, particularly as you stated earlier in the abstract that you focus on 3D data. Therefore e.g. the AFFC/CMC algorithm developed by John Song at NIST would be more appropriate.

> We have included the additional reference to the CMC algorithm.

- Abstract: "… enabling virtual comparisons…" Technically speaking, this is also possible with 2D data.

> We agree, thanks for the correction. The abstract has been modified to make it clear that the 3D microscopy allows for more precise virtual comparisons.

- Abstract: What do you mean with "external tests"?

> An external test set is one which was not used to train the algorithm - that is, data which the algorithm has not "seen" before. We have modified the language to avoid the terminology as it appears it is not common outside machine learning and statistics.

- Abstract: "We are presenting a discussion of the performance of the matching algorithm [2] in three studies". This is misleading I think. Basically what you do is to compare the performance of three similarity metrics, your own random forest approach, cross correlation and CMS. The alignment is based on the cross correlation function, which is not novel or specific to your algorithm.

> The matching algorithm encompassess more than the alignment based on cross-correlation. It also involves several cleaning and smoothing steps, the selection of an appropriately stable crosscut, and the calculation of other metrics (e.g. peak identification, CMS calculation) after sequences are aligned based on maximal cross-correlation. While we evaluate different scoring metrics, all of these scoring metrics are calculated conditional on the alignment using the algorithm described in Hare et al. (2015). Thus, it is important to distinguish between the matching algorithm and the scoring mechanism.

- Page 2: "… to assess the strength of a match …" The cross correlation is merely a similarity metric, that doesn't say anything about the strength of the match. For that you need reference data. An LR is a measure of the strength of a match. Same three lines down "… assessing match strength…". Please check whole document and revise.

> We have (throughout the paper) changed the language to distinguish between similarity metrics and match strength; we appreciate the importance of this clarification.

- Methods: What is a "stable crosscut"?

> We have added a clause describing stable ccrosscuts as an area where striae are similar in the region above and below the crosscut. This excludes regions with damage that interfere with similarity score calculation

- Figure 1 is pretty small, particularly the caption font size. Furthermore I think it would be more illustrative to put two corresponding profiles in the same plot.

> We created a replacement for Figure 1 with the signatures overlaid, but found it difficult to read because the sequences are not (at this stage of the algorithm) aligned, so much of the similarity is difficult to see when the unaligned details are overlaid. We include the modified Figure 1 here as a demonstration, but ultimately feel that the original Fig 1 (with labels of increased size) allows for easier visual comparison of the peaks in the sequence.

![](fig1-alt.png)

- Page 4: "Note that with automatic alignment…. consecutive matching striae of zero". I don't understand this sentence.

> We have added some clarifying language. Essentially, because examiners align bullets based on peaks, there will always be at least one matching striation mark; we align based on cross-correlation, which does not guarantee that there will be any matching striae. So the measure of CMS used in this paper differs slightly from the measure of CMS used by examiners.

- Page 4: "… and with it may indicate that the two signatures come from the same source". Same issue as above. Based on the similarity metric only, you can't say anything about the strength of the evidence. For that you need reference distributions. Same eight lines down on the same page.

> The goal of this sentence and the sentence below it is to provide some intuition about the direction of the scoring mechanism: 0 indicates low similarity, 1 indicates extremely high correspondence. We have modified the language somewhat to make this clearer.

- Page 4: "An automated algorithm selects a cutoff…". That is a choice and should as such be mentioned in the article. Some authors do indeed calculate a threshold, but many these days prefer a system that does not require to use a threshold and rather calculate a continuous measure like the LR.

> We have added some language to contrast our approach with SLRs: "Of course, in jurisdictions which utilize score-based likelihood ratios, this threshold system is not necessary because likelihood ratios are continuous. Even with likelihood ratios, however, there is a natural threshold at 1 which functions similarly to the selected threshold in the binary decision case."

- Page 4: What do you mean with stability? If your reference distribution is the same, the conclusion will always be the same. Or do you mean that automated systems should not include a random element that might lead to variability in the outcome?

> For each test set, we assemble a distribution of scores of known matches and non-matches that might serve as a reference distribution for similar firearms. Ideally, the threshold for separating a match score and a non-match score would be the same across different test sets. We have added some language to make this clearer when defining the term.

- Page 5: "When imaging bullets, operators scan one land at a time…" These days there are a couple of systems around that measure the whole bullet at once, e.g. IBIS 3D or Laboratory Imaging BalScan.

> We have added some language to this effect.

- Figure 2: What does RF stand for? Random Forest score?

> We have defined RF in the figure legend.

- Page 6: What is a in the first formula?

> We have added some clarifying language indicating that the a's are the elements of the matrix A.

- Figure 3 is too small?

> We have increased the text size.

- Page 7: "\*\*resolution\*\*". Was this intended?

> Thanks for catching that; we have fixed it.

- I like the fact that the authors used three different test sets. Are these also publicly available or do the authors intend to do that? That would be great.

> We are working on gaining approval to publish a data paper with the data used in this paper, to make the land scans publicly available for additional analysis; we strongly agree with the reviewer that the community would benefit from access to these scans. At the moment, two of the three sets have been uploaded to NBTRD at a lower resolution than the resolution used in our calculations, but before we provide bulk downloads of the higher-resolution data, we must consult with CSAFE collaborators at NIST and other institutions. We are in the process of creating a [metadata-only figshare collection](https://doi.org/10.25380/iastate.c.4781447) with contact information to request the data (it will be available at once it is released by our librarians, but can be seen [here](https://figshare.com/s/bb22f76c9cb196a4f9d3) until that approval is granted). Once we have collaborator approval, we hope to submit a Data in Brief paper and upload the full datasets for public consumption.

- Ruger barrels are typically leaving very distinct and good marks on bullets. That might be the reason why the performance for random forest score and cross correlation is similar (Figure 7). There I suggest to extend the comparison study also to barrels that do not mark that well. In those scenarios the random forest approach should demonstrate to perform much better than simply using cross correlation. In other words, for Ruger bullets you don't need a random forest approach and can simply use cross correlation and that'll be sufficient. Looking at figure 10, the random forest approach is actually worse for most cases.

> We appreciate the reviewer's discussion about the Ruger vs. non-Ruger bullets; we have included some of the suggestions in the discussion section of the paper, but do not currently have sufficient data to include a fourth case study in this paper. We agree that this would be an interesting test of the random forest score compared to the cross correlation and are planning several future studies which explore the random forest score's use on non-Ruger bullets. As this paper is already fairly lengthy, as noted by the reviewer, we feel that the addition of a non-Ruger case study is best reserved for one of these future papers.

- Page 13: "… we would need to considerably expand the database used for training". Same as in the previous comment. I'm afraid that simply training on Ruger data will not be enough as these are basically among the best marking barrels around. Fine tuning with less well marking barrels will probably necessary.

> We are aware that the random forest will need to be re-trained on a broader set of bullets, including non-Ruger bullets; this is another reason why we are hesitant to apply the random forest scoring algorithm to non-Ruger bullets - we need to re-train the model on more diverse data.

- Page 13: It would of course be great to have perfectly separated distributions. However, if we calculate a score based LR (SLR) for each land and can show that the marks are independent, which they should be, then it is possible to multiply all individual SLRs. Even if we end up with a very low SLR for each individual land, say only SLR = 10, this would yield a bullet SLR of 10^6.

> We cannot recommend this course of action in part because the land to land scores are not independent, rather, they result from pairwise comparisons that are inherently dependent. Further, we can't assume that striae observed on different lands from the same physical bullet are independent, because they were all part of the same firing event. While this assumption is often made for practical purposes, the assumption is highly suspect and should not be further promoted without considering the covariance structures within the data. For this reason, we feel that the use of the sequence average match method for combining scores is more reasonable because it accounts for the spatial relationships between the set of lands belonging to the same bullet, without requiring assumptions of independence.

- Figure 10 is too small, particularly the caption fonts. Also specify which symbol is which score on the right column.

> We have modified the figure text size. The caption for the figure is located at the bottom between the two graphs, and does show which shapes are used for each score; however, each score is also explicitly labeled on the left-hand side of figure B.

- It is a bit strange that the CMS is not part of the figures 7 and 9. You state that you compare the three scores, but you only do that in figure 10 and 11.

> We have fixed that - thank you for catching it!

- Peak valley detection may be tricky. Which algorithm did you use to determine the CMS?

> This is part of the matching algorithm described in Hare et al. (2017). We have made the reference to the algorithm more explicit, but do not feel we can completely explain the peak detection and matching process here without significantly lengthening the paper. As it is already part of the published record, citing the previous paper seems most appropriate.

- Figure 11 is too small.

> We have adjusted the text size in figures throughout the paper.

- Page 17: Known match and known non-match reference distributions for whole LEA marks and partial LEA marks are different, as the SLR is dependent on the width of the mark. The KM scores stay more or less the same, but the KNM scores get higher if the mark width get smaller. As your approach also includes cross correlation I assume the random forest score is also sensitive to mark width. As a result, probably LEAs with partial marks should just be discarded, as we luckily have a couple of LEAs left. Otherwise one would have to have reference distributions of different mark lengths available.

> Hare, Hofmann, and Carriquiry (2017) showed how to use the random forest score to match degraded and partial land impressions; the random forest score remained fairly stable up to 50% signature length, suggesting that the random forest score, at least, is relatively robust to degradation and partial mark size. For the reasons stated above, we do not carry this conclusion all the way to the calculation of SLRs, but given the previously cited paper, we do not expect to require a separate comparison distribution when the partial LEA includes at least 50% of the length of a full LEA signature.

- It will be very interesting to see how your scores compare to score given by forensic examiners but even more interesting would be to compare the actual conclusions of the examiners with automatically determined conclusions (in the form of an SLR for example).

> We agree. The Houston FSC study data shown here were part of a larger examiner study; we are preparing a paper which compares the algorithm's performance using the decision criteria shown in this paper to the performance of firearm examiners and expect to submit this paper for review in mid 2020.

### 2020-01-17 index-blinded.Rnw
updated filfes, changes, etc.

@@
 \begin{abstract}
-Recent advances in microscopy have made it possible to collect 3D topographic data, enabling virtual comparisons based on the collected 3D data as a supplement to traditional comparison microscopy and 2D photography.
+Recent advances in microscopy have made it possible to collect 3D topographic data, enabling more precise virtual comparisons based on the collected 3D data as a supplement to traditional comparison microscopy and 2D photography.
 Automatic comparison algorithms have been introduced for various scenarios, such as matching cartridge cases \citep{tai2018fully,song3DTopographyMeasurements2014} or matching bullet striae \citep{aoas2, chuAutomaticIdentificationBullet2013, dekinderAutomatedComparisonsBullet1999}.
@@
   fig.path = "figures/",
-  echo = FALSE
+  echo = FALSE,
+  dev = "png"
 )
@@
 Cross correlation varies between -1 and 1, but as the alignment is based on the maximized cross correlation value, in practice the cross correlations for aligned signatures are generally positive.
-Cross correlation values which are close to 1 indicate similarity between signatures, and may consequentially indicate that the two signatures come from the same source.
-Low cross correlation values indicate signatures which may originate from different sources.
+Cross correlation values which are close to 1 indicate similarity between signatures, with the possibility that the two signatures come from the same source.
+Low cross correlation values indicate signatures which are different and thus may originate from different sources.
 
 The random forest presented in \citet{aoas} is based on a combination of multiple characteristics, including cross correlation, number of matching striae, and number of consecutively matching striae.
-The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the similarity of the two signatures, where scores close to 1 indicate it to be highly likely that the aligned signatures come from the same source, while scores close to 0 indicate that the aligned signatures are highly likely to originate from different sources.
+The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the similarity of the two signatures, where scores close to 1 indicate signatures which are similar and may have originated from the same source, while scores close to 0 indicate that the aligned signatures are different and may originate from different sources.
 
@@
 
-<<sam-sketch, out.width='\\textwidth', fig.width=8, fig.height = 3, fig.cap="Sketch of all six land-to-land sequences between two bullets with six lands.">>=
+<<sam-sketch, out.width='\\textwidth', fig.width=6, fig.height = 2.5, fig.cap="Sketch of all six land-to-land sequences between two bullets with six lands.">>=
 df2 <- data.frame(expand.grid(x = 1:6, y = 1:6))

### 2020-01-17 index.tex
updated filfes, changes, etc.

--- 2019-12-09 index.Rnw
@@
-\documentclass[doubleblind]{elsarticle}
+\documentclass[doubleblind]{elsarticle}\usepackage[]{graphicx}\usepackage[]{color}
+% maxwidth is the original width if it is less than linewidth
+% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\let\hlipl\hlkwb
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 
-<<echo = FALSE, message = FALSE, warning = FALSE>>=
-library(knitr)
-knitr::opts_chunk$set(
-  message = FALSE,
-  warning = FALSE,
-  error = FALSE,
-  collapse = TRUE,
-  comment = "",
-  fig.height = 6,
-  fig.width = 6,
-  dpi = 600,
-  fig.align = "center",
-  out.width = "\\textwidth",
-  cache = TRUE,
-  fig.path = "figures/",
-  echo = FALSE
-)
-options(knitr.table.format = "latex")
-
-library(stringr)
-
-library(tidyverse)
-library(scales)
-library(multidplyr) # install_github("hadley/multidplyr")
-library(bulletxtrctr) # install_github("csafe-isu/bulletr")
-library(gridExtra)
-library(kableExtra)
-@
-
 In current practice, firearms and toolmark examiners (FTE) evaluate the similarity of striae on bullets by placing the evidence together with another bullet under a comparison microscope.
@@
 \autoref{fig:signatures} shows a set of six signatures corresponding to the six land engraved areas of bullets 1 and 2 from barrel 1 of set 44 of the Hamby study \citep{hamby}.
-<<signatures, fig.cap="Signatures of all six lands of bullets 1 and 2 from barrel 1 of the Hamby set 44.
-Matching lands between the two bullets have been placed above each other.", fig.height=4, fig.width=10>>=
-# signatures for one bullet
-# h44cc <- readRDS("data/h44-ccdata.rda")
-# br1 <- h44cc[1:12]
-# dframe <- 1:12 %>% purrr::map_df(.f = function(i) {
-#   d <- br1[[i]]
-#   d <- d %>% group_by(y) %>% summarize(
-#     value = mean(value, na.rm=TRUE)
-#   )
-#   d$id <- i
-#   d %>% ungroup(y) %>% rename(
-#     x = y
-#   )
-# })
-#
-# dframe$bullet <- rep(c("Bullet 1", "Bullet 2"), each = 6)[dframe$id]
-# dframe$land <- paste("Land", (dframe$id %% 6 + 1))
-#
-# dframe <- dframe %>% group_by(bullet, land) %>% nest()
-# names(dframe)[3] <- "ccdata"
-#
-# library(bulletxtrctr)
-# dframe <- dframe %>% mutate(
-#   grooves = ccdata %>%
-#     purrr::map(.f = cc_locate_grooves, method = "middle",
-#                adjust = 30, return_plot = TRUE)
-# )
-# dframe <- dframe %>% mutate(
-#   sigs = purrr::map2(
-#     .x = ccdata, .y = grooves,
-#     .f = function(x, y) {
-#       cc_get_signature(
-#         ccdata = x, grooves = y, span1 = 0.75, span2 = 0.03)
-#     })
-# )
-# dframe$barrelland <- dframe$land
-# idx <- which(dframe$bullet=="Bullet 2")
-# dframe$landid <- parse_number(dframe$land)
-# dframe$landid[idx] <- (dframe$landid[idx] +1) %% 6 + 1
-# dframe$land <- paste("Land", dframe$landid)
-# signatures <- dframe %>% select(bullet, land, sigs) %>% tidyr::unnest()
-
-signatures <- read.csv("data/signatures.csv")
-signatures %>%
-  ggplot(aes(x = x / 1000)) +
-#  geom_line(aes(y = raw_sig), colour = "grey70") +
-  geom_line(aes(y = sig), colour = "grey30") +
-  facet_grid(bullet ~ land) +
-  ylim(c(-5, 5)) +
-  theme_bw() +
-  ylab("Surface measurement (in microns)") +
-  xlab("Relative Location (in millimeters)")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/signatures-1}
+
+}
+
+\caption[Signatures of all six lands of bullets 1 and 2 from barrel 1 of the Hamby set 44]{Signatures of all six lands of bullets 1 and 2 from barrel 1 of the Hamby set 44.
+Matching lands between the two bullets have been placed above each other.}\label{fig:signatures}
+\end{figure}
+
+\end{knitrout}
 
@@
 Cross correlation varies between -1 and 1, but as the alignment is based on the maximized cross correlation value, in practice the cross correlations for aligned signatures are generally positive.
-Cross correlation values which are close to 1 indicate similarity between signatures, and may consequentially indicate that the two signatures come from the same source.
-Low cross correlation values indicate signatures which may originate from different sources.
-
-The random forest presented in \citet{aoas2} is based on a combination of multiple characteristics, including cross correlation, number of matching striae, and number of consecutively matching striae.
-The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the similarity of the two signatures, where scores close to 1 indicate it to be highly likely that the aligned signatures come from the same source, while scores close to 0 indicate that the aligned signatures are highly likely to originate from different sources.
+Cross correlation values which are close to 1 indicate similarity between signatures, with the possibility that the two signatures come from the same source.
+Low cross correlation values indicate signatures which are different and thus may originate from different sources.
+
+The random forest presented in \citet{aoas} is based on a combination of multiple characteristics, including cross correlation, number of matching striae, and number of consecutively matching striae.
+The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the similarity of the two signatures, where scores close to 1 indicate signatures which are similar and may have originated from the same source, while scores close to 0 indicate that the aligned signatures are different and may originate from different sources.
 
@@
 
-<<demo-l2lscores, echo=FALSE, results = 'hide', warning = FALSE, message = FALSE>>=
-features <- read.csv("data/fsi-features.csv.gz")
-features <- features %>% mutate(
-  bullet1 = gsub("ullet ", "", bullet1),
-  bullet2 = gsub("ullet ", "", bullet2)
-)
-b12 <- features %>% filter(
-  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
-  group2 == "G1", barrel2 == "KA", bullet2 == "B2"
-)
-# 
-# p1 <- b12 %>%
-#   ggplot(aes(x = land1, y = land2, fill = rfscore)) +
-#   geom_tile() + ggtitle("Same barrel") +
-#   scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
-#   theme_bw() +
-#   xlab("Bullet 2") + ylab("Bullet 1") +
-#   theme(legend.position = "bottom") + coord_equal()
-# 
-bab <- features %>% filter(
-  group1 == "G1", barrel1 == "KA", bullet1 == "B1",
-  group2 == "G1", barrel2 == "KB", bullet2 == "B1"
-)
-# p2 <- bab %>%
-#   ggplot(aes(x = land1, y = land2, fill = rfscore)) +
-#   geom_tile() + ggtitle("Different barrel") +
-#   scale_fill_gradient2("RF score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
-#   theme_bw() +
-#   xlab("Bullet 2") + ylab("Bullet 1") +
-#   theme(legend.position = "bottom") + coord_equal()
-
-l2lscores <- bind_rows(
-  mutate(b12, type = "Same barrel"),
-  mutate(bab, type = "Different barrel")
-) %>% 
-  mutate(type = factor(type, levels = c("Same barrel", "Different barrel"), ordered = T))
-
-@
-
 Land-to-land comparisons lead to a whole set of scores to evaluate for bullet-to-bullet comparisons.
@@
 
-<<b2b, echo = FALSE, fig.width = 5.4, fig.height = 3, out.width = ".6\\textwidth", fig.cap="Overview of land-land matching scores for two pairs of bullet-bullet comparisons.
-On the left the two bullets are known to come from the same source (barrel) on the right, the bullets are from two different sources (barrels).">>=
-# grid.arrange(p1, p2, ncol = 2, padding = 2)
-ggplot(data = l2lscores, aes(x = land1, y = land2, fill = rfscore)) +
-  geom_tile() + 
-  scale_fill_gradient2("Random Forest\n(RF) score: ", midpoint = 0.3, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
-  theme_bw() +
-  xlab("Bullet 2") + ylab("Bullet 1") +
-  theme(legend.position = "right") + coord_equal() + facet_wrap(~type)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+
+{\centering \includegraphics[width=.6\textwidth]{figures/b2b-1}
+
+}
+
+\caption[Overview of land-land matching scores for two pairs of bullet-bullet comparisons]{Overview of land-land matching scores for two pairs of bullet-bullet comparisons.
+On the left the two bullets are known to come from the same source (barrel) on the right, the bullets are from two different sources (barrels).}\label{fig:b2b}
+\end{figure}
+
+\end{knitrout}
 
@@
 
-<<sam-sketch, out.width='\\textwidth', fig.width=8, fig.height = 3, fig.cap="Sketch of all six land-to-land sequences between two bullets with six lands.">>=
-df2 <- data.frame(expand.grid(x = 1:6, y = 1:6))
-df2$case1 <- df2$x == ((df2$y + 5) %% 6) + 1
-df2$case2 <- df2$x == ((df2$y + 0) %% 6) + 1
-df2$case3 <- df2$x == ((df2$y + 1) %% 6) + 1
-df2$case4 <- df2$x == ((df2$y + 2) %% 6) + 1
-df2$case5 <- df2$x == ((df2$y + 3) %% 6) + 1
-df2$case6 <- df2$x == ((df2$y + 4) %% 6) + 1
-
-df2 %>%
-  gather("case", "values", starts_with("case")) %>%
-  mutate(case = gsub("case", "", case)) %>%
-  mutate(case = paste0("k = ", as.numeric(case) - 1)) %>%
-  ggplot(aes(x = factor(x), y = factor(y, levels = 1:6), fill = values)) +
-  geom_tile(colour = "grey20") +
-  facet_wrap(~case, ncol = 6) +
-  xlab("Bullet 2") + ylab("Bullet 1") +
-  scale_x_discrete(labels = paste0("L", 1:6)) +
-  scale_y_discrete(labels = paste0("L", 1:6)) +
-  scale_fill_manual("Cell included in SA(A,k)", values = c("white", "grey20")) +
-  coord_equal() +
-  theme(legend.position = "bottom")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/sam-sketch-1}
+
+}
+
+\caption[Sketch of all six land-to-land sequences between two bullets with six lands]{Sketch of all six land-to-land sequences between two bullets with six lands.}\label{fig:sam-sketch}
+\end{figure}
+
+\end{knitrout}
 
@@
 
-<<h44, echo=FALSE, fig.width = 8, fig.height=4, fig.cap="Hamby set 44: overview of all bullet-to-bullet matches between all pairs of questioned bullets (y-axis) and known test fires from ten barrels (x axis).">>=
-h44features <- read.csv("data/h44-features.csv")
-h44features <- h44features %>% mutate(
-  lys = overlap*signature_length*1000/.645,
-  cms = cms2 *lys*.645/1000
-)
-# tank rash lands:
-#h44features <- h44features %>% mutate(
-#  rfscore = replace(rfscore, (bullet1 == "I" & land1 == 6) | (bullet2 == #"I" & land2 == 6), NA)
-#)
-h44features <- h44features %>% mutate(
-  rfscore = replace(rfscore, (barrel1 == 8 & bullet1 == "1" & land1 == 6) | (barrel2 == 8 & bullet2 == "1" & land2 == 6), NA)
-)
-
-h44nest <- h44features %>%
-  filter(barrel1 == "Unk") %>%
-  filter(barrel1 != barrel2 | bullet1 != bullet2) %>%
-  mutate(
-    barrel1 = "Unknown",
-    barrel2 = as.character(barrel2),
-    barrel2 = replace(barrel2, barrel2 == "Unk", "Unknown")
-  ) %>%
-  group_by(bullet1, barrel2, bullet2) %>%
-  nest()
-
-h44nest <- h44nest %>% mutate(
-  sam_ccf = data %>% purrr::map_dbl(.f = function(d) {
-    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$ccf)
-    max(scores)
-  }),
-  sam_rf = data %>% purrr::map_dbl(.f = function(d) {
-    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore)
-    max(scores)
-  }),
-  sam_cms = data %>% purrr::map_dbl(.f = function(d) {
-    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$cms)
-    max(scores)
-  })
-)
-h44nest <- h44nest %>%
-  mutate(
-    barrel2 = factor(barrel2, levels = c(1:10, "Unknown")),
-    bullet1 = factor(bullet1, levels = rev(c(
-      "K", "O", "L", "P",
-      "J",
-      "H", "I", "Y", "G",
-      "E", "U", "X", "F",
-      "T", "S"
-    ))),
-    bullet2 = factor(bullet2, levels = c(1:2, rev(levels(bullet1))))
-  )
-h44 <- h44nest %>%
-  ggplot(aes(
-    y = bullet1, x = bullet2,
-    fill = sam_rf
-  )) +
-  geom_tile(size = 1) +
-  facet_grid(. ~ barrel2, space = "free", scales = "free") +
-  ylab("Questioned bullets") +
-  xlab("Test fires from barrels 1 to 10 (left), questioned bullets (right)") +
-  scale_fill_gradient2("RF score ",
-    low = "darkgrey",
-    high = "darkorange", midpoint = 0.35, limits = c(0, 1)
-  ) +
-  scale_colour_manual("same source", values = "darkorange") +
-
-  #  scale_x_discrete(position = "top") +
-  scale_y_discrete() +
-  geom_tile(aes(colour = TRUE), size = .5, data = filter(h44nest, sam_rf > 0.3)) +
-  guides(colour = guide_legend(override.aes = list(fill = NA))) + ggtitle("Hamby set 44") +
-  theme_bw() +
-  theme(legend.position = "bottom")
-
-h44nest$samesource <- h44nest$sam_rf > 0.3
-
-h44
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/h44-1}
+
+}
+
+\caption[Hamby set 44]{Hamby set 44: overview of all bullet-to-bullet matches between all pairs of questioned bullets (y-axis) and known test fires from ten barrels (x axis).}\label{fig:h44}
+\end{figure}
+
+\end{knitrout}
 
@@
 
-<<close, eval=FALSE, fig.width=11.25, fig.height=4.2, fig.cap="Examples of land-to-land comparisons from the Hamby-44 set.
-Along the top are the three bullets from different source with the highest RF scores, along the bottom are the three bullets from same source with the lowest RF score.", fig.subcap=c("Ground truth: different source", "Ground truth: same-source"), fig.align="middle", fig.ncol=1>>=
-# ex2 <- h44nest %>% filter(samesource == FALSE & sam_rf >= 0.287)
-# rfs2 <- ex2$sam_rf
-# ex2 <- ex2 %>% mutate(
-#   data = data %>% purrr::map(.f = function(d) {
-#  #   browser()
-#     index <- which.max(bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore))
-#     d$same <- ((d$land2 - d$land1) %% 6)  == index-1
-#
-#   #   d %>% ggplot(aes(x = land1, y=land2, fill=rfscore)) +
-#   #                  geom_tile() +
-#   #     geom_tile(colour="darkorange", data = d %>% filter(same), size=0.5) +
-#   #     scale_fill_gradient2(low="darkgrey", high = "darkorange",
-#   #                      midpoint = 0.5) +
-#   #     scale_colour_manual(values=c( "darkorange")) +
-#   # scale_x_discrete("") +
-#   # scale_y_discrete("")
-#     d
-#   })
-# )
-# ex2$example <- with(ex2, paste0(bullet1, " vs ", barrel2, "-", bullet2))
-# ex2 <- unnest(ex2, data)
-# ex2 %>%
-#   ggplot(aes(x = land1, y = land2, fill=rfscore)) +
-#   geom_tile() +
-#   theme_bw() +
-#   scale_fill_gradient2("RF score", low="darkgrey", high = "darkorange",
-#                        midpoint = 0.5) +
-# #  geom_tile(colour="darkgrey", data = ex2 %>% filter(same), size=0.5) +
-#   facet_grid(.~example) +
-#   scale_x_discrete("") +
-#   scale_y_discrete("") +
-#   annotate(geom="text", x = 0.5, y = 6.75, label=sprintf("SAM RF: %.3f", rfs2), hjust = 0)
-#
-#
-#
-# ex1 <- h44nest %>% filter(samesource == TRUE & sam_rf < 0.4)
-# rfs <- ex1$sam_rf
-# ex1 <- ex1 %>% mutate(
-#   data = data %>% purrr::map(.f = function(d) {
-#  #   browser()
-#     index <- which.max(bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore))
-#     d$same <- ((d$land2 - d$land1) %% 6)  == index-1
-#
-#   #   d %>% ggplot(aes(x = land1, y=land2, fill=rfscore)) +
-#   #                  geom_tile() +
-#   #     geom_tile(colour="darkorange", data = d %>% filter(same), size=0.5) +
-#   #     scale_fill_gradient2(low="darkgrey", high = "darkorange",
-#   #                      midpoint = 0.5) +
-#   #     scale_colour_manual(values=c( "darkorange")) +
-#   # scale_x_discrete("") +
-#   # scale_y_discrete("")
-#     d
-#   })
-# )
-# ex1$example <- with(ex1, paste0(bullet1, " vs ", barrel2, "-", bullet2))
-# ex1 <- unnest(ex1, data)
-# ex1 %>%
-#   ggplot(aes(x = land1, y = land2, fill=rfscore)) +
-#   geom_tile() +
-#   theme_bw() +
-#   scale_fill_gradient2("RF score", low="darkgrey", high = "darkorange",
-#                        midpoint = 0.5, na.value = NA) +
-#   geom_tile(colour="darkorange", data = ex1 %>% filter(same), size=0.5) +
-#   facet_grid(.~example) +
-#   scale_x_discrete("") +
-#   scale_y_discrete("") +
-#   annotate(geom="text", x = 0.5, y = 6.75, label=sprintf("SAM RF: %.3f", rfs), hjust = 0)
-
-# \autoref{fig:close} shows an example of six comparisons of bullet pairs at the land-to-land level.
-The panels along the top of the figure show three bullet pairs that are known to be from different barrels, the three bullet pairs along the bottom are known to be from the same barrel.
-The light orange frame in the bottom three panels indicate which lands correspond to one-another in the matching.
-It is obvious that not all of the scores are as high as we would usually see in same-source comparisons.
-@
-
 \subsection{Phoenix PD}
 
-<<pd, echo=FALSE, fig.width = 8, fig.height=3.56, fig.cap="Phoenix study: overview of all bullet-to-bullet matches between all pairs of questioned bullets (y-axis) and known test fires from ten barrels (x axis).
-The order of the bullets on the y axis is determined by matching barrel.">>=
-features <- read.csv("data/pd-features.csv.gz")
-features <- features %>% mutate(
-  lys = overlap*signature_length*1000/.645,
-  cms = cms2 *lys*.645/1000
-)
-# change labels for land 3 and land 4 of bullet B3 in P7
-features <- features %>% mutate(
-  first = as.character(first),
-  first = replace(first, first=="Gun 1-P7/B3/L4.dat", "XXX"),
-  first = replace(first, first=="Gun 1-P7/B3/L3.dat", "Gun 1-P7/B3/L4.dat"),
-  first = replace(first, first=="XXX", "Gun 1-P7/B3/L3.dat")
-)
-
-features <- features %>% mutate(
-  KM = replace(KM, first %in% c("Gun 1-P7/B3/L3.dat", "Gun 1-P7/B3/L4.dat"), FALSE)
-)
-features <- features %>% mutate(
-  KM = ifelse(first == "Gun 1-P7/B3/L3.dat", (second == "Unknown 1-J/L3.dat"), KM),
-  KM = ifelse(first == "Gun 1-P7/B3/L4.dat", (second == "Unknown 1-J/L4.dat"), KM)
-)
-
-f2 <- features %>%
-  separate(first, into = c("foo1", "foo2", "barrel1", "bullet1", "land1", "foo3"), remove = FALSE) %>%
-  separate(second, into = c("foo4", "foo5", "bullet2", "land2", "foo6"), remove = FALSE) %>%
-  select(-foo1, -foo2, -foo3, -foo4, -foo5, -foo6)
-
-unknowns <- read.csv("data/pd-features-unknown.csv")
-unknowns <- unknowns %>% filter(bullet1 != bullet2)
-unknowns <- unknowns %>% mutate(
-  bullet2 = gsub("Unknown 1-", "", bullet2),
-  bullet1 = gsub("Unknown 1-", "", bullet1)
-)
-f3 <- rbind(
-  f2 %>%
-    select(
-      b1, b2, barrel1, bullet1,
-      land1, bullet2, land2, ccf, rfscore, cms, KM
-    ) %>%
-    mutate(barrel2 = "Unknown"),
-  unknowns %>%
-    select(
-      b1, b2, bullet1,
-      land1, bullet2, land2, ccf, rfscore, cms, 
-    ) %>%
-    mutate(barrel1 = "Unknown", barrel2 = "Unknown", KM = FALSE)
-)
-
-pdnest <- f3 %>% group_by(bullet2, barrel1, bullet1) %>% nest()
-
-pdnest <- pdnest %>% mutate(
-  sam_ccf = data %>% purrr::map_dbl(.f = function(d) {
-    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$ccf)
-    max(scores)
-  }),
-  sam_rf = data %>% purrr::map_dbl(.f = function(d) {
-    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore)
-    max(scores)
-  }),
-  KM = data %>% purrr::map_lgl(.f = function(d) {
-    any(d$KM)
-  }),
-  sam_cms = data %>% purrr::map_dbl(.f = function(d) {
-    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$cms)
-    max(scores)
-  })
-)
-
-pdnest <- pdnest %>%
-  mutate(
-    bullet2 = factor(bullet2, levels = rev(c("N", "B", "E", "T", "H", "J", "K", "Q", "Y", "Z"))),
-    bullet1 = factor(bullet1, levels = c("B1", "B2", "B3", rev(levels(bullet2))))
-  )
-p3 <- pdnest %>%
-  ggplot(aes(y = bullet2, x = bullet1, fill = sam_rf)) +
-  geom_tile(size = 1) +
-  facet_grid(. ~ barrel1, scales = "free", space = "free") +
-  ylab("Questioned bullets") +
-  xlab("Known bullets and Questioned bullets") +
-  scale_fill_gradient2("RF score ",
-    low = "darkgrey",
-    high = "darkorange", midpoint = 0.45, limits = c(0, 1)
-  ) +
-  scale_colour_manual("same source", values = "darkorange") +
-  scale_y_discrete() +
-  theme_bw() +
-  theme(legend.position = "bottom") +
-  #  coord_equal() +
-  geom_tile(aes(colour = TRUE),
-    size = .5,
-    data = filter(pdnest, KM)
-  ) +
-  guides(colour = guide_legend(override.aes = list(fill = NA)))
-
-pdnest$samesource <- pdnest$KM
-
-p3 + ggtitle("Phoenix PD set")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/pd-1}
+
+}
+
+\caption[Phoenix study]{Phoenix study: overview of all bullet-to-bullet matches between all pairs of questioned bullets (y-axis) and known test fires from ten barrels (x axis).
+The order of the bullets on the y axis is determined by matching barrel.}\label{fig:pd}
+\end{figure}
+
+\end{knitrout}
 
@@
 
-<<hou, fig.width=8, fig.height = 2.67, fig.cap="Overview of matching scores for all pairs of questioned bullets to known bullets from five barrels (on the left) and questioned bullets to themselves (tiles on the right).", fig.subcap=c("Set 1", "Set 2", "Set 3"), fig.align="middle", fig.ncol=1>>=
-features <- read.csv("data/fsi-features.csv.gz")
-features <- features %>% mutate(
-  lys = overlap*signature_length*1000/.645,
-  cms = cms2 *lys*.645/1000
-)
-
-features <- features %>% mutate(
-  bullet1 = gsub("ullet ", "", bullet1),
-  bullet2 = gsub("ullet ", "", bullet2)
-)
-
-features <- features %>%
-  filter(group1 == group2) %>%
-  filter(barrel1 == "Unknowns") %>%
-  filter(barrel1 != barrel2 | bullet1 != bullet2)
-
-groups <- features %>%
-  group_by(group1, barrel1, bullet1, barrel2, bullet2) %>%
-  nest()
-
-hounest <- groups %>% mutate(
-  sam_ccf = data %>% purrr::map_dbl(.f = function(d) {
-    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$ccf)
-    max(scores)
-  }),
-  sam_rf = data %>% purrr::map_dbl(.f = function(d) {
-    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$rfscore)
-    max(scores)
-  }),
-  sam_cms = data %>% purrr::map_dbl(.f = function(d) {
-    scores <- bulletxtrctr::compute_average_scores(d$land1, d$land2, d$cms)
-    max(scores)
-  })
-)
-
-hounest$samesource <- hounest$sam_rf > 0.3
-
-hou1nest <- hounest %>%
-  filter(group1 == "G1") %>%
-  mutate(
-    bullet1 = factor(bullet1, levels = rev(c("U40", "U36", "U42", "U10", "U77", "U37", "U28", "U15"))),
-    bullet2 = factor(bullet2, levels = c("B1", "B2", "B3", levels(bullet1)))
-  )
-
-hou1 <- hou1nest %>%
-  ggplot(aes(y = bullet1, x = bullet2, fill = sam_rf)) +
-  geom_tile(size = 1) +
-  facet_grid(. ~ barrel2, scales = "free", space = "free") +
-  ylab("Questioned bullets") +
-  xlab("Known bullets and Questioned bullets") +
-  scale_fill_gradient2("RF score ",
-    low = "darkgrey",
-    high = "darkorange", midpoint = 0.375, limits = c(0, 1)
-  ) +
-  scale_y_discrete() +
-  scale_colour_manual("same source", values = "darkorange") +
-  geom_tile(aes(colour = TRUE), size = .5, data = filter(hou1nest, sam_rf > 0.3)) +
-  guides(colour = guide_legend(override.aes = list(fill = NA))) + ggtitle("Houston set 1")
-
-hou2nest <- hounest %>%
-  filter(group1 == "G2") %>%
-  mutate(
-    bullet1 = factor(bullet1, levels = rev(c("U23", "U41", "U61", "U98", "U73", "U34", "U63", "U66"))),
-    bullet2 = factor(bullet2, levels = c("B1", "B2", "B3", levels(bullet1)))
-  )
-hou2 <- hou2nest %>%
-  ggplot(aes(y = bullet1, x = bullet2, fill = sam_rf)) +
-  geom_tile(size = 1) +
-  facet_grid(. ~ barrel2, scales = "free", space = "free") +
-  ylab("Questioned bullets") +
-  xlab("Known bullets and Questioned bullets") +
-  scale_fill_gradient2("RF score ",
-    low = "darkgrey",
-    high = "darkorange", midpoint = 0.375, limits = c(0, 1)
-  ) +
-  scale_y_discrete() +
-  scale_colour_manual("same source", values = "darkorange") +
-  geom_tile(aes(colour = TRUE), size = .5, data = filter(hou2nest, sam_rf > 0.3)) +
-  guides(colour = guide_legend(override.aes = list(fill = NA))) + ggtitle("Houston set 2")
-
-hou3nest <- hounest %>%
-  filter(group1 == "G3") %>%
-  mutate(
-    bullet1 = factor(bullet1, levels = rev(c("U27", "U33", "U36", "U49", "U56", "U65", "U14", "U45"))),
-    bullet2 = factor(bullet2, levels = c("B1", "B2", "B3", levels(bullet1)))
-  )
-
-hou3 <- hou3nest %>%
-  ggplot(aes(y = bullet1, x = bullet2, fill = sam_rf)) +
-  geom_tile(size = 1) +
-  facet_grid(. ~ barrel2, scales = "free", space = "free") +
-  ylab("Questioned bullets") +
-  xlab("Known bullets and Questioned bullets") +
-  scale_fill_gradient2("RF score ",
-    low = "darkgrey",
-    high = "darkorange", midpoint = 0.375, limits = c(0, 1)
-  ) +
-  scale_y_discrete() +
-  scale_colour_manual("same source", values = "darkorange") +
-  geom_tile(aes(colour = TRUE), size = .5, data = filter(hou3nest, sam_rf > 0.3)) +
-  guides(colour = guide_legend(override.aes = list(fill = NA))) + ggtitle("Houston set 3")
-
-hou1
-
-hou2
-
-hou3
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+\subfloat[Set 1\label{fig:hou1}]{\includegraphics[width=\textwidth]{figures/hou-1} }\newline
+\subfloat[Set 2\label{fig:hou2}]{\includegraphics[width=\textwidth]{figures/hou-2} }\newline
+\subfloat[Set 3\label{fig:hou3}]{\includegraphics[width=\textwidth]{figures/hou-3} }\caption[Overview of matching scores for all pairs of questioned bullets to known bullets from five barrels (on the left) and questioned bullets to themselves (tiles on the right)]{Overview of matching scores for all pairs of questioned bullets to known bullets from five barrels (on the left) and questioned bullets to themselves (tiles on the right).}\label{fig:hou}
+\end{figure}
+
+\end{knitrout}
 
@@
 
-<<compare, fig.width = 8, fig.height = 7, fig.cap="Density curves of similarity scores from random forest scores (top), cross correlation (middle), and consecutive matching striae (bottom).
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/compare-1}
+
+}
+
+\caption[Density curves of similarity scores from random forest scores (top), cross correlation (middle), and consecutive matching striae (bottom)]{Density curves of similarity scores from random forest scores (top), cross correlation (middle), and consecutive matching striae (bottom).
 Different colors indicate same source versus different source.
-Ideally all scores of different source pairs should be much lower than scores for same source pairs.">>=
-nests <- rbind(
-  h44nest %>%
-    select(sam_rf, sam_ccf, sam_cms, samesource) %>%
-    mutate(study = "Hamby 44", group1 = NA),
-  pdnest %>%
-    select(sam_rf, sam_ccf, sam_cms, samesource) %>%
-    mutate(study = "Phoenix PD", group1 = NA),
-  hounest %>%
-    select(sam_rf, sam_ccf, sam_cms, samesource, group1) %>%
-    mutate(study = "Houston FSI")
-)
-nests$study <- factor(nests$study, levels = c("Hamby 44", "Phoenix PD", "Houston FSI"))
-nestslong <- nests %>% gather(key = "measure", value = "value", sam_rf, sam_ccf, sam_cms)
-nestslong$measure <- factor(nestslong$measure)
-levels(nestslong$measure) <- c("Cross correlation", "CMS", "RF score")
-
-labels <- h44nest %>% filter(samesource, sam_rf < 0.5)
-labels$comparison <- with(labels, paste(bullet1, "vs", paste(barrel2, bullet2, sep = "-")))
-labels$measure <- "RF score"
-labels$study <- "Hamby 44"
-labels$y <- 5 + 4:1
-p1 <- nestslong %>%
-  filter(measure %in% c("RF score")) %>%
-  ggplot(aes(
-    x = value, fill = samesource,
-    colour = samesource
-  )) +
-  geom_density(alpha = 0.6) +
-  geom_rug(alpha = 0.3, show.legend = F) +
-  facet_grid(measure ~ study) +
-  theme_bw() +
-  scale_fill_manual("Same source", values = c("darkgrey", "darkorange"), guide = F) +
-  scale_colour_manual("Same source", values = c("darkgrey", "darkorange"), guide = F) +
-  scale_x_continuous("SAM scores", limits=c(0,1)) +
-  theme(legend.position = "bottom") +
-  geom_hline(yintercept = 0, show.legend = F) +
-  geom_text(aes(x = sam_rf, y = y, label = comparison),
-            show.legend = F,
-            hjust = 0,
-            data = labels
-  ) +
-  geom_segment(aes(x = sam_rf, xend = sam_rf, y = 0, yend = y),
-               show.legend = F,
-               size = 0.1,
-               colour = "grey20",
-               data = labels
-  ) + 
-  theme(axis.title.x = element_blank(), axis.text.y = element_text(hjust = .5, angle = 90))
-
-p2 <- nestslong %>%
-  filter(measure %in% c("Cross correlation")) %>% 
-  ggplot(aes(
-    x = value, fill = samesource,
-    colour = samesource
-  )) +
-  geom_density(alpha = 0.6) +
-  geom_rug(alpha = 0.3, show.legend = F) +
-  facet_grid(measure ~ study) +
-  theme_bw() +
-  theme(legend.position = "bottom") +
-  scale_fill_manual("Same source", values = c("darkgrey", "darkorange"), guide = F) +
-  scale_colour_manual("Same source", values = c("darkgrey", "darkorange"), guide = F) +
-  scale_x_continuous("SAM scores") + 
-  theme(axis.title.x = element_blank(), axis.text.y = element_text(hjust = .5, angle = 90))
-
-p3 <- nestslong %>%
-  filter(measure %in% c("CMS")) %>% 
-  ggplot(aes(
-    x = value, fill = samesource,
-    colour = samesource
-  )) +
-  geom_density(alpha = 0.6) +
-  geom_rug(aes(x = jitter(value)), alpha = 0.3, show.legend = F) +
-  facet_grid(measure ~ study) +
-  theme_bw() +
-  theme(legend.position = "bottom") +
-  scale_fill_manual("Same source", values = c("darkgrey", "darkorange")) +
-  scale_colour_manual("Same source", values = c("darkgrey", "darkorange")) +
-  scale_x_continuous("SAM scores")  + 
-  theme(axis.text.y = element_text(hjust = .5, angle = 90))
-
-grid.arrange(p1, p2, p3, nrow = 3, heights = c(1, 1, 1.4))
-@
+Ideally all scores of different source pairs should be much lower than scores for same source pairs.}\label{fig:compare}
+\end{figure}
+
+\end{knitrout}
 
@@
 These issues affect the algorithm's ability to identify a stable signature, which impacts the subsequent similarity scores at the land-to-land and bullet-to-bullet level.
-<<tankrash, fig.cap="Overview of bullet lands with prominent deficiency such as tank rash or extreme pitting in the Hamby-44 study.", fig.width=4, fig.height = 2, out.width='47.5%', fig.subcap=c("Br1-B1-L6 (tank rask)", "Br2-B2-L5 (tank rash)", "Br3-B1-L5 (tank rash)", "Br8-B1-L6 (tank rash)", "Br8-B2-L2 (pitting)", "Br8-B2-L6 (tank rash)", "Unk-E-L6 (tank rash)", "Unk-I-L6 (tank rash and break-off)"), fig.align="middle", fig.ncol=2>>=
-knitr::include_graphics(
-  c(
-    "./images/B1-B1-L6.png", "./images/B2-B2-L5.png",
-    "./images/B3-B1-L5.png", "./images/B8-B1-L6.png",
-    "./images/B8-B2-L2.png", "./images/B8-B2-L6.png",
-    "./images/NA-BE-L6.png", "./images/NA-BI-L6.png"
-  )
-)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+\subfloat[Br1-B1-L6 (tank rask)\label{fig:tankrash1}]{\includegraphics[width=0.475\linewidth]{./images/B1-B1-L6} }
+\subfloat[Br2-B2-L5 (tank rash)\label{fig:tankrash2}]{\includegraphics[width=0.475\linewidth]{./images/B2-B2-L5} }\newline
+\subfloat[Br3-B1-L5 (tank rash)\label{fig:tankrash3}]{\includegraphics[width=0.475\linewidth]{./images/B3-B1-L5} }
+\subfloat[Br8-B1-L6 (tank rash)\label{fig:tankrash4}]{\includegraphics[width=0.475\linewidth]{./images/B8-B1-L6} }\newline
+\subfloat[Br8-B2-L2 (pitting)\label{fig:tankrash5}]{\includegraphics[width=0.475\linewidth]{./images/B8-B2-L2} }
+\subfloat[Br8-B2-L6 (tank rash)\label{fig:tankrash6}]{\includegraphics[width=0.475\linewidth]{./images/B8-B2-L6} }\newline
+\subfloat[Unk-E-L6 (tank rash)\label{fig:tankrash7}]{\includegraphics[width=0.475\linewidth]{./images/NA-BE-L6} }
+\subfloat[Unk-I-L6 (tank rash and break-off)\label{fig:tankrash8}]{\includegraphics[width=0.475\linewidth]{./images/NA-BI-L6} }\caption[Overview of bullet lands with prominent deficiency such as tank rash or extreme pitting in the Hamby-44 study]{Overview of bullet lands with prominent deficiency such as tank rash or extreme pitting in the Hamby-44 study.}\label{fig:tankrash}
+\end{figure}
+
+\end{knitrout}
 
@@
 
-<<cache=FALSE>>=
-hounest <- hounest %>% mutate(
-  data = data %>% purrr::map(
-    .f = function(d) {
-      d$sameland = bullet_to_land_predict(
-        land1 = d$land1, land2 = d$land2, 
-        d$rfscore, difference=0.05)
-      d
-    })
-)
-houlands <- hounest %>% unnest(data)
-
-set.seed(20140501)
-h44nest$samesource <- h44nest$sam_rf > 0.3
-h44nest <- h44nest %>% 
-  mutate(
-    data = purrr::map2(.x = data, .y = samesource, .f = function(d, y) {
-      d$samesource <- y
-      d
-  }
-  ))
-
-h44nest <- h44nest %>% mutate(
-  data = data %>% purrr::map(
-    .f = function(d) {
-      if (any(d$samesource)) {
-      d$sameland = bullet_to_land_predict(
-        land1 = d$land1, land2 = d$land2, 
-        d$rfscore, difference=0.01)
-      } else d$sameland <- FALSE
-      d <- d %>% select(-samesource)
-      d
-    })
-)
-h44lands <- h44nest %>% unnest(data)
-
-pdnest <- pdnest %>% mutate(
-  data = data %>% purrr::map(
-    .f = function(d) {
-      d$sameland = bullet_to_land_predict(
-        land1 = d$land1, land2 = d$land2, 
-        d$rfscore, difference=0)
-      d
-    })
-)
-pdlands <- pdnest %>%
-  rename(bullet_KM = KM) %>% 
-  unnest(data)
-
-all_lands <- rbind(
-  h44lands %>%
-    select(rfscore, ccf, cms, sameland, samesource) %>%
-    mutate(study = "Hamby 44", group1 = ""),
-  pdlands %>%
-    select(rfscore, ccf, cms, sameland, samesource) %>%
-    mutate(study = "Phoenix PD", group1 = ""),
-  houlands %>%
-    select(rfscore, ccf, cms, sameland, samesource, group1) %>%
-    mutate(study = "Houston FSI")
-)
-all_lands$study <- factor(all_lands$study, levels = c("Hamby 44", "Phoenix PD", "Houston FSI"))
-all_lands <- all_lands %>%
-  mutate(
-    sameland = replace(sameland, !samesource, FALSE)
-  )
-@
-
-<<check44, eval=FALSE>>=
-h44lands <- h44lands %>%
-  mutate(
-    sameland = replace(sameland, !samesource, FALSE)
-  )
-h44lands %>% 
-  ggplot(aes(x = land1, y = land2, fill = ccf)) +
-  geom_tile() +
-  geom_tile(aes(colour = sameland), size = 1, 
-            data = h44lands %>% filter(sameland)) +
-  scale_fill_gradient2("RF score: ", midpoint=0.5, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
-  scale_colour_manual(values="darkorange") +
-  facet_grid(bullet1~barrel2+bullet2)
-
-@
-
-<<checkpd, eval=FALSE>>=
-pdlands <- pdlands %>%
-  mutate(
-    sameland = replace(sameland, !samesource, FALSE)
-  )
-pdlands %>% 
-  ggplot(aes(x = land1, y = land2, fill = ccf)) +
-  geom_tile() +
-  geom_tile(aes(colour = sameland), size = 1, 
-            data = pdlands %>% filter(sameland)) +
-  scale_fill_gradient2("RF score: ", midpoint=0.5, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
-  scale_colour_manual(values="darkorange") +
-  facet_grid(bullet1~barrel2+bullet2)
-
-@
-
-<<checkhou, eval=FALSE>>=
-houlands <- houlands %>%
-  mutate(
-    sameland = replace(sameland, !samesource, FALSE)
-  )
-hou1 <- houlands %>% filter(group1==group2, group1=="G1")
-
-hou1 %>%
-  ggplot(aes(x = land1, y = land2, fill = ccf)) +
-  geom_tile() +
-  geom_tile(aes(colour = sameland), size = 1, 
-            data = hou1 %>% filter(sameland)) +
-  scale_fill_gradient2("RF score: ", midpoint=0.5, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
-  scale_colour_manual(values="darkorange") +
-  facet_grid(bullet1~barrel2+bullet2)
-
-hou2 <- houlands %>% filter(group1==group2, group1=="G2")
-
-hou2 %>%
-  ggplot(aes(x = land1, y = land2, fill = ccf)) +
-  geom_tile() +
-  geom_tile(aes(colour = sameland), size = 1, 
-            data = hou2 %>% filter(sameland)) +
-  scale_fill_gradient2("RF score: ", midpoint=0.5, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
-  scale_colour_manual(values="darkorange") +
-  facet_grid(bullet1~barrel2+bullet2)
-
-hou3 <- houlands %>% filter(group1==group2, group1=="G3")
-
-hou3 %>%
-  ggplot(aes(x = land1, y = land2, fill = ccf)) +
-  geom_tile() +
-  geom_tile(aes(colour = sameland), size = 1, 
-            data = hou3 %>% filter(sameland)) +
-  scale_fill_gradient2("RF score: ", midpoint=0.5, low = "darkgrey", high = "darkorange", limits = c(0, 1)) +
-  scale_colour_manual(values="darkorange") +
-  facet_grid(bullet1~barrel2+bullet2)  
-@
-
-<<compare-land-to-land, fig.width = 8, fig.height = 6.5, fig.cap="Density curves of land-to-land similarity scores from RF scores (top), cross-correlation (middle), and consecutive matching striae (CMS)(bottom).
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/compare-land-to-land-1}
+
+}
+
+\caption[Density curves of land-to-land similarity scores from RF scores (top), cross-correlation (middle), and consecutive matching striae (CMS)(bottom)]{Density curves of land-to-land similarity scores from RF scores (top), cross-correlation (middle), and consecutive matching striae (CMS)(bottom).
 Different colors indicate same source versus different source for each land.
 RF scores for different source comparisons are generally well below 0.5.
-Some same-source comparisons are also below 0.5 for RF scores, indicating potential problems with at least one of the lands involved in the comparison.">>=
-# all_lands %>% 
-#   gather(key = "measure", value = "value", ccf, rfscore) %>%
-#   mutate(measure = factor(measure, levels = c("ccf", "rfscore"),
-#                           labels = c("Cross correlation", "RF score")),
-#          study = paste(study, group1) %>% str_trim() %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"))) %>%
-#   ggplot(aes(
-#     x = value, fill = sameland,
-#     colour = sameland
-#   )) +
-#   geom_density(alpha = 0.6) +
-# #  geom_rug(alpha = 0.5/12, show.legend = F) +
-#   facet_grid(measure ~ study) +
-#   theme_bw() +
-#   scale_fill_manual("Same source", values = c("darkgrey", "darkorange")) +
-#   scale_colour_manual("Same source", values = c("darkgrey", "darkorange")) +
-#   scale_x_continuous("Land-to-Land scores", limits = c(0, 1), labels = c(0, 0.25, 0.5, 0.75, 1)) +
-#   theme(legend.position = "bottom") +
-#   geom_hline(yintercept = 0, show.legend = F)
-p1 <- all_lands %>% 
-  gather(key = "measure", value = "value", ccf, cms, rfscore) %>%
-  mutate(measure = factor(measure, levels = c("ccf", "cms", "rfscore"),
-                          labels = c("Cross correlation", "CMS", "RF score")),
-         study = paste(study, group1) %>% str_trim() %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"))) %>%
-  filter(measure == "RF score") %>%
-  ggplot(aes(
-    x = value, fill = sameland,
-    colour = sameland
-  )) +
-  geom_density(alpha = 0.6) +
-#  geom_rug(alpha = 0.5/12, show.legend = F) +
-  facet_grid(measure ~ study) +
-  theme_bw() +
-  scale_fill_manual("Same source", values = c("darkgrey", "darkorange"), guide = F) +
-  scale_colour_manual("Same source", values = c("darkgrey", "darkorange"), guide = F) +
-  scale_x_continuous("Land-to-Land scores", limits = c(0, 1), labels = c(0, 0.25, 0.5, 0.75, 1)) +
-  theme(legend.position = "bottom", axis.title.x = element_blank(), axis.text.y = element_text(hjust = .5, angle = 90)) +
-  geom_hline(yintercept = 0, show.legend = F)
-
-p2 <- all_lands %>% 
-  gather(key = "measure", value = "value", ccf, cms, rfscore) %>%
-  mutate(measure = factor(measure, levels = c("ccf", "cms", "rfscore"),
-                          labels = c("Cross correlation", "CMS", "RF score")),
-         study = paste(study, group1) %>% str_trim() %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"))) %>%
-  filter(measure == "Cross correlation") %>%
-  ggplot(aes(
-    x = value, fill = sameland,
-    colour = sameland
-  )) +
-  geom_density(alpha = 0.6) +
-#  geom_rug(alpha = 0.5/12, show.legend = F) +
-  facet_grid(measure ~ study) +
-  theme_bw() +
-  scale_fill_manual("Same source", values = c("darkgrey", "darkorange"), guide = F) +
-  scale_colour_manual("Same source", values = c("darkgrey", "darkorange"), guide = F) +
-  scale_x_continuous("Land-to-Land scores", limits = c(0, 1), labels = c(0, 0.25, 0.5, 0.75, 1)) +
-  theme(legend.position = "bottom", axis.title.x = element_blank(), axis.text.y = element_text(hjust = .5, angle = 90)) +
-  geom_hline(yintercept = 0, show.legend = F)
-
-p3data <- all_lands %>% 
-  gather(key = "measure", value = "value", ccf, cms, rfscore) %>%
-  mutate(measure = factor(measure, levels = c("ccf", "cms", "rfscore"),
-                          labels = c("Cross correlation", "CMS", "RF score")),
-         study = paste(study, group1) %>% str_trim() %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"))) %>%
-  filter(measure == "CMS")
-p3bw <- bw.nrd0(p3data$value[p3data$sameland & p3data$study == "Houston FSI G1"])
-p3 <- ggplot(p3data, aes(
-    x = value, fill = sameland,
-    colour = sameland
-  )) +
-  geom_density(alpha = 0.6, bw = p3bw) +
-  facet_grid(measure ~ study) +
-  theme_bw() +
-  scale_fill_manual("Same source", values = c("darkgrey", "darkorange")) +
-  scale_colour_manual("Same source", values = c("darkgrey", "darkorange")) +
-  scale_x_continuous("Land-to-Land scores") +
-  theme(legend.position = "bottom", axis.text.y = element_text(hjust = .5, angle = 90)) +
-  geom_hline(yintercept = 0, show.legend = F)
-
-grid.arrange(p1, p2, p3, nrow = 3, heights = c(1, 1, 1.4))
-@
+Some same-source comparisons are also below 0.5 for RF scores, indicating potential problems with at least one of the lands involved in the comparison.}\label{fig:compare-land-to-land}
+\end{figure}
+
+\end{knitrout}
 
@@
 
-<<rocs>>=
-library(plotROC)
-library(pROC)
-rocpic <- all_lands %>% gather(measure, score, rfscore, ccf, cms) %>%
-  mutate(
-    measure = factor(measure, labels = c("CC", "CMS", "RF")),
-    studyg = paste(study, group1) %>% str_trim %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"))
-  ) %>%
-  ggplot(aes(linetype = measure, colour = studyg, shape = measure)) +
-  geom_roc(aes(m = score, d = as.numeric(sameland)), 
-           labels = FALSE, pointsize = 0, alpha = 0.8, size = 1) +
-  theme_bw(base_size = 16) +
-  # coord_fixed() + 
-  facet_wrap(~studyg, nrow = 3) +
-  xlab("Percent Wrong Identifications") +
-  ylab("Percent Correct Identifications") +
-  scale_colour_manual("Study", values = c("#fe9929", "#aa0000",
-                                         "#74a9cf", "#2b8cbe", "#045a8d"),
-                      guide=FALSE) +
-  theme(legend.position= c(1, 0), legend.justification = c(1, 0), legend.direction = "vertical",
-        legend.key.width = unit(1, "cm"), axis.text.y = element_text(angle = 90, hjust = .5)) +
-  guides(linetype = guide_legend(override.aes = c(listsize = .5))) + 
-  scale_linetype_manual("Measure", values=c(1,3,2)) +
-  scale_shape_discrete("Measure") +
-  scale_x_continuous(breaks=seq(0,1, by=0.25), labels=c("0", "25%", "50%", "75%", "100%")) +
-  scale_y_continuous(breaks=seq(0,1, by=0.25), labels=c("0", "25%", "50%", "75%", "100%"))
-
-@
-
-<<aucs-pic, include = F>>=
-eer <- function(roc){
-  dframe <- data.frame(sens = roc$sensitivities, 
-                       spec = roc$specificities,
-                       threshold = roc$thresholds)
-  dframe$err1 <- 1 - dframe$sens
-  dframe$err2 <- 1 - dframe$spec
-  dframe$diff <- abs(dframe$err1-dframe$err2)
-  dframe$eer <- (dframe$err1+dframe$err2)/2
-  idx <- which.min(dframe$diff)
-  dframe[idx,c("eer", "threshold")]
-}
-
-aucs <- all_lands %>% gather(measure, score, rfscore, ccf, cms) %>% 
-  group_by(study, group1, measure) %>% nest() %>%
-  mutate(
-    auc = data %>% purrr::map_dbl(function(d) {
-      auc(d$sameland, d$score)
-    }),
-    auc_ci_low = data %>% purrr::map_dbl(function(d) {
-      ci.auc(d$sameland, d$score)[1]
-    }),
-    auc_ci_high = data %>% purrr::map_dbl(function(d) {
-      ci.auc(d$sameland, d$score)[3]
-    }),
-    roc = data %>% purrr::map(function(d) {
-      roc(d$sameland, d$score)
-    }),
-    errors = roc %>% purrr::map(.f = eer)
-  ) %>% 
-  ungroup() %>% 
-  mutate(
-    studyg = paste(study, group1) %>% str_trim %>% factor(levels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"),
-                                                          labels = c("Hamby 44", "Phoenix PD", "Houston FSI\nG1", "Houston FSI\nG2", "Houston FSI\nG3")),
-    measure = factor(measure, levels = c("cms", "ccf", "rfscore"), labels = c("CMS", "CC", "RF")),
-    measurenum = as.numeric(measure),
-    measurechar = as.character(measure)
-  )
-aucs <- aucs %>% unnest(errors)
-
-aucpic <- aucs %>%
-  ggplot(aes(x = auc, y = measurenum, colour = studyg)) +
-  facet_grid(studyg~.) +
-  geom_segment(aes(x = auc_ci_low, xend = auc_ci_low, y = measurenum - .125, yend = measurenum + .125), size = 1) + 
-  geom_segment(aes(x = auc_ci_high, xend = auc_ci_high, y = measurenum - .125, yend = measurenum + .125), size = 1) + 
-  geom_segment(aes(x=auc_ci_low, xend=auc_ci_high, y = measurenum, yend = measurenum, linetype=measure), size=1) +
-  geom_point(size=3.5, aes(shape = measure)) +
-  theme_bw(base_size = 16) +
-  xlab("AUC and 95% confidence intervals") +
-  xlim(c(0.5, 1)) +
-  scale_y_continuous("", breaks = unique(aucs$measurenum), labels = unique(aucs$measurechar), limits=c(0.5, 3.5)) + 
-  scale_colour_manual("Study", values = c("#fe9929", "#aa0000",
-                                         "#74a9cf", "#2b8cbe", "#045a8d"),
-                        guide = FALSE) +
-  theme(axis.text.y = element_text(angle = 90, hjust = 0.5)) + 
-  # theme(legend.position = "bottom",
-  #       legend.key.width = unit(1.5, "cm")) +
-  scale_linetype_manual("Measure", values = c(3,1,2), guide = F) +
-  scale_shape_discrete(guide=F)
-
-@
-
-<<roc-auc, echo = FALSE, out.width = c("0.51\\textwidth", ".4725\\textwidth"), out.height = c("0.63\\textwidth", "0.63\\textwidth"), fig.width = c(5, 6), fig.height = c(7.5, 7.5), fig.subcap=c("ROC curves", "Area under the curve (AUC)"), fig.cap="ROC curves and AUC for each (set) of the studies for the random forest score (RF), cross correlation (CC) and consecutively matching striae (CMS).
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+
+{\centering \subfloat[ROC curves\label{fig:roc-auc1}]{\includegraphics[width=0.51\textwidth,height=0.63\textwidth]{figures/roc-auc-1} }
+\subfloat[Area under the curve (AUC)\label{fig:roc-auc2}]{\includegraphics[width=.4725\textwidth,height=0.63\textwidth]{figures/roc-auc-2} }
+
+}
+
+\caption[ROC curves and AUC for each (set) of the studies for the random forest score (RF), cross correlation (CC) and consecutively matching striae (CMS)]{ROC curves and AUC for each (set) of the studies for the random forest score (RF), cross correlation (CC) and consecutively matching striae (CMS).
 As seen in the ROC curves, the algorithm performs the least well on set 2 of the Houston FSI study.
 Based on AUC, the overall performance on all sets is very good to excellent for both the cross correlation and the random forest score, and moderate for CMS scores.
-At the land level there is no significant difference in prediction power between RF score and cross correlation.", cache=FALSE>>=
-  rocpic 
-  aucpic
-@
+At the land level there is no significant difference in prediction power between RF score and cross correlation.}\label{fig:roc-auc}
+\end{figure}
+
+\end{knitrout}
 
@@
 
-<<eval=FALSE>>=
-xtabs(~sameland+I(round(cms)>=6), data = all_lands)
-@
-
-<<eer, fig.width=6, fig.height = 2.6, fig.cap="Equal error (in Percent)  and corresponding thresholds for all measures (random forest score, cross correlation and CMS) under all five (sub-)studies.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+
+{\centering \includegraphics[width=\textwidth]{figures/eer-1}
+
+}
+
+\caption[Equal error (in Percent)  and corresponding thresholds for all measures (random forest score, cross correlation and CMS) under all five (sub-)studies]{Equal error (in Percent)  and corresponding thresholds for all measures (random forest score, cross correlation and CMS) under all five (sub-)studies.
 CMS has an equal error threshold of 2 for all studies, and perform signifcantly worse than both cross correlation and the random forest score.
-At best, Cross correlation and RF score have equal errors for land-to-land comparisons of around 5\\%.
-The performance of the random forest score is slightly better than the cross-correlation on sets which have higher equal error rates (Houston FSI G2, Hamby 44).", out.width = '\\textwidth'>>=
-aucs %>% 
-  mutate(
-    eer = 100*eer,
-    studyg = factor(studyg, labels = c("Hamby 44", "Phoenix PD", "Houston FSI G1", "Houston FSI G2", "Houston FSI G3"))
-  ) %>%
-  gather(eer_thresh, value, eer, threshold) %>%
-  filter(!(measure=="CMS" & eer_thresh=="threshold")) %>%
-  mutate(
-    eer_thresh = factor(eer_thresh, labels=c("Equal Error (in Percent)", "Threshold for Equal Error"))
-  ) %>%
-  ggplot(aes(y = studyg, x = value, colour=measure, shape = measure)) + 
-  geom_point(size = 3.5) +
-  theme_bw() +
-  xlab("Equal Error % (left) and corresponding thresholds (right)") +
-  ylab("") +
-  scale_color_brewer("Measure", palette = "Set2") +
-  scale_shape_discrete("Measure", solid = F) + 
-  theme(legend.position = "right") +
-  facet_grid(.~eer_thresh, scales="free") +
-  xlim(c(0, NA))
-@
+At best, Cross correlation and RF score have equal errors for land-to-land comparisons of around 5\%.
+The performance of the random forest score is slightly better than the cross-correlation on sets which have higher equal error rates (Houston FSI G2, Hamby 44).}\label{fig:eer}
+\end{figure}
+
+\end{knitrout}
 
@@
 
-<<output-all-figs, include = F>>=
-thisdoc <- readLines("index.Rmd")
-
-# First, get figures that aren't in code chunks
-# Haven't really tested this part - regex may need work
-raw_figures <- data_frame(
-  line_start = which(str_detect(thisdoc, "\\begin\\{figure\\}")), 
-  line_end = which(str_detect(thisdoc, "\\end\\{figure\\}")),
-  subfloats = purrr::map2(line_start, line_end, 
-                          ~str_extract(thisdoc[.x:.y], "\\includegraphics(?:\\[.*\\])?\\{(.*)\\}") %>%
-                            str_replace("\\includegraphics(?:\\[.*\\])?\\{(.*)\\}", "\\1"))
-)
-
-# Code chunk figures
-chunk_figures <- data_frame(
-  line_start = which(str_detect(thisdoc, "^```\\{r.*\\}$")),
-  line_end = which(str_detect(thisdoc, "^``` *$")),
-  chunk_name = str_replace(thisdoc[line_start], "(```\\{r ?)([0-9A-z-]*)(,?.*\\})", "\\2")
-) %>% 
-  mutate(
-    chunk_name = ifelse(chunk_name == "", "unnamed-chunk", chunk_name),
-  ) %>% 
-  mutate(
-    subfloats = purrr::map2(line_start, chunk_name, ~thisdoc[.x] %>%
-                              str_extract("fig\\.subcap ?= ?.*?\\),") %>%
-                              str_replace("fig\\.subcap ?= ?", "") %>%
-                              str_replace(",$", "") %>%
-                              parse(text=.) %>% eval %>% length() %>% 
-                              seq(1, ., by = 1) %>%
-                              sprintf("figures/%s-%d.pdf", .y, .))
-  )
-
-# Knitr included graphics figures
-chunk_incl_figures <- data_frame(
-  line_start = which(str_detect(thisdoc, "^```\\{r.*\\}$")),
-  line_end = which(str_detect(thisdoc, "^``` *$")),
-  chunk_name = str_replace(thisdoc[line_start], "(```\\{r ?)([0-9A-z-]*)(,?.*\\})", "\\2")
-) %>% 
-  mutate(
-    chunk_name = ifelse(chunk_name == "", "unnamed-chunk", chunk_name),
-  ) %>% 
-  mutate(
-    subfloats = purrr::map2(line_start, line_end, function(.x, .y) {
-      thisdoc[.x:.y] %>%
-        paste(collapse = " ") %>%
-        str_extract(., "include_graphics\\(c?\\(?.*?\\)?\\)") %>%
-        str_replace("include_graphics\\(", "") %>%
-        str_extract_all('\\"[\\.A-z0-9 /-]*?\\"') %>%
-        parse(text = .) %>% eval() %>%
-        str_remove_all("\\\\|\"")
-    })
-  )
-
-# All figures
-figs <- bind_rows(raw_figures, chunk_figures, chunk_incl_figures) %>%
-  arrange(line_start) %>%
-  mutate(subfloats = purrr::map(subfloats, function(x) data_frame(file = x, exists = file.exists(x)))) %>%
-  tidyr::unnest() %>%
-  filter(exists) %>%
-  mutate(fig_num = group_indices(., line_start, line_end)) %>%
-  group_by(fig_num) %>%
-  mutate(rn = 1:n(),
-         maxrn = n(),
-         fig_label = letters[rn],
-         fig_label = ifelse(maxrn > 1, fig_label, ""),
-         fileext = tools::file_ext(file),
-         fig_label = sprintf("Figure_%d%s.%s", fig_num, fig_label, fileext))
-
-if (!dir.exists("separate_fig_files")) dir.create("separate_fig_files")
-file.copy(figs$file, file.path("separate_fig_files", figs$fig_label), overwrite = T)
-
-@
 \section{References}
@@
 % <!--include marginal distributions of training data and compare to test data-->
-
-<< echo=FALSE, eval=FALSE>>=
-training <- read.csv("data/features-hamby.csv")
-gtrain <- training %>% 
-  select(profile1_id, profile2_id, study1, study2, match, ccf:sum_peaks) %>%
-  gather(feature, value, ccf:sum_peaks)
-# now compare to test data
-@

### 2020-01-17 old-index.tex
updated filfes, changes, etc.

--- 2020-01-17 index.tex
@@
 
-\title{Comparison of three similarity scores for bullet LEA matching\tnoteref{t1}\tnoteref{t2}}
+\title{Case study validations of automatic bullet matching\tnoteref{t1}\tnoteref{t2}}
 
@@
 \begin{abstract}
-Recent advances in microscopy have made it possible to collect 3D topographic data, enabling more precise virtual comparisons based on the collected 3D data as a supplement to traditional comparison microscopy and 2D photography.
-Automatic comparison algorithms have been introduced for various scenarios, such as matching cartridge cases \citep{tai2018fully,song3DTopographyMeasurements2014} or matching bullet striae \citep{aoas2, chuAutomaticIdentificationBullet2013, dekinderAutomatedComparisonsBullet1999}.
-One key aspect of validating these automatic comparison algorithms is to evaluate the performance of the algorithm on external tests, that is, using data which were not used to train the algorithm.
-Here, we present a discussion of the performance of the matching algorithm \citep{aoas2} in three studies conducted using different Ruger weapons.
-We consider the performance of three scoring measures: random forest score, cross correlation, and consecutive matching striae (CMS) at the land-to-land level and, using Sequential Average Maxima scores, also at the bullet-to bullet level.
-Cross correlation and random forest scores both result in perfect discrimination of same-source and different-source bullets.
-At the land-to-land level, discrimination for both cross correlation and random forest scores (based on area under the curve, AUC) is excellent ($\ge$ 0.90).
+Recent advances in microscopy have made it possible to collect 3D topographic data, enabling virtual comparisons based on the collected 3D data next to traditional comparison microscopy.
+Automatic matching algorithms have been introduced for various scenarios, such as matching cartridge cases \citep{tai2018fully} or matching bullet striae \citep{aoas2, chuAutomaticIdentificationBullet2013, dekinderAutomatedComparisonsBullet1999}.
+One key aspect of validating automatic matching algorithms is to evaluate the performance of the algorithm on external tests.
+Here, we are presenting a discussion of the performance of the matching algorithm \citep{aoas2} in three studies.
+We are considering matching performance based on the Random forest score, cross correlation, and consecutive matching striae (CMS) at the land-to-land level and, using Sequential Average Maxima scores, also at the bullet-to bullet level.
+Cross correlation and Random Forest scores both result in perfect discrimination of same-source and different-source bullets.
+At the land-to-land level, discrimination (based on area under the curve, AUC) is excellent ($\ge$ 0.90).
 \end{abstract}
@@
 Exact guidelines for this classification vary from lab to lab; some labs will exclude only on the basis of non-matching class characteristics, such as direction of the twist in rifling, land length or number of lands, or type of rifling.
-In other labs, CMS (consecutively matching striae) as defined by \citeauthor{biasotti} \citep{biasotti} is used as a measure to quantify the similarity of two lands.
+In other labs, CMS (consecutively matching striae) as defined by \citeauthor{biasotti} \citep{biasotti} is used as a measure to quantify the strength of a match.
 In virtually all labs, individual characteristics used to identify matching bullets are derived from visual assessment; some class characteristics may be directly measured, but these are not sufficient for individualization.
@@
 In many cases, these approaches also provide some level of automation of the comparison process, with the goal of reducing human biases by augmenting the visual information with 3D measurements.
-Utilizing the 3D measurements allows for examination of both peaks and valleys in LEAs.
-It also allows us to take the depth of striae into account; something that is difficult, if not impossible, from visual inspection.
+Utilizing the 3D measurements allows for examination of both peaks and valleys in LEAs, it also allows us to take the depth of striae into account, something that is difficult, if not impossible, from visual inspection.
 
@@
 Individual characteristics, which are not shared by all members of a class, can also be automatically assessed from 3D surface measurements.
-One of the most common features used to describe the similarity of two surfaces is the cross correlation function, which is utilized in several studies \citep{maNISTBulletSignature2004, vorburgerApplicationsCrosscorrelationFunctions2011, chuPilotStudyAutomated2010}.
-Additional features proposed for automatically assessing similarity also include signature distance \citep{maNISTBulletSignature2004}, striae depth and width \citep{ComputerIdentificationBullets1978}, and consecutive matching striae (CMS) \citep{chuAutomaticIdentificationBullet2013}.
+One of the most common features used to assess the strength of a match based on individual characteristics is the cross correlation function, which is utilized in several studies \citep{maNISTBulletSignature2004, vorburgerApplicationsCrosscorrelationFunctions2011, chuPilotStudyAutomated2010}.
+Additional features proposed for automatically assessing match strength also include signature distance \citep{maNISTBulletSignature2004}, striae depth and width \citep{ComputerIdentificationBullets1978}, and consecutive matching striae (CMS) \citep{chuAutomaticIdentificationBullet2013}.
 
@@
 \begin{itemize}
-\item[(a)] Identify an area on the LEA with expressed striae.
-Locate a stable crosscut in the identified area; that is, an area where the striae are similar in the region above and below the crosscut.
-This excludes regions with extreme pitting, breakoff, tank rash, and other flaws that would interfere with similarity score calculation.
+\item[(a)] Identify an area on the LEA with expressed striae. Locate a stable crosscut in the identified area.
 \item[(b)] Discard extraneous/contaminated data, such as data from groove engraved area and areas affected by break-off or contact with objects after the bullet exited the barrel  (``tank rash").
@@
 
-Once signatures have been extracted from stable regions with expressed striae, they must be aligned in order to assess their similarity, just as examiners would manually align two bullets under a comparison microscope.
+Once signatures have been extracted from stable regions with expressed striae, they must be aligned in order to assess their match strength.
 Maximized cross correlation is used to pair-wise align signatures \citep{aoas2,vorburgerApplicationsCrosscorrelationFunctions2011}.
@@
 Using the lag determined for aligning two signatures, other quantitative features describing the similarity of the two signatures can be extracted.
-Numerical features such as cross correlation and Euclidean distance can be computed from the aligned signatures alone.
-Additional features are modeled after visual assessment methods used by examiners, including striae depth, total number of matching striae, and the number of consecutively matching striae (CMS).
+Numerical features such as cross correlation and Euclidean distance
+can be computed from the aligned signatures alone.
+Other features include characteristics of the signature that firearms and toolmark examiners base their visual assessment on, such as striae depth, number of matching striae, and the number of consecutively matching striae (CMS), one of the measures firearms examiners use to describe the strength of a match quantitatively.
 
 In order to evaluate CMS, we must first identify peaks and valleys in each of the signatures, then determine whether these peaks and valleys overlap sufficiently.
-Features which firearms examiners assess visually generally depend on the identification of these extrema, as striation marks and the corresponding peaks are the most salient feature when viewing a bullet using a microscope.
-If there are at least six consecutive matching striae, the bullets are considered to be similar in practice \citep{biasotti_firearms_1997,nichols_scientific_2006}.
-CMS, as measured by examiners, is the number of consecutively matching peaks in signatures of two aligned lands, and, because examiner typically lines up two distinctive markings, resulting in at least one matching striation mark.
-Peaks -- rather than peaks and valleys -- are used in part because it is difficult to visually assess valleys except as relative to peaks.
-In contrast, the matching algorithm can identify both peaks and valleys, and signatures are aligned based on maximum cross-correlation (as opposed to matching striation marks), there is the possibility that two lands will have zero consecutive matching striae.
-Thus, the matching algorithm uses a slightly different measure of CMS than examiners, but the underlying principle is the same.
-The algorithms used to identify peaks and valleys in each signature and determine CMS from the peak and valley identifications are described in more detail in \citep[p.
-2340, step 3 -- 4]{aoas2}.
+Features which firearms examiners assess visually depend on the identification of these extrema.
+If there are at least six consecutive matching striae, the bullets are considered to be a good match in practice \citep{biasotti_firearms_1997,nichols_scientific_2006}.
+CMS in this setting are counted as the number of consecutively matching peaks in signatures of two aligned lands.
+Peaks -- rather than peaks and valleys -- are used for this determination in part because it is difficult to visually assess valleys except as relative to peaks.
+Note that with automatic alignment based on the maximum cross correlation rather than matching striation marks, it is possible for two lands to have a consecutive matching striae of zero.
 
@@
 
-Cross correlation can also be used to assess similarity between two aligned signatures.
+Cross correlation can also be used to assess match strength between two aligned signatures.
 Cross correlation varies between -1 and 1, but as the alignment is based on the maximized cross correlation value, in practice the cross correlations for aligned signatures are generally positive.
-Cross correlation values which are close to 1 indicate similarity between signatures, with the possibility that the two signatures come from the same source.
-Low cross correlation values indicate signatures which are different and thus may originate from different sources.
+Cross correlation values which are close to 1 indicate similarity between signatures and with it may indicate that the two signatures come from the same source.
+Low cross correlation values indicate signatures which may originate from different sources.
 
-The random forest presented in \citet{aoas} is based on a combination of multiple characteristics, including cross correlation, number of matching striae, and number of consecutively matching striae.
-The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the similarity of the two signatures, where scores close to 1 indicate signatures which are similar and may have originated from the same source, while scores close to 0 indicate that the aligned signatures are different and may originate from different sources.
+The random forest presented in \citet{aoas2} is based on a combination of multiple characteristics, including cross correlation, number of matching striae, and number of consecutively matching striae.
+The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where scores close to 1 indicate it to be highly likely that the aligned signatures come from the same source, while scores close to 0 indicate that the aligned signatures are highly likely to originate from different sources.
 
@@
 The features described above are each individually designed to separate same-source and different-source signatures: for instance, same-source signatures would have high cross correlation values and different-source signatures would have low cross correlation values.
-For each feature, or an aggregate score composed of many features, the end result for any two signatures is a number which must be compared to a distribution of other scores from similar situations.
+An automated algorithm selects a cutoff value, introducing a barrier between identification and exclusion.
 
-Using this distribution of scores and training data, an automated algorithm selects a cutoff value, introducing a barrier between identification and exclusion.
 This threshold system introduces a binary classification: identification or exclusion.
@@
 This increased rigor may increase the error rate of the model when compared to examiner error rates, but we expect that the increased information available from the 3D measurement data will compensate for this loss.
-A binary decision model is also easier to interpret and provides more clear-cut, definitive results than the AFTE Theory of Identification system used in most forensic laboratories in the United States.
-Of course, in jurisdictions which utilize score-based likelihood ratios, this threshold system is not necessary because likelihood ratios are continuous.
-Even with likelihood ratios, however, there is a natural threshold at 1 which functions similarly to the selected threshold in the binary decision case.
+A binary decision model is also easier to interpret and provides more clear-cut, definitive results.
 
-The remainder of this paper will utilize the binary decision model which is compatible with the legal framework commonly used in US jurisdictions.
-With this model, we can enumerate characteristics of an ideal similarity scoring mechanism:
-
+We can enumerate characteristics of an ideal similarity scoring mechanism: 
 \begin{description}
 \item[(R1)] {\bf Monotonicity:} a higher score is indicative of higher similarity between a pair of bullets, in particular, similarity scores of same-source pairs of bullets are higher than different-source pairs.
-\item[(R2)] {\bf Stability:} the same score leads to the same conclusion in all situations and under separately assembled reference distributions. 
+\item[(R2)] {\bf Stability:} the same score leads to the same conclusion. 
 \end{description}
@@
 
-\subsection{Assessing Bullet-to-Bullet Similarity}
+\subsection{Assessing Bullet-to-Bullet Match Strength}
 
@@
 While these comparisons are useful, the question of interest typically involves the entire physical object (e.g.
-all lands on a single bullet), and a conclusion on same or different source should be reached based on the evidence of all lands of the two bullets.
+all lands on a single bullet), and a conclusion on same or different source should be reached based on the evidence of all lands of the bullets.
 
-Land-to-land comparisons lead to a whole set of scores to evaluate for bullet-to-bullet comparisons.
+Land-to-land comparisons lead to a whole set of scores for bullet-to-bullet comparisons.
 \autoref{fig:b2b} shows two matrices of scores for two pairs of bullet-to-bullet matches.
@@
 
-While some imaging systems, such as BulletTrax 3D\footnote{https://www.ultra-forensictechnology.com/en/our-products/ballistic-identification/bullettrax/} or BalScan\footnote{https://www.forensic.cz/en/products/balscan}, may capture lower-resolution scans of an entire bullet at one time, confocal light microscopes do not have 360$^\circ$ capability.
-Thus, when imaging bullets using a confocal light microscope, operators scan one land at a time in a clockwise (left twisted rifling) or anti-clockwise (right twisted rifling) sequence.
+When imaging bullets, operators scan one land at a time in a clockwise (left twisted rifling) or anti-clockwise (right twisted rifling) sequence.
 The order in which scans are acquired is kept as meta-information.
-
 Let us assume that lands on a bullet are labelled $\ell_i$ with $i = 1, .., p$, where $p$ indicates the number of lands a bullet has, as determined by the rifling of a barrel.
-For all of the bullets considered here the number of lands, $p$, is 6.
+For all of the bullets considered here the number of lands $p$ is 6.
 A match between  a pair of lands from two bullets therefore results in an expected additional $p-1$ matches between pairs of lands.
@@
 \begin{definition}{Sequence Average and its Maximum}
-Let $A$ be a square real-valued matrix of dimensions $p \times p$.
-For the purpose of this paper, $A$ consists of scores describing the similarity between two sets of land engraved areas.
-The entries in $A$ are represented as $a_{i,j}$, where $i$ and $j$ are row and column indexes.
-
+Let $A$ be a square real-valued matrix of dimensions $p \times p$. For the purpose of this paper, $A$ consists of scores describing the similarity between two sets of land engraved areas.
 The $k$th \emph{sequence average} $SA(A, k)$ for $k = 0, ... p-1$ is defined  as 
@@
 
+In this paper, we validate the algorithm described in \citet{aoas2} on three external test sets.
+We assess the results on each set and evaluate whether the minimal requirements described above are fulfilled for bullet-to-bullet SAM scores.
+We compare the random forest algorithm's performance to the performance of other suggested measures for quantitative assessment of match strength, such as cross correlation and consecutive matching striae.
+
+\section{Validation Sets}
+
 Random forests \citep{breiman} have a built-in internal testing mechanism to prevent potential overfitting.
-Errors reported by random forest algorithms are based on these internal test sets, though there is some debate about the bias of these errors \citep{rfOver2, rfOver}.
+Errors reported by random forest algorithms are based on these internal test sets.
+However, there is some discussion on whether or in which situations these errors are biased \citep{rfOver2, rfOver}.
 Neither of these papers addresses another issue with internal test sets: internal test sets are constructed to have the same distribution as the training data (apart from sampling variability).
@@
 
-In this paper, we validate the algorithm described in \citet{aoas2} on three external test sets.
-We assess the results on each set and evaluate whether the minimal requirements described above are fulfilled for bullet-to-bullet SAM scores.
-We compare the random forest algorithm's performance to the performance of other suggested measures for quantitative assessment of bullet similarity, such as cross correlation and consecutive matching striae.
-
-\section{Validation Sets}
-
 The algorithm in \citet{aoas2} is trained on scans from Hamby sets 252 and 173 made available through \citet{nist}.
-Set 173 was originally published as Hamby set 44, but the mis-labeling has been recently corrected.
+Set 173 was originally published as Hamby set 44.
+This mis-labeling has been corrected now.
 The scans used to train the model were taken at 20 fold magnification for a resolution of 1.5625 microns per pixel.
-While magnification is generally of interest in microscopy, \emph{resolution} -- usually measured in microns per pixel -- is of more interest for 3D topographic measurements, as it determines the operative level of available data.
+While magnification is generally of interest in microscopy, **resolution** -- usually measured in microns per pixel -- is of more interest for 3D topographic measurements, as it determines the operative level of available data.
 
@@
 
-Case studies were chosen such that theoretical difficulty for the matching algorithm increases with each case study: Hamby set 44 is part of the Hamby study.
+Case studies were chosen such that difficulty for the matching algorithm increases with each case study: Hamby set 44 is part of the Hamby study.
 The algorithm in \citet{aoas2} was trained on sets 173 and 252, so the bullets in Hamby 44 are of the same type of ammunition and are fired through the same barrels as the bullets in the training set.
@@
 These barrels are first rifled traditionally, i.e.
-similar to the Ruger P-85, but are treated with a secondary round of heating after rifling, which may introduce some subclass characteristics.
+similar to the Ruger P-85, but are treated to a secondary round of heating after rifling.
+This second round of heat exposure has the potential to introduce some subclass characteristics, i.e.\ anomalies that are shared between different barrels manufactured in this manner.
 This should make the automatic matching harder, and in particular, is expected to complicate the classification as different-source land engraved areas/bullets.
@@
 Each barrel was test-fired twice, so each of the questioned bullets (shown along the $y$ axis) matches two bullets fired from a known barrel.
-Colored tiles are used to show the similarity score: light grey colors correspond to low similarity scores, dark colors correspond to high similarity scores.
+Colored tiles are used to show the strength of the comparison; light grey colors correspond to low similarity scores, dark colors correspond to high similarity scores.
 Ground truth is encoded in this figure as a thin, dark, colored frame for all pairs of same-source bullets.
@@
 \section{Evaluating the Random Forest Algorithm}
-The random forest scores correctly separate known matches from known non-matches, using SAM scores to aggregate land-to-land scores into a bullet-to-bullet comparison.
+The random forest SAM scores correctly separate known matches from known non-matches, using SAM scores to aggregate land-to-land scores into a bullet-to-bullet comparison.
 With no errors at the bullet-to-bullet level, we can now evaluate the scores in light of (R1) and (R2).
-In this section, we will primarily compare the random forest scores to cross correlation scores; cross correlation is one of the components of the random forest, but has also been used to quantify the similarity of two signatures in its own right \citep{chuPilotStudyAutomated2010}.
-In addition, both scoring methods use the same scale (0 to 1) and are continuous along that interval; other evaluation methods, such as consecutive matching striae, are discrete and more difficult to directly compare to continuous measurements.
+In this section, we will compare the random forest scores to cross correlation scores; cross correlation is one of the components of the random forest, but has also been used to quantify the strength of a match in its own right \citep{chuPilotStudyAutomated2010}.
+In addition, both scoring methods use the same scale (0 to 1) and are continuous along that interval; other evaluation methods, such as consecutive matching striae are discrete and more difficult to directly compare to continuous measurements.
 
@@
 Color indicates ground truth.
-Small vertical lines below the $x$-axis indicate observed scores in a \emph{rug-plot}.
+Small vertical lines below the $x$-axis indicate observed scores in a **rug-plot**.
 The plot shows that for all three case studies and both continuous measures, i.e.\ SAM scores based on cross correlation as well as SAM scores based on the random forest scores, all similarity scores are higher for pairs of bullets from the same source, i.e.\ bullets fired through the same barrel, than pairs of bullets from different sources (fired through different barrels).
-There is considerably more ambiguity in the SAM CMS scores, where the same-source distribution significantly overlaps the different-source distribution.
 
@@
 
-\caption[Density curves of similarity scores from random forest scores (top), cross correlation (middle), and consecutive matching striae (bottom)]{Density curves of similarity scores from random forest scores (top), cross correlation (middle), and consecutive matching striae (bottom).
+\caption[Density curves of similarity scores from cross correlation (top) and RF scores (bottom)]{Density curves of similarity scores from cross correlation (top) and RF scores (bottom).
 Different colors indicate same source versus different source.
@@
 
-According to \autoref{fig:compare}, neither cross correlation nor random forest (RF) score fulfill both of the minimal requirements laid out in the introduction.
+According to \autoref{fig:compare}, neither cross correlation nor random forest score fulfill both of the minimal requirements laid out in the introduction.
 Both measures fulfill (R1), i.e.\ all  scores  of  different-source pairs are lower than scores of same-source pairs for SAM scores based on cross correlation and RF scores.
@@
 
-Regarding requirement (R2), \autoref{fig:compare} shows that a single cutoff value does not exist that separates same-source scores from different-source scores across all three studies and all three scoring measures.
+Regarding requirement (R2), \autoref{fig:compare} shows that a single cutoff value does not exist that separates same-source scores from different-source scores across all three studies without introducing errors.
 
@@
 The downside of either of these options is that we would need to considerably expand the database used for training.
-Another solution would be to augment the aggregation used to get from land-to-land scores to bullet scores: SAM scores only take the scores of the maximum sequence into account and ignore scores associated with pairs of land engraved areas which are off the matching sequence.
+Another solution would be to augment the aggregation used to get from land-to-land scores to bullet scores: SAM scores only take the scores of the maximum sequence into account and ignore scores associated with pairs of land engraved areas from different sources.
 Those scores might be useful in establishing a baseline that better expresses similarity between pairs of bullets.
@@
 There is some overlap between known matching and known non-matching land-to-land scores; in many cases the random forest scores for these overlapping values are more extreme than the corresponding cross correlation scores.
-This suggests that the matching algorithm is sensitive to the presence of LEAs which have low land-to-land similarity scores, such as the matches of bullets 'I' and 'F' in the Hamby 44 study.
+This suggests that the matching algorithm is sensitive to the presence of LEAs which have low land-to-land match scores, such as the matches of bullets 'I' and 'F' in the Hamby 44 study.
 Comparing this to the bullet-to-bullet score density plot shown in \autoref{fig:compare}, we see that there is much greater separation between the same-source and different-source densities for each study in the bullet-to-bullet comparisons than in the land-to-land comparisons, that is, one or two poorly matching lands does not prevent the bullet from showing a matching score, though if there are weak matches on several lands, such as the marked Hamby 44 comparisons in \autoref{fig:compare}, the overall score may be affected.
 
-Several lands in Hamby set 44 have major deficiencies,  such as `tank rash' (a collision of the bullet after exiting the barrel with a surface causing markings on top of the striations from the barrel) or extreme pitting (holes caused by direct contact with burning gun powder).
-\autoref{fig:tankrash} shows rendered scans of all affected lands in the Hamby-44 set.
-These issues affect the algorithm's ability to identify a stable signature, which impacts the subsequent similarity scores at the land-to-land and bullet-to-bullet level.
+Several lands in Hamby set 44 have major deficiencies,  such as 'tank rash' (a collision of the bullet after exiting the barrel with a surface causing markings on top of the striations from the barrel) or extreme pitting (holes caused by direct contact with burning gun powder).
+\autoref{fig:tankrash} gives an overview of all lands of affected bullets in the Hamby-44 set.
+These issues affect the algorithm's ability to identify a stable signature, which impacts the subsequent match scores  at the land-to-land and bullet-to-bullet level.
 \begin{knitrout}
@@
 
-\caption[Density curves of land-to-land similarity scores from RF scores (top), cross-correlation (middle), and consecutive matching striae (CMS)(bottom)]{Density curves of land-to-land similarity scores from RF scores (top), cross-correlation (middle), and consecutive matching striae (CMS)(bottom).
+\caption[Density curves of land-to-land similarity scores from cross correlation (top) and RF scores (bottom)]{Density curves of land-to-land similarity scores from cross correlation (top) and RF scores (bottom).
 Different colors indicate same source versus different source for each land.
@@
 When evaluating classifier performance, it is common to use a receiver operating characteristic (ROC) curve, which plots the percent of wrong identifications against the percent correct identifications for each possible value of the cutoff between the two classes (in this case, known match and known non-match).
-As the land-to-land scores are not perfectly separated, we can use ROC curves to distinguish between the performance of the different methods and studies.
+Land-to-land scores are not perfectly separated, we can therefore use ROC curves to distinguish between the performance of the different methods and studies.
 The ROC curves for the land-to-land scores are shown in \autoref{fig:roc-auc} (a).
@@
 As seen in the ROC curves, the algorithm performs the least well on set 2 of the Houston FSI study.
-Based on AUC, the overall performance on all sets is very good to excellent for both the cross correlation and the random forest score, and moderate for CMS scores.
+Based on AUC, the overall performance on all sets is very good to excellent for cross correlation and the random forest score.
 At the land level there is no significant difference in prediction power between RF score and cross correlation.}\label{fig:roc-auc}
@@
 AUC values are useful for differentiating between poor, good, and excellent model performance, but are not particularly useful when determining which of several models with approximately the same level of performance should be used \citep{marzban_roc_2004}.
-%Hamby 44 and Phoenix PD are the only studies with a somewhat noticeable difference between the AUC score for RF compared to cross correlation, but even those differences are not statistically significant.
+Hamby 44 and Phoenix PD are the only studies with a somewhat noticeable difference between the AUC score for RF compared to cross correlation, but even those differences are not statistically significant.
 
 \autoref{fig:eer} shows an overview of Equal Error and their thresholds based on the ROC curves of \autoref{fig:roc-auc}.
-Equal error rates are the error rates when the sensitivity and specificity of a test are equal, i.e.\ we see the same percentage of missed and wrong identifications for a land-to-land comparison.
+Equal Errors are errors when sensitivity and specificity of a test are equal, i.e.\ we see the same percentage of missed and wrong identifications for a land-to-land comparison.
 Equal errors based on CMS of two lands are significantly higher than equal errors based on cross correlation or the RF score.
 At best, RF score and cross correlation have an equal error of around 5\% for land-to-land comparisons.
-% Interestingly, for consecutively matching striae, equal errors are reached for just two matching striae, which from a practical point of view is not sustainable.
+Interestingly, for consecutively matching striae, equal errors are reached for just two matching striae, which from a practical point of view is not sustainable.
 A  CMS of  6 or above is an established cutoff used in identifications; when applied to the land-to-land comparisons, it is much more likely to result in a missed identification than a wrong identification.
@@
 
-{\centering \includegraphics[width=\textwidth]{figures/eer-1}
+{\centering \includegraphics[width=.8\textwidth]{figures/eer-1}
 
@@
 
-\caption[Equal error (in Percent)  and corresponding thresholds for all measures (random forest score, cross correlation and CMS) under all five (sub-)studies]{Equal error (in Percent)  and corresponding thresholds for all measures (random forest score, cross correlation and CMS) under all five (sub-)studies.
-CMS has an equal error threshold of 2 for all studies, and perform signifcantly worse than both cross correlation and the random forest score.
-At best, Cross correlation and RF score have equal errors for land-to-land comparisons of around 5\%.
-The performance of the random forest score is slightly better than the cross-correlation on sets which have higher equal error rates (Houston FSI G2, Hamby 44).}\label{fig:eer}
+\caption[Equal Error (in Percent)  and corresponding thresholds for all measures (RF score, Cross correlation and CMS) under all five (sub-)studies]{Equal Error (in Percent)  and corresponding thresholds for all measures (RF score, Cross correlation and CMS) under all five (sub-)studies.
+CMS has an equal error threshold of 2 for all studies.
+CMS is signifcantly worse than both Cross correlation and the RF score.
+At best, Cross correlation and RF score have equal errors for land-to-land comparison of around 5\%.}\label{fig:eer}
 \end{figure}
@@
 
-At the beginning of this study, we anticipated Hamby 44 would be the easiest set to evaluate because of its similarity to the sets used to train the random forest algorithm.
-In a surprise turn of events, it was the hardest, in part because of damage to the bullets that obscured LEA striae.
-It has been shown \citet{lpr2} that parts of lands can be used for successful identifications, if at least 50\% of the land is present (using full-length scores as the reference distribution).
-The random forest algorithm proposed in \citet{aoas2} is not capable of automatically detecting parts of lands with well-expressed striae.
+At the beginning of this study, we anticipated Hamby 44 would be the easiest set to work with because of its similarity to the sets used to train the random forest algorithm.
+In a surprise turn of events, it was the hardest to solve, in part because of damage to the bullets that obscured LEA striae.
+It has been shown \citet{lpr2} that parts of lands can be used for successful identifications.
+The algorithm proposed by \citet{aoas2} is not capable of  automatically detecting parts of lands with well-expressed striae.
 It may be useful to couple the \citet{aoas2} algorithm with an algorithm which assesses the quality of the input data and determines which portions of the data to use for comparison.
-This would emulate the process used by examiners, who first assess the quality of the evidence and whether there is enough information present to attempt a comparison.
+This would emulate the process used by examiners, who first assess the quality of the evidence and whether there is enough information present to attempt a match.
 One major disadvantage of an automated algorithm is that all of the decisions humans make (recognizing degraded land areas, excluding those areas from consideration, matching only the remaining areas) must be explicitly characterized; however, this explicit characterization means that the process can be scientifically validated to a much higher degree than the human perceptual process.
@@
 Rugers are among the firearms that mark very well, so we would correspondingly expect to see lowered similarity scores for other firearms.
-While the performance of cross-correlation and random forest scores are similar when evaluated on Rugers, it is likely that the additional information used to generate the random forest score may be useful when comparing bullets from barrels which produce fewer distinctive marks.
-A future study will compare the performance of the random forest score, cross-correlation score, and CMS on non-Ruger barrels.
-The random forest algorithm is trained on scans taken at the land level, so barrels with polygonal rifling, such as Glocks, which do not introduce well defined lands on bullets, cannot be compared using the random forest algorithm examined here.
-In addition, we plan to re-fit the random forest using a wider set of scans taken at different resolutions, with different ammunition, fired from barrels of different models.
-We expect that a random forest trained on a wider set of input data will be more likely to produce scores which meet the criteria for monotonicity and stability.
+The matching algorithm works on matching lands.
+Barrels with polygonal rifling, such as Glocks, do not introduce well defined lands on bullets, which makes an application of the matching algorithm impossible.
 
@@
 For a future round of the Houston study, we are planning to compare the algorithmic performance directly with scores given by forensic toolmark examiners.
-The algorithm's score is not intended to replace an examiner's judgment; instead, it provides a complementary tool to measure, quantify, and compare the similarity of two bullets in an identification.
-With more research into the behavior of score-based likelihood ratios, the random forest score may also be used to compute a SLR to explicitly quantify the strength of the match between two bullets relative to other matches in the population.
+The algorithm's score is not intended to replace an examiner's judgment; instead, it provides a complementary tool to measure and quantify the strength of an identification.
 By validating the algorithm on external test sets, we have demonstrated that the method can be generalized to different types of ammunition and is not overly sensitive to small differences in rifling procedure.

### 2020-02-21 index.Rnw
Fix citations

@@
 Cross correlation varies between -1 and 1, but as the alignment is based on the maximized cross correlation value, in practice the cross correlations for aligned signatures are generally positive.
-Cross correlation values which are close to 1 indicate similarity between signatures, and may consequentially indicate that the two signatures come from the same source.
-Low cross correlation values indicate signatures which may originate from different sources.
+Cross correlation values which are close to 1 indicate similarity between signatures, with the possibility that the two signatures come from the same source.
+Low cross correlation values indicate signatures which are different and thus may originate from different sources.
 
 The random forest presented in \citet{aoas2} is based on a combination of multiple characteristics, including cross correlation, number of matching striae, and number of consecutively matching striae.
-The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the similarity of the two signatures, where scores close to 1 indicate it to be highly likely that the aligned signatures come from the same source, while scores close to 0 indicate that the aligned signatures are highly likely to originate from different sources.
+The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the similarity of the two signatures, where scores close to 1 indicate signatures which are similar and may have originated from the same source, while scores close to 0 indicate that the aligned signatures are different and may originate from different sources.
 
